### 2023-10-13 Demo_marrow_facs.Rmd
Demo_marrow_facs

---
title: "Tutorial for Runing Mouse Bone Marrow Dataset"
author: "Ruiqi Li"
date: '2023-10-12'
output: pdf_document
bibliography: references.bib
---

# Install & Load Packages
```{r}
# Install necessary tools if not already present
if (!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")
if (!requireNamespace("remotes", quietly = TRUE)) install.packages("remotes")
# Define package lists
bioc_packages <- c("Seurat")
github_packages <- c("satijalab/SeuratWrappers")

# Install packages
sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
sapply(github_packages, function(pkg) if (!requireNamespace(strsplit(pkg, "/")[[1]][2], quietly = TRUE)) remotes::install_github(pkg))

library("Seurat")
library("SeuratWrappers")
```

# Preparing input data
```{r}
# If load url doesn't work, please download it and load from your working directory
load(url("https://figshare.com/ndownloader/files/13092380"))
tiss <- UpdateSeuratObject(tiss)
```
We obtain the following objects from this Seurat Obj:
* `feature_space`: The output of PCA. This matrix contains the first 20 pricipal components (PCs).
* `visual_space`: a data.frame with t-SNE coordinates (2D) based on the first 20 PCs.
* `dat`: a matrix object with the log-normalized expression of genes (rows) in each cell (columns)
* `cell_label`: Any metadata you wish to visualize and examine related to cells (here we take the cell type annotation)
```{r}
DefaultAssay(tiss) <- "RNA"
n_dim = dim(tiss@reductions$pca@cell.embeddings)[2]
feature_space = as.matrix(tiss@reductions$pca@cell.embeddings[,1:n_dim])
visual_space = data.frame(tiss@reductions$tsne@cell.embeddings)
dat = as.matrix(tiss[[DefaultAssay(tiss)]]@data)
cell_label = tiss$cell_ontology_class
```
We process the `dat` by only keep the genes detected in more than 10 cells and less than 50% of cells (genes is "detected" in a cell only if its expression level is greater than the median expression level of that cell)
```{r}
Gene_detected_count <- apply(dat > apply(dat,2,median),1,sum)
selected_genes = (Gene_detected_count >= 10) & (Gene_detected_count <= ncol(dat) * 0.5)
dat = dat[selected_genes,,drop = FALSE]
```

# Running LMD Step by Step
This provides a step-by-step tutorial for a better understanding of the algorithm. If you want to generate results directly, please skip to 'Running LMD in One Step'
```{r}
source("./LMD_function.R", echo = F)
```
## Step1: Build KNN Graph
```{r}
# Construct knn graph
W = Symmetric_KNN_graph(knn = 5, feature_space = feature_space)[[1]]

# Plot knn graph
plot_knn_graph(affinity_m = W, label = cell_label, layout = visual_space)
```

## Step2: Diffuse Genes on the cell graph
```{r}
# Generate $P^{t}, t = 2,4,8,...,max_time$
P_ls = Obtain_Pls(W = W, max_time = 2^10)
# Generate initial gene distribution
rho = Rowwise_normalize(dat[,colnames(W)])
# Check One Gene Diffusion on Cell graph by time
gene = "Fcnb"
print(FeaturePlot_diffusion(coord = visual_space,init_state = rho[gene,],P_ls = P_ls,check_time = c(2,16,128),gene_name = gene))
```

## Step3: Obtain Diffusion KL score & Cumulative score
```{r}
# Obtain diffusion KL score
res = fast_calculate_multi_score(W = W, init_state = rho, P_ls = P_ls)

# Check Diffusion KL pattern for several genes
genes = c("Tlr11","Fcnb","Cd79a","4933425H06Rik","Uchl4","Dock11")
genes_label = rep(c("good","bad"),each = 3)
Visualize_score_pattern(res, genes = genes, label_class = genes_label, facet_class = NULL)

# Show result
show_result_lmd(res,n = 10)
```

# Running LMD in one step
```{r}
res = LMD(dat,feature_space)
```

# Identify Gene Modules
## Step1: Deriving Gene Modules
```{r}
# Select top1000 genes as local genes
local_gene = show_result_lmd(res, n = 1000)
local_gene = rownames(local_gene)

# Denoising
#' We use ALRA @linderman2022zero to Denoise our data
tiss <- RunALRA(tiss, assay = "RNA")
DefaultAssay(tiss) <- "RNA"
dat_alra = as.matrix(tiss[["alra"]]@data)[local_gene,]

# Calculate Gene-gene co-localization distance
distance_method = "jaccard"
dist = Calculate_distance(dat_alra, method = distance_method)

# Hierarchical Tree
clustering_method = "average"
gene_hree = hclust(dist, method = clustering_method)

# Obtain Gene modules
gene_partition = dynamicTreeCut::cutreeDynamic(
  dendro = gene_hree, 
  distM = as.matrix(dist), deepSplit = 2,
  pamStage = TRUE, pamRespectsDendro = TRUE,
  minClusterSize = 5)
names(gene_partition) = labels(dist)

# Re-name gene modules based on the order of htree
gene_partition = as.factor(gene_partition)
module_order = order(sapply(levels(gene_partition), function(mod) {
  median(which(gene_partition[gene_hree$order] == mod))
}))
levels(gene_partition)[module_order] = seq(nlevels(gene_partition))
gene_partition = factor(gene_partition,levels = seq(nlevels(gene_partition)))
rm(module_order)
```
### Visualize Heatmap
```{r}
p = pheatmap(as.matrix(dist),
              annotation_col = data.frame(Module = gene_partition), 
              cluster_rows = gene_hree, 
              cluster_cols = gene_hree,
              ann_colors = list(Module = setNames(colorRampPalette(brewer.pal(12, "Paired"))(nlevels(gene_partition)),levels(gene_partition))),
              treeheight_row = 0,
              col=colorRampPalette(c("red","white","blue4"))(99),
              show_colnames = FALSE, 
              show_rownames = FALSE,
              main = paste(distance_method,
                           clustering_method,sep = "-"),
              silent = TRUE )
as.ggplot(p)
```

## Step2: Computing Per-Cell Module Activity Scores
```{r}
knn = 10; major_vote = 5
A = Symmetric_KNN_graph(knn = knn, feature_space = feature_space)[[2]]
cell_block_loop = lapply(1:100,function(loop){
  cat("\n",loop,"\n")
  GMM_subsampling(seed = loop, gene_partition, expr_dat = dat, cell_kNN_graph = A, major_vote)
})
cell_block_prop = Reduce(`+`, cell_block_loop) / length(cell_block_loop)
cell_block = cell_block_prop
```
### Visualize Module on Cell embeddings
```{r}
pl = lapply(levels(gene_partition),function(module){
  FeaturePlot_custom(value = cell_block[,module], coord = visual_space, value_name = "Module Score",title_name = sprintf("Module %s (%d genes)",module, sum(gene_partition == module)))
})
plot_grid(plotlist = pl,ncol = 10) # may takes long time to plot
```
## Step3: Identify Celltype-specific activated Modules
```{r}
# Create one-hot vector of Cell Type
one_hot_matrix <- sapply(unique(cell_label), function(x) as.integer(cell_label == x))

# Order Cell Type based on its centroid in Featurespace
Idents(tiss) = cell_label
ordered_cell_types <- unique(cell_label)[hclust(dist(t(sapply(unique(cell_label), function(x) colMeans(Embeddings(tiss, reduction = "pca")[WhichCells(tiss, idents = x), ,drop = FALSE]) ))))$order]

# Calculate Jaccard index between module activity & cell type
jaccard_index <- apply(cell_block, 2, function(module_col) {
  apply(one_hot_matrix, 2, function(one_hot_col) {
    1 - philentropy::distance(rbind(module_col,one_hot_col), method = "jaccard")
  })
})
```

### Visualize Heatmap of Jaccard Index Matrix
```{r}
pheatmap(jaccard_index[rev(ordered_cell_types),], 
         cluster_rows=FALSE, cluster_cols=FALSE, 
         scale = "none", 
         color = colorRampPalette(rev(c("firebrick3", "white", "navy")))(100))
```

### Visualize How CellType-Module is Matched on Cell Embeddings
For each cell type, select the most specific modules for visualization (maximum jaccard index)
```{r}
tiss@meta.data = cbind(tiss@meta.data,one_hot_matrix)
jaccard_thred = 0.4

# Get Module plot
pl_module = lapply(levels(gene_partition),function(module){
  FeaturePlot_custom(value = cell_block[,module], coord = visual_space, value_name = "Module Score",title_name = sprintf("Module %s (%d genes)",module, sum(gene_partition == module)))
})
names(pl_module) = levels(gene_partition)

# Get CellType One-hot plot
pl_ct = lapply(colnames(one_hot_matrix),function(ct){
  FeaturePlot_custom(value = as.factor(one_hot_matrix[,ct]), coord = visual_space, value_name = "Indicator",title_name = ct)
})
names(pl_ct) = colnames(one_hot_matrix)

pl <- lapply(rownames(jaccard_index)[rev(order(apply(jaccard_index,1,max)))], function(celltype){
  if(max(jaccard_index[celltype,]) > jaccard_thred){
    p1 = pl_ct[[celltype]]
    module_name = colnames(jaccard_index)[which.max(jaccard_index[celltype,])]
    p2 = pl_module[[module_name]]
    p = as.ggplot(p1) + as.ggplot(p2) + plot_annotation(paste0("Jaccard Index: ",round(jaccard_index[celltype,module_name],2)))
    p
  }
})
```
### Annotate Each Celltype-specific activated Modules with its corresponding Celltypes
```{r}
module_celltype = apply(jaccard_index,2,function(x){paste(names(which(x > 0.4)),collapse = ",")})
```

## Step4: Module Enrichment Analysis
### GO enrichment Analysis
```{r}
if (!requireNamespace("clusterProfiler", quietly = TRUE)) BiocManager::install("clusterProfiler")
library(clusterProfiler)

# Set all genes(after filtering) in the expression matrix as background,
# such that the enrichment result of any module is not attributed to its high expression levels
universe_df = data.frame("symbol" = rownames(dat),"entrez" = mapIds(org.Mm.eg.db, keys=rownames(dat), column="ENTREZID", keytype="SYMBOL"))

# Filter GO terms based on p values < 0.05
ego_result <- lapply(1:nlevels(gene_partition), function(i){
  enrichGO(gene = names(gene_partition)[gene_partition == i],
           OrgDb = 'org.Mm.eg.db', # mouse
           keyType = "SYMBOL",
           ont = "ALL",
           pAdjustMethod = "BH",
           universe = universe_df$symbol,
           pvalueCutoff = 0.05)
})

# Only keep modules which has at lease one GO term having adjusted p-value < 0.05
top_go_clusters <- which(sapply(ego_result, function(x)
  (x@result$p.adjust[1] < 0.05)))

# Annotate each modules with top5 GO terms
top_enrich_go <- lapply(top_go_clusters, function(i) ego_result[[i]]@result[1:5,])
```

### Reactome enrichment Analysis
```{r}
if (!requireNamespace("ReactomePA", quietly = TRUE)) BiocManager::install("ReactomePA")
if (!requireNamespace("org.Mm.eg.db", quietly = TRUE)) BiocManager::install("org.Mm.eg.db")
library("clusterProfiler")
library("ReactomePA")
library("org.Mm.eg.db")

# Set all genes(after filtering) in the expression matrix as background,
# such that the enrichment result of any module is not attributed to its high expression levels
universe_df = data.frame("symbol" = rownames(dat),"entrez" = mapIds(org.Mm.eg.db, keys=rownames(dat), column="ENTREZID", keytype="SYMBOL"))
universe_df = universe_df[!is.na(universe_df$entrez),]

# Filter Pathway based on p values < 0.05
epathway_result <- lapply(1:nlevels(gene_partition), function(i){
  enrichPathway(gene=universe_df[universe_df$symbol %in% names(gene_partition)[gene_partition == i],"entrez"],
                organism = "mouse",
                pAdjustMethod = "BH",
                pvalueCutoff = 0.05,
                universe = universe_df$entrez)
})

# Only keep modules which has at lease one Pathway having adjusted p-value < 0.05
top_pathway_clusters <- which(sapply(epathway_result, function(x) 
  (x@result$p.adjust[1] < 0.05)))

# Annotate each modules with top5 Pathway terms
top_enrich_pathway <- lapply(top_pathway_clusters, function(i) epathway_result[[i]]@result[1:5,])
```

## Step5: Generate Gene Module Annotation Table
```{r}
df_module = data.frame(Module = levels(gene_partition))
df_module[,"Gene"] = unlist(lapply(levels(gene_partition),function(x){paste(names(gene_partition)[gene_partition == x],collapse = ",")}))
df_module[,"celltype"] = module_celltype[levels(gene_partition)]
df_module[top_go_clusters,"GO_description"] = unlist(lapply(top_enrich_go,function(x) paste(x[['Description']],collapse = ",")))
df_module[top_pathway_clusters,"Pathway_description"] = unlist(lapply(top_enrich_pathway,function(x) paste(x[['Description']],collapse = ",")))
df_module[is.na(df_module)] = ""
write.csv(df_module,file = "module_description.csv",row.names = FALSE)
```

### 2023-10-13 LMD_function.R

@@ -1,0 +1,440 @@
+cran_packages <- c("igraph","ggplot2", "cowplot", "RColorBrewer", 
+                   "data.table", "dplyr", "patchwork", 
+                   "pheatmap", "ggplotify", "ggraph", 
+                   "ClusterR","Rcpp","RcppArmadillo")
+sapply(cran_packages, function(pkg) if(!requireNamespace(pkg, quietly = TRUE)){install.packages(pkg)})
+lapply(cran_packages, require, character.only = TRUE)
+
+# Mtx operation===========
+# Enable C++11
+Sys.setenv("PKG_CXXFLAGS"="-std=c++11")
+# Create the C++ function
+cppFunction('arma::mat fastMatMult(arma::mat A, arma::mat B) {
+  arma::mat C = A * B;
+  return C;
+}', depends="RcppArmadillo")
+
+
+Rcpp::sourceCpp(code='
+#include <Rcpp.h>
+using namespace Rcpp;
+
+NumericVector calcKL(NumericVector p, NumericVector q) {
+  double sum = 0;
+  for (int j = 0; j < p.size(); ++j) {
+    if(p[j] == 0) continue;
+    double val = p[j] * log(p[j] / q[j]);
+    if (NumericVector::is_na(val) || std::isinf(val)) val = 0.0;
+    sum += val;
+  }
+  return NumericVector::create(sum);
+}
+
+// [[Rcpp::export]]
+NumericVector fastKLMatrix(NumericMatrix x, NumericMatrix y) {
+  int n = x.nrow();
+  NumericVector out(n);
+
+  for (int i = 0; i < n; ++i) {
+    out[i] = calcKL(x.row(i), y.row(i))[0];
+  }
+  return out;
+}
+
+// [[Rcpp::export]]
+NumericVector fastKLVector(NumericMatrix x, NumericVector y) {
+  int n = x.nrow();
+  NumericVector out(n);
+
+  for (int i = 0; i < n; ++i) {
+    out[i] = calcKL(x.row(i), y)[0];
+  }
+  return out;
+}
+')
+Rowwise_normalize <- function(x){
+  return( sweep(x, 1, rowSums(x), FUN = "/") )
+}
+
+# Doubly Stochastic
+l2_norm = function(x) sqrt(sum(x^2))
+sinkhorn_knopp = function(A, sums = rep(1, nrow(A)),
[…]

### 2023-10-13 benchmark.R

@@ -1,0 +1,1 @@
+package_ls = "singleCellHaystack"

### 2023-10-13 paper_figure.R

@@ -1,0 +1,57 @@
+# Sensitivity Test----------
+## Rank Stability-----------
+score_df = lapply(c(3:10,20,30,40,50),function(kNN){
+  try({
+    W = Symmetric_KNN_graph(knn = kNN, feature_space = feature_space)[[1]]
+    rho = Rowwise_normalize(dat[,colnames(W)])
+    marker = fast_calculate_multi_score(W = W, init_state = rho)
+    
+    sub_score0 = grep("score0",colnames(marker))
+    sub_score1 = grep("score1_0",colnames(marker))
+    accu_score1 = apply(marker,1,function(x){
+      sum(x[sub_score0]/x[sub_score1])
+    })
+    accu_score1
+  },silent = TRUE)
+})
+names(score_df) = paste0("kNN",c(3:10,20,30,40,50))
+score_df = do.call(cbind,score_df)
+score_ranked <- data.frame(score_df) %>% mutate(across(everything(), rank))
+base = "kNN5"
+top_genes_list <- lapply(c(50,seq(100,1000,100)), 
+                         function(x) head(score_ranked[order(score_ranked[,base]),], x))
+average_ranks <- lapply(top_genes_list, function(top_genes) {
+  colMeans(top_genes)
+})
+average_ranks <- data.frame(
+  Top_Genes = c(50,seq(100,1000,100)),
+  do.call(rbind, average_ranks)
+)
+average_ranks <- average_ranks %>% tidyr::pivot_longer(cols = -Top_Genes, names_to = "kNN", values_to = "Average_Rank")
+average_ranks$kNN = as.numeric(gsub("kNN","",average_ranks$kNN))
+ggplot(data = average_ranks, aes(x = kNN, y = Average_Rank, colour = factor(Top_Genes), group = Top_Genes)) +
+  geom_line() +
+  geom_point() +
+  xlab("Top Genes") +
+  ylab(sprintf("Average Rank of Top # Genes\n (base on %s)",base)) +
+  theme_minimal() + facet_wrap(~Top_Genes,scales = "free")
+
+## AUC stability------------
+gt_genes_list <- c(lapply(c(50,seq(100,400,100)), function(x){
+  names(max_logfc)[1:x]
+}),list(true_marker))
+auc_df = apply(score_ranked,2,function(x){
+  unlist(lapply(gt_genes_list,function(genes){
[…]

### 2023-10-13 Demo_marrow_facs.Rmd
adjacency_mtx

@@
 # Construct knn graph
-W = Symmetric_KNN_graph(knn = 5, feature_space = feature_space)[[1]]
+res = Symmetric_KNN_graph(knn = 5, feature_space = feature_space)
+
+A = res$adj_matrix # Adjacency Matrix
+W = res$graph # Symmetrized Graph ((A + AT) / 2)
 

### 2023-10-13 Demo_marrow_facs.Rmd
added AnnotationDbi, moved requirements to top

@@
 # Install & Load Packages
+
 ```{r}
@@
 # Define package lists
-bioc_packages <- c("Seurat")
-github_packages <- c("satijalab/SeuratWrappers")
+bioc_packages <- c("Seurat",
+                   "clusterProfiler", "AnnotationDbi", "ReactomePA", "org.Mm.eg.db")
+github_packages <- c("satijalab/seurat-wrappers")
+# Note: repeating packages from LMD_function.R as knitr does not install them properly
+cran_packages <- c("igraph","ggplot2", "cowplot", "RColorBrewer",
+                   "data.table", "dplyr", "patchwork",
+                   "pheatmap", "ggplotify", "ggraph",
+                   "ClusterR", "Rcpp", "RcppArmadillo", "tictoc")
 
@@
 sapply(github_packages, function(pkg) if (!requireNamespace(strsplit(pkg, "/")[[1]][2], quietly = TRUE)) remotes::install_github(pkg))
+sapply(cran_packages, function(pkg) if(!requireNamespace(pkg, quietly = TRUE)){install.packages(pkg)})
 
@@
 # Preparing input data
+
 ```{r}
 # If load url doesn't work, please download it and load from your working directory
-load(url("https://figshare.com/ndownloader/files/13092380"))
+#load(url("https://figshare.com/ndownloader/files/13092380"))
+load('facs_Marrow_seurat_tiss.Robj')
 tiss <- UpdateSeuratObject(tiss)
@@
 ```
+
 We process the `dat` by only keep the genes detected in more than 10 cells and less than 50% of cells (genes is "detected" in a cell only if its expression level is greater than the median expression level of that cell)
+
 ```{r}
@@
 # Running LMD Step by Step
+
 This provides a step-by-step tutorial for a better understanding of the algorithm. If you want to generate results directly, please skip to 'Running LMD in One Step'
+
 ```{r}
@@
 ```
+
 ## Step1: Build KNN Graph
+
 ```{r}
@@
 ## Step2: Diffuse Genes on the cell graph
+
 ```{r}
@@
 ## Step3: Obtain Diffusion KL score & Cumulative score
+
 ```{r}
@@
 # Running LMD in one step
+
 ```{r}
@@
 # Identify Gene Modules
+
 ## Step1: Deriving Gene Modules
+
 ```{r}
@@
 ```
+
 ### Visualize Heatmap
+
 ```{r}
@@
 ## Step2: Computing Per-Cell Module Activity Scores
+
 ```{r}
@@
 ```
+
 ### Visualize Module on Cell embeddings
+
 ```{r}
@@
 ```
+
 ## Step3: Identify Celltype-specific activated Modules
+
 ```{r}
@@
 ### Visualize Heatmap of Jaccard Index Matrix
+
 ```{r}
@@
 ### Visualize How CellType-Module is Matched on Cell Embeddings
+
 For each cell type, select the most specific modules for visualization (maximum jaccard index)
+
 ```{r}
@@
 ```
+
 ### Annotate Each Celltype-specific activated Modules with its corresponding Celltypes
+
 ```{r}
@@
 ## Step4: Module Enrichment Analysis
+
 ### GO enrichment Analysis
+
 ```{r}
-if (!requireNamespace("clusterProfiler", quietly = TRUE)) BiocManager::install("clusterProfiler")
-library(clusterProfiler)
+library("clusterProfiler")
+library("ReactomePA")
+library("org.Mm.eg.db")
+library("AnnotationDbi")
 
@@
 ### Reactome enrichment Analysis
+
 ```{r}
-if (!requireNamespace("ReactomePA", quietly = TRUE)) BiocManager::install("ReactomePA")
-if (!requireNamespace("org.Mm.eg.db", quietly = TRUE)) BiocManager::install("org.Mm.eg.db")
-library("clusterProfiler")
 library("ReactomePA")
@@
 ## Step5: Generate Gene Module Annotation Table
+
 ```{r}

### 2023-10-15 Demo_marrow_facs.Rmd
file download + moved imports

@@
 # Install packages
+sapply(cran_packages, function(pkg) if(!requireNamespace(pkg, quietly = TRUE)){install.packages(pkg)})
 sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
 sapply(github_packages, function(pkg) if (!requireNamespace(strsplit(pkg, "/")[[1]][2], quietly = TRUE)) remotes::install_github(pkg))
-sapply(cran_packages, function(pkg) if(!requireNamespace(pkg, quietly = TRUE)){install.packages(pkg)})
 
@@
 ```{r}
+if (!file.exists('facs_Marrow_seurat_tiss.Robj')) {
    # If load url doesn't work, please download it and load from your working directory
-#load(url("https://figshare.com/ndownloader/files/13092380"))
+  options(timeout=6000) 
+  download.file("https://figshare.com/ndownloader/files/13092380", 
+                destfile = 'facs_Marrow_seurat_tiss.Robj', method = 'libcurl')
+ }
 load('facs_Marrow_seurat_tiss.Robj')
@@
 We obtain the following objects from this Seurat Obj:
-* `feature_space`: The output of PCA. This matrix contains the first 20 pricipal components (PCs).
+* `feature_space`: The output of PCA. This matrix contains the first 20 principal components (PCs).
 * `visual_space`: a data.frame with t-SNE coordinates (2D) based on the first 20 PCs.

### 2023-10-16 Demo_marrow_facs.Rmd
removed functions that print a lot

@@
 cell_block_loop = lapply(1:100,function(loop){
-  cat("\n",loop,"\n")
   GMM_subsampling(seed = loop, gene_partition, expr_dat = dat, cell_kNN_graph = A, major_vote)
@@
   apply(one_hot_matrix, 2, function(one_hot_col) {
-    1 - philentropy::distance(rbind(module_col,one_hot_col), method = "jaccard")
+    1 - philentropy::distance(rbind(module_col,one_hot_col), method = "jaccard", mute.message=TRUE)
   })

### 2023-10-16 Demo_marrow_facs.Rmd
changed variable names for results, removed middle color for heatmaps

@@
 # Construct knn graph
-res = Symmetric_KNN_graph(knn = 5, feature_space = feature_space)
+knn_result = Symmetric_KNN_graph(knn = 5, feature_space = feature_space)
 
-A = res$adj_matrix # Adjacency Matrix
-W = res$graph # Symmetrized Graph ((A + AT) / 2)
+A = knn_result$adj_matrix # Adjacency Matrix
+W = knn_result$graph # Symmetrized Graph ((A + AT) / 2)
 
@@
 # Obtain diffusion KL score
-res = fast_calculate_multi_score(W = W, init_state = rho, P_ls = P_ls)
+score_result = fast_calculate_multi_score(W = W, init_state = rho, P_ls = P_ls)
 
@@
 genes_label = rep(c("good","bad"),each = 3)
-Visualize_score_pattern(res, genes = genes, label_class = genes_label, facet_class = NULL)
+Visualize_score_pattern(score_result, genes = genes, label_class = genes_label, facet_class = NULL)
 
 # Show result
-show_result_lmd(res,n = 10)
+show_result_lmd(score_result,n = 10)
 ```
@@
 ```{r}
-res = LMD(dat,feature_space)
+lmd_result = LMD(dat, feature_space)
 ```
@@
 # Select top1000 genes as local genes
-local_gene = show_result_lmd(res, n = 1000)
+local_gene = show_result_lmd(lmd_result, n = 1000)
 local_gene = rownames(local_gene)
@@
               treeheight_row = 0,
-              col=colorRampPalette(c("red","white","blue4"))(99),
+              col=colorRampPalette(c("red","blue4"))(99),
               show_colnames = FALSE, 

### 2023-10-17 Demo_marrow_facs.Rmd
added figwidth

@@
 
-```{r}
+```{r, fig.width=20}
 # Generate $P^{t}, t = 2,4,8,...,max_time$
@@
 
-```{r}
+```{r, fig.width=40}
 pl = lapply(levels(gene_partition),function(module){

### 2023-10-18 Demo_marrow_facs.Rmd
benchmark&figure_code

--- 2023-10-16 Demo_marrow_facs.Rmd
@@
 # Install & Load Packages
-
 ```{r}
@@
 # Define package lists
-bioc_packages <- c("Seurat",
-                   "clusterProfiler", "AnnotationDbi", "ReactomePA", "org.Mm.eg.db")
-github_packages <- c("satijalab/seurat-wrappers")
-# Note: repeating packages from LMD_function.R as knitr does not install them properly
-cran_packages <- c("igraph","ggplot2", "cowplot", "RColorBrewer",
-                   "data.table", "dplyr", "patchwork",
-                   "pheatmap", "ggplotify", "ggraph",
-                   "ClusterR", "Rcpp", "RcppArmadillo", "tictoc")
+bioc_packages <- c("Seurat")
+github_packages <- c("satijalab/SeuratWrappers")
 
 # Install packages
-sapply(cran_packages, function(pkg) if(!requireNamespace(pkg, quietly = TRUE)){install.packages(pkg)})
 sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
@@
 # Preparing input data
-
 ```{r}
-if (!file.exists('facs_Marrow_seurat_tiss.Robj')) {
 # If load url doesn't work, please download it and load from your working directory
-  options(timeout=6000) 
-  download.file("https://figshare.com/ndownloader/files/13092380", 
-                destfile = 'facs_Marrow_seurat_tiss.Robj', method = 'libcurl')
- }
-load('facs_Marrow_seurat_tiss.Robj')
+load(url("https://figshare.com/ndownloader/files/13092380"))
 tiss <- UpdateSeuratObject(tiss)
@@
 We obtain the following objects from this Seurat Obj:
-* `feature_space`: The output of PCA. This matrix contains the first 20 principal components (PCs).
+* `feature_space`: The output of PCA. This matrix contains the first 20 pricipal components (PCs).
 * `visual_space`: a data.frame with t-SNE coordinates (2D) based on the first 20 PCs.
@@
 ```
-
 We process the `dat` by only keep the genes detected in more than 10 cells and less than 50% of cells (genes is "detected" in a cell only if its expression level is greater than the median expression level of that cell)
-
 ```{r}
@@
 # Running LMD Step by Step
-
 This provides a step-by-step tutorial for a better understanding of the algorithm. If you want to generate results directly, please skip to 'Running LMD in One Step'
-
 ```{r}
@@
 ```
-
 ## Step1: Build KNN Graph
-
 ```{r}
@@
 ## Step2: Diffuse Genes on the cell graph
-
 ```{r}
@@
 ## Step3: Obtain Diffusion KL score & Cumulative score
-
 ```{r}
@@
 # Running LMD in one step
-
 ```{r}
@@
 # Identify Gene Modules
-
 ## Step1: Deriving Gene Modules
-
 ```{r}
@@
 ```
-
 ### Visualize Heatmap
-
 ```{r}
@@
 ## Step2: Computing Per-Cell Module Activity Scores
-
 ```{r}
 knn = 10; major_vote = 5
-A = Symmetric_KNN_graph(knn = knn, feature_space = feature_space)[[2]]
+A = Symmetric_KNN_graph(knn = knn, feature_space = feature_space)$'adj_matrix'
 cell_block_loop = lapply(1:100,function(loop){
+  cat("\n",loop,"\n")
   GMM_subsampling(seed = loop, gene_partition, expr_dat = dat, cell_kNN_graph = A, major_vote)
@@
 ```
-
 ### Visualize Module on Cell embeddings
-
 ```{r}
@@
 ## Step3: Identify Celltype-specific activated Modules
-
 ```{r}
@@
   apply(one_hot_matrix, 2, function(one_hot_col) {
-    1 - philentropy::distance(rbind(module_col,one_hot_col), method = "jaccard", mute.message=TRUE)
+    1 - philentropy::distance(rbind(module_col,one_hot_col), method = "jaccard")
   })
 })
+jaccard_index = jaccard_index[rev(ordered_cell_types),]
+rm(ordered_cell_types)
 ```
@@
 ### Visualize Heatmap of Jaccard Index Matrix
-
 ```{r}
-pheatmap(jaccard_index[rev(ordered_cell_types),], 
+pheatmap(jaccard_index, 
          cluster_rows=FALSE, cluster_cols=FALSE, 
@@
 ### Visualize How CellType-Module is Matched on Cell Embeddings
-
 For each cell type, select the most specific modules for visualization (maximum jaccard index)
-
 ```{r}
@@
 ```
-
 ### Annotate Each Celltype-specific activated Modules with its corresponding Celltypes
-
 ```{r}
@@
 
+## Save Intermediate data
+```{r}
+save(gene_partition,dist,gene_hree,cell_block,jaccard_index,file = "./marrow_result/module_result.rdata")
+```
+
 ## Step4: Module Enrichment Analysis
-
 ### GO enrichment Analysis
-
 ```{r}
-library("clusterProfiler")
-library("ReactomePA")
-library("org.Mm.eg.db")
-library("AnnotationDbi")
+if (!requireNamespace("clusterProfiler", quietly = TRUE)) BiocManager::install("clusterProfiler")
+library(clusterProfiler)
 
@@
 ### Reactome enrichment Analysis
-
 ```{r}
+if (!requireNamespace("ReactomePA", quietly = TRUE)) BiocManager::install("ReactomePA")
+if (!requireNamespace("org.Mm.eg.db", quietly = TRUE)) BiocManager::install("org.Mm.eg.db")
+library("clusterProfiler")
 library("ReactomePA")
@@
 ## Step5: Generate Gene Module Annotation Table
-
 ```{r}

### 2023-10-18 Marker_identfication_benchmark.Rmd
benchmark&figure_code

---
title: "Marker_identification_Benchmark"
output: pdf_document
date: '2023-10-16'
---
# Load Packages
```{r}
if (!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")
if (!requireNamespace("BiocManager", quietly = TRUE)) BiocManager::install("Seurat")
if (!requireNamespace("readxl", quietly = TRUE)) install.packages("readxl")
if (!requireNamespace("dplyr", quietly = TRUE)) install.packages("dplyr")

library("Seurat")
library("dplyr")
```

# Load Data & Process
## Load Data
```{r}
path <- "/data/ruiqi/local_marker/dataset/tabular_muris"
tiss <- readRDS(paste0(path,"/marrow_facs.rds"))
```
## Process
```{r}
DefaultAssay(tiss) <- "RNA"
n_dim = dim(tiss@reductions$pca@cell.embeddings)[2]
feature_space = as.matrix(tiss@reductions$pca@cell.embeddings[,1:n_dim])
visual_space = data.frame(tiss@reductions$tsne@cell.embeddings)
dat = as.matrix(tiss[[DefaultAssay(tiss)]]@data)
cell_label = tiss$cell_ontology_class

Gene_detected_count <- apply(dat > apply(dat,2,median),1,sum)
selected_genes = (Gene_detected_count >= 10) & (Gene_detected_count <= ncol(dat) * 0.5)
selected_genes = names(selected_genes)[selected_genes]
dat = dat[selected_genes,,drop = FALSE]
raw_count = as.matrix(tiss@assays$RNA@counts)[selected_genes,,drop = FALSE]
total_count = tiss@meta.data$nCount_RNA
```

# Define Ground truth Marker
## Criterion1: Fold-change
For each gene, sorting cell types by mean expression for each gene and computing fold changes between consecutive types, take the maximum value among the N-1 fold change values, given N cell types.
```{r}
path = "./benchmark_result/"
tissue_name = "marrow_facs"
```

```{r}
avg_exp = AverageExpression(subset(tiss,features = selected_genes), assays = "RNA", slot = "counts", group.by = "cell_ontology_class") %>% as.data.frame()

avg_exp_ordered <- avg_exp %>%
  rowwise() %>%
  mutate(cell_order_val = list(sort(c_across(everything()), decreasing = FALSE))) %>%
  ungroup()

avg_exp_ordered = avg_exp_ordered %>%
  rowwise() %>%
  mutate(log_fold_change = list(c(NA, diff(log2(unlist(cell_order_val)+1))))) %>%
  ungroup()

max_logfc = unlist( lapply(avg_exp_ordered$log_fold_change,function(vec){max(vec,na.rm = TRUE)}) )
names(max_logfc) = rownames(avg_exp)

max_logfc = sort(max_logfc,decreasing = TRUE)
write.table(max_logfc,file = paste0(path,tissue_name,"_ground_truth_c1.txt"))
```
## Criterion2: CellMarkerDB
Downloaded marker from http://bio-bigdata.hrbmu.edu.cn/CellMarker/CellMarker_download_files/file/Cell_marker_Seq.xlsx
PanglaoDB: https://panglaodb.se/markers/PanglaoDB_markers_27_Mar_2020.tsv.gz and exclude genes that reported only once
```{r}
# Load CellMarkerDB
cell_marker_db = readxl::read_xlsx("/data/ruiqi/local_marker/dataset/Cell_Marker/Cell_marker_All.xlsx") %>% filter(species == "Mouse") %>% filter(tissue_class == "Bone marrow") %>% filter(!is.na(cellontology_id)) %>% filter(!is.na(PMID)) %>% mutate(gene_name1 = marker) %>% mutate(gene_name2 = Symbol) %>% select(c("gene_name1","gene_name2")) %>% group_by(across(everything())) %>% dplyr::summarise(count = n(), .groups = "drop") %>% filter(count > 1)

# Match gene name
celldb_marker = union(cell_marker_db$gene_name1,cell_marker_db$gene_name2)
celldb_marker = unlist( lapply(celldb_marker,function(i){
  id = grep(paste0("^", i, "$"), selected_genes, ignore.case = TRUE)
  if(length(id)==0){return(NA)}
  else{return(selected_genes[id])}
}) )
celldb_marker = celldb_marker[!is.na(celldb_marker)] %>% unique()
write.table(celldb_marker,file = paste0(path,tissue_name,"_ground_truth_c2.txt"))
```

# Run Each Method and save results
```{r}
path = "./benchmark_result/"
tissue_name = "marrow_facs"
tissue_name = "lung_facs"
```

## LocalMarkerDetector
Genes were ranked by the increasing order of Cumulative Diffusion-KL score
```{r}
method = "lmd"
source("./LMD_function.R", echo = F)
res = LMD(dat,feature_space,max_time = 2^10)
marker = show_result_lmd(res, n = length(selected_genes))
marker[order(marker$score),'rank'] = 1:nrow(marker)
write.table(marker,file = sprintf("%s%s_%s.csv",path,method,tissue_name),row.names = TRUE)
```

## Seurat v3 Wilcoxon unfiltered
Genes were ranked by their lowest adjusted P-value in any cluster (increasing order), if two genes had the same, prioritize genes with larger FC.
```{r}
tiss <- FindNeighbors(tiss, dims = 1:n_dim)
tiss <- FindClusters(tiss, resolution = 1.5, algorithm = 3)
method = "wilcox_no_filter"
marker <- FindAllMarkers(tiss, features = rownames(dat), only.pos = FALSE, test.use = "wilcox", logfc.threshold = 0, min.pct = 0)
gene_rank = marker %>% arrange(.,p_val_adj,desc(avg_log2FC)) %>% distinct(.,gene, .keep_all = TRUE) %>% select(gene)
gene_rank$'rank' = 1:nrow(gene_rank)
marker <- merge(marker,gene_rank,by = "gene")
write.table(marker,file = sprintf("%s%s_%s.csv",path,method,tissue_name),row.names = TRUE)
```
## Seurat v3 default
Genes were ranked by their lowest adjusted P-value in any cluster (increasing order), if two genes had the same, prioritize genes with larger FC.
```{r}
tiss <- FindNeighbors(tiss, dims = 1:n_dim)
tiss <- FindClusters(tiss, resolution = 1.5, algorithm = 3)

method = "wilcox_default"
marker <- FindAllMarkers(tiss, features = rownames(dat), only.pos = FALSE, test.use = "wilcox", logfc.threshold = 0.25,min.pct = 0.1)

gene_rank = marker %>% arrange(.,p_val_adj,desc(avg_log2FC)) %>% distinct(.,gene, .keep_all = TRUE) %>% select(gene)
gene_rank$'rank' = 1:nrow(gene_rank)
marker <- merge(marker,gene_rank,by = "gene")
write.table(marker,file = sprintf("%s%s_%s.csv",path,method,tissue_name),row.names = TRUE)
```
## singleCellHaystack
https://alexisvdb.github.io/singleCellHaystack/articles/examples/a02_example_scRNAseq.html
Genes were ranked based on the increasing order of log(adjusted P-value)
```{r}
method = "haystack"
library(singleCellHaystack)
set.seed(123)
res <- haystack(x = feature_space, expression = dat)
marker <- show_result_haystack(res.haystack = res)
marker[order(marker$log.p.adj),"rank"] = 1:nrow(marker)

write.table(marker,file = sprintf("%s%s_%s.csv",path,method,tissue_name),row.names = TRUE)
```
## SEMITONES
https://github.com/ohlerlab/SEMITONES/tree/master
https://github.com/ohlerlab/SEMITONES_paper/blob/main/notebooks/10_SEMITONES_benchmark_corrected.ipynb

Genes were ranked by their lowest adjusted P-value in any reference cell (increasing order), if two genes had the same p.adj, prioritize genes with larger absolute enrichment score in any reference cell.
```{python}
import SEMITONES
import numpy as np
import pandas as pd
import matplotlib.pyplot as plt
from SEMITONES.cell_selection import from_knn_dist
from SEMITONES.cell_selection import from_gui
from SEMITONES.cell_selection import from_2D_embedding
from SEMITONES.cell_selection import get_cells_from_gui
from sklearn.metrics.pairwise import pairwise_kernels
from SEMITONES.enrichment_scoring import calculate_escores
from SEMITONES.enrichment_scoring import permute
from SEMITONES.enrichment_scoring import sig_interval
from SEMITONES.support_funcs import pairwise_similarities
from SEMITONES.support_funcs import sig_dictionary
from SEMITONES.support_funcs import sig_bool
from SEMITONES.enrichment_scoring import pvals_per_cell

# Load data
data = pd.DataFrame(r.dat.T)  # counts
PC20 = pd.DataFrame(r.feature_space)
# Reference Cell
g = 8.6e-4  
S = pairwise_kernels(PC20, metric="rbf", gamma=g) 
median = np.median(S, axis=0)
start = int(np.argmin(median))
num_refcell = int(data.shape[1]*0.005)
dd_rcells = from_knn_dist(X=PC20, 
                          n_ret=num_refcell,
                          start=start, 
                          metric="rbf",
                          metric_params={"gamma": g})
S = pairwise_similarities(PC20,
                          query=dd_rcells,
                          metric="rbf",
                          metric_params={"gamma": g})
dd_rcells = data.index[dd_rcells]
escores = calculate_escores(data.values, query=dd_rcells, S=S)

# semi_ranks = escores.abs().rank(ascending=False, method="dense").min(1)

P = permute(data.values)
pscores = calculate_escores(P, query=dd_rcells, S=S)
pvals = pvals_per_cell(escores,pscores,ret = "q")
# semi_ranks = pvals.min(1).rank(ascending=True, method="dense")

result = pd.DataFrame({'escore': escores.abs().max(1), 'padj': pvals.min(1)})
```
```{r}
method = "semi"
marker = reticulate::py$result
rownames(marker) = rownames(dat)
marker = marker %>% arrange(.,padj,desc(escore))
marker$'rank' = 1:nrow(marker)

write.table(marker,file = sprintf("%s%s_%s.csv",path,method,tissue_name),row.names = TRUE)
```

## Hotspot
https://hotspot.readthedocs.io/en/latest/CD4_Tutorial.html
Genes were ranked by the increasing order FDR value (if two genes had the same, prioritize genes with larger Z-score).
```{r}
cell_id = colnames(raw_count)
gene_id = rownames(raw_count)
```
```{python}
import hotspot
import scanpy as sc
import pandas as pd
from scipy.sparse import csc_matrix
adata = sc.AnnData(X = r.raw_count.T,obs = pd.DataFrame(index = r.cell_id), var = pd.DataFrame(index = r.gene_id))
adata.obs['total_counts'] = r.total_count
adata.layers["counts"] = adata.X.copy()
adata.obsm['X_pca'] = r.feature_space
adata.layers["counts_csc"] = csc_matrix(adata.layers["counts"])

hs = hotspot.Hotspot(
    adata,
    layer_key="counts_csc",
    model='danb',
    latent_obsm_key="X_pca",
    umi_counts_obs_key="total_counts"
)
hs.create_knn_graph(weighted_graph=False, n_neighbors=30)

# Determining informative genes
hs_results = hs.compute_autocorrelations()
```
```{r}
method = "hotspot"
marker = reticulate::py$hs_results
marker[order(marker$FDR),"rank"] = 1:nrow(marker)
write.table(marker,file = sprintf("%s%s_%s.csv",path,method,tissue_name),row.names = TRUE)
```

## Marcopolo
The rank of genes is given by `Marcopolo_rank`
```{python}
import pickle
import numpy as np
import pandas as pd
import torch
import anndata as ad
import scanpy as sc
import matplotlib.pyplot as plt
import MarcoPolo

adata = sc.AnnData(X = r.raw_count.T)

if "size_factor" not in adata.obs.columns:
    norm_factor = sc.pp.normalize_total(adata, exclude_highly_expressed=True, max_fraction= 0.2, inplace=False)["norm_factor"]
    adata.obs["size_factor"] = norm_factor/norm_factor.mean()
    print("size factor was calculated")

regression_result = MarcoPolo.run_regression(adata=adata, size_factor_key="size_factor",
                         num_threads=1, device="cpu")

result = MarcoPolo.find_markers(adata=adata, regression_result=regression_result)
```
```{r}
method = "marcopolo"
marker = reticulate::py$result
marker$'rank' = marker$MarcoPolo_rank
write.table(marker,file = sprintf("%s%s_%s.csv",path,method,tissue_name),row.names = TRUE)
```

## Highly Variable Genes
```{r}
method = "hvg"

DefaultAssay(tiss) <- "RNA"
# tiss <- NormalizeData(tiss, normalization.method = "LogNormalize", scale.factor = 1e6)
marker <- subset(tiss,features = selected_genes) %>% ScaleData(features = selected_genes) %>% 
  FindVariableFeatures(nfeatures = 5000) %>% VariableFeatures() %>% as.data.frame() %>% setNames(c("gene")) %>% mutate(rank = 1:n())

write.table(marker,file = sprintf("%s%s_%s.csv",path,method,tissue_name),row.names = FALSE)
```

# Integrate rank from all methods into one dataframe
```{r}
path = "./benchmark_result/"
tissue_name = "marrow_facs"
```
```{r}
method_ls = c("lmd","haystack","hotspot","marcopolo","semi","wilcox_default","wilcox_no_filter","hvg")
method_fullname = setNames(c("LMD","singleCellHaystack","Hotspot","Marcopolo","SEMITONES","Seurat v3 default", "Seurat v3 Wilcoxon unfiltered","HighlyVariableGenes"),method_ls)

df_benchmark = do.call(cbind,lapply(method_ls, function(method){
  marker = read.table(sprintf("%s%s_%s.csv",path,method,tissue_name),header = TRUE)
  if("gene" %in% colnames(marker)){
    return(marker$rank[match(selected_genes,marker$gene)])
  }else{
    return(marker$rank[match(selected_genes,rownames(marker))])
  }
}))
df_benchmark[is.na(df_benchmark)] = nrow(df_benchmark)
df_benchmark = data.frame(df_benchmark,row.names = selected_genes)
colnames(df_benchmark) = method_ls

write.table(df_benchmark,file = paste0(path,tissue_name,"_benchmark_rank_table.csv"),row.names = TRUE)
```

# AUROC
```{r}
path = "./benchmark_result/"
tissue_name = "marrow_facs"
df_benchmark = read.table(paste0(path,tissue_name,"_benchmark_rank_table.csv"))
```
```{r}
max_logfc = read.table(paste0(path,tissue_name,"_ground_truth_c1.txt")) %>% rownames()
celldb_marker = read.table(paste0(path,tissue_name,"_ground_truth_c2.txt"))[,1]

gt_list = c(lapply(seq(50,1000,50),function(x) max_logfc[1:x]),list(celldb_marker))
names(gt_list) = c(paste0("Top",seq(50,1000,50)),"CellMarkerDB")
df = do.call(rbind,lapply(1:length(gt_list),function(i){
  true_marker = gt_list[[i]]
  df_benchmark$"gt" = 0
  df_benchmark[true_marker,"gt"] = 1

  library(pROC)
  roc_df = do.call(rbind,lapply(setdiff(colnames(df_benchmark),"gt"), function(method){
    roc = roc(df_benchmark$gt, df_benchmark[,method], direction = ">")
    data.frame(FPR = 1 - roc$specificities, TPR = roc$sensitivities, Method = method, AUROC = as.numeric(auc(roc)))
  }) )

  df = roc_df %>% select(c("Method","AUROC")) %>% distinct()
  df$'gt_set' = names(gt_list)[i]
  df
}))
df$Method = as.factor(df$Method)
levels(df$Method) = method_fullname[levels(df$Method)]
df$gt_set = factor(df$gt_set,levels = names(gt_list))
write.table(df,file = paste0(path,tissue_name,"_auroc.csv"),row.names = FALSE)
```

# Density Index
```{r}
path = "./benchmark_result/"
tissue_name = "marrow_facs"
```
```{r}
df_benchmark = read.table(paste0(path,tissue_name,"_benchmark_rank_table.csv"))

score_measure =do.call(rbind,lapply(c(method_ls,"All genes"),function(method){
  if(method == "All genes"){
    top_gs = rownames(df_benchmark)
    top_n = length(top_gs)
  }else{
    top_gs = rownames(df_benchmark)[order(df_benchmark[,method])]
    top_n = c(seq(100,900,100),seq(1000,5000,1000))
  }
  result = unlist(lapply(top_n, function(top){
    tiss <- RunPCA(tiss, features = top_gs[1:top], npcs = 20)
    d = dist(Embeddings(tiss, reduction = "pca"))
    c = as.integer(factor(tiss$cell_ontology_class[labels(d)]))

    # # Silhouette score
    # s1 = mean(as.data.frame(cluster::silhouette(c, d)[,]) %>%
    #   group_by(cluster) %>%
    #   summarize(mean_silhouette = mean(sil_width)) %>%.$mean_silhouette)

    # Density-index
    nn.dist = FNN::get.knn(Embeddings(tiss, reduction = "pca"), k = 10, algorithm = "kd_tree")$nn.dist
    sdVec <- stats::na.omit(tiss@reductions$pca@stdev)
    length_scale <- sqrt(sum(sdVec^2))
    s2 = length_scale/mean(nn.dist)

    s2

  }))
  # rownames(result) = c("Mean_SI","DI")
  result = data.frame(TopGene = top_n,Method = method,DensityIndex = result)
}) )

write.table(score_measure,file = paste0(path,tissue_name,"_density_index.csv"),row.names = FALSE)
```

# Sensitivity Test (Try different K)
```{r}
path = "./marrow_result/"
tissue_name = "marrow_facs"
```
```{r}
# RunLMD under differen K
score_df = do.call(cbind,lapply(c(3:10,20,30,40,50),function(kNN){
  try({
    res = LMD(dat,feature_space,knn = kNN)
    return(res$cumulative_score)
  },silent = TRUE)
}))
colnames(score_df) = paste0("kNN",c(3:10,20,30,40,50))
write.table(score_df,file = paste0(path,tissue_name,"_different_knn.csv"),row.names = FALSE)
```
```{r}
# Obtain AUROC
score_ranked <- data.frame(score_df) %>% mutate(across(everything(), rank))

max_logfc = read.table(paste0(path,tissue_name,"_ground_truth_c1.txt")) %>% rownames()
celldb_marker = read.table(paste0(path,tissue_name,"_ground_truth_c2.txt"))[,1]

gt_list = c(lapply(seq(50,1000,50),function(x) max_logfc[1:x]),list(celldb_marker))
names(gt_list) = c(paste0("Top",seq(50,1000,50)),"CellMarkerDB")

auc_df = apply(score_ranked,2,function(x){
  unlist(lapply(gt_genes_list,function(genes){
    roc = pROC::roc(as.integer(rownames(score_ranked) %in% genes), x, direction = ">")
    pROC::auc(roc)
  }))
})
auc_df = data.frame(auc_df,gt = names(gt_list))
write.table(auc_df,file = paste0(path,tissue_name,"_sensitivity_auroc.csv"),row.names = FALSE)
```

### 2023-10-18 paper_figure.Rmd
benchmark&figure_code

---
title: "Paper_Figure"
output: pdf_document
date: '2023-10-16'
---
# Load Data & Preprocess
```{r}
library("Seurat")
path <- "/data/ruiqi/local_marker/dataset/tabular_muris/"
tissue_name = "marrow_facs"
tiss <- readRDS(sprintf("%s%s.rds",path,tissue_name))

DefaultAssay(tiss) <- "RNA"
n_dim = dim(tiss@reductions$pca@cell.embeddings)[2]
feature_space = as.matrix(tiss@reductions$pca@cell.embeddings[,1:n_dim])
visual_space = data.frame(tiss@reductions$tsne@cell.embeddings)
dat = as.matrix(tiss[[DefaultAssay(tiss)]]@data)
cell_label = tiss$cell_ontology_class

Gene_detected_count <- apply(dat > apply(dat,2,median),1,sum)
selected_genes = (Gene_detected_count >= 10) & (Gene_detected_count <= ncol(dat) * 0.5)
selected_genes = names(selected_genes)[selected_genes]
dat = dat[selected_genes,,drop = FALSE]
raw_count = as.matrix(tiss@assays$RNA@counts)[selected_genes,,drop = FALSE]
total_count = tiss@meta.data$nCount_RNA
dat_alra = as.matrix(tiss[["alra"]]@data)[selected_genes,]
```

# Fig - Pipeline
```{r}
source("./LMD_function.R", echo = F)

# Construct knn graph
res = Symmetric_KNN_graph(knn = 5, feature_space = feature_space)

A = res$adj_matrix # Adjacency Matrix
W = res$graph # Symmetrized Graph ((A + AT) / 2)

P_ls = Obtain_Pls(W = W, max_time = 2^10)
rho = Rowwise_normalize(dat[,colnames(W)])
res = fast_calculate_multi_score(W = W, init_state = rho, P_ls = P_ls)
```
```{r}
# Plot knn graph
plot_knn_graph(affinity_m = W, label = cell_label, layout = visual_space)
```
```{r}
# Check Diffusion KL pattern for several genes
genes = c("Fcnb","Uchl4")
genes_label = c("localized","non_localized")
check_time = c(0,8,128,1024)
p = Visualize_score_pattern(res, genes = genes, label_class = genes_label, facet_class = NULL, add_point = check_time)

# Check Gene Diffusion on Cell graph by time
pl = lapply(genes,function(gene){
  FeaturePlot_diffusion(coord = visual_space,init_state = rho[gene,],P_ls = P_ls,check_time = check_time,gene_name = gene)
})
plot_grid(plotlist = pl,ncol = 1)
```

# Fig - the AUROC given the rank and the ground truth
```{r}
library("RColorBrewer")
library("dplyr")
library("ggplot2")
path = "./benchmark_result/"
tissue_name = "marrow_facs"

df_benchmark = read.table(paste0(path,tissue_name,"_benchmark_rank_table.csv"))
method_ls = colnames(df_benchmark)
method_fullname = setNames(c("LMD","singleCellHaystack","Hotspot","Marcopolo","SEMITONES","Seurat v3 default", "Seurat v3 Wilcoxon unfiltered","HighlyVariableGenes"),method_ls)
coldef = setNames(
  colorRampPalette(brewer.pal(7, "Set1"))(length(method_fullname)),
  method_fullname)
```
```{r}
df_auc <- read.table(paste0(path,tissue_name,"_auroc.csv"),header = TRUE)
df_auc = df_auc %>% filter(gt_set %in% c("Top100","CellMarkerDB")) %>% filter(!Method %in% "HighlyVariableGenes")
df_auc$Method <- with(df_auc, reorder(Method, -AUROC, FUN = median))
df_auc$gt_set <- factor(df_auc$gt_set,levels = c("Top100","CellMarkerDB"))
levels(df_auc$gt_set) = c("Criterion1 - MaxFC", "Criterion2 - CellMarkerDB")

# Create the plot
fig <- ggplot(df_auc, aes(x = gt_set, y = AUROC, fill = Method)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = round(AUROC, 2)), position = position_dodge(width = 0.85), vjust = -0.25, size = 2) +
  theme_minimal() +
  labs(x = "Ground-Truth Genesets", y = "AUROC", fill = "Method") + coord_cartesian(ylim = c(0.5, 1)) +
  theme(legend.position = "bottom")  +  scale_fill_manual(values= coldef[names(coldef)%in%df_auc$Method])
fig
```

# Fig - Enrichment Overlap of Prioritized Genes with CellMarker Database Confidence Markers
```{r}
path = "./benchmark_result/"
tissue_name = "marrow_facs"

library(Seurat)
cell_marker_db = readxl::read_xlsx("/data/ruiqi/local_marker/dataset/Cell_Marker/Cell_marker_All.xlsx") %>% filter(species == "Mouse") %>% 
  filter(tissue_class == "Bone marrow") %>% filter(!is.na(cellontology_id)) %>% 
  filter(!is.na(PMID)) %>% mutate(gene_name1 = marker) %>% mutate(gene_name2 = Symbol) %>% select(c("gene_name1","gene_name2"))

df_benchmark = read.table(paste0(path,tissue_name,"_benchmark_rank_table.csv"))

method = "lmd"
gene_bin = data.frame(gene = rownames(df_benchmark)[order(df_benchmark[,method])][1:1000])
gene_bin$freq = unlist(lapply(gene_bin$gene,function(i){
  id = which(grepl(paste0("^", i, "$"), cell_marker_db$gene_name1, ignore.case = TRUE) + grepl(paste0("^", i, "$"), cell_marker_db$gene_name2, ignore.case = TRUE) >=1)
  return(length(id))
}) )
gene_bin$'freq_class' = cut(gene_bin$freq, breaks = c(-Inf,1,3,Inf), labels = c("none", "low", "high"), right = FALSE)

gene_bin$bin <- cut(seq(1:nrow(gene_bin)), breaks = 40, labels = FALSE)
gene_bin$'bin_label' = paste0(nrow(gene_bin)/max(gene_bin$bin)*(gene_bin$bin-1)+1,"-",nrow(gene_bin)/max(gene_bin$bin)*(gene_bin$bin))
gene_bin$'bin_label' = factor(gene_bin$'bin_label',levels = unique(gene_bin$'bin_label'))

gene_freq_prop <- gene_bin %>%
  group_by(bin_label, freq_class) %>%
  summarise(n = n()) %>%
  mutate(freq = n / sum(n))
gene_freq_prop$freq_class = factor(gene_freq_prop$freq_class,levels = c("none","low","high"),labels = c("None(=0)","Low(1,2)","High(>=3)"))

coldef = setNames(c("gray71","firebrick1","firebrick4"),levels(gene_freq_prop$freq_class))

fig = ggplot(gene_freq_prop, aes(fill=freq_class, y=freq, x=bin_label)) +
  geom_bar(position="fill", stat="identity") +
  ylab("Frequency") +
  xlab("Gene Rank") +
  labs(fill = "Confidence") +
  scale_fill_manual(values= coldef) +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

fig
```
```{r}
genes = c("Ctla2a","Fpr2","Mmp8")
pl = lapply(genes,function(g) FeaturePlot(tiss, g, order = TRUE) + labs(color = "Expression") + ggtitle(sprintf("%s (rank %d)",g,which(gene_bin == g))))
fig = cowplot::plot_grid(plotlist = pl, ncol = 3)
fig
```

# Fig - Gene Ranking Variations between LMD and Other Methods
```{r}
path = "./benchmark_result/"
tissue_name = "marrow_facs"

df_benchmark = read.table(paste0(path,tissue_name,"_benchmark_rank_table.csv"))

method_vs = "lmd"
method_vs_set = c("hotspot","haystack","wilcox_default")
pl = lapply(method_vs_set,function(method){
  p = ggplot(df_benchmark, aes(x=!!as.name(method), y=!!as.name(method_vs))) + geom_point(size = 1,col = "blue") +
    geom_abline(intercept = 0, slope = 1, col = "grey", linetype = "dashed") + scale_x_log10(breaks = 10^(0:4), labels = expression(10^0, 10^1, 10^2, 10^3, 10^4)) +
    scale_y_log10(breaks = 10^(0:4), labels = expression(10^0, 10^1, 10^2, 10^3, 10^4)) +
    expand_limits(x = c(1, 10^4), y = c(1, 10^4)) +
    labs(x = paste0(method_fullname[method]," Rank"), y = paste0(method_fullname[method_vs]," Rank")) + theme_minimal()

  # annotate gene labels
  library(ggrepel)
  df_label = data.frame(df_benchmark[,c(method,method_vs)],label = rownames(df_benchmark))
  sub = c(which(df_label[,method] <= 100 & df_label[,method_vs] > 500),
          which(df_label[,method] > 500 & df_label[,method_vs] <= 100) )
  df_label = df_label[sub,]
  df_label = df_label[order(df_label[,1]),]
  p = p + geom_text_repel(data = df_label, aes(label=label),col = "red",arrow = arrow(type = "closed", length = unit(0.01, "inches")),max.overlaps = 200)
  p
})
fig = cowplot::plot_grid(plotlist = pl, ncol = length(method_vs_set))
fig
```
```{r}
sub = do.call(c,lapply(method_vs_set,function(method){
  sub = c(which(df_benchmark[,method] <= 100 & df_benchmark[,method_vs] > 500),
          which(df_benchmark[,method] > 500 & df_benchmark[,method_vs] <= 100) )
  sub
}))
df_label_all = df_benchmark[unique(sub),c(method_vs_set,method_vs)]
df_label_all$'legend' = sapply(seq(nrow(df_label_all)),function(i){
  paste(paste(method_fullname[colnames(df_label_all)[c(4,1:3)]],df_label_all[i,c(4,1:3)],sep = ":"),collapse = "\n")
})

g = c("Mogat2","Vpreb2","Cd19","Mup3","Fyb","Naaa")
pl = lapply(g,function(x){
  FeaturePlot(tiss,features = x,order =TRUE) + annotation_custom(grob = grid::textGrob(df_label_all[x,"legend"], x = 1, y = 1, just = c("right", "top"),gp = grid::gpar(col = "red", fontsize = 10)))
})
fig = cowplot::plot_grid(plotlist = pl,ncol = 3)
fig
```

# Fig - Gene Co-localization Module Benchmark
```{r}
library(cowplot)
library(patchwork)
source("./LMD_function.R", echo = F)

path = "./benchmark_result/"
tissue_name = "marrow_facs"

df_benchmark = read.table(paste0(path,tissue_name,"_benchmark_rank_table.csv"))
method = "lmd"
topn = 2000
# Gene prioritized by lmd
pl1 = lapply(c("wilcox_default","haystack","hotspot","hvg"),function(method_vs){
  local_gene = setdiff(rownames(df_benchmark)[order(df_benchmark[,method_vs])][1:topn], rownames(df_benchmark)[order(df_benchmark[,method])][1:topn])

  ## Gene-gene distance(similarity)---------
  distance_method = "jaccard"
  clustering_method = "average"
  dist = Calculate_distance(dat_alra[local_gene,], method = distance_method)

  gene_hree = hclust(dist, method = clustering_method)
  gene_partition = cutree(gene_hree,k=5)
  names(gene_partition) = labels(dist)

  top5 <- names(sort(table(gene_partition),decreasing = TRUE))[1:5]
  annotation_df = data.frame(Module = gene_partition)
  ann_colors = list(
    Module = c(setNames(colorRampPalette(brewer.pal(9, "Set1"))(5),1:5)))

  p_heatmap = pheatmap(as.matrix(dist),
                annotation_col = annotation_df,
                cluster_rows = gene_hree,
                cluster_cols = gene_hree,
                treeheight_row = 0,
                annotation_colors = ann_colors, 
                col=colorRampPalette(c("red","white","blue4"))(99),
                show_colnames = FALSE,
                show_rownames = FALSE,
                annotation_legend = TRUE,
                silent = TRUE)
  pl_module = FeaturePlot_meta(dat,visual_space,gene_partition)
  legend <- get_legend(pl_module[[1]])
  prow <- cowplot::plot_grid(plotlist = lapply(pl_module,function(x) x + theme(legend.position="none")), ncol = ceiling(length(pl_module)/2))

  as.ggplot(p_heatmap) + cowplot::plot_grid(prow, legend, rel_widths = c(ceiling(length(pl_module)/2), .2)) + plot_annotation(title = sprintf("Genes in top 2000 by %s but not by %s (N = %d)",method_fullname[method_vs],method_fullname[method],length(local_gene))) + plot_layout(ncol = 2, widths = c(2, 3))
})
# The other way around
pl2 = lapply(c("wilcox_default","haystack","hotspot","hvg"),function(method_vs){
  local_gene = setdiff(rownames(df_benchmark)[order(df_benchmark[,method])][1:topn], rownames(df_benchmark)[order(df_benchmark[,method_vs])][1:topn])

  ## Gene-gene distance(similarity)---------
  distance_method = "jaccard"
  clustering_method = "average"
  dist = Calculate_distance(dat_alra[local_gene,], method = distance_method)

  gene_hree = hclust(dist, method = clustering_method)
  gene_partition = cutree(gene_hree,k=5)
  names(gene_partition) = labels(dist)

  top5 <- names(sort(table(gene_partition),decreasing = TRUE))[1:5]
  annotation_df = data.frame(Module = gene_partition)
  ann_colors = list(
    Module = c(setNames(colorRampPalette(brewer.pal(9, "Set1"))(5),1:5)))

  p_heatmap = pheatmap(as.matrix(dist),
                annotation_col = annotation_df,
                cluster_rows = gene_hree,
                cluster_cols = gene_hree,
                treeheight_row = 0,
                annotation_colors = ann_colors, 
                col=colorRampPalette(c("red","white","blue4"))(99),
                show_colnames = FALSE,
                show_rownames = FALSE,
                annotation_legend = TRUE,
                silent = TRUE)
  pl_module = FeaturePlot_meta(dat,visual_space,gene_partition)
  legend <- get_legend(pl_module[[1]])
  prow <- cowplot::plot_grid(plotlist = lapply(pl_module,function(x) x + theme(legend.position="none")), ncol = ceiling(length(pl_module)/2))

  as.ggplot(p_heatmap) + cowplot::plot_grid(prow, legend, rel_widths = c(ceiling(length(pl_module)/2), .2)) + plot_annotation(title = sprintf("Genes in top 2000 by %s but not by %s (N = %d)",method_fullname[method],method_fullname[method_vs],length(local_gene))) + plot_layout(ncol = 2, widths = c(2, 3))
})

pdf(width = 12.5,height = 5)
pl1
dev.off()

pdf(width = 12.5,height = 5)
pl2
dev.off()
```

# Fig - Density Index
```{r}
path = "./benchmark_result/"
tissue_name = "marrow_facs"
score_measure = read.table(paste0(path,tissue_name,"_density_index.csv"),header = TRUE)
score_measure$Method = ifelse(score_measure$Method %in% names(method_fullname),method_fullname[score_measure$Method],score_measure$Method)

baseline = score_measure %>% filter(Method == "all genes") %>%.$DensityIndex
max_x = score_measure %>% filter(Method!="all genes") %>%.$TopGene %>% max()

fig = ggplot(score_measure %>% filter(Method!="all genes"), aes(x = TopGene, y = DensityIndex, color = Method, group = Method)) + geom_line() + scale_color_manual(values= coldef[names(coldef)%in%score_measure$Method]) + geom_hline(yintercept = baseline, linetype = "dashed", color = "grey1", alpha = 0.6) + annotate("text", x = max_x, y = baseline, label = "All_genes", vjust = -1, color = "grey1")  + theme_minimal()
fig
```

# Fig - Sensitivity Test
```{r}
## Rank Stability
score_df = read.table(paste0(path,tissue_name,"_different_knn.csv"),header = TRUE)
score_ranked <- data.frame(score_df) %>% mutate(across(everything(), rank))
base = "kNN5"
top_genes_list <- lapply(c(50,seq(100,1000,100)), 
                         function(x) head(score_ranked[order(score_ranked[,base]),], x))
average_ranks <- lapply(top_genes_list, function(top_genes) {
  colMeans(top_genes)
})
average_ranks <- data.frame(
  Top_Genes = c(50,seq(100,1000,100)),
  do.call(rbind, average_ranks)
)
average_ranks <- average_ranks %>% tidyr::pivot_longer(cols = -Top_Genes, names_to = "kNN", values_to = "Average_Rank")
average_ranks$kNN = as.numeric(gsub("kNN","",average_ranks$kNN))
fig = ggplot(data = average_ranks, aes(x = kNN, y = Average_Rank, colour = factor(Top_Genes), group = Top_Genes)) +
  geom_line() +
  geom_point() +
  xlab("Top Genes") +
  ylab(sprintf("Average Rank of Top # Genes\n (base on %s)",base)) +
  theme_minimal() + facet_wrap(~Top_Genes,scales = "free")
fig
```
```{r}
## AUC stability
gt_genes_list <- c(lapply(c(50,seq(100,400,100)), function(x){
  names(max_logfc)[1:x]
}),list(true_marker))
auc_df = apply(score_ranked,2,function(x){
  unlist(lapply(gt_genes_list,function(genes){
    roc = pROC::roc(as.integer(rownames(score_ranked) %in% genes), x, direction = ">")
    pROC::auc(roc)
  }))
})
auc_df = data.frame(auc_df,gt = c(paste0("top",c(50,seq(100,400,100))),"cellMarkerDB"))
auc_df <- auc_df %>% tidyr::pivot_longer(cols = -gt, names_to = "kNN", values_to = "AUC")
auc_df$kNN = as.numeric(gsub("kNN","",auc_df$kNN))
ggplot(data = auc_df, aes(x = kNN, y = AUC, colour = factor(gt,levels = unique(gt)), group = gt)) +
  geom_line() +
  geom_point() +
  xlab("kNN") +
  ylab("AUC") +
  theme_minimal()

```

# Fig - Module HeatMap
```{r}
load("./marrow_result/module_result.rdata")
module_color = setNames(colorRampPalette(brewer.pal(12, "Paired"))(nlevels(gene_partition)),levels(gene_partition))
celltype_color = setNames(colorRampPalette(brewer.pal(9, "Set1"))(length(unique(cell_label))),unique(cell_label))

p = pheatmap(as.matrix(dist),
              annotation_col = data.frame(Module = gene_partition),
              cluster_rows = gene_hree, 
              cluster_cols = gene_hree,
              ann_colors = list(Module = module_color),
              treeheight_row = 0,
              col=colorRampPalette(c("red","white","blue4"))(99),
              show_colnames = FALSE, 
              show_rownames = FALSE,
              annotation_legend = FALSE,
              # main = paste(distance_method,
              #              clustering_method,sep = "-"),
              silent = TRUE)
as.ggplot(p)
```
# Fig - Module Jaccard
```{r}
# Heatmap
pheatmap(jaccard_index, 
         cluster_rows=FALSE, cluster_cols=FALSE, 
         scale = "none", 
         color = colorRampPalette(rev(c("firebrick3", "white", "navy")))(100))
```
```{r}
# Bipartite Plot
df_edge <- as.data.frame(as.table(jaccard_index))
colnames(df_edge) = c("from","to","weight")
df_edge <- subset(df_edge, weight > 0.4)
df_edge$width = ifelse(df_edge$weight > 0.8,2,0.5)
df_node = data.frame(name=c(names(module_color),names(celltype_color)),color = c(module_color,celltype_color))
df_node$type = ifelse(df_node$name %in% names(module_color),TRUE,FALSE)

g <- graph_from_data_frame(df_edge, directed=FALSE, vertices=df_node)
V(g)$shape <- ifelse(V(g)$type, "circle", "square")
E(g)$color <- ifelse(E(g)$weight > 0.8, "red", "grey")
l <- layout_as_bipartite(g, types=!V(g)$type)

plot(g, layout=l, vertex.size=2, vertex.label=ifelse(V(g)$type, V(g)$name, NA), vertex.frame.color = NA, asp=0.2, 
     vertex.label.cex=0.5,
     vertex.label.dist=-1,
     vertex.label.degree=pi/2)
coords = as.data.frame(l) %>% filter(V2 == 0) %>% mutate(labels = names(celltype_color))
text(x = coords$V1, y = coords$V2, labels = coords$labels, srt = 90, adj = c(1.5, 0.5),cex = 0.5,col = "black")
# Add legend for colors
legend("bottom", legend=names(celltype_color), fill=celltype_color, bty="n", cex=0.5)
```
# Fig - Module TSNE
```{r}
pl_module = lapply(levels(gene_partition),function(module){
  FeaturePlot_custom(value = cell_block[,module], coord = visual_space, value_name = "Module Score",title_name = sprintf("Module %s (%d genes)",module, sum(gene_partition == module)))
})
names(pl_module) = levels(gene_partition)
legend <- get_legend(pl_module[[1]])
prow <- cowplot::plot_grid(plotlist = lapply(pl_module,function(x) x + theme(legend.position="none")), ncol = 10)
pdf(width = 30, height = 18)
cowplot::plot_grid(prow, legend, rel_widths = c(10, .2))
dev.off()
```
# Fig - Module-CellType
```{r}
jaccard_thred = 0.4
one_hot_matrix <- sapply(unique(cell_label), function(x) as.integer(cell_label == x))
# Get CellType One-hot plot
pl_ct = lapply(colnames(one_hot_matrix),function(ct){
  FeaturePlot_custom(value = as.factor(one_hot_matrix[,ct]), coord = visual_space, value_name = "Indicator",title_name = ct)
})
names(pl_ct) = colnames(one_hot_matrix)

pl <- lapply(rownames(jaccard_index)[rev(order(apply(jaccard_index,1,max)))], function(celltype){
  if(max(jaccard_index[celltype,]) > jaccard_thred){
    p1 = pl_ct[[celltype]]
    module_name = colnames(jaccard_index)[which.max(jaccard_index[celltype,])]
    p2 = pl_module[[module_name]]
    p = as.ggplot(p1) + as.ggplot(p2) + plot_annotation(paste0("Jaccard Index: ",round(jaccard_index[celltype,module_name],2)))
    p
  }
})
sub = unlist(lapply(pl,function(x)is.null(x)))
pl = pl[-which(sub)]
pdf(width = 15, height = 20)
cowplot::plot_grid(plotlist = pl,ncol = 3)
dev.off()
```

### 2023-10-18 paper_figure.R

@@ -1,57 +1,0 @@
-# Sensitivity Test----------
-## Rank Stability-----------
-score_df = lapply(c(3:10,20,30,40,50),function(kNN){
-  try({
-    W = Symmetric_KNN_graph(knn = kNN, feature_space = feature_space)[[1]]
-    rho = Rowwise_normalize(dat[,colnames(W)])
-    marker = fast_calculate_multi_score(W = W, init_state = rho)
-    
-    sub_score0 = grep("score0",colnames(marker))
-    sub_score1 = grep("score1_0",colnames(marker))
-    accu_score1 = apply(marker,1,function(x){
-      sum(x[sub_score0]/x[sub_score1])
-    })
-    accu_score1
-  },silent = TRUE)
-})
-names(score_df) = paste0("kNN",c(3:10,20,30,40,50))
-score_df = do.call(cbind,score_df)
-score_ranked <- data.frame(score_df) %>% mutate(across(everything(), rank))
-base = "kNN5"
-top_genes_list <- lapply(c(50,seq(100,1000,100)), 
-                         function(x) head(score_ranked[order(score_ranked[,base]),], x))
-average_ranks <- lapply(top_genes_list, function(top_genes) {
-  colMeans(top_genes)
-})
-average_ranks <- data.frame(
-  Top_Genes = c(50,seq(100,1000,100)),
-  do.call(rbind, average_ranks)
-)
-average_ranks <- average_ranks %>% tidyr::pivot_longer(cols = -Top_Genes, names_to = "kNN", values_to = "Average_Rank")
-average_ranks$kNN = as.numeric(gsub("kNN","",average_ranks$kNN))
-ggplot(data = average_ranks, aes(x = kNN, y = Average_Rank, colour = factor(Top_Genes), group = Top_Genes)) +
-  geom_line() +
-  geom_point() +
-  xlab("Top Genes") +
-  ylab(sprintf("Average Rank of Top # Genes\n (base on %s)",base)) +
-  theme_minimal() + facet_wrap(~Top_Genes,scales = "free")
-
-## AUC stability------------
-gt_genes_list <- c(lapply(c(50,seq(100,400,100)), function(x){
-  names(max_logfc)[1:x]
-}),list(true_marker))
-auc_df = apply(score_ranked,2,function(x){
-  unlist(lapply(gt_genes_list,function(genes){
[…]

### 2023-10-18 Marker_identfication_benchmark.Rmd
robustness test

@@
 ```{r}
-path <- "/data/ruiqi/local_marker/dataset/tabular_muris"
-tiss <- readRDS(paste0(path,"/marrow_facs.rds"))
+path <- "/data/ruiqi/local_marker/dataset/tabular_muris/"
+tissue_name = "marrow_facs" #lung_facs, pancreas_facs
+tiss <- readRDS(paste0(path,tissue_name,".rds"))
 ```
@@
 path = "./benchmark_result/"
-tissue_name = "marrow_facs"
+# tissue_name = "marrow_facs"
 ```
@@
 path = "./benchmark_result/"
-tissue_name = "marrow_facs"
-tissue_name = "lung_facs"
+# tissue_name = "marrow_facs"
 ```
@@
 source("./LMD_function.R", echo = F)
-res = LMD(dat,feature_space,max_time = 2^10)
+res = LMD(dat,feature_space,max_time = 2^10,knn = 7)
 marker = show_result_lmd(res, n = length(selected_genes))
@@
 path = "./benchmark_result/"
-tissue_name = "marrow_facs"
+# tissue_name = "marrow_facs"
 df_benchmark = read.table(paste0(path,tissue_name,"_benchmark_rank_table.csv"))
@@
 path = "./marrow_result/"
-tissue_name = "marrow_facs"
+# tissue_name = "marrow_facs"
 ```
@@
 # RunLMD under differen K
+source("./LMD_function.R", echo = F)
 score_df = do.call(cbind,lapply(c(3:10,20,30,40,50),function(kNN){
@@
 score_ranked <- data.frame(score_df) %>% mutate(across(everything(), rank))
-
+path = "./benchmark_result/"
 max_logfc = read.table(paste0(path,tissue_name,"_ground_truth_c1.txt")) %>% rownames()
@@
 auc_df = apply(score_ranked,2,function(x){
-  unlist(lapply(gt_genes_list,function(genes){
+  unlist(lapply(gt_list,function(genes){
     roc = pROC::roc(as.integer(rownames(score_ranked) %in% genes), x, direction = ">")
@@
 })
+path = "./marrow_result/"
 auc_df = data.frame(auc_df,gt = names(gt_list))

### 2023-10-18 paper_figure.Rmd
robustness test

@@
   theme(legend.position = "bottom")  +  scale_fill_manual(values= coldef[names(coldef)%in%df_auc$Method])
-fig
+fig + ggtitle("Lung")
 ```
@@
 average_ranks$kNN = as.numeric(gsub("kNN","",average_ranks$kNN))
-fig = ggplot(data = average_ranks, aes(x = kNN, y = Average_Rank, colour = factor(Top_Genes), group = Top_Genes)) +
+
+fig = ggplot(data = average_ranks, aes(x = kNN, y = Average_Rank, group = Top_Genes)) +
   geom_line() +
-  geom_point() +
-  xlab("Top Genes") +
-  ylab(sprintf("Average Rank of Top # Genes\n (base on %s)",base)) +
-  theme_minimal() + facet_wrap(~Top_Genes,scales = "free")
+  geom_point() + labs(x = "kNN", y = sprintf("Average Rank of Top # Genes\n (base on %s)",base)) + geom_vline(xintercept = 5, color="red", linetype="dashed") +
+  theme_minimal() + facet_wrap(~Top_Genes,scales = "free",labeller = as_labeller(setNames(paste0("Top",unique(average_ranks$Top_Genes)),unique(average_ranks$Top_Genes)))) + NoLegend()
+pdf(width = 8,height = 5)
 fig
+dev.off()
 ```
@@
 ## AUC stability
-gt_genes_list <- c(lapply(c(50,seq(100,400,100)), function(x){
-  names(max_logfc)[1:x]
-}),list(true_marker))
-auc_df = apply(score_ranked,2,function(x){
-  unlist(lapply(gt_genes_list,function(genes){
-    roc = pROC::roc(as.integer(rownames(score_ranked) %in% genes), x, direction = ">")
-    pROC::auc(roc)
-  }))
-})
-auc_df = data.frame(auc_df,gt = c(paste0("top",c(50,seq(100,400,100))),"cellMarkerDB"))
-auc_df <- auc_df %>% tidyr::pivot_longer(cols = -gt, names_to = "kNN", values_to = "AUC")
+auc_df = read.table(paste0(path,tissue_name,"_sensitivity_auroc.csv"),header = TRUE)
+auc_df <- auc_df %>% tidyr::pivot_longer(cols = -gt, names_to = "kNN", values_to = "AUC") %>% filter(gt %in% c("Top100","CellMarkerDB"))
+auc_df$gt = factor(auc_df$gt); levels(auc_df$gt) = c("CellMarkerDB","MacFC(Top100)")
 auc_df$kNN = as.numeric(gsub("kNN","",auc_df$kNN))
-ggplot(data = auc_df, aes(x = kNN, y = AUC, colour = factor(gt,levels = unique(gt)), group = gt)) +
+fig = ggplot(data = auc_df, aes(x = kNN, y = AUC), group = gt) +
   geom_line() +
-  geom_point() +
-  xlab("kNN") +
-  ylab("AUC") +
+  geom_point() + labs(x = "kNN", y = "AUROC") + facet_wrap(~gt, scales = "free") + 
   theme_minimal()
-
+pdf(width = 6,height = 3)
+fig
+dev.off()
 ```

### 2023-10-19 Demo_marrow_facs.Rmd
figwidth and color gradient

--- 2023-10-17 Demo_marrow_facs.Rmd
@@
 
-```{r}
+```{r, fig.width=20}
 # Construct knn graph
@@
               treeheight_row = 0,
-              col=colorRampPalette(c("red","blue4"))(99),
+              col=colorRampPalette(c("firebrick3","white"))(99),
               show_colnames = FALSE, 
@@
 
-```{r, fig.width=40}
+```{r, fig.width=20, fig.asp=1}
 pl = lapply(levels(gene_partition),function(module){
@@
 })
-plot_grid(plotlist = pl,ncol = 10) # may takes long time to plot
+plot_grid(plotlist = pl,ncol = 5) # may takes long time to plot
 ```
@@
          scale = "none", 
-         color = colorRampPalette(rev(c("firebrick3", "white", "navy")))(100))
+         color = colorRampPalette(rev(c("firebrick3", "white")))(100))
 ```

### 2023-10-19 Demo_marrow_facs.Rmd
figure colors

--- 2023-10-17 Demo_marrow_facs.Rmd
@@
               treeheight_row = 0,
-              col=colorRampPalette(c("red","blue4"))(99),
+              col=colorRampPalette(c("firebrick3", "white"))(99),
               show_colnames = FALSE, 
@@
          scale = "none", 
-         color = colorRampPalette(rev(c("firebrick3", "white", "navy")))(100))
+         color = colorRampPalette(rev(c("firebrick3", "white")))(99))
 ```

### 2023-10-19 Demo_marrow_facs.Rmd

@@
 knn = 10; major_vote = 5
-A = Symmetric_KNN_graph(knn = knn, feature_space = feature_space)[[2]]
+A = Symmetric_KNN_graph(knn = knn, feature_space = feature_space)$'adj_matrix'
 cell_block_loop = lapply(1:100,function(loop){
@@
 })
+jaccard_index = jaccard_index[rev(ordered_cell_types),]
+rm(ordered_cell_types)
 ```
@@
 ```{r}
-pheatmap(jaccard_index[rev(ordered_cell_types),], 
+pheatmap(jaccard_index,
          cluster_rows=FALSE, cluster_cols=FALSE, 
@@
 module_celltype = apply(jaccard_index,2,function(x){paste(names(which(x > 0.4)),collapse = ",")})
+```
+
+## Save Intermediate data
+```{r}
+save(gene_partition,dist,gene_hree,cell_block,jaccard_index,file = "./marrow_result/module_result.rdata")
 ```

### 2023-10-20 Demo_marrow_facs.Rmd
gene_ranking

--- 2023-10-19 Demo_marrow_facs.Rmd
@@
 # Preparing input data
-
 ```{r}
-if (!file.exists('facs_Marrow_seurat_tiss.Robj')) {
 # If load url doesn't work, please download it and load from your working directory
-  options(timeout=6000)
-  download.file("https://figshare.com/ndownloader/files/13092380",
-                destfile = 'facs_Marrow_seurat_tiss.Robj', method = 'libcurl')
- }
-load('facs_Marrow_seurat_tiss.Robj')
+load(url("https://figshare.com/ndownloader/files/13092380"))
 tiss <- UpdateSeuratObject(tiss)
@@
 ```
-
 ## Step1: Build KNN Graph
-
 ```{r}
 # Construct knn graph
-knn_result = Symmetric_KNN_graph(knn = 5, feature_space = feature_space)
+res = Symmetric_KNN_graph(knn = 5, feature_space = feature_space)
 
-A = knn_result$adj_matrix # Adjacency Matrix
-W = knn_result$graph # Symmetrized Graph ((A + AT) / 2)
+A = res$adj_matrix # Adjacency Matrix
+W = res$graph # Symmetrized Graph ((A + AT) / 2)
 
@@
 ## Step2: Diffuse Genes on the cell graph
-
-```{r, fig.width=20}
+```{r}
 # Generate $P^{t}, t = 2,4,8,...,max_time$
@@
 ## Step3: Obtain Diffusion KL score & Cumulative score
-
 ```{r}
 # Obtain diffusion KL score
-score_result = fast_calculate_multi_score(W = W, init_state = rho, P_ls = P_ls)
+res = fast_calculate_multi_score(W = W, init_state = rho, P_ls = P_ls)
 
@@
 genes_label = rep(c("good","bad"),each = 3)
-Visualize_score_pattern(score_result, genes = genes, label_class = genes_label, facet_class = NULL)
+Visualize_score_pattern(res, genes = genes, label_class = genes_label, facet_class = NULL)
 
 # Show result
-show_result_lmd(score_result,n = 10)
+show_result_lmd(res,n = 10)
 ```
@@
 # Running LMD in one step
-
 ```{r}
-lmd_result = LMD(dat, feature_space)
+res = LMD(dat,feature_space)
 ```
@@
 # Identify Gene Modules
-
 ## Step1: Deriving Gene Modules
-
 ```{r}
 # Select top1000 genes as local genes
-local_gene = show_result_lmd(lmd_result, n = 1000)
+local_gene = show_result_lmd(res, n = 1000)
 local_gene = rownames(local_gene)
@@
               treeheight_row = 0,
-              col=colorRampPalette(c("firebrick3", "white"))(99),
+              col=colorRampPalette(c("red","white","blue4"))(99),
               show_colnames = FALSE, 
@@
 ## Step2: Computing Per-Cell Module Activity Scores
-
 ```{r}
@@
 cell_block_loop = lapply(1:100,function(loop){
+  cat("\n",loop,"\n")
   GMM_subsampling(seed = loop, gene_partition, expr_dat = dat, cell_kNN_graph = A, major_vote)
@@
 ```
-
 ### Visualize Module on Cell embeddings
-
-```{r, fig.width=40}
+```{r}
 pl = lapply(levels(gene_partition),function(module){
@@
   apply(one_hot_matrix, 2, function(one_hot_col) {
-    1 - philentropy::distance(rbind(module_col,one_hot_col), method = "jaccard", mute.message=TRUE)
+    1 - philentropy::distance(rbind(module_col,one_hot_col), method = "jaccard")
   })
@@
 ### Visualize Heatmap of Jaccard Index Matrix
-
 ```{r}
@@
          scale = "none", 
-         color = colorRampPalette(rev(c("firebrick3", "white")))(99))
+         color = colorRampPalette(rev(c("firebrick3", "white", "navy")))(100))
 ```
@@
 ## Step4: Module Enrichment Analysis
-
 ### GO enrichment Analysis
-
 ```{r}
+if (!requireNamespace("clusterProfiler", quietly = TRUE)) BiocManager::install("clusterProfiler")
 library("clusterProfiler")
-library("ReactomePA")
 library("org.Mm.eg.db")
-library("AnnotationDbi")
 
@@
 df_module[,"celltype"] = module_celltype[levels(gene_partition)]
-df_module[top_go_clusters,"GO_description"] = unlist(lapply(top_enrich_go,function(x) paste(x[['Description']],collapse = ",")))
-df_module[top_pathway_clusters,"Pathway_description"] = unlist(lapply(top_enrich_pathway,function(x) paste(x[['Description']],collapse = ",")))
+df_module[top_go_clusters,"GO_description"] = unlist(lapply(top_enrich_go,function(x){
+  des = x[["Description"]]
+  paste(des[!is.na(des)],collapse = ",")
+}
+))
+df_module[top_pathway_clusters,"Pathway_description"] = unlist(lapply(top_enrich_pathway,function(x){
+  des = x[["Description"]]
+  paste(des[!is.na(des)],collapse = ",")
+}))
 df_module[is.na(df_module)] = ""
-write.csv(df_module,file = "module_description.csv",row.names = FALSE)
+write.csv(df_module,file = "./marrow_result/module_description.csv",row.names = FALSE)
 ```
+
+## Cell Cycle Modules
+```{r}
+library(gprofiler2)
+mmus_s = gorth(cc.genes.updated.2019$s.genes, source_organism = "hsapiens", target_organism = "mmusculus")$ortholog_name
+mmus_g2m = gorth(cc.genes.updated.2019$g2m.genes, source_organism = "hsapiens", target_organism = "mmusculus")$ortholog_name
+tiss = CellCycleScoring(tiss, s.features = mmus_s, 
+                        g2m.features = mmus_g2m, set.ident = FALSE)
+DimPlot(tiss,group.by = "Phase")
+
+pdf(width = 30, height = 5)
+pl
+dev.off()
+```

### 2023-10-20 Marker_identfication_benchmark.Rmd

@@
 method = "wilcox_no_filter"
-marker <- FindAllMarkers(tiss, features = rownames(dat), only.pos = FALSE, test.use = "wilcox", logfc.threshold = 0, min.pct = 0)
+marker <- FindAllMarkers(tiss, features = selected_genes, only.pos = FALSE, test.use = "wilcox", logfc.threshold = 0, min.pct = 0)
 gene_rank = marker %>% arrange(.,p_val_adj,desc(avg_log2FC)) %>% distinct(.,gene, .keep_all = TRUE) %>% select(gene)
@@
 method = "wilcox_default"
-marker <- FindAllMarkers(tiss, features = rownames(dat), only.pos = FALSE, test.use = "wilcox", logfc.threshold = 0.25,min.pct = 0.1)
+marker <- FindAllMarkers(tiss, features = selected_genes, only.pos = FALSE, test.use = "wilcox", logfc.threshold = 0.25,min.pct = 0.1)
 
@@
 marker = reticulate::py$result
-marker$'rank' = marker$MarcoPolo_rank
+marker$'rank' = marker$MarcoPolo_rank + 1
 write.table(marker,file = sprintf("%s%s_%s.csv",path,method,tissue_name),row.names = TRUE)

### 2023-10-20 paper_figure.Rmd

@@
 coldef = setNames(
-  colorRampPalette(brewer.pal(7, "Set1"))(length(method_fullname)),
+  colorRampPalette(brewer.pal(9, "Set1"))(length(method_fullname)),
   method_fullname)
@@
   theme(legend.position = "bottom")  +  scale_fill_manual(values= coldef[names(coldef)%in%df_auc$Method])
-fig + ggtitle("Lung")
+fig + ggtitle(tissue_name)
 ```
 
-# Fig - Enrichment Overlap of Prioritized Genes with CellMarker Database Confidence Markers
-```{r}
-path = "./benchmark_result/"
-tissue_name = "marrow_facs"
+<!-- # Fig - Enrichment Overlap of Prioritized Genes with CellMarker Database Confidence Markers -->
+<!-- ```{r} -->
+<!-- path = "./benchmark_result/" -->
+<!-- tissue_name = "marrow_facs" -->
 
-library(Seurat)
-cell_marker_db = readxl::read_xlsx("/data/ruiqi/local_marker/dataset/Cell_Marker/Cell_marker_All.xlsx") %>% filter(species == "Mouse") %>% 
-  filter(tissue_class == "Bone marrow") %>% filter(!is.na(cellontology_id)) %>% 
-  filter(!is.na(PMID)) %>% mutate(gene_name1 = marker) %>% mutate(gene_name2 = Symbol) %>% select(c("gene_name1","gene_name2"))
+<!-- library(Seurat) -->
+<!-- cell_marker_db = readxl::read_xlsx("/data/ruiqi/local_marker/dataset/Cell_Marker/Cell_marker_All.xlsx") %>% filter(species == "Mouse") %>%  -->
+<!--   filter(tissue_class == "Bone marrow") %>% filter(!is.na(cellontology_id)) %>%  -->
+<!--   filter(!is.na(PMID)) %>% mutate(gene_name1 = marker) %>% mutate(gene_name2 = Symbol) %>% select(c("gene_name1","gene_name2")) -->
 
-df_benchmark = read.table(paste0(path,tissue_name,"_benchmark_rank_table.csv"))
+<!-- df_benchmark = read.table(paste0(path,tissue_name,"_benchmark_rank_table.csv")) -->
 
-method = "lmd"
-gene_bin = data.frame(gene = rownames(df_benchmark)[order(df_benchmark[,method])][1:1000])
-gene_bin$freq = unlist(lapply(gene_bin$gene,function(i){
-  id = which(grepl(paste0("^", i, "$"), cell_marker_db$gene_name1, ignore.case = TRUE) + grepl(paste0("^", i, "$"), cell_marker_db$gene_name2, ignore.case = TRUE) >=1)
-  return(length(id))
-}) )
-gene_bin$'freq_class' = cut(gene_bin$freq, breaks = c(-Inf,1,3,Inf), labels = c("none", "low", "high"), right = FALSE)
+<!-- method = "lmd" -->
+<!-- gene_bin = data.frame(gene = rownames(df_benchmark)[order(df_benchmark[,method])][1:1000]) -->
+<!-- gene_bin$freq = unlist(lapply(gene_bin$gene,function(i){ -->
+<!--   id = which(grepl(paste0("^", i, "$"), cell_marker_db$gene_name1, ignore.case = TRUE) + grepl(paste0("^", i, "$"), cell_marker_db$gene_name2, ignore.case = TRUE) >=1) -->
+<!--   return(length(id)) -->
+<!-- }) ) -->
+<!-- gene_bin$'freq_class' = cut(gene_bin$freq, breaks = c(-Inf,1,3,Inf), labels = c("none", "low", "high"), right = FALSE) -->
 
-gene_bin$bin <- cut(seq(1:nrow(gene_bin)), breaks = 40, labels = FALSE)
-gene_bin$'bin_label' = paste0(nrow(gene_bin)/max(gene_bin$bin)*(gene_bin$bin-1)+1,"-",nrow(gene_bin)/max(gene_bin$bin)*(gene_bin$bin))
-gene_bin$'bin_label' = factor(gene_bin$'bin_label',levels = unique(gene_bin$'bin_label'))
+<!-- gene_bin$bin <- cut(seq(1:nrow(gene_bin)), breaks = 40, labels = FALSE) -->
+<!-- gene_bin$'bin_label' = paste0(nrow(gene_bin)/max(gene_bin$bin)*(gene_bin$bin-1)+1,"-",nrow(gene_bin)/max(gene_bin$bin)*(gene_bin$bin)) -->
+<!-- gene_bin$'bin_label' = factor(gene_bin$'bin_label',levels = unique(gene_bin$'bin_label')) -->
 
-gene_freq_prop <- gene_bin %>%
-  group_by(bin_label, freq_class) %>%
-  summarise(n = n()) %>%
-  mutate(freq = n / sum(n))
-gene_freq_prop$freq_class = factor(gene_freq_prop$freq_class,levels = c("none","low","high"),labels = c("None(=0)","Low(1,2)","High(>=3)"))
+<!-- gene_freq_prop <- gene_bin %>% -->
+<!--   group_by(bin_label, freq_class) %>% -->
+<!--   summarise(n = n()) %>% -->
+<!--   mutate(freq = n / sum(n)) -->
+<!-- gene_freq_prop$freq_class = factor(gene_freq_prop$freq_class,levels = c("none","low","high"),labels = c("None(=0)","Low(1,2)","High(>=3)")) -->
 
-coldef = setNames(c("gray71","firebrick1","firebrick4"),levels(gene_freq_prop$freq_class))
+<!-- coldef = setNames(c("gray71","firebrick1","firebrick4"),levels(gene_freq_prop$freq_class)) -->
 
-fig = ggplot(gene_freq_prop, aes(fill=freq_class, y=freq, x=bin_label)) +
-  geom_bar(position="fill", stat="identity") +
-  ylab("Frequency") +
-  xlab("Gene Rank") +
-  labs(fill = "Confidence") +
-  scale_fill_manual(values= coldef) +
-  theme_minimal() + 
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+<!-- fig = ggplot(gene_freq_prop, aes(fill=freq_class, y=freq, x=bin_label)) + -->
+<!--   geom_bar(position="fill", stat="identity") + -->
+<!--   ylab("Frequency") + -->
+<!--   xlab("Gene Rank") + -->
+<!--   labs(fill = "Confidence") + -->
+<!--   scale_fill_manual(values= coldef) + -->
+<!--   theme_minimal() +  -->
+<!--   theme(axis.text.x = element_text(angle = 45, hjust = 1)) -->
 
-fig
-```
-```{r}
-genes = c("Ctla2a","Fpr2","Mmp8")
-pl = lapply(genes,function(g) FeaturePlot(tiss, g, order = TRUE) + labs(color = "Expression") + ggtitle(sprintf("%s (rank %d)",g,which(gene_bin == g))))
-fig = cowplot::plot_grid(plotlist = pl, ncol = 3)
-fig
-```
+<!-- fig -->
+<!-- ``` -->
+<!-- ```{r} -->
+<!-- genes = c("Ctla2a","Fpr2","Mmp8") -->
+<!-- pl = lapply(genes,function(g) FeaturePlot(tiss, g, order = TRUE) + labs(color = "Expression") + ggtitle(sprintf("%s (rank %d)",g,which(gene_bin == g)))) -->
+<!-- fig = cowplot::plot_grid(plotlist = pl, ncol = 3) -->
+<!-- fig -->
+<!-- ``` -->
 
@@
 df_benchmark = read.table(paste0(path,tissue_name,"_benchmark_rank_table.csv"))
-
 method_vs = "lmd"
-method_vs_set = c("hotspot","haystack","wilcox_default")
-pl = lapply(method_vs_set,function(method){
-  p = ggplot(df_benchmark, aes(x=!!as.name(method), y=!!as.name(method_vs))) + geom_point(size = 1,col = "blue") +
-    geom_abline(intercept = 0, slope = 1, col = "grey", linetype = "dashed") + scale_x_log10(breaks = 10^(0:4), labels = expression(10^0, 10^1, 10^2, 10^3, 10^4)) +
-    scale_y_log10(breaks = 10^(0:4), labels = expression(10^0, 10^1, 10^2, 10^3, 10^4)) +
-    expand_limits(x = c(1, 10^4), y = c(1, 10^4)) +
-    labs(x = paste0(method_fullname[method]," Rank"), y = paste0(method_fullname[method_vs]," Rank")) + theme_minimal()
-
-  # annotate gene labels
-  library(ggrepel)
-  df_label = data.frame(df_benchmark[,c(method,method_vs)],label = rownames(df_benchmark))
-  sub = c(which(df_label[,method] <= 100 & df_label[,method_vs] > 500),
-          which(df_label[,method] > 500 & df_label[,method_vs] <= 100) )
-  df_label = df_label[sub,]
-  df_label = df_label[order(df_label[,1]),]
-  p = p + geom_text_repel(data = df_label, aes(label=label),col = "red",arrow = arrow(type = "closed", length = unit(0.01, "inches")),max.overlaps = 200)
-  p
-})
-fig = cowplot::plot_grid(plotlist = pl, ncol = length(method_vs_set))
-fig
+method_vs_set = c("wilcox_default","hotspot","marcopolo","semi","haystack")
 ```
@@
 }))
-df_label_all = df_benchmark[unique(sub),c(method_vs_set,method_vs)]
+df_label_all = df_benchmark[unique(sub),c(method_vs,method_vs_set)]
 df_label_all$'legend' = sapply(seq(nrow(df_label_all)),function(i){
-  paste(paste(method_fullname[colnames(df_label_all)[c(4,1:3)]],df_label_all[i,c(4,1:3)],sep = ":"),collapse = "\n")
+  paste(paste(method_fullname[colnames(df_label_all)],df_label_all[i,],sep = ":"),collapse = "\n")
 })
-
-g = c("Mogat2","Vpreb2","Cd19","Mup3","Fyb","Naaa")
+g = c("Mogat2","Vpreb2","Cd19","Abca13","Il2rb","Mup3","Fyb","Naaa","Zf12","Phgdh")
 pl = lapply(g,function(x){
@@
 })
-fig = cowplot::plot_grid(plotlist = pl,ncol = 3)
+fig = cowplot::plot_grid(plotlist = pl,ncol = length(method_vs_set))
+fig
+```
+```{r}
+pl_rank = lapply(method_vs_set,function(method){
+  df_label = data.frame(df_benchmark[,c(method,method_vs)],label = rownames(df_benchmark))
+  df_label[c(which(df_label[,method] <= 100 & df_label[,method_vs] > 500)),"color"] = sprintf("Favor by %s (rank <= 100)",method)
+  df_label[c(which(df_label[,method] > 500 & df_label[,method_vs] <= 100) ),"color"] = sprintf("Favor by %s (rank <= 100)",method_vs)
+  df_label$'color' = as.factor(df_label$'color')
+  p = ggplot(df_label, aes(x=!!as.name(method), y=!!as.name(method_vs),color = color)) + geom_point(size = 1) + 
+  scale_color_manual(values = c("red", "blue", "grey50"),
+                     breaks = levels(df_label$'color')) + 
+    geom_abline(intercept = 0, slope = 1, col = "grey", linetype = "dashed") + scale_x_log10(breaks = 10^(0:4), labels = expression(10^0, 10^1, 10^2, 10^3, 10^4)) +
+    scale_y_log10(breaks = 10^(0:4), labels = expression(10^0, 10^1, 10^2, 10^3, 10^4)) +
+    expand_limits(x = c(1, 10^4), y = c(1, 10^4)) +
+    labs(x = paste0(method_fullname[method]," Rank"), y = paste0(method_fullname[method_vs]," Rank"), color = "Highlighted Genes") + theme_minimal() 
+  # annotate gene labels
+  # library(ggrepel)
+  p = p + geom_text(data = df_label %>% filter(!is.na(color)) %>% filter(label %in% g), aes(label=label))
+  p
+})
+fig = cowplot::plot_grid(plotlist = pl_rank, ncol = length(method_vs_set))
 fig
@@
 load("./marrow_result/module_result.rdata")
+source("./LMD_function.R", echo = F)
+library(RColorBrewer)
+library(dplyr)
 module_color = setNames(colorRampPalette(brewer.pal(12, "Paired"))(nlevels(gene_partition)),levels(gene_partition))
@@
 # Bipartite Plot
+jaccard_thred = 0.4
+
+library(igraph)
+library("grid") 
+library("gridExtra")
+
 df_edge <- as.data.frame(as.table(jaccard_index))
 colnames(df_edge) = c("from","to","weight")
-df_edge <- subset(df_edge, weight > 0.4)
+df_edge <- subset(df_edge, weight > jaccard_thred)
 df_edge$width = ifelse(df_edge$weight > 0.8,2,0.5)
@@
 df_node$type = ifelse(df_node$name %in% names(module_color),TRUE,FALSE)
+
+# Define specific color for module
+rownames(df_node) = df_node$name
+ct_module = names(which(apply(jaccard_index,2,max) > jaccard_thred))
+ct_subtype_module = c(4,5,6,30,39,42,43)
+ct_multype_module = c(2,13,19,33,36,54)
+cc_module = c(16,17,18,31)
+df_node[names(module_color),"color"] = "grey"
+df_node[ct_module,"color"] = "red"
+df_node[ct_multype_module,"color"] = "purple"
+df_node[ct_subtype_module,"color"] = "yellow"
+df_node[cc_module,"color"] = "green"
 
@@
      vertex.label.degree=pi/2)
-coords = as.data.frame(l) %>% filter(V2 == 0) %>% mutate(labels = names(celltype_color))
-text(x = coords$V1, y = coords$V2, labels = coords$labels, srt = 90, adj = c(1.5, 0.5),cex = 0.5,col = "black")
-# Add legend for colors
-legend("bottom", legend=names(celltype_color), fill=celltype_color, bty="n", cex=0.5)
+
+# Create Node Legend & Edge Legend
+color_label = c("green" = "cell cycle",purple = "multi cell type",red = "cell type",yellow = "cell subtype")
+df_node$subtype = color_label[df_node$color]
+
+p <- ggplot(df_node %>% filter(type), aes(x = 0, y = 0,color = subtype)) + geom_point(size = 7) + scale_color_manual(values= setNames(names(color_label),color_label)) + theme_minimal() + labs(color = "Module Type") + theme(legend.direction="horizontal")
+grid.newpage()
+grid.draw(get_legend(p))
+
+p <- ggplot(df_node %>% filter(!type), aes(x = 0, y = 0,color = name)) + geom_point(size = 7,shape = 15) + scale_color_manual(values= df_node %>% filter(!type) %>% pull(color,name)) + theme_minimal() + labs(color = "Cell Type") + theme(legend.direction="horizontal")
+grid.newpage()
+grid.draw(get_legend(p))
 ```

### 2023-10-27 Demo_marrow_facs.Rmd
cellmarkerdb_change

--- 2023-10-19 Demo_marrow_facs.Rmd
@@
 library("SeuratWrappers")
+library("pheatmap")
+library("ggplot2")
+library("ggplotify")
 ```
@@
 
-### Visualize Module on Cell embeddings
+## Step3: Identify Celltype-specific activated Modules
+```{r}
+cell_label = as.factor(tiss$cell_ontology_class)
+abbrev = levels(cell_label)
+abbrev[grepl("T cell",abbrev)] = "T cell"
+abbrev[grepl("natural killer",abbrev)] = "NK cell"
+abbrev[abbrev == "B cell"] = "Cd3e+ Klrb1+ B cell"
+abbrev[grepl("Slamf1-positive",abbrev)] = "Slamf1+ MPC"
+abbrev[grepl("Slamf1-negative",abbrev)] = "Slamf1- MPC"
+abbrev[grepl("common lymphoid",abbrev)] = "CLP"
+abbrev[grepl("megakaryocyte",abbrev)] = "MEP"
+abbrev[grepl("hematopoietic",abbrev)] = "HSC"
+abbrev[grepl("hematopoietic",abbrev)] = "HSC"
+abbrev[grepl("granulocyte monocyte",abbrev)] = "GMP"
+# abbrev[grepl("granulocytopoietic",abbrev)] = "G-poiesis"
+names(abbrev) = levels(cell_label)
+levels(cell_label) = abbrev[levels(cell_label)]
 
-```{r, fig.width=20, fig.asp=1}
-pl = lapply(levels(gene_partition),function(module){
-  FeaturePlot_custom(value = cell_block[,module], coord = visual_space, value_name = "Module Score",title_name = sprintf("Module %s (%d genes)",module, sum(gene_partition == module)))
-})
-plot_grid(plotlist = pl,ncol = 5) # may takes long time to plot
-```
-
-## Step3: Identify Celltype-specific activated Modules
-
-```{r}
 # Create one-hot vector of Cell Type
@@
 
-# Order Cell Type based on its centroid in Featurespace
-Idents(tiss) = cell_label
-ordered_cell_types <- unique(cell_label)[hclust(dist(t(sapply(unique(cell_label), function(x) colMeans(Embeddings(tiss, reduction = "pca")[WhichCells(tiss, idents = x), ,drop = FALSE]) ))))$order]
+# # Order Cell Type based on its centroid in Featurespace
+# Idents(tiss) = cell_label
+# ordered_cell_types <- unique(cell_label)[hclust(dist(t(sapply(unique(cell_label), function(x) colMeans(Embeddings(tiss, reduction = "pca")[WhichCells(tiss, idents = x), ,drop = FALSE]) ))))$order]
 
@@
 })
-jaccard_index = jaccard_index[rev(ordered_cell_types),]
-rm(ordered_cell_types)
+# jaccard_index = jaccard_index[rev(ordered_cell_types),]
+# rm(ordered_cell_types)
 ```
 
-### Visualize Heatmap of Jaccard Index Matrix
+### Jaccard Index Heatmap
+```{r}
+thred = 0.4
+jaccard_index_pruned = jaccard_index[apply(jaccard_index,1,max) > thred,]
+jaccard_index_pruned = jaccard_index_pruned[,apply(jaccard_index_pruned,2,max) > thred]
 
-```{r}
-pheatmap(jaccard_index,
+colnames(jaccard_index_pruned) = paste0("M",colnames(jaccard_index_pruned))
+jaccard_index_pruned = jaccard_index_pruned[names(sort(apply(jaccard_index_pruned,1,which.max))),]
+p = pheatmap(t(jaccard_index_pruned),
          cluster_rows=FALSE, cluster_cols=FALSE, 
          scale = "none", 
-         color = colorRampPalette(rev(c("firebrick3", "white")))(99))
+         color = colorRampPalette(rev(c("firebrick3", "white")))(99),
+         angle_col = 45,legend_title = "Jaccard Index",silent = TRUE,
+         cellwidth = 15, cellheight = 10)
+p_jaccard = as.ggplot(p) + coord_fixed()
+ggsave(filename = "jaccard_heatmap.png", plot = p_jaccard, width = 5, height = 10)
 ```
 
-### Visualize How CellType-Module is Matched on Cell Embeddings
-
+### CellType-Module Matched on Cell Embeddings
 For each cell type, select the most specific modules for visualization (maximum jaccard index)
-
 ```{r}
-tiss@meta.data = cbind(tiss@meta.data,one_hot_matrix)
-jaccard_thred = 0.4
-
 # Get Module plot
 pl_module = lapply(levels(gene_partition),function(module){
-  FeaturePlot_custom(value = cell_block[,module], coord = visual_space, value_name = "Module Score",title_name = sprintf("Module %s (%d genes)",module, sum(gene_partition == module)))
+  FeaturePlot_custom(value = cell_block[,module], coord = visual_space, value_name = "Module Score",title_name = sprintf("Module %s (%d genes)",module, sum(gene_partition == module))) + scale_color_gradient(limits = c(0, 1),low = "lightgrey", high = "blue")
 })
@@
 pl_ct = lapply(colnames(one_hot_matrix),function(ct){
-  FeaturePlot_custom(value = as.factor(one_hot_matrix[,ct]), coord = visual_space, value_name = "Indicator",title_name = ct)
+  FeaturePlot_custom(value = as.factor(one_hot_matrix[,ct]), coord = visual_space, value_name = "Indicator",title_name = ct) + scale_color_manual(values = c("0" = "lightgrey", "1" = scales::hue_pal()(1)))
 })
@@
 
-pl <- lapply(rownames(jaccard_index)[rev(order(apply(jaccard_index,1,max)))], function(celltype){
-  if(max(jaccard_index[celltype,]) > jaccard_thred){
-    p1 = pl_ct[[celltype]]
-    module_name = colnames(jaccard_index)[which.max(jaccard_index[celltype,])]
-    p2 = pl_module[[module_name]]
-    p = as.ggplot(p1) + as.ggplot(p2) + plot_annotation(paste0("Jaccard Index: ",round(jaccard_index[celltype,module_name],2)))
-    p
-  }
+ct_module_match = apply(jaccard_index_pruned,1,function(x) colnames(jaccard_index_pruned)[which.max(x)])
+ct_module_match = gsub("M","",ct_module_match)
+
+pl <- lapply(seq(length(ct_module_match)),function(i){
+    (pl_module[[ct_module_match[i]]] + NoLegend())/
+    (pl_ct[[names(ct_module_match)[i]]] + NoLegend())
 })
+
+ncol = 7
+legend_module = get_legend(pl_module[[1]])
+legend_ct = get_legend(pl_ct[[1]])
+bottom_row = plot_grid(
+  plot_grid(plotlist = pl[(ncol+1):length(pl)],ncol = length(pl) - ncol), 
+  plot_grid(legend_module, legend_ct, ncol = 1),
+  rel_widths = c((ncol-1),1),nrow = 1)
+p = plot_grid(
+  plot_grid(plotlist = pl[1:ncol], ncol = ncol),
+  bottom_row,
+  ncol = 1)
+rm(bottom_row)
+
+ggsave(filename = "ct-specific-module.png", plot = p, width = 35, height = 20)
 ```
@@
 ### Annotate Each Celltype-specific activated Modules with its corresponding Celltypes
-
 ```{r}
@@
 
+## Transferability of Modules
+### Load Data & Calculate Module Activity
+```{r}
+if (!file.exists('droplet_Marrow_seurat_tiss.Robj')) {
+   # If load url doesn't work, please download it and load from your working directory
+  options(timeout=6000)
+  download.file("https://figshare.com/ndownloader/files/13092380",
+                destfile = 'droplet_Marrow_seurat_tiss.Robj', method = 'libcurl')
+ }
+load('droplet_Marrow_seurat_tiss.Robj')
+tiss_droplet <- UpdateSeuratObject(tiss_droplet)
+
+# path <- "/data/ruiqi/local_marker/dataset/tabular_muris/"
+# tissue_name = "marrow_droplet"
+# tiss_droplet <- readRDS(sprintf("%s%s.rds",path,tissue_name))
+
+feature_space_droplet = as.matrix(tiss_droplet@reductions$pca@cell.embeddings[,1:n_dim])
+dat_droplet = as.matrix(tiss_droplet[["RNA"]]@data)
+
+knn = 10; major_vote = 5
+A = Symmetric_KNN_graph(knn = knn, feature_space = feature_space_droplet)$'adj_matrix'
+cell_block_loop = lapply(1:100,function(loop){
+  GMM_subsampling(seed = loop, gene_partition, expr_dat = dat_droplet, cell_kNN_graph = A, major_vote)
+})
+cell_block_prop = Reduce(`+`, cell_block_loop) / length(cell_block_loop)
+cell_block_droplet = cell_block_prop
+```
+
+### Calculate Jaccard Index
+```{r}
+cell_label = tiss_droplet$cell_ontology_class
+cell_label[!cell_label %in% names(abbrev)] = NA
+cell_label = setNames(abbrev[cell_label],names(cell_label))
+
+one_hot_matrix <- sapply(unique(na.omit(cell_label)), function(x) as.integer(cell_label == x))
+one_hot_matrix[is.na(one_hot_matrix)] = 0
+
+jaccard_index_droplet <- apply(cell_block_droplet, 2, function(module_col) {
+  apply(one_hot_matrix, 2, function(one_hot_col) {
+    1 - philentropy::distance(rbind(module_col,one_hot_col), method = "jaccard", mute.message=TRUE)
+  })
+})
+```
+### Check whether the modules specific in FACs still specific in droplet
+```{r}
+module_selected = unlist(apply(jaccard_index[rownames(jaccard_index_droplet),],1,function(x) colnames(jaccard_index)[which(x > 0.4)]))
+module_selected = module_selected[order(as.numeric(module_selected))]
+ct_selected = rownames(jaccard_index_pruned)[rownames(jaccard_index_pruned)%in% rownames(jaccard_index_droplet)] 
+jaccard_index_pruned = jaccard_index_droplet[ct_selected,module_selected]
+colnames(jaccard_index_pruned) = paste0("M",colnames(jaccard_index_pruned))
+
+p = pheatmap(t(jaccard_index_pruned),
+         cluster_rows=FALSE, cluster_cols=FALSE, 
+         scale = "none", 
+         color = colorRampPalette(rev(c("firebrick3", "white")))(99),
+         angle_col = 45,legend_title = "Jaccard Index",silent = TRUE,
+         cellwidth = 15, cellheight = 10)
+p_jaccard = as.ggplot(p) + coord_fixed()
+ggsave(filename = "fig-cross-dataset-jaccard.png", plot = p_jaccard, width = 5, height = 10)
+```
+
 ## Save Intermediate data
 ```{r}
-save(gene_partition,dist,gene_hree,cell_block,jaccard_index,file = "./marrow_result/module_result.rdata")
+save(gene_partition,dist,gene_hree,cell_block,cell_block_droplet,jaccard_index,file = "./marrow_result/module_result.rdata")
 ```
@@
 df_module[,"celltype"] = module_celltype[levels(gene_partition)]
-df_module[top_go_clusters,"GO_description"] = unlist(lapply(top_enrich_go,function(x) paste(x[['Description']],collapse = ",")))
-df_module[top_pathway_clusters,"Pathway_description"] = unlist(lapply(top_enrich_pathway,function(x) paste(x[['Description']],collapse = ",")))
+df_module[top_go_clusters,"GO_description"] = unlist(lapply(top_enrich_go,function(x){
+  des = x[["Description"]]
+  paste(des[!is.na(des)],collapse = ",")
+}
+))
+df_module[top_pathway_clusters,"Pathway_description"] = unlist(lapply(top_enrich_pathway,function(x){
+  des = x[["Description"]]
+  paste(des[!is.na(des)],collapse = ",")
+}))
 df_module[is.na(df_module)] = ""
-write.csv(df_module,file = "module_description.csv",row.names = FALSE)
+write.csv(df_module,file = "./marrow_result/module_description.csv",row.names = FALSE)
 ```
+
+## Module-analysis
+```{r}
+df_module = read.csv("./marrow_result/module_description.csv")
+df_module$'multi-celltype' = NA
+df_module$'note' = NA
+```
+
+### Cell Cycle Modules
+```{r}
+library(gprofiler2)
+mmus_s = gorth(cc.genes.updated.2019$s.genes, source_organism = "hsapiens", target_organism = "mmusculus")$ortholog_name
+mmus_g2m = gorth(cc.genes.updated.2019$g2m.genes, source_organism = "hsapiens", target_organism = "mmusculus")$ortholog_name
+tiss = CellCycleScoring(tiss, s.features = mmus_s, 
+                        g2m.features = mmus_g2m, set.ident = FALSE)
+p = as.ggplot(DimPlot(tiss,group.by = "Phase", cols = c("grey",scales::hue_pal()(3)[2:3])))
+
+# df_module[c(16,18,17,31,32),'note'] = "cell_cycle"
+
+p = plot_grid(plotlist = c(list(p),pl_module[c(16,18,17,31,32)]),ncol = 3)
+ggsave(filename = "fig-cellcycle.png", plot = p, width = 15, height = 10)
+```
+
+### Multi-resolution Modules
+```{r}
+ct_prob_per_module = do.call(rbind,lapply(unique(cell_label),function(x){apply(cell_block[cell_label == x,],2,median)}))
+rownames(ct_prob_per_module) = unique(cell_label)
+ct_num_per_module = sort(apply(ct_prob_per_module,2,function(x)sum(x > 0.5)))
+
+m = 19
+df_module[m,"celltype"] = paste0(names(which(ct_prob_per_module[,m] > 0.5)),collapse = ",")
+
+df_module[54,"multi-celltype"] = c("T/NK/NKT")
+df_module[27,"multi-celltype"] = c("Progenitor/NK/NKT")
+df_module[33,"multi-celltype"] = c("Progenitor/basophil/granulocytopoietic/half-granulocyte")
+df_module[46,"multi-celltype"] = c("monocyte/granulocytopoietic/half-granulocyte")
+df_module[48,"multi-celltype"] = c("monocyte/granulocytopoietic/granulocyte")
+df_module[49,"multi-celltype"] = c("monocyte/macrophage/granulocytopoietic/granulocyte")
+df_module[55,"multi-celltype"] = c("NK")
+df_module[56,"multi-celltype"] = c("NK")
+df_module[13,"multi-celltype"] = c("B")
+df_module[50,"multi-celltype"] = c("monocyte/granulocytopoietic/granulocyte/basophil")
+df_module[28,"multi-celltype"] = c("Progenitor")
+df_module[9,"multi-celltype"] = c("immature/naive B")
+df_module[11,"multi-celltype"] = c("late pro/pre B")
+df_module[7,"multi-celltype"] = c("late pro/pre B")
+df_module[44,"multi-celltype"] = c("granulocytopoietic/half-granulocyte")
+df_module[47,"multi-celltype"] = c("granulocytopoietic/granulocyte")
+df_module[52,"multi-celltype"] = c("T")
+df_module[53,"multi-celltype"] = c("macrophage/NK")
+df_module[1,"multi-celltype"] = c("monocyte/macrophage/granulocyte monocyte progenitor cell")
+df_module[20,"multi-celltype"] = c("granulocytopoietic/Progenitor")
+df_module[c(42,35),"multi-celltype"] = c("below-half-granulocyte/monocyte")
+df_module[c(36,38),"multi-celltype"] = c("granulocyte monocyte progenitor cell/monocyte")
+df_module[2,"multi-celltype"] = c("granulocyte monocyte progenitor cell/macrophage")
+df_module[4,"celltype"] = c("granulocyte monocyte progenitor cell")
+df_module[c(5,6,43),"note"] = "subtype"
+```
+
+### Joint-CT Module visualization
+```{r}
+level0_module = c(46,13,54,28,38)
+# level1_module = c(45,41,51,24,14,4,37,3,8,10,52,56)
+ncol = length(level0_module)
+legend_module <- get_legend(pl_module[[1]])
+prow <- cowplot::plot_grid(plotlist = lapply(pl_module[level0_module],function(x) x + NoLegend()), ncol = ncol)
+p = plot_grid(prow, legend_module, rel_widths = c(ncol, .2))
+
+ggsave(filename = "fig-joint-module.png", plot = p, width = 25, height = 5)
+```
+
+### Subtype Investigation
+```{r}
+# coldef = setNames(colorRampPalette(brewer.pal(12, "Paired"))(length(unique(cell_label))),unique(cell_label))
+# 
+# centroids <- aggregate(. ~ cell_label, data = as.data.frame(Embeddings(tiss, "tsne")), FUN = mean)
+# p <- DimPlot(tiss, group.by = "cell_ontology_class",label = FALSE, cols = coldef) + theme_void() + NoLegend() + ggtitle("Bone Marrow FACS")
+# p = p + 
+#   geom_text_repel(data = centroids, aes(x = tSNE_1, y = tSNE_2, label = cell_label, color = cell_label),
+#                    box.padding = 0.3, point.padding = 0.3, segment.color = 'grey50',max.overlaps = 100)
+
+p <- DimPlot(tiss, group.by = "cell_ontology_class",label = FALSE, cols = setNames(scales::hue_pal()(2),c("late pro-B cell","granulocyte"))) + ggtitle("pre-annotated cell types")
+pl_module_subtype =
+
+pl1 = FeaturePlot(tiss,features = names(gene_partition)[gene_partition %in% c(5,6)][c(1:3,7,9)],ncol = 5,order = TRUE)
+pl2 = FeaturePlot(tiss,features = names(gene_partition)[gene_partition == 43][1:5],ncol = 5,order = TRUE)
+pl_module_subtype = lapply(pl_module[c(5,6,43)],function(x) x + scale_color_gradient(low = "lightgrey", high = "blue"))
+
+combined_p = plot_grid(p,plotlist = pl_module_subtype,ncol = 2) + as.ggplot(pl1) + as.ggplot(pl2) + plot_layout(design = "\nAAAABBBBBBB\nAAAACCCCCCC\n")
+
+ggsave(filename = "fig-subtype.png", plot = combined_p, width = 35, height = 10)
+```

### 2023-10-27 Marker_identfication_benchmark.Rmd
cellmarkerdb_change

@@
 path = "./benchmark_result/"
-# tissue_name = "marrow_facs"
+tissue_name = "marrow_facs"
 ```
@@
 Downloaded marker from http://bio-bigdata.hrbmu.edu.cn/CellMarker/CellMarker_download_files/file/Cell_marker_Seq.xlsx
-PanglaoDB: https://panglaodb.se/markers/PanglaoDB_markers_27_Mar_2020.tsv.gz and exclude genes that reported only once
+<!-- PanglaoDB: https://panglaodb.se/markers/PanglaoDB_markers_27_Mar_2020.tsv.gz and exclude genes that reported only once -->
 ```{r}
 # Load CellMarkerDB
-cell_marker_db = readxl::read_xlsx("/data/ruiqi/local_marker/dataset/Cell_Marker/Cell_marker_All.xlsx") %>% filter(species == "Mouse") %>% filter(tissue_class == "Bone marrow") %>% filter(!is.na(cellontology_id)) %>% filter(!is.na(PMID)) %>% mutate(gene_name1 = marker) %>% mutate(gene_name2 = Symbol) %>% select(c("gene_name1","gene_name2")) %>% group_by(across(everything())) %>% dplyr::summarise(count = n(), .groups = "drop") %>% filter(count > 1)
+cell_marker_db = readxl::read_xlsx("/data/ruiqi/local_marker/dataset/Cell_Marker/Cell_marker_All.xlsx") %>% filter(species == "Mouse") %>% filter(tissue_class == "Bone marrow") %>% filter(!is.na(cellontology_id)) %>% filter(!is.na(PMID)) %>% mutate(gene_name1 = marker) %>% mutate(gene_name2 = Symbol) %>% select(c("gene_name1","gene_name2")) %>% group_by(across(everything())) %>% dplyr::summarise(count = n(), .groups = "drop") %>% filter(count >= 1) # Lung, Pancreas
 
@@
 df_benchmark = read.table(paste0(path,tissue_name,"_benchmark_rank_table.csv"))
+
+method_ls = c("lmd","haystack","hotspot","marcopolo","semi","wilcox_default","wilcox_no_filter","hvg")
+method_fullname = setNames(c("LMD","singleCellHaystack","Hotspot","Marcopolo","SEMITONES","Seurat v3 default", "Seurat v3 Wilcoxon unfiltered","HighlyVariableGenes"),method_ls)
 ```
@@
 celldb_marker = read.table(paste0(path,tissue_name,"_ground_truth_c2.txt"))[,1]
-
 gt_list = c(lapply(seq(50,1000,50),function(x) max_logfc[1:x]),list(celldb_marker))
 names(gt_list) = c(paste0("Top",seq(50,1000,50)),"CellMarkerDB")
+
+# AUC table
 df = do.call(rbind,lapply(1:length(gt_list),function(i){
@@
 write.table(df,file = paste0(path,tissue_name,"_auroc.csv"),row.names = FALSE)
+```
+
+```{r}
+# Concordance table
+do.call(cbind,lapply(gt_list[c(2,21)],function(true_marker){
+  df_benchmark$"gt" = 0
+  df_benchmark[true_marker,"gt"] = 1
+  apply(df_benchmark,2,function(x){
+    sum(df_benchmark[order(x),"gt"][1:100])
+  })
+}))
+
+# AUC curve
+p = ggplot(roc_df, aes(x = FPR, y = TPR, color = label)) +
+  geom_line(size = 0.5) +
+  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray") + labs(title = sprintf("CellMarkerDB (%d genes)",length(true_marker)),
+     x = "False Positive Rate",
+     y = "True Positive Rate") +  scale_colour_manual(values= coldef)
 ```

### 2023-10-27 paper_figure.Rmd
cellmarkerdb_change

@@
 library("ggplot2")
-path = "./benchmark_result/"
-tissue_name = "marrow_facs"
 
-df_benchmark = read.table(paste0(path,tissue_name,"_benchmark_rank_table.csv"))
-method_ls = colnames(df_benchmark)
+# Set the color code for all methods
+method_ls = c("lmd","haystack","hotspot","marcopolo","semi","wilcox_default","wilcox_no_filter","hvg")
 method_fullname = setNames(c("LMD","singleCellHaystack","Hotspot","Marcopolo","SEMITONES","Seurat v3 default", "Seurat v3 Wilcoxon unfiltered","HighlyVariableGenes"),method_ls)
@@
 ```{r}
-df_auc <- read.table(paste0(path,tissue_name,"_auroc.csv"),header = TRUE)
+path = "./benchmark_result/"
+tissue_ls = c("marrow_facs","pancreas_facs","lung_facs")
+df_auc = do.call(rbind,lapply(tissue_ls,function(tissue_name){df = read.table(paste0(path,tissue_name,"_auroc.csv"),header = TRUE); df$'Tissue' = tissue_name; df}))
+
 df_auc = df_auc %>% filter(gt_set %in% c("Top100","CellMarkerDB")) %>% filter(!Method %in% "HighlyVariableGenes")
 df_auc$Method <- with(df_auc, reorder(Method, -AUROC, FUN = median))
-df_auc$gt_set <- factor(df_auc$gt_set,levels = c("Top100","CellMarkerDB"))
-levels(df_auc$gt_set) = c("Criterion1 - MaxFC", "Criterion2 - CellMarkerDB")
-
+df_auc$gt_set <- factor(df_auc$gt_set,levels = c("CellMarkerDB","Top100"))
+levels(df_auc$gt_set) = c("Criterion1 - CellMarkerDB","Criterion2 - MaxFC")
+df_auc$Tissue = factor(df_auc$Tissue,levels = tissue_ls)
+levels(df_auc$Tissue) = c("Bone Marrow","Pancreas","Lung")
 # Create the plot
-fig <- ggplot(df_auc, aes(x = gt_set, y = AUROC, fill = Method)) +
+fig <- ggplot(df_auc, aes(x = Tissue, y = AUROC, fill = Method)) +
   geom_bar(stat = "identity", position = "dodge") +
@@
   theme_minimal() +
-  labs(x = "Ground-Truth Genesets", y = "AUROC", fill = "Method") + coord_cartesian(ylim = c(0.5, 1)) +
-  theme(legend.position = "bottom")  +  scale_fill_manual(values= coldef[names(coldef)%in%df_auc$Method])
-fig + ggtitle(tissue_name)
+  labs(x = "Tissue", y = "AUROC", fill = "Method") + coord_cartesian(ylim = c(0.5, 1)) +
+  theme(legend.position = "bottom")  +  scale_fill_manual(values= coldef[names(coldef)%in%df_auc$Method]) + facet_wrap(~gt_set)
+
+ggsave(filename = "fig-known_marker_comparison.png", plot = fig, width = 10, height = 5)
 ```
@@
 })
-g = c("Mogat2","Vpreb2","Cd19","Abca13","Il2rb","Mup3","Fyb","Naaa","Zf12","Phgdh")
+g = c("Cd19","Naaa","Mogat2","Tifa","Abca13","Zf12","Il2rb","Phgdh","Vpreb2","Fyb")
 pl = lapply(g,function(x){
@@
 })
-fig = cowplot::plot_grid(plotlist = pl,ncol = length(method_vs_set))
-fig
+fig = cowplot::plot_grid(plotlist = pl[as.numeric(t(matrix(1:length(pl),nrow = 2)))],ncol = length(method_vs_set))
+ggsave(filename = "fig-ranking_comparison.png", plot = fig, width = 25, height = 10)
 ```
@@
   # library(ggrepel)
-  p = p + geom_text(data = df_label %>% filter(!is.na(color)) %>% filter(label %in% g), aes(label=label))
+  p = p + geom_text_repel(data = df_label %>% filter(!is.na(color)) %>% filter(label %in% g), aes(label=label),min.segment.length = unit(0, 'lines'),nudge_y = -0.5)
   p
 })
-fig = cowplot::plot_grid(plotlist = pl_rank, ncol = length(method_vs_set))
-fig
+pdf(width = 25,height = 5)
+legend = get_legend(pl_rank[[1]] +  guides(color=guide_legend(direction="horizontal")))
+prow = plot_grid(plotlist = lapply(pl_rank,function(x) x + NoLegend()), ncol = length(method_vs_set))
+plot_grid(prow, legend, ncol = 1, rel_heights = c(1, .2))
+dev.off()
 ```
 
-# Fig - Gene Co-localization Module Benchmark
-```{r}
-library(cowplot)
-library(patchwork)
-source("./LMD_function.R", echo = F)
+<!-- # Fig - Gene Co-localization Module Benchmark -->
+<!-- ```{r} -->
+<!-- library(cowplot) -->
+<!-- library(patchwork) -->
+<!-- source("./LMD_function.R", echo = F) -->
 
-path = "./benchmark_result/"
-tissue_name = "marrow_facs"
+<!-- path = "./benchmark_result/" -->
+<!-- tissue_name = "marrow_facs" -->
 
-df_benchmark = read.table(paste0(path,tissue_name,"_benchmark_rank_table.csv"))
-method = "lmd"
-topn = 2000
-# Gene prioritized by lmd
-pl1 = lapply(c("wilcox_default","haystack","hotspot","hvg"),function(method_vs){
-  local_gene = setdiff(rownames(df_benchmark)[order(df_benchmark[,method_vs])][1:topn], rownames(df_benchmark)[order(df_benchmark[,method])][1:topn])
+<!-- df_benchmark = read.table(paste0(path,tissue_name,"_benchmark_rank_table.csv")) -->
+<!-- method = "lmd" -->
+<!-- topn = 2000 -->
+<!-- # Gene prioritized by lmd -->
+<!-- pl1 = lapply(c("wilcox_default","haystack","hotspot","hvg"),function(method_vs){ -->
+<!--   local_gene = setdiff(rownames(df_benchmark)[order(df_benchmark[,method_vs])][1:topn], rownames(df_benchmark)[order(df_benchmark[,method])][1:topn]) -->
 
-  ## Gene-gene distance(similarity)---------
-  distance_method = "jaccard"
-  clustering_method = "average"
-  dist = Calculate_distance(dat_alra[local_gene,], method = distance_method)
+<!--   ## Gene-gene distance(similarity)--------- -->
+<!--   distance_method = "jaccard" -->
+<!--   clustering_method = "average" -->
+<!--   dist = Calculate_distance(dat_alra[local_gene,], method = distance_method) -->
 
-  gene_hree = hclust(dist, method = clustering_method)
-  gene_partition = cutree(gene_hree,k=5)
-  names(gene_partition) = labels(dist)
+<!--   gene_hree = hclust(dist, method = clustering_method) -->
+<!--   gene_partition = cutree(gene_hree,k=5) -->
+<!--   names(gene_partition) = labels(dist) -->
 
-  top5 <- names(sort(table(gene_partition),decreasing = TRUE))[1:5]
-  annotation_df = data.frame(Module = gene_partition)
-  ann_colors = list(
-    Module = c(setNames(colorRampPalette(brewer.pal(9, "Set1"))(5),1:5)))
+<!--   top5 <- names(sort(table(gene_partition),decreasing = TRUE))[1:5] -->
+<!--   annotation_df = data.frame(Module = gene_partition) -->
+<!--   ann_colors = list( -->
+<!--     Module = c(setNames(colorRampPalette(brewer.pal(9, "Set1"))(5),1:5))) -->
 
-  p_heatmap = pheatmap(as.matrix(dist),
-                annotation_col = annotation_df,
-                cluster_rows = gene_hree,
-                cluster_cols = gene_hree,
-                treeheight_row = 0,
-                annotation_colors = ann_colors, 
-                col=colorRampPalette(c("red","white","blue4"))(99),
-                show_colnames = FALSE,
-                show_rownames = FALSE,
-                annotation_legend = TRUE,
-                silent = TRUE)
-  pl_module = FeaturePlot_meta(dat,visual_space,gene_partition)
-  legend <- get_legend(pl_module[[1]])
-  prow <- cowplot::plot_grid(plotlist = lapply(pl_module,function(x) x + theme(legend.position="none")), ncol = ceiling(length(pl_module)/2))
+<!--   p_heatmap = pheatmap(as.matrix(dist), -->
+<!--                 annotation_col = annotation_df, -->
+<!--                 cluster_rows = gene_hree, -->
+<!--                 cluster_cols = gene_hree, -->
+<!--                 treeheight_row = 0, -->
+<!--                 annotation_colors = ann_colors,  -->
+<!--                 col=colorRampPalette(c("red","white","blue4"))(99), -->
+<!--                 show_colnames = FALSE, -->
+<!--                 show_rownames = FALSE, -->
+<!--                 annotation_legend = TRUE, -->
+<!--                 silent = TRUE) -->
+<!--   pl_module = FeaturePlot_meta(dat,visual_space,gene_partition) -->
+<!--   legend <- get_legend(pl_module[[1]]) -->
+<!--   prow <- cowplot::plot_grid(plotlist = lapply(pl_module,function(x) x + theme(legend.position="none")), ncol = ceiling(length(pl_module)/2)) -->
 
-  as.ggplot(p_heatmap) + cowplot::plot_grid(prow, legend, rel_widths = c(ceiling(length(pl_module)/2), .2)) + plot_annotation(title = sprintf("Genes in top 2000 by %s but not by %s (N = %d)",method_fullname[method_vs],method_fullname[method],length(local_gene))) + plot_layout(ncol = 2, widths = c(2, 3))
-})
-# The other way around
-pl2 = lapply(c("wilcox_default","haystack","hotspot","hvg"),function(method_vs){
-  local_gene = setdiff(rownames(df_benchmark)[order(df_benchmark[,method])][1:topn], rownames(df_benchmark)[order(df_benchmark[,method_vs])][1:topn])
+<!--   as.ggplot(p_heatmap) + cowplot::plot_grid(prow, legend, rel_widths = c(ceiling(length(pl_module)/2), .2)) + plot_annotation(title = sprintf("Genes in top 2000 by %s but not by %s (N = %d)",method_fullname[method_vs],method_fullname[method],length(local_gene))) + plot_layout(ncol = 2, widths = c(2, 3)) -->
+<!-- }) -->
+<!-- # The other way around -->
+<!-- pl2 = lapply(c("wilcox_default","haystack","hotspot","hvg"),function(method_vs){ -->
+<!--   local_gene = setdiff(rownames(df_benchmark)[order(df_benchmark[,method])][1:topn], rownames(df_benchmark)[order(df_benchmark[,method_vs])][1:topn]) -->
 
-  ## Gene-gene distance(similarity)---------
-  distance_method = "jaccard"
-  clustering_method = "average"
-  dist = Calculate_distance(dat_alra[local_gene,], method = distance_method)
+<!--   ## Gene-gene distance(similarity)--------- -->
+<!--   distance_method = "jaccard" -->
+<!--   clustering_method = "average" -->
+<!--   dist = Calculate_distance(dat_alra[local_gene,], method = distance_method) -->
 
-  gene_hree = hclust(dist, method = clustering_method)
-  gene_partition = cutree(gene_hree,k=5)
-  names(gene_partition) = labels(dist)
+<!--   gene_hree = hclust(dist, method = clustering_method) -->
+<!--   gene_partition = cutree(gene_hree,k=5) -->
+<!--   names(gene_partition) = labels(dist) -->
 
-  top5 <- names(sort(table(gene_partition),decreasing = TRUE))[1:5]
-  annotation_df = data.frame(Module = gene_partition)
-  ann_colors = list(
-    Module = c(setNames(colorRampPalette(brewer.pal(9, "Set1"))(5),1:5)))
+<!--   top5 <- names(sort(table(gene_partition),decreasing = TRUE))[1:5] -->
+<!--   annotation_df = data.frame(Module = gene_partition) -->
+<!--   ann_colors = list( -->
+<!--     Module = c(setNames(colorRampPalette(brewer.pal(9, "Set1"))(5),1:5))) -->
 
-  p_heatmap = pheatmap(as.matrix(dist),
-                annotation_col = annotation_df,
-                cluster_rows = gene_hree,
-                cluster_cols = gene_hree,
-                treeheight_row = 0,
-                annotation_colors = ann_colors, 
-                col=colorRampPalette(c("red","white","blue4"))(99),
-                show_colnames = FALSE,
-                show_rownames = FALSE,
-                annotation_legend = TRUE,
-                silent = TRUE)
-  pl_module = FeaturePlot_meta(dat,visual_space,gene_partition)
-  legend <- get_legend(pl_module[[1]])
-  prow <- cowplot::plot_grid(plotlist = lapply(pl_module,function(x) x + theme(legend.position="none")), ncol = ceiling(length(pl_module)/2))
+<!--   p_heatmap = pheatmap(as.matrix(dist), -->
+<!--                 annotation_col = annotation_df, -->
+<!--                 cluster_rows = gene_hree, -->
+<!--                 cluster_cols = gene_hree, -->
+<!--                 treeheight_row = 0, -->
+<!--                 annotation_colors = ann_colors,  -->
+<!--                 col=colorRampPalette(c("red","white","blue4"))(99), -->
+<!--                 show_colnames = FALSE, -->
+<!--                 show_rownames = FALSE, -->
+<!--                 annotation_legend = TRUE, -->
+<!--                 silent = TRUE) -->
+<!--   pl_module = FeaturePlot_meta(dat,visual_space,gene_partition) -->
+<!--   legend <- get_legend(pl_module[[1]]) -->
+<!--   prow <- cowplot::plot_grid(plotlist = lapply(pl_module,function(x) x + theme(legend.position="none")), ncol = ceiling(length(pl_module)/2)) -->
 
-  as.ggplot(p_heatmap) + cowplot::plot_grid(prow, legend, rel_widths = c(ceiling(length(pl_module)/2), .2)) + plot_annotation(title = sprintf("Genes in top 2000 by %s but not by %s (N = %d)",method_fullname[method],method_fullname[method_vs],length(local_gene))) + plot_layout(ncol = 2, widths = c(2, 3))
-})
+<!--   as.ggplot(p_heatmap) + cowplot::plot_grid(prow, legend, rel_widths = c(ceiling(length(pl_module)/2), .2)) + plot_annotation(title = sprintf("Genes in top 2000 by %s but not by %s (N = %d)",method_fullname[method],method_fullname[method_vs],length(local_gene))) + plot_layout(ncol = 2, widths = c(2, 3)) -->
+<!-- }) -->
 
-pdf(width = 12.5,height = 5)
-pl1
-dev.off()
+<!-- pdf(width = 12.5,height = 5) -->
+<!-- pl1 -->
+<!-- dev.off() -->
 
-pdf(width = 12.5,height = 5)
-pl2
-dev.off()
-```
+<!-- pdf(width = 12.5,height = 5) -->
+<!-- pl2 -->
+<!-- dev.off() -->
+<!-- ``` -->
 
@@
 
-fig = ggplot(score_measure %>% filter(Method!="all genes"), aes(x = TopGene, y = DensityIndex, color = Method, group = Method)) + geom_line() + scale_color_manual(values= coldef[names(coldef)%in%score_measure$Method]) + geom_hline(yintercept = baseline, linetype = "dashed", color = "grey1", alpha = 0.6) + annotate("text", x = max_x, y = baseline, label = "All_genes", vjust = -1, color = "grey1")  + theme_minimal()
+fig = ggplot(score_measure %>% filter(Method!="all genes"), aes(x = TopGene, y = DensityIndex, color = Method, group = Method)) + geom_line(aes(linetype = Method)) + guides(linetype="none") + scale_color_manual(values= coldef[names(coldef)%in%score_measure$Method]) + geom_hline(yintercept = baseline, linetype = "dashed", color = "grey1", alpha = 0.6) + annotate("text", x = max_x, y = baseline, label = "All_genes", vjust = -1, color = "grey1")  + theme_minimal() + geom_text_repel(data = score_measure %>%  filter(Method!="all genes") %>% group_by(Method) %>% slice_head(n = 2) %>% slice_tail(n = 1), aes(label = Method), hjust = 0)
+pdf(width = 8,height = 5)
 fig
+dev.off()
 ```
@@
 
-# Fig - Module HeatMap
+# Fig - Gene-Gene Colocolization HeatMap
 ```{r}
@@
               treeheight_row = 0,
-              col=colorRampPalette(c("red","white","blue4"))(99),
+              col=colorRampPalette(c("firebrick3", "white"))(99),
               show_colnames = FALSE, 
@@
               silent = TRUE)
-as.ggplot(p)
+p = as.ggplot(p)
+ggsave(filename = "fig-moduleheatmap.png", plot = p, width = 10, height = 10)
 ```
@@
 ```{r}
-# Heatmap
-pheatmap(jaccard_index, 
+jaccard_index_pruned = jaccard_index[apply(jaccard_index,1,max) > 0.4,]
+jaccard_index_pruned = jaccard_index_pruned[,apply(jaccard_index_pruned,2,max) > 0.4]
+
+colnames(jaccard_index_pruned) = paste0("M",colnames(jaccard_index_pruned))
+p = pheatmap(t(jaccard_index_pruned)[,names(sort(apply(jaccard_index_pruned,1,which.max)))],
          cluster_rows=FALSE, cluster_cols=FALSE, 
          scale = "none", 
-         color = colorRampPalette(rev(c("firebrick3", "white", "navy")))(100))
+         color = colorRampPalette(rev(c("firebrick3", "white")))(99),
+         angle_col = 45,legend_title = "Jaccard Index",silent = TRUE,
+         cellwidth = 15, cellheight = 10)
+p = as.ggplot(p)
+ggsave(filename = "jaccard_heatmap.png", plot = p, width = 5, height = 10)
 ```
 ```{r}
-# Bipartite Plot
-jaccard_thred = 0.4
-
-library(igraph)
-library("grid") 
-library("gridExtra")
-
-df_edge <- as.data.frame(as.table(jaccard_index))
-colnames(df_edge) = c("from","to","weight")
-df_edge <- subset(df_edge, weight > jaccard_thred)
-df_edge$width = ifelse(df_edge$weight > 0.8,2,0.5)
-df_node = data.frame(name=c(names(module_color),names(celltype_color)),color = c(module_color,celltype_color))
-df_node$type = ifelse(df_node$name %in% names(module_color),TRUE,FALSE)
-
-# Define specific color for module
-rownames(df_node) = df_node$name
-ct_module = names(which(apply(jaccard_index,2,max) > jaccard_thred))
-ct_subtype_module = c(4,5,6,30,39,42,43)
-ct_multype_module = c(2,13,19,33,36,54)
-cc_module = c(16,17,18,31)
-df_node[names(module_color),"color"] = "grey"
-df_node[ct_module,"color"] = "red"
-df_node[ct_multype_module,"color"] = "purple"
-df_node[ct_subtype_module,"color"] = "yellow"
-df_node[cc_module,"color"] = "green"
-
-g <- graph_from_data_frame(df_edge, directed=FALSE, vertices=df_node)
-V(g)$shape <- ifelse(V(g)$type, "circle", "square")
-E(g)$color <- ifelse(E(g)$weight > 0.8, "red", "grey")
-l <- layout_as_bipartite(g, types=!V(g)$type)
-
-plot(g, layout=l, vertex.size=2, vertex.label=ifelse(V(g)$type, V(g)$name, NA), vertex.frame.color = NA, asp=0.2, 
-     vertex.label.cex=0.5,
-     vertex.label.dist=-1,
-     vertex.label.degree=pi/2)
-
-# Create Node Legend & Edge Legend
-color_label = c("green" = "cell cycle",purple = "multi cell type",red = "cell type",yellow = "cell subtype")
-df_node$subtype = color_label[df_node$color]
-
-p <- ggplot(df_node %>% filter(type), aes(x = 0, y = 0,color = subtype)) + geom_point(size = 7) + scale_color_manual(values= setNames(names(color_label),color_label)) + theme_minimal() + labs(color = "Module Type") + theme(legend.direction="horizontal")
-grid.newpage()
-grid.draw(get_legend(p))
-
-p <- ggplot(df_node %>% filter(!type), aes(x = 0, y = 0,color = name)) + geom_point(size = 7,shape = 15) + scale_color_manual(values= df_node %>% filter(!type) %>% pull(color,name)) + theme_minimal() + labs(color = "Cell Type") + theme(legend.direction="horizontal")
-grid.newpage()
-grid.draw(get_legend(p))
+# # Bipartite Plot
+# jaccard_thred = 0.4
+# 
+# library(igraph)
+# library("grid") 
+# library("gridExtra") 
+# 
+# df_edge <- as.data.frame(as.table(jaccard_index))
+# colnames(df_edge) = c("from","to","weight")
+# df_edge <- subset(df_edge, weight > jaccard_thred)
+# df_edge$width = ifelse(df_edge$weight > 0.8,2,0.5)
+# df_node = data.frame(name=c(names(module_color),names(celltype_color)),color = c(module_color,celltype_color))
+# df_node$type = ifelse(df_node$name %in% names(module_color),TRUE,FALSE)
+# 
+# # Define specific color for module
+# rownames(df_node) = df_node$name
+# ct_module = names(which(apply(jaccard_index,2,max) > jaccard_thred))
+# ct_subtype_module = c(4,5,6,30,39,42,43)
+# ct_multype_module = c(2,13,19,33,36,54)
+# cc_module = c(16,17,18,31)
+# df_node[names(module_color),"color"] = "grey"
+# df_node[ct_module,"color"] = "red"
+# df_node[ct_multype_module,"color"] = "purple"
+# df_node[ct_subtype_module,"color"] = "yellow"
+# df_node[cc_module,"color"] = "green"
+# 
+# g <- graph_from_data_frame(df_edge, directed=FALSE, vertices=df_node)
+# V(g)$shape <- ifelse(V(g)$type, "circle", "square")
+# E(g)$color <- ifelse(E(g)$weight > 0.8, "red", "grey")
+# l <- layout_as_bipartite(g, types=!V(g)$type)
+# 
+# plot(g, layout=l, vertex.size=2, vertex.label=ifelse(V(g)$type, V(g)$name, NA), vertex.frame.color = NA, asp=0.2, 
+#      vertex.label.cex=0.5,
+#      vertex.label.dist=-1,
+#      vertex.label.degree=pi/2)
+# 
+# # Create Node Legend & Edge Legend
+# color_label = c("green" = "cell cycle",purple = "multi cell type",red = "cell type",yellow = "cell subtype")
+# df_node$subtype = color_label[df_node$color]
+# 
+# p <- ggplot(df_node %>% filter(type), aes(x = 0, y = 0,color = subtype)) + geom_point(size = 7) + scale_color_manual(values= setNames(names(color_label),color_label)) + theme_minimal() + labs(color = "Module Type") + theme(legend.direction="horizontal")
+# grid.newpage()
+# grid.draw(get_legend(p))
+# 
+# p <- ggplot(df_node %>% filter(!type), aes(x = 0, y = 0,color = name)) + geom_point(size = 7,shape = 15) + scale_color_manual(values= df_node %>% filter(!type) %>% pull(color,name)) + theme_minimal() + labs(color = "Cell Type") + theme(legend.direction="horizontal")
+# grid.newpage()
+# grid.draw(get_legend(p))
 ```
@@
 pl_module = lapply(levels(gene_partition),function(module){
-  FeaturePlot_custom(value = cell_block[,module], coord = visual_space, value_name = "Module Score",title_name = sprintf("Module %s (%d genes)",module, sum(gene_partition == module)))
+  FeaturePlot_custom(value = cell_block[,module], coord = visual_space, value_name = "Module Score",title_name = sprintf("Module %s (%d genes)",module, sum(gene_partition == module))) + scale_color_gradient(limits = c(0, 1),low = "lightgrey", high = "blue")
 })
 names(pl_module) = levels(gene_partition)
-legend <- get_legend(pl_module[[1]])
+legend_module <- get_legend(pl_module[[1]])
 prow <- cowplot::plot_grid(plotlist = lapply(pl_module,function(x) x + theme(legend.position="none")), ncol = 10)
-pdf(width = 30, height = 18)
-cowplot::plot_grid(prow, legend, rel_widths = c(10, .2))
-dev.off()
+p = cowplot::plot_grid(prow, legend_module, rel_widths = c(10, 1))
+ggsave(filename = "fig-moduletsne.png", plot = p, width = 30, height = 18)
 ```
-# Fig - Module-CellType
-```{r}
-jaccard_thred = 0.4
-one_hot_matrix <- sapply(unique(cell_label), function(x) as.integer(cell_label == x))
-# Get CellType One-hot plot
-pl_ct = lapply(colnames(one_hot_matrix),function(ct){
-  FeaturePlot_custom(value = as.factor(one_hot_matrix[,ct]), coord = visual_space, value_name = "Indicator",title_name = ct)
-})
-names(pl_ct) = colnames(one_hot_matrix)
-
-pl <- lapply(rownames(jaccard_index)[rev(order(apply(jaccard_index,1,max)))], function(celltype){
-  if(max(jaccard_index[celltype,]) > jaccard_thred){
-    p1 = pl_ct[[celltype]]
-    module_name = colnames(jaccard_index)[which.max(jaccard_index[celltype,])]
-    p2 = pl_module[[module_name]]
-    p = as.ggplot(p1) + as.ggplot(p2) + plot_annotation(paste0("Jaccard Index: ",round(jaccard_index[celltype,module_name],2)))
-    p
-  }
-})
-sub = unlist(lapply(pl,function(x)is.null(x)))
-pl = pl[-which(sub)]
-pdf(width = 15, height = 20)
-cowplot::plot_grid(plotlist = pl,ncol = 3)
-dev.off()
-```

### 2023-10-27 Marker_identfication_benchmark.Rmd
cellMarkerDB_count

@@
 ## Criterion2: CellMarkerDB
-Downloaded marker from http://bio-bigdata.hrbmu.edu.cn/CellMarker/CellMarker_download_files/file/Cell_marker_Seq.xlsx
+Downloaded marker from 
+http://117.50.127.228/CellMarker/CellMarker_download_files/file/Cell_marker_Seq.xlsx
 <!-- PanglaoDB: https://panglaodb.se/markers/PanglaoDB_markers_27_Mar_2020.tsv.gz and exclude genes that reported only once -->
@@
 # Load CellMarkerDB
-cell_marker_db = readxl::read_xlsx("/data/ruiqi/local_marker/dataset/Cell_Marker/Cell_marker_All.xlsx") %>% filter(species == "Mouse") %>% filter(tissue_class == "Bone marrow") %>% filter(!is.na(cellontology_id)) %>% filter(!is.na(PMID)) %>% mutate(gene_name1 = marker) %>% mutate(gene_name2 = Symbol) %>% select(c("gene_name1","gene_name2")) %>% group_by(across(everything())) %>% dplyr::summarise(count = n(), .groups = "drop") %>% filter(count >= 1) # Lung, Pancreas
+cell_marker_db = readxl::read_xlsx("/data/ruiqi/local_marker/dataset/Cell_Marker/Cell_marker_All.xlsx") %>% filter(species == "Mouse") %>% filter(tissue_class == "Bone marrow") %>% filter(!is.na(cellontology_id)) %>% filter(!is.na(PMID)) %>% mutate(gene_name1 = marker) %>% mutate(gene_name2 = Symbol) %>% select(c("gene_name1","gene_name2"))
+# %>% group_by(dplyr::across(dplyr::everything())) %>% dplyr::summarise(count = n(), .groups = "drop") %>% filter(count >= 1) # Lung, Pancreas
 

### 2023-10-29 Demo_marrow_facs.Rmd
fig-adjustment

--- 2023-10-20 Demo_marrow_facs.Rmd
@@
 bioc_packages <- c("Seurat",
-                   "clusterProfiler", "AnnotationDbi", "ReactomePA", "org.Mm.eg.db")
+                   "clusterProfiler", "AnnotationDbi", "ReactomePA", "org.Mm.eg.db","gprofiler2")
 github_packages <- c("satijalab/seurat-wrappers")
@@
 ## Step1: Deriving Gene Modules
-
 ```{r}
@@
 ### Visualize Heatmap
-
 ```{r}
@@
 
-### Visualize Module on Cell embeddings
+## Step3: Identify Celltype-specific activated Modules
+```{r}
+cell_label = as.factor(tiss$cell_ontology_class)
+abbrev = levels(cell_label)
+abbrev[grepl("T cell",abbrev)] = "T cell"
+abbrev[grepl("natural killer",abbrev)] = "NK cell"
+abbrev[abbrev == "B cell"] = "Cd3e+ Klrb1+ B cell"
+abbrev[grepl("Slamf1-positive",abbrev)] = "Slamf1+ MPC"
+abbrev[grepl("Slamf1-negative",abbrev)] = "Slamf1- MPC"
+abbrev[grepl("common lymphoid",abbrev)] = "CLP"
+abbrev[grepl("megakaryocyte",abbrev)] = "MEP"
+abbrev[grepl("hematopoietic",abbrev)] = "HSC"
+abbrev[grepl("hematopoietic",abbrev)] = "HSC"
+abbrev[grepl("granulocyte monocyte",abbrev)] = "GMP"
+# abbrev[grepl("granulocytopoietic",abbrev)] = "G-poiesis"
+names(abbrev) = levels(cell_label)
+levels(cell_label) = abbrev[levels(cell_label)]
 
-```{r, fig.width=20, fig.asp=1}
-pl = lapply(levels(gene_partition),function(module){
-  FeaturePlot_custom(value = cell_block[,module], coord = visual_space, value_name = "Module Score",title_name = sprintf("Module %s (%d genes)",module, sum(gene_partition == module)))
-})
-plot_grid(plotlist = pl,ncol = 5) # may takes long time to plot
-```
-
-## Step3: Identify Celltype-specific activated Modules
-
-```{r}
 # Create one-hot vector of Cell Type
 one_hot_matrix <- sapply(unique(cell_label), function(x) as.integer(cell_label == x))
+colnames(one_hot_matrix) = unique(cell_label)
+rownames(one_hot_matrix) = names(cell_label)
 
-# Order Cell Type based on its centroid in Featurespace
-Idents(tiss) = cell_label
-ordered_cell_types <- unique(cell_label)[hclust(dist(t(sapply(unique(cell_label), function(x) colMeans(Embeddings(tiss, reduction = "pca")[WhichCells(tiss, idents = x), ,drop = FALSE]) ))))$order]
+# # Order Cell Type based on its centroid in Featurespace
+# Idents(tiss) = cell_label
+# ordered_cell_types <- unique(cell_label)[hclust(dist(t(sapply(unique(cell_label), function(x) colMeans(Embeddings(tiss, reduction = "pca")[WhichCells(tiss, idents = x), ,drop = FALSE]) ))))$order]
 
@@
 })
-jaccard_index = jaccard_index[rev(ordered_cell_types),]
-rm(ordered_cell_types)
-```
-
-### Visualize Heatmap of Jaccard Index Matrix
-
-```{r}
-pheatmap(jaccard_index,
-         cluster_rows=FALSE, cluster_cols=FALSE, 
-         scale = "none", 
-         color = colorRampPalette(rev(c("firebrick3", "white")))(99))
-```
-
-### Visualize How CellType-Module is Matched on Cell Embeddings
-
-For each cell type, select the most specific modules for visualization (maximum jaccard index)
-
-```{r}
-tiss@meta.data = cbind(tiss@meta.data,one_hot_matrix)
-jaccard_thred = 0.4
-
-# Get Module plot
-pl_module = lapply(levels(gene_partition),function(module){
-  FeaturePlot_custom(value = cell_block[,module], coord = visual_space, value_name = "Module Score",title_name = sprintf("Module %s (%d genes)",module, sum(gene_partition == module)))
-})
-names(pl_module) = levels(gene_partition)
-
-# Get CellType One-hot plot
-pl_ct = lapply(colnames(one_hot_matrix),function(ct){
-  FeaturePlot_custom(value = as.factor(one_hot_matrix[,ct]), coord = visual_space, value_name = "Indicator",title_name = ct)
-})
-names(pl_ct) = colnames(one_hot_matrix)
-
-pl <- lapply(rownames(jaccard_index)[rev(order(apply(jaccard_index,1,max)))], function(celltype){
-  if(max(jaccard_index[celltype,]) > jaccard_thred){
-    p1 = pl_ct[[celltype]]
-    module_name = colnames(jaccard_index)[which.max(jaccard_index[celltype,])]
-    p2 = pl_module[[module_name]]
-    p = as.ggplot(p1) + as.ggplot(p2) + plot_annotation(paste0("Jaccard Index: ",round(jaccard_index[celltype,module_name],2)))
-    p
-  }
-})
+# jaccard_index = jaccard_index[rev(ordered_cell_types),]
+# rm(ordered_cell_types)
 ```
@@
 
+## Transferability of Modules
+### Load Data & Calculate Module Activity
+```{r}
+if (!file.exists('droplet_Marrow_seurat_tiss.Robj')) {
+   # If load url doesn't work, please download it and load from your working directory
+  options(timeout=6000)
+  download.file("https://figshare.com/ndownloader/files/13092380",
+                destfile = 'droplet_Marrow_seurat_tiss.Robj', method = 'libcurl')
+ }
+load('droplet_Marrow_seurat_tiss.Robj')
+tiss_droplet <- UpdateSeuratObject(tiss_droplet)
+
+# path <- "/data/ruiqi/local_marker/dataset/tabular_muris/"
+# tissue_name = "marrow_droplet"
+# tiss_droplet <- readRDS(sprintf("%s%s.rds",path,tissue_name))
+
+feature_space_droplet = as.matrix(tiss_droplet@reductions$pca@cell.embeddings[,1:n_dim])
+dat_droplet = as.matrix(tiss_droplet[["RNA"]]@data)
+
+knn = 10; major_vote = 5
+A = Symmetric_KNN_graph(knn = knn, feature_space = feature_space_droplet)$'adj_matrix'
+cell_block_loop = lapply(1:100,function(loop){
+  GMM_subsampling(seed = loop, gene_partition, expr_dat = dat_droplet, cell_kNN_graph = A, major_vote)
+})
+cell_block_prop = Reduce(`+`, cell_block_loop) / length(cell_block_loop)
+cell_block_droplet = cell_block_prop
+```
+
+### Calculate Jaccard Index
+```{r}
+cell_label = tiss_droplet$cell_ontology_class
+cell_label[!cell_label %in% names(abbrev)] = NA
+cell_label = setNames(abbrev[cell_label],names(cell_label))
+
+one_hot_matrix_droplet <- sapply(unique(na.omit(cell_label)), function(x) as.integer(cell_label == x))
+one_hot_matrix_droplet[is.na(one_hot_matrix_droplet)] = 0
+
+jaccard_index_droplet <- apply(cell_block_droplet, 2, function(module_col) {
+  apply(one_hot_matrix_droplet, 2, function(one_hot_col) {
+    1 - philentropy::distance(rbind(module_col,one_hot_col), method = "jaccard", mute.message=TRUE)
+  })
+})
+```
 ## Save Intermediate data
 ```{r}
-save(gene_partition,dist,gene_hree,cell_block,jaccard_index,file = "./marrow_result/module_result.rdata")
+save(gene_partition,dist,gene_hree,cell_block,cell_block_droplet,one_hot_matrix,one_hot_matrix_droplet,jaccard_index,jaccard_index_droplet,file = "./marrow_result/module_result.rdata")
 ```
@@
 ## Step4: Module Enrichment Analysis
-
 ### GO enrichment Analysis
-
 ```{r}
@@
 ## Cell Cycle Modules
+### Reactome analysis
+```{r}
+library("ReactomePA")
+library("org.Mm.eg.db")
+universe_df = data.frame("symbol" = rownames(dat),"entrez" = mapIds(org.Mm.eg.db, keys=rownames(dat), column="ENTREZID", keytype="SYMBOL"))
+universe_df = universe_df[!is.na(universe_df$entrez),]
+module_cc = c(16,18,17,31,32)
+epathway_result_cc <- lapply(module_cc, function(i){
+  enrichPathway(gene=universe_df[universe_df$symbol %in% names(gene_partition)[gene_partition == i],"entrez"],
+                organism = "mouse",
+                pAdjustMethod = "BH",
+                pvalueCutoff = 0.05,
+                universe = universe_df$entrez)
+})
+names(epathway_result_cc) = module_cc
+# Annotate each modules with top5 Pathway terms
+top_enrich_pathway_cc <- do.call(rbind,lapply(1:length(epathway_result_cc), function(i) data.frame(epathway_result_cc[[i]]@result[1:5,],module = module_cc[i])))
+top_enrich_pathway_cc <- top_enrich_pathway_cc[complete.cases(top_enrich_pathway_cc), ]
+top_enrich_pathway_cc = top_enrich_pathway_cc %>% filter(p.adjust < 0.05)
+
+write.csv(top_enrich_pathway_cc,file = "./marrow_result/reactome_pathway_cc.csv")
+```
+### Annotate Cell Cycle
 ```{r}
@@
                         g2m.features = mmus_g2m, set.ident = FALSE)
-DimPlot(tiss,group.by = "Phase")
+cell_cycle_table = tiss@meta.data[,c("S.Score","G2M.Score","Phase")]
+write.csv(cell_cycle_table,file = "./marrow_result/annotated_cell_cycle.csv")
+```
 
-pdf(width = 30, height = 5)
-pl
-dev.off()
-```
+<!-- ## Multi-resolution Modules - analysis -->
+<!-- ```{r} -->
+<!-- df_module = read.csv("./marrow_result/module_description.csv") -->
+<!-- df_module$'multi-celltype' = NA -->
+<!-- df_module$'note' = NA -->
+
+<!-- ct_prob_per_module = do.call(rbind,lapply(unique(cell_label),function(x){apply(cell_block[cell_label == x,],2,median)})) -->
+<!-- rownames(ct_prob_per_module) = unique(cell_label) -->
+<!-- ct_num_per_module = sort(apply(ct_prob_per_module,2,function(x)sum(x > 0.5))) -->
+
+<!-- m = 19 -->
+<!-- df_module[m,"celltype"] = paste0(names(which(ct_prob_per_module[,m] > 0.5)),collapse = ",") -->
+
+<!-- df_module[54,"multi-celltype"] = c("T/NK/NKT") -->
+<!-- df_module[27,"multi-celltype"] = c("Progenitor/NK/NKT") -->
+<!-- df_module[33,"multi-celltype"] = c("Progenitor/basophil/granulocytopoietic/half-granulocyte") -->
+<!-- df_module[46,"multi-celltype"] = c("monocyte/granulocytopoietic/half-granulocyte") -->
+<!-- df_module[48,"multi-celltype"] = c("monocyte/granulocytopoietic/granulocyte") -->
+<!-- df_module[49,"multi-celltype"] = c("monocyte/macrophage/granulocytopoietic/granulocyte") -->
+<!-- df_module[55,"multi-celltype"] = c("NK") -->
+<!-- df_module[56,"multi-celltype"] = c("NK") -->
+<!-- df_module[13,"multi-celltype"] = c("B") -->
+<!-- df_module[50,"multi-celltype"] = c("monocyte/granulocytopoietic/granulocyte/basophil") -->
+<!-- df_module[28,"multi-celltype"] = c("Progenitor") -->
+<!-- df_module[9,"multi-celltype"] = c("immature/naive B") -->
+<!-- df_module[11,"multi-celltype"] = c("late pro/pre B") -->
+<!-- df_module[7,"multi-celltype"] = c("late pro/pre B") -->
+<!-- df_module[44,"multi-celltype"] = c("granulocytopoietic/half-granulocyte") -->
+<!-- df_module[47,"multi-celltype"] = c("granulocytopoietic/granulocyte") -->
+<!-- df_module[52,"multi-celltype"] = c("T") -->
+<!-- df_module[53,"multi-celltype"] = c("macrophage/NK") -->
+<!-- df_module[1,"multi-celltype"] = c("monocyte/macrophage/granulocyte monocyte progenitor cell") -->
+<!-- df_module[20,"multi-celltype"] = c("granulocytopoietic/Progenitor") -->
+<!-- df_module[c(42,35),"multi-celltype"] = c("below-half-granulocyte/monocyte") -->
+<!-- df_module[c(36,38),"multi-celltype"] = c("granulocyte monocyte progenitor cell/monocyte") -->
+<!-- df_module[2,"multi-celltype"] = c("granulocyte monocyte progenitor cell/macrophage") -->
+<!-- df_module[4,"celltype"] = c("granulocyte monocyte progenitor cell") -->
+<!-- df_module[c(5,6,43),"note"] = "subtype" -->
+
+<!-- df_module[c(16,18,17,31,32),'note'] = "cell_cycle" -->
+<!-- ``` -->

### 2023-10-29 Marker_identfication_benchmark.Rmd
fig-adjustment

@@
     top_gs = rownames(df_benchmark)[order(df_benchmark[,method])]
-    top_n = c(seq(100,900,100),seq(1000,5000,1000))
+    top_n = c(seq(50,90,10),seq(100,900,100),seq(1000,3000,500),seq(4000,9000,1000),seq(10000,max(df_benchmark),2000))
   }
@@
 colnames(score_df) = paste0("kNN",c(3:10,20,30,40,50))
-write.table(score_df,file = paste0(path,tissue_name,"_different_knn.csv"),row.names = FALSE)
+write.table(score_df,file = paste0(path,tissue_name,"_different_knn.csv"),row.names = TRUE)
 ```

### 2023-10-29 paper_figure.Rmd
fig-adjustment

@@
 
+# SupFig - Bone-marrow Cell Embedding
+```{r}
+coldef = setNames(colorRampPalette(brewer.pal(12, "Paired"))(length(unique(cell_label))),unique(cell_label))
+p = DimPlot(tiss, group.by = "cell_ontology_class",label = FALSE, cols = coldef) + guides(color = guide_legend(ncol = 1,override.aes = list(size = 3))) + labs(title = NULL)
+
+ggsave(filename = "supfig-bonemarrow.png", plot = p, width = 8, height = 5, dpi = 300)
+# 
+# centroids <- aggregate(. ~ cell_label, data = as.data.frame(Embeddings(tiss, "tsne")), FUN = mean)
+# p <- DimPlot(tiss, group.by = "cell_ontology_class",label = FALSE, cols = coldef) + theme_void() + NoLegend() + ggtitle("Bone Marrow FACS")
+# p = p + 
+#   geom_text_repel(data = centroids, aes(x = tSNE_1, y = tSNE_2, label = cell_label, color = cell_label),
+#                    box.padding = 0.3, point.padding = 0.3, segment.color = 'grey50',max.overlaps = 100)
+```
+
 # Fig - Pipeline
@@
 # Plot knn graph
-plot_knn_graph(affinity_m = W, label = cell_label, layout = visual_space)
+p = plot_knn_graph(affinity_m = W, layout = visual_space)
+ggsave(filename = "fig-pipeline1.png", plot = p, width = 6, height = 5, dpi = 300)
 ```
@@
 # Check Diffusion KL pattern for several genes
-genes = c("Fcnb","Uchl4")
-genes_label = c("localized","non_localized")
-check_time = c(0,8,128,1024)
-p = Visualize_score_pattern(res, genes = genes, label_class = genes_label, facet_class = NULL, add_point = check_time)
+# good gene: Igll1,Fcnb # garbage marker: Fkbp1b, Tnks1bp1, Magi2,Uchl4
+genes = c("Igll1","Tnks1bp1") 
+# genes_label = c("localized","non_localized")
+check_time = c(0,2^3,2^10,2^15)
+p = Visualize_score_pattern(res, genes = genes, add_point = check_time)
+
+ggsave(filename = "fig-pipeline2.png", plot = p, width = 8, height = 5)
 
 # Check Gene Diffusion on Cell graph by time
-pl = lapply(genes,function(gene){
-  FeaturePlot_diffusion(coord = visual_space,init_state = rho[gene,],P_ls = P_ls,check_time = check_time,gene_name = gene)
+pl = lapply(1:length(genes),function(i){
+  gene = genes[i]
+  FeaturePlot_diffusion(coord = visual_space,init_state = rho[gene,],P_ls = P_ls,W = W,check_time = check_time,gene_name = gene,gene_color = scales::hue_pal()(2)[i])
 })
-plot_grid(plotlist = pl,ncol = 1)
+p = as.ggplot(pl[[1]]) / as.ggplot(pl[[2]] & labs(title = NULL))
+ggsave(filename = "fig-pipeline3.png", plot = p, width = 16, height = 8)
 ```
@@
   geom_bar(stat = "identity", position = "dodge") +
-  geom_text(aes(label = round(AUROC, 2)), position = position_dodge(width = 0.85), vjust = -0.25, size = 2) +
-  theme_minimal() +
+  geom_text(aes(label = round(AUROC, 2)), position = position_dodge(width = 0.85), vjust = -0.25, size = 3) +
   labs(x = "Tissue", y = "AUROC", fill = "Method") + coord_cartesian(ylim = c(0.5, 1)) +
   theme(legend.position = "bottom")  +  scale_fill_manual(values= coldef[names(coldef)%in%df_auc$Method]) + facet_wrap(~gt_set)
+fig = fig + theme(
+          legend.title = element_text(size = 20),
+          legend.text = element_text(size = 15),
+          axis.title.x = element_text(size = 15),
+          axis.title.y = element_text(size = 15),
+          axis.text.x = element_text(size = 12),
+          axis.text.y = element_text(size = 10),
+          panel.grid = element_blank(),
+          axis.line = element_line(colour = "black"),
+          panel.background = element_blank(),
+          strip.background = element_rect(fill = "white", color = "white") ,
+          strip.text = element_text(size = 15, color = "black"))
 
@@
 df_label_all = df_benchmark[unique(sub),c(method_vs,method_vs_set)]
-df_label_all$'legend' = sapply(seq(nrow(df_label_all)),function(i){
-  paste(paste(method_fullname[colnames(df_label_all)],df_label_all[i,],sep = ":"),collapse = "\n")
+# df_label_all$'legend' = sapply(seq(nrow(df_label_all)),function(i){
+#   paste(paste(method_fullname[colnames(df_label_all)],df_label_all[i,],sep = ":"),collapse = "\n")
+# })
+g = c("Cd19","Naaa","Mogat2","Tifa","Abca13","Zf12","Il2rb","Phgdh","Vpreb2","Fyb")
+names(g) = rep(method_vs_set,each = 2)
+pl = lapply(1:length(g),function(i){
+  x = g[i]
+  subtitle = sprintf("%s:%d, %s:%d",method_fullname[method_vs],df_label_all[x,method_vs],method_fullname[names(g)[i]],df_label_all[x,names(g)[i]])
+  color_label = ifelse(df_label_all[x,method_vs] > df_label_all[x,names(g)[i]],"red","blue")
+  FeaturePlot(tiss,features = x,order =TRUE) + labs(color = "Expression") + labs(title = x,subtitle = subtitle) + 
+    theme(
+      plot.title = element_text(face="bold", size = 30,color = color_label),
+      plot.subtitle = element_text(size = 20),
+      legend.title = element_text(size = 20),
+      legend.text = element_text(size = 15),
+      axis.title = element_blank(),
+      axis.text = element_blank(),
+      axis.ticks=element_blank(),
+      axis.line = element_blank())
+  # + annotation_custom(grob = grid::textGrob(df_label_all[x,"legend"], x = 1, y = 1, just = c("right", "top"),gp = grid::gpar(col = "red", fontsize = 10)))
 })
-g = c("Cd19","Naaa","Mogat2","Tifa","Abca13","Zf12","Il2rb","Phgdh","Vpreb2","Fyb")
-pl = lapply(g,function(x){
-  FeaturePlot(tiss,features = x,order =TRUE) + annotation_custom(grob = grid::textGrob(df_label_all[x,"legend"], x = 1, y = 1, just = c("right", "top"),gp = grid::gpar(col = "red", fontsize = 10)))
-})
-fig = cowplot::plot_grid(plotlist = pl[as.numeric(t(matrix(1:length(pl),nrow = 2)))],ncol = length(method_vs_set))
-ggsave(filename = "fig-ranking_comparison.png", plot = fig, width = 25, height = 10)
+pl = pl[order(seq(length(g))%%2,decreasing = TRUE)]
+pl[[1]] = pl[[1]] + labs(y = "Favored by LMD") + theme(axis.title.y = element_text(size = 20,face="bold", colour = "blue"))
+pl[[(length(pl)/2)+1]] = pl[[(length(pl)/2)+1]] + labs(y = "Favored by Other methods") + theme(axis.title.y = element_text(size = 20,face="bold", colour = "red"))
+
+fig = wrap_plots(pl,nrow = 2)
+
+ggsave(filename = "fig-ranking_comparison2.png", plot = fig, width = 25, height = 10)
 ```
@@
   df_label = data.frame(df_benchmark[,c(method,method_vs)],label = rownames(df_benchmark))
-  df_label[c(which(df_label[,method] <= 100 & df_label[,method_vs] > 500)),"color"] = sprintf("Favor by %s (rank <= 100)",method)
-  df_label[c(which(df_label[,method] > 500 & df_label[,method_vs] <= 100) ),"color"] = sprintf("Favor by %s (rank <= 100)",method_vs)
+  fullmethod = method_fullname[method]
+  fullmethodvs = method_fullname[method_vs]
+  df_label[c(which(df_label[,method] <= 100 & df_label[,method_vs] > 500)),"color"] = fullmethod
+  df_label[c(which(df_label[,method] > 500 & df_label[,method_vs] <= 100) ),"color"] = fullmethodvs
   df_label$'color' = as.factor(df_label$'color')
-  p = ggplot(df_label, aes(x=!!as.name(method), y=!!as.name(method_vs),color = color)) + geom_point(size = 1) + 
-  scale_color_manual(values = c("red", "blue", "grey50"),
-                     breaks = levels(df_label$'color')) + 
-    geom_abline(intercept = 0, slope = 1, col = "grey", linetype = "dashed") + scale_x_log10(breaks = 10^(0:4), labels = expression(10^0, 10^1, 10^2, 10^3, 10^4)) +
+  rank_cor = cor(log(df_label[,1:2],10))[1,2]
+  colnames(df_label)[1:2] = c("x","y")
+  p = ggplot(df_label, aes(x=x, y=y,color = color)) + geom_point(size = 1) + 
+  scale_color_manual(values = c("red","blue","grey50"),
+                     breaks = c(fullmethod,fullmethodvs),
+                     labels = c("Favored by LMD","Favored by Other methods")) + 
+    geom_abline(intercept = 0, slope = 1, col = "grey", linetype = "dashed") + 
+    scale_x_log10(breaks = 10^(0:4), labels = expression(10^0, 10^1, 10^2, 10^3, 10^4)) +
     scale_y_log10(breaks = 10^(0:4), labels = expression(10^0, 10^1, 10^2, 10^3, 10^4)) + 
     expand_limits(x = c(1, 10^4), y = c(1, 10^4)) +
-    labs(x = paste0(method_fullname[method]," Rank"), y = paste0(method_fullname[method_vs]," Rank"), color = "Highlighted Genes") + theme_minimal() 
+    labs(x = paste0(method_fullname[method]," Rank"), 
+         y = paste0(method_fullname[method_vs]," Rank"),
+         color = "Highlighted Genes",
+         title = paste("cor =", round(rank_cor, 2)))
   # annotate gene labels
-  # library(ggrepel)
-  p = p + geom_text_repel(data = df_label %>% filter(!is.na(color)) %>% filter(label %in% g), aes(label=label),min.segment.length = unit(0, 'lines'),nudge_y = -0.5)
-  p
+  p = p + ggrepel::geom_text_repel(data = df_label %>% filter(!is.na(color)) %>% filter(label %in% g), aes(label=label),min.segment.length = unit(0, 'lines'),nudge_y = -0.5,show.legend = FALSE,size = 5)
+  p = p + theme(
+    plot.title = element_text(size = 15),
+    legend.title = element_text(size = 20),
+    legend.text = element_text(size = 15),
+    axis.title.x = element_text(size = 20),
+    axis.title.y = element_text(size = 20),
+    axis.text.x = element_text(size = 15),
+    axis.text.y = element_text(size = 15),
+    panel.grid = element_blank(),
+    panel.background = element_blank(),
+    axis.line = element_line(colour = "black"))
 })
-pdf(width = 25,height = 5)
-legend = get_legend(pl_rank[[1]] +  guides(color=guide_legend(direction="horizontal")))
-prow = plot_grid(plotlist = lapply(pl_rank,function(x) x + NoLegend()), ncol = length(method_vs_set))
-plot_grid(prow, legend, ncol = 1, rel_heights = c(1, .2))
-dev.off()
+for(i in 2:length(pl_rank)){
+  pl_rank[[i]] <- pl_rank[[i]] + theme(axis.title.y=element_blank())
+}
+p = (wrap_plots(pl_rank,nrow = 1) & theme(legend.position="bottom")) + plot_layout(guides = "collect")
+
+ggsave(filename = "fig-ranking_comparison1.png", plot = p, width = 25, height = 6)
 ```
@@
 
-baseline = score_measure %>% filter(Method == "all genes") %>%.$DensityIndex
-max_x = score_measure %>% filter(Method!="all genes") %>%.$TopGene %>% max()
+baseline = score_measure %>% filter(Method == "All genes") %>%.$DensityIndex
+max_x = score_measure %>% filter(Method!="All genes") %>%.$TopGene %>% max()
 
-fig = ggplot(score_measure %>% filter(Method!="all genes"), aes(x = TopGene, y = DensityIndex, color = Method, group = Method)) + geom_line(aes(linetype = Method)) + guides(linetype="none") + scale_color_manual(values= coldef[names(coldef)%in%score_measure$Method]) + geom_hline(yintercept = baseline, linetype = "dashed", color = "grey1", alpha = 0.6) + annotate("text", x = max_x, y = baseline, label = "All_genes", vjust = -1, color = "grey1")  + theme_minimal() + geom_text_repel(data = score_measure %>%  filter(Method!="all genes") %>% group_by(Method) %>% slice_head(n = 2) %>% slice_tail(n = 1), aes(label = Method), hjust = 0)
-pdf(width = 8,height = 5)
-fig
-dev.off()
+line_type = replace(coldef,names(coldef),"solid")
+line_type["Seurat v3 Wilcoxon unfiltered"] = "twodash"
+line_type["All genes"] = "dashed"
+
+fig = ggplot(score_measure %>% filter(Method!="All genes"), aes(x = TopGene, y = DensityIndex, color = Method, group = Method, linetype = Method),linewidth = 2) + 
+  geom_line(lwd = 1) +
+  scale_x_log10(breaks = 10^(2:4), labels = expression(10^2, 10^3, 10^4)) +
+    # expand_limits(x = c(50, 10^4)) +
+  geom_hline(aes(yintercept = baseline, color = "All genes", linetype = "All genes")) +
+  scale_color_manual(values= c(coldef[names(coldef)%in%score_measure$Method],"All genes" = "black")) + 
+  scale_linetype_manual(values = line_type) +
+  theme(
+    legend.title = element_text(size = 20),
+    legend.text = element_text(size = 15),
+    axis.title.x = element_text(size = 20),
+    axis.title.y = element_text(size = 20),
+    axis.text.x = element_text(size = 15),
+    axis.text.y = element_text(size = 15),
+    panel.grid = element_blank(),
+    panel.background = element_blank(),
+    axis.line = element_line(colour = "black"))
+  # geom_text_repel(data = score_measure %>%  filter(Method!="all genes") %>% group_by(Method) %>% slice_head(n = 2) %>% slice_tail(n = 1), aes(label = Method), hjust = 0)
+
+ggsave(filename = "fig-densityindex.png", plot = fig, width = 8, height = 5)
 ```
 
-# Fig - Sensitivity Test
+# SupFig - Sensitivity Test
 ```{r}
@@
 base = "kNN5"
-top_genes_list <- lapply(c(50,seq(100,1000,100)), 
-                         function(x) head(score_ranked[order(score_ranked[,base]),], x))
-average_ranks <- lapply(top_genes_list, function(top_genes) {
-  colMeans(top_genes)
+
+top_genes = c(50,seq(100,1000,100))
+jaccard_index_bw_sets = do.call(cbind,lapply(top_genes,function(top_gene)
+    {
+     top_gene_ls = apply(score_ranked,2,function(x) rownames(score_ranked)[order(x)][1:top_gene])
+     apply(top_gene_ls,2,function(x){
+       length(intersect(x,top_gene_ls[,base]))/length(union(x,top_gene_ls[,base]))
      })
-average_ranks <- data.frame(
-  Top_Genes = c(50,seq(100,1000,100)),
-  do.call(rbind, average_ranks)
-)
-average_ranks <- average_ranks %>% tidyr::pivot_longer(cols = -Top_Genes, names_to = "kNN", values_to = "Average_Rank")
-average_ranks$kNN = as.numeric(gsub("kNN","",average_ranks$kNN))
+  }))
+colnames(jaccard_index_bw_sets) = top_genes
+df = reshape2::melt(jaccard_index_bw_sets,value.name = "JaccardIndex")
+colnames(df)[1:2] = c("kNN","Top_Genes")
+df$kNN = as.factor(as.numeric(gsub("kNN","",df$kNN)))
+df$Top_Genes = as.factor(df$Top_Genes)
 
-fig = ggplot(data = average_ranks, aes(x = kNN, y = Average_Rank, group = Top_Genes)) +
+fig = ggplot(data = df, aes(x = kNN, y = JaccardIndex, group = Top_Genes, color = Top_Genes)) +
   geom_line() +
-  geom_point() + labs(x = "kNN", y = sprintf("Average Rank of Top # Genes\n (base on %s)",base)) + geom_vline(xintercept = 5, color="red", linetype="dashed") +
-  theme_minimal() + facet_wrap(~Top_Genes,scales = "free",labeller = as_labeller(setNames(paste0("Top",unique(average_ranks$Top_Genes)),unique(average_ranks$Top_Genes)))) + NoLegend()
-pdf(width = 8,height = 5)
-fig
-dev.off()
+  geom_point() + labs(x = "# of kNN used", y = sprintf("Jaccard Index\n (with %s)",base),color = "Top Genes") + geom_vline(xintercept = 3, color="red", linetype="dashed") + scale_y_continuous(limits = c(0,1)) +
+  theme(
+    legend.title = element_text(size = 20),
+    legend.text = element_text(size = 15),
+    axis.title.x = element_text(size = 20),
+    axis.title.y = element_text(size = 20),
+    axis.text.x = element_text(size = 15),
+    axis.text.y = element_text(size = 15),
+    panel.grid = element_blank(),
+    panel.background = element_blank(),
+    axis.line = element_line(colour = "black"))
+
+ggsave(filename = "supfig-sensitivity1.png", plot = fig, width = 8, height = 5)
 ```
@@
 auc_df$gt = factor(auc_df$gt); levels(auc_df$gt) = c("CellMarkerDB","MacFC(Top100)")
-auc_df$kNN = as.numeric(gsub("kNN","",auc_df$kNN))
-fig = ggplot(data = auc_df, aes(x = kNN, y = AUC), group = gt) +
+auc_df$kNN = as.factor(as.numeric(gsub("kNN","",auc_df$kNN)))
+
+fig = ggplot(data = auc_df, aes(x = kNN, y = AUC, group = gt, color = gt)) +
   geom_line() +
-  geom_point() + labs(x = "kNN", y = "AUROC") + facet_wrap(~gt, scales = "free") + 
-  theme_minimal()
-pdf(width = 6,height = 3)
-fig
-dev.off()
+  geom_point() + labs(x = "# of kNN used", y = "AUROC", color = "Gold-Standard Genesets") + scale_y_continuous(limits = c(0,1)) + theme(
+    legend.title = element_text(size = 20),
+    legend.text = element_text(size = 15),
+    axis.title.x = element_text(size = 20),
+    axis.title.y = element_text(size = 20),
+    axis.text.x = element_text(size = 15),
+    axis.text.y = element_text(size = 15),
+    panel.grid = element_blank(),
+    panel.background = element_blank(),
+    axis.line = element_line(colour = "black"))
+
+ggsave(filename = "supfig-sensitivity2.png", plot = fig, width = 8, height = 5)
 ```
 
-# Fig - Gene-Gene Colocolization HeatMap
+# SupFig - Gene-Gene Colocolization HeatMap
 ```{r}
@@
 p = as.ggplot(p)
-ggsave(filename = "fig-moduleheatmap.png", plot = p, width = 10, height = 10)
+ggsave(filename = "supfig-moduleheatmap.png", plot = p, width = 10, height = 10)
 ```
-# Fig - Module Jaccard
+
+# SupFig - Module TSNE
 ```{r}
-jaccard_index_pruned = jaccard_index[apply(jaccard_index,1,max) > 0.4,]
-jaccard_index_pruned = jaccard_index_pruned[,apply(jaccard_index_pruned,2,max) > 0.4]
+pl_module = lapply(levels(gene_partition),function(module){
+  FeaturePlot_custom(value = cell_block[,module], coord = visual_space, value_name = "Module Score",title_name = sprintf("Module %s (%d genes)",module, sum(gene_partition == module))) + scale_color_gradient(limits = c(0, 1),low = "lightgrey", high = "blue")
+})
+names(pl_module) = levels(gene_partition)
+p = (wrap_plots(pl_module,ncol = 10) & theme(plot.title = element_text(face="bold", size = 10),legend.title = element_text(size = 30),legend.text = element_text(size = 20))) + plot_layout(guides = "collect")
+
+ggsave(filename = "supfig-moduletsne.png", plot = p, width = 30, height = 18)
+```
+
+# SupFig-CellType-specific Module
+##  Matched on Cell Embeddings
+For each cell type, select the most specific modules for visualization (maximum jaccard index)
+```{r}
+# Get CellType One-hot plot
+pl_ct = lapply(colnames(one_hot_matrix),function(ct){
+  FeaturePlot_custom(value = as.factor(one_hot_matrix[,ct]), coord = visual_space, value_name = "Indicator",title_name = ct) + scale_color_manual(values = c("0" = "lightgrey", "1" = scales::hue_pal()(1)))
+})
+names(pl_ct) = colnames(one_hot_matrix)
+
+ct_module_match = apply(jaccard_index_pruned,1,function(x) colnames(jaccard_index_pruned)[which.max(x)])
+ct_module_match = gsub("M","",ct_module_match)
+
+pl <- lapply(seq(length(ct_module_match)),function(i){
+    pl_module[[ct_module_match[i]]]/
+    pl_ct[[names(ct_module_match)[i]]]
+})
+
+p = (wrap_plots(pl,ncol = 7) & theme(plot.title = element_text(face="bold", size = 10),legend.title = element_text(size = 15),legend.text = element_text(size = 10))) + plot_layout(guides = "collect")
+
+ggsave(filename = "supfig-ct-specific-module1.png", plot = p, width = 3*7, height = 3*4)
+```
+
+## Jaccard Index Heatmap
+```{r}
+thred = 0.4
+jaccard_index_pruned = jaccard_index[apply(jaccard_index,1,max) > thred,]
+jaccard_index_pruned = jaccard_index_pruned[,apply(jaccard_index_pruned,2,max) > thred]
 
 colnames(jaccard_index_pruned) = paste0("M",colnames(jaccard_index_pruned))
-p = pheatmap(t(jaccard_index_pruned)[,names(sort(apply(jaccard_index_pruned,1,which.max)))],
-         cluster_rows=FALSE, cluster_cols=FALSE, 
-         scale = "none", 
-         color = colorRampPalette(rev(c("firebrick3", "white")))(99),
-         angle_col = 45,legend_title = "Jaccard Index",silent = TRUE,
-         cellwidth = 15, cellheight = 10)
-p = as.ggplot(p)
-ggsave(filename = "jaccard_heatmap.png", plot = p, width = 5, height = 10)
+jaccard_index_pruned = jaccard_index_pruned[names(sort(apply(jaccard_index_pruned,1,which.max))),]
+p = Visualize_jaccard_mtx(jaccard_index_pruned)
+p = as.ggplot(p) + coord_fixed()
+ggsave(filename = "supfig-ct-specific-module2.png", plot = p, width = 8, height = 10)
 ```
@@
 ```
-# Fig - Module TSNE
+
+## Droplet Jaccard Comparison
 ```{r}
-pl_module = lapply(levels(gene_partition),function(module){
-  FeaturePlot_custom(value = cell_block[,module], coord = visual_space, value_name = "Module Score",title_name = sprintf("Module %s (%d genes)",module, sum(gene_partition == module))) + scale_color_gradient(limits = c(0, 1),low = "lightgrey", high = "blue")
+module_selected = unlist(apply(jaccard_index[rownames(jaccard_index_droplet),],1,function(x) colnames(jaccard_index)[which(x > thred)]))
+module_selected = module_selected[order(as.numeric(module_selected))]
+
+ct_selected = unique(unlist(apply(jaccard_index[rownames(jaccard_index_droplet),module_selected],2,function(x) rownames(jaccard_index_droplet)[which(x > 0.4)])))
+ct_selected = names(sort(apply(jaccard_index[ct_selected,],1,which.max)))
+
+pl = lapply(list(jaccard_index,jaccard_index_droplet),function(mtx){
+  jaccard_index_pruned = mtx[ct_selected,module_selected]
+  colnames(jaccard_index_pruned) = paste0("M",colnames(jaccard_index_pruned))
+  Visualize_jaccard_mtx(jaccard_index_pruned)
 })
-names(pl_module) = levels(gene_partition)
-legend_module <- get_legend(pl_module[[1]])
-prow <- cowplot::plot_grid(plotlist = lapply(pl_module,function(x) x + theme(legend.position="none")), ncol = 10)
-p = cowplot::plot_grid(prow, legend_module, rel_widths = c(10, 1))
-ggsave(filename = "fig-moduletsne.png", plot = p, width = 30, height = 18)
+titles = c("FACS","Droplet")
+pl <- lapply(1:length(pl), function(i){pl[[i]] + labs(title = titles[i]) + theme(plot.title = element_text(size = 20, face = "bold"))})
+p = wrap_plots(pl,ncol = 2) + plot_layout(guides = "collect")
+ggsave(filename = "supfig-ct-specific-module3.png", plot = p, width = 10, height = 8)
 ```
+# Fig - Cell Cycle
+## Annotate CC
+```{r}
+cell_cycle = read.csv("./marrow_result/annotated_cell_cycle.csv")[["Phase"]]
+tiss$'Phase' = cell_cycle
+p1 = DimPlot(tiss,group.by = "Phase", cols = c("grey",scales::hue_pal()(3)[2:3])) + labs(title = "Phase annotated by Seurat",color = "Phase") + theme(
+  axis.title = element_blank(),
+      axis.text = element_blank(),
+      axis.ticks=element_blank(),
+      legend.title = element_text(size = 20),
+      legend.text = element_text(size = 15),
+      axis.line = element_blank())
+p = wrap_plots(p1,(wrap_plots(pl_module[c(16,17,18,31,32)],ncol = 3) & theme(plot.title = element_text(size = 15))) + plot_layout(guides = "collect"),nrow = 1,design = "AABBB")
+
+ggsave(filename = "fig-module-cellcycle.png", plot = p, width = 3*5, height = 3*2)
+```
+## Reactome enrichment
+```{r}
+df = read.csv("./marrow_result/reactome_pathway_cc.csv")
+df$module = as.factor(df$module)
+df$GeneRatio = as.numeric(df$Count / table(gene_partition)[df$module])
+df$Description = factor(df$Description, levels = names(sort(table(df$Description))))
+levels(df$Description)[6] = "HDR through Homologous Recombination\n or Single Strand Annealing (SSA)"
+p = ggplot(df, aes(x=module, y=Description)) +
+  geom_point(aes(size=GeneRatio, color=p.adjust), alpha=0.7) +
+  scale_color_gradientn(colors=colorRampPalette(c("red", "blue"))(99), name="p.adjust") +
+  theme_minimal() +
+  labs(size="Gene Ratio") +
+  theme(legend.position="right",
+        axis.title = element_text(size = 20),
+        axis.text.x = element_text(size = 15),
+      axis.text.y = element_text(size = 10),
+      legend.title = element_text(size = 20),
+      legend.text = element_text(size = 15))
+ggsave(filename = "fig-module-cellcycle2.png", plot = p, width = 10, height = 6)
+```
+
+# SupFig - Joint-CT Module visualization
+```{r}
+level0_module = c(13,27,38,46,54)
+p = (wrap_plots(pl_module[level0_module],ncol = 5) &  theme(plot.title = element_text(size = 15))) + plot_layout(guides = "collect")
+
+ggsave(filename = "supfig-joint-module.png", plot = p, width = 5*3, height = 3)
+```
+
+# Fig - Subtype Investigation
+```{r}
+p <- DimPlot(tiss, group.by = "cell_ontology_class",label = FALSE, cols = setNames(scales::hue_pal()(2),c("late pro-B cell","granulocyte"))) + ggtitle("Pre-annotated cell types") + theme(
+  axis.title = element_blank(),
+      axis.text = element_blank(),
+      axis.ticks=element_blank(),
+      legend.title = element_text(size = 20),
+      legend.text = element_text(size = 15),
+      axis.line = element_blank())
+ggsave(filename = "fig-subtype1.png", plot = p, width = 6, height = 5)
+
+pl1 = FeaturePlot(tiss,features = names(gene_partition)[gene_partition %in% c(5,6)][c(1:3,7,9)],ncol = 5,order = TRUE)
+pl2 = FeaturePlot(tiss,features = names(gene_partition)[gene_partition == 43][1:5],ncol = 5,order = TRUE)
+# pl_module_subtype = lapply(pl_module[c(5,6,43)],function(x) x + scale_color_gradient(low = "lightgrey", high = "blue"))
+
+pl = wrap_plots(pl1 & theme(plot.title = element_text(color = scales::hue_pal()(2)[1])),
+                pl2 & theme(plot.title = element_text(color = scales::hue_pal()(2)[2])),nrow = 2) & labs(color = "Expression") & theme(
+      plot.title = element_text(face="bold", size = 30),
+      legend.title = element_text(size = 20),
+      legend.text = element_text(size = 15),
+      axis.title = element_blank(),axis.text = element_blank(), axis.ticks=element_blank(),axis.line = element_blank())
+
+ggsave(filename = "fig-subtype2.png", plot = pl, width = 4*5 + 3, height = 4*2)
+```

### 2023-10-29 LMD_function.R

@@ -166,24 +166,34 @@ Obtain_Pls <- function(W, max_time){
   names(P_ls) = 2^seq(1,floor(log(max_step,2)))
   return(P_ls)
 }
-plot_knn_graph <- function(affinity_m, label, layout){
+plot_knn_graph <- function(affinity_m, label = NULL, layout){
   layout <- as.matrix(layout)
-  g <- graph_from_adjacency_matrix(affinity_m, 'undirected')
+  g <- graph_from_adjacency_matrix(affinity_m, 'undirected', weighted = TRUE)
   V(g)$frame.color <- NA
   
-  # coldef <- c(brewer.pal(6,"Blues"),brewer.pal(6,"Reds"))
-  
-  coldef <- setNames(
-    colorRampPalette(brewer.pal(12, "Paired"))(length(unique(label))),
-    unique(label))
-  
-  col <- coldef[label]
-  names(col) <- label
-  
-  plot.igraph(g, layout = layout, vertex.size=1,
-              vertex.color=col,vertex.label=NA)
-  legend("topright",names(coldef),col=coldef,
-         pch=16,cex=0.5,bty='n')
+  if(!is.null(label)){
+    V(g)$color <- label
+    coldef <- setNames(
+      colorRampPalette(brewer.pal(12, "Paired"))(length(unique(label))),
+      unique(label))
+    p = ggraph(g,layout = layout) + 
+      geom_edge_link(aes(width = weight),color = "grey") + 
+      geom_node_point(aes(color = color), size = 1) + 
+      scale_edge_width_continuous(range = c(0.5, 1), breaks = c(0.5, 1)) + 
+      scale_color_manual(values = coldef) +
+      theme_graph()
+  }else{
+    p = ggraph(g,layout = layout) + 
+      geom_edge_link(aes(width = weight),color = "grey") + 
+      geom_node_point(size = 1) + 
+      scale_edge_width_continuous(range = c(0.5, 1), breaks = c(0.5, 1)) + 
+      theme_graph()
+  }
+  p = p + theme(
+    plot.title = element_text(face="bold", size = 30),
+    legend.title = element_text(size = 20),
+    legend.text = element_text(size = 15))
+  return(p)
 }
 
 # Calculating Diffusion score =========
@@ -238,21 +248,28 @@ LMD <- function(expression, feature_space, knn = 5, max_time = 2^10){
 }
 
 # Visualization =========
-FeaturePlot_custom <- function(value,coord,value_name = NULL,title_name = NULL){
[…]

### 2023-11-03 Demo_marrow_facs.Rmd
path adjustment

@@
 ---
-title: "Tutorial for Runing Mouse Bone Marrow Dataset"
+title: "Tutorial for Runing LMD on Mouse Bone Marrow Dataset"
 author: "Ruiqi Li"
@@
 sapply(github_packages, function(pkg) if (!requireNamespace(strsplit(pkg, "/")[[1]][2], quietly = TRUE)) remotes::install_github(pkg))
-
 library("Seurat")
@@
 ```
+# Create a folder to save results
+```{r}
+folder_path = "marrow_result"
+if (!dir.exists(path)) {
+  # If it doesn't exist, create it
+  dir.create(path)
+} else {
+}
+```
 
 # Preparing input data
-
 ```{r}
@@
 ```{r}
-source("./LMD_function.R", echo = F)
+source("LMD_function.R", echo = F)
 ```
@@
 # Denoising
-#' We use ALRA @linderman2022zero to Denoise our data
+# We use ALRA @linderman2022zero to Denoise our data
 tiss <- RunALRA(tiss, assay = "RNA")
@@
 abbrev[grepl("granulocyte monocyte",abbrev)] = "GMP"
-# abbrev[grepl("granulocytopoietic",abbrev)] = "G-poiesis"
 names(abbrev) = levels(cell_label)
@@
 
-# # Order Cell Type based on its centroid in Featurespace
-# Idents(tiss) = cell_label
-# ordered_cell_types <- unique(cell_label)[hclust(dist(t(sapply(unique(cell_label), function(x) colMeans(Embeddings(tiss, reduction = "pca")[WhichCells(tiss, idents = x), ,drop = FALSE]) ))))$order]
-
 # Calculate Jaccard index between module activity & cell type
@@
 })
-# jaccard_index = jaccard_index[rev(ordered_cell_types),]
-# rm(ordered_cell_types)
 ```
@@
 load('droplet_Marrow_seurat_tiss.Robj')
-tiss_droplet <- UpdateSeuratObject(tiss_droplet)
-
-# path <- "/data/ruiqi/local_marker/dataset/tabular_muris/"
-# tissue_name = "marrow_droplet"
-# tiss_droplet <- readRDS(sprintf("%s%s.rds",path,tissue_name))
+tiss_droplet <- UpdateSeuratObject(tiss)
+saveRDS(tiss_droplet, file = "processed_droplet_Marrow_srat.rds")
 
@@
 ```{r}
-save(gene_partition,dist,gene_hree,cell_block,cell_block_droplet,one_hot_matrix,one_hot_matrix_droplet,jaccard_index,jaccard_index_droplet,file = "./marrow_result/module_result.rdata")
+save(gene_partition,dist,gene_hree,cell_block,cell_block_droplet,one_hot_matrix,one_hot_matrix_droplet,jaccard_index,jaccard_index_droplet,file = file.path(folder_path, "module_result.rdata"))
+
+saveRDS(tiss, file = "processed_facs_Marrow_srat.rds")
 ```
@@
 
-# Set all genes(after filtering) in the expression matrix as background,
-# such that the enrichment result of any module is not attributed to its high expression levels
+# Set all genes(after filtering) in the expression matrix as background
 universe_df = data.frame("symbol" = rownames(dat),"entrez" = mapIds(org.Mm.eg.db, keys=rownames(dat), column="ENTREZID", keytype="SYMBOL"))
@@
 df_module[is.na(df_module)] = ""
-write.csv(df_module,file = "./marrow_result/module_description.csv",row.names = FALSE)
+write.csv(df_module,file = file.path(folder_path, "module_description.csv"),row.names = FALSE)
 ```
@@
 
-write.csv(top_enrich_pathway_cc,file = "./marrow_result/reactome_pathway_cc.csv")
+write.csv(top_enrich_pathway_cc,file = file.path(folder_path, "reactome_pathway_cc.csv"))
 ```
+
 ### Annotate Cell Cycle
 ```{r}
-library(gprofiler2)
-mmus_s = gorth(cc.genes.updated.2019$s.genes, source_organism = "hsapiens", target_organism = "mmusculus")$ortholog_name
-mmus_g2m = gorth(cc.genes.updated.2019$g2m.genes, source_organism = "hsapiens", target_organism = "mmusculus")$ortholog_name
+mmus_s = gprofiler2::gorth(cc.genes.updated.2019$s.genes, source_organism = "hsapiens", target_organism = "mmusculus")$ortholog_name
+mmus_g2m = gprofiler2::gorth(cc.genes.updated.2019$g2m.genes, source_organism = "hsapiens", target_organism = "mmusculus")$ortholog_name
 tiss = CellCycleScoring(tiss, s.features = mmus_s, 
                         g2m.features = mmus_g2m, set.ident = FALSE)
-cell_cycle_table = tiss@meta.data[,c("S.Score","G2M.Score","Phase")]
-write.csv(cell_cycle_table,file = "./marrow_result/annotated_cell_cycle.csv")
+cell_cycle_table = data.frame(tiss@meta.data[,c("S.Score","G2M.Score","Phase")],"Dataset" = "FACS")
 ```
 
-<!-- ## Multi-resolution Modules - analysis -->
-<!-- ```{r} -->
-<!-- df_module = read.csv("./marrow_result/module_description.csv") -->
-<!-- df_module$'multi-celltype' = NA -->
-<!-- df_module$'note' = NA -->
+### Transferablity
+```{r}
+tiss_droplet = CellCycleScoring(tiss_droplet, s.features = mmus_s, 
+                        g2m.features = mmus_g2m, set.ident = FALSE)
+cell_cycle_table = rbind(cell_cycle_table,data.frame(tiss_droplet@meta.data[,c("S.Score","G2M.Score","Phase")],"Dataset" = "Droplet"))
 
-<!-- ct_prob_per_module = do.call(rbind,lapply(unique(cell_label),function(x){apply(cell_block[cell_label == x,],2,median)})) -->
-<!-- rownames(ct_prob_per_module) = unique(cell_label) -->
-<!-- ct_num_per_module = sort(apply(ct_prob_per_module,2,function(x)sum(x > 0.5))) -->
-
-<!-- m = 19 -->
-<!-- df_module[m,"celltype"] = paste0(names(which(ct_prob_per_module[,m] > 0.5)),collapse = ",") -->
-
-<!-- df_module[54,"multi-celltype"] = c("T/NK/NKT") -->
-<!-- df_module[27,"multi-celltype"] = c("Progenitor/NK/NKT") -->
-<!-- df_module[33,"multi-celltype"] = c("Progenitor/basophil/granulocytopoietic/half-granulocyte") -->
-<!-- df_module[46,"multi-celltype"] = c("monocyte/granulocytopoietic/half-granulocyte") -->
-<!-- df_module[48,"multi-celltype"] = c("monocyte/granulocytopoietic/granulocyte") -->
-<!-- df_module[49,"multi-celltype"] = c("monocyte/macrophage/granulocytopoietic/granulocyte") -->
-<!-- df_module[55,"multi-celltype"] = c("NK") -->
-<!-- df_module[56,"multi-celltype"] = c("NK") -->
-<!-- df_module[13,"multi-celltype"] = c("B") -->
-<!-- df_module[50,"multi-celltype"] = c("monocyte/granulocytopoietic/granulocyte/basophil") -->
-<!-- df_module[28,"multi-celltype"] = c("Progenitor") -->
-<!-- df_module[9,"multi-celltype"] = c("immature/naive B") -->
-<!-- df_module[11,"multi-celltype"] = c("late pro/pre B") -->
-<!-- df_module[7,"multi-celltype"] = c("late pro/pre B") -->
-<!-- df_module[44,"multi-celltype"] = c("granulocytopoietic/half-granulocyte") -->
-<!-- df_module[47,"multi-celltype"] = c("granulocytopoietic/granulocyte") -->
-<!-- df_module[52,"multi-celltype"] = c("T") -->
-<!-- df_module[53,"multi-celltype"] = c("macrophage/NK") -->
-<!-- df_module[1,"multi-celltype"] = c("monocyte/macrophage/granulocyte monocyte progenitor cell") -->
-<!-- df_module[20,"multi-celltype"] = c("granulocytopoietic/Progenitor") -->
-<!-- df_module[c(42,35),"multi-celltype"] = c("below-half-granulocyte/monocyte") -->
-<!-- df_module[c(36,38),"multi-celltype"] = c("granulocyte monocyte progenitor cell/monocyte") -->
-<!-- df_module[2,"multi-celltype"] = c("granulocyte monocyte progenitor cell/macrophage") -->
-<!-- df_module[4,"celltype"] = c("granulocyte monocyte progenitor cell") -->
-<!-- df_module[c(5,6,43),"note"] = "subtype" -->
-
-<!-- df_module[c(16,18,17,31,32),'note'] = "cell_cycle" -->
-<!-- ``` -->
+write.csv(cell_cycle_table,file = file.path(folder_path, "annotated_cell_cycle.csv"))
+```

### 2023-11-03 Marker_identfication_benchmark.Rmd
path adjustment

@@
 title: "Marker_identification_Benchmark"
+author: "Ruiqi Li"
 output: pdf_document
@@
 ```{r}
-path <- "/data/ruiqi/local_marker/dataset/tabular_muris/"
-tissue_name = "marrow_facs" #lung_facs, pancreas_facs
-tiss <- readRDS(paste0(path,tissue_name,".rds"))
+tissue_fullname = "Marrow"
+
+if(tissue_fullname == "Bone marrow"){
+  tissue_name = "marrow_facs"
+  if (!file.exists('facs_Marrow_seurat_tiss.Robj')) {
+   # If load url doesn't work, please download it and load from your working directory
+  options(timeout=6000)
+  download.file("https://figshare.com/ndownloader/files/13092380",
+                destfile = 'facs_Marrow_seurat_tiss.Robj', method = 'libcurl')
+  }
+  load('facs_Marrow_seurat_tiss.Robj')
+  tiss <- UpdateSeuratObject(tiss)
+}if(tissue_fullname == "Pancreas"){
+  tissue_name = "pancreas_facs"
+  if (!file.exists('facs_Pancreas_seurat_tiss.Robj')) {
+     # If load url doesn't work, please download it and load from your working directory
+    options(timeout=6000)
+    download.file("https://figshare.com/ndownloader/files/13092386",
+                  destfile = 'facs_Pancreas_seurat_tiss.Robj', method = 'libcurl')
+  }
+  load('facs_Pancreas_seurat_tiss.Robj')
+  tiss <- UpdateSeuratObject(tiss)
+}if(tissue_fullname == "Lung"){
+  tissue_name = "lung_facs"
+  if (!file.exists('facs_Lung_seurat_tiss.Robj')) {
+     # If load url doesn't work, please download it and load from your working directory
+    options(timeout=6000)
+    download.file("https://figshare.com/ndownloader/files/13092194",
+                  destfile = 'facs_Lung_seurat_tiss.Robj', method = 'libcurl')
+   }
+  load('facs_Lung_seurat_tiss.Robj')
+  tiss <- UpdateSeuratObject(tiss)
+}
+
 ```
@@
 ```{r}
-path = "./benchmark_result/"
-tissue_name = "marrow_facs"
+folder_path = "benchmark_result"
+# Check if the folder exists
+if (!dir.exists(folder_path)) {
+  # If it doesn't exist, create it
+  dir.create(folder_path)
+} else {
+}
 ```
@@
 max_logfc = sort(max_logfc,decreasing = TRUE)
-write.table(max_logfc,file = paste0(path,tissue_name,"_ground_truth_c1.txt"))
+write.table(max_logfc,file = file.path(folder_path, paste0(tissue_name,"_ground_truth_c1.txt")))
 ```
@@
 http://117.50.127.228/CellMarker/CellMarker_download_files/file/Cell_marker_Seq.xlsx
-<!-- PanglaoDB: https://panglaodb.se/markers/PanglaoDB_markers_27_Mar_2020.tsv.gz and exclude genes that reported only once -->
 ```{r}
@@
 cell_marker_db = readxl::read_xlsx("/data/ruiqi/local_marker/dataset/Cell_Marker/Cell_marker_All.xlsx") %>% filter(species == "Mouse") %>% filter(tissue_class == "Bone marrow") %>% filter(!is.na(cellontology_id)) %>% filter(!is.na(PMID)) %>% mutate(gene_name1 = marker) %>% mutate(gene_name2 = Symbol) %>% select(c("gene_name1","gene_name2"))
-# %>% group_by(dplyr::across(dplyr::everything())) %>% dplyr::summarise(count = n(), .groups = "drop") %>% filter(count >= 1) # Lung, Pancreas
+# Lung, Pancreas
 
@@
 celldb_marker = celldb_marker[!is.na(celldb_marker)] %>% unique()
-write.table(celldb_marker,file = paste0(path,tissue_name,"_ground_truth_c2.txt"))
+write.table(celldb_marker,file = file.path(folder_path, paste0(tissue_name,"_ground_truth_c2.txt")))
 ```
@@
 # Run Each Method and save results
-```{r}
-path = "./benchmark_result/"
-# tissue_name = "marrow_facs"
-```
-
 ## LocalMarkerDetector
@@
 method = "lmd"
-source("./LMD_function.R", echo = F)
+source("LMD_function.R", echo = F)
 res = LMD(dat,feature_space,max_time = 2^10,knn = 7)
@@
 marker[order(marker$score),'rank'] = 1:nrow(marker)
-write.table(marker,file = sprintf("%s%s_%s.csv",path,method,tissue_name),row.names = TRUE)
+write.table(marker,file = file.path(folder_path, paste0(method,"_",tissue_name,".csv")),row.names = TRUE)
 ```
 
-## Seurat v3 Wilcoxon unfiltered
+## Seurat v4 Wilcoxon unfiltered
 Genes were ranked by their lowest adjusted P-value in any cluster (increasing order), if two genes had the same, prioritize genes with larger FC.
@@
 marker <- merge(marker,gene_rank,by = "gene")
-write.table(marker,file = sprintf("%s%s_%s.csv",path,method,tissue_name),row.names = TRUE)
+write.table(marker,file = file.path(folder_path, paste0(method,"_",tissue_name,".csv")),row.names = TRUE)
 ```
-## Seurat v3 default
+## Seurat v4 default
 Genes were ranked by their lowest adjusted P-value in any cluster (increasing order), if two genes had the same, prioritize genes with larger FC.
@@
 marker <- merge(marker,gene_rank,by = "gene")
-write.table(marker,file = sprintf("%s%s_%s.csv",path,method,tissue_name),row.names = TRUE)
+write.table(marker,file = file.path(folder_path, paste0(method,"_",tissue_name,".csv")),row.names = TRUE)
 ```
@@
 
-write.table(marker,file = sprintf("%s%s_%s.csv",path,method,tissue_name),row.names = TRUE)
+write.table(marker,file = file.path(folder_path, paste0(method,"_",tissue_name,".csv")),row.names = TRUE)
 ```
@@
 marker[order(marker$FDR),"rank"] = 1:nrow(marker)
-write.table(marker,file = sprintf("%s%s_%s.csv",path,method,tissue_name),row.names = TRUE)
+write.table(marker,file = file.path(folder_path, paste0(method,"_",tissue_name,".csv")),row.names = TRUE)
 ```
@@
 marker$'rank' = marker$MarcoPolo_rank + 1
-write.table(marker,file = sprintf("%s%s_%s.csv",path,method,tissue_name),row.names = TRUE)
+write.table(marker,file = file.path(folder_path, paste0(method,"_",tissue_name,".csv")),row.names = TRUE)
 ```
@@
 
-write.table(marker,file = sprintf("%s%s_%s.csv",path,method,tissue_name),row.names = FALSE)
+write.table(marker,file = file.path(folder_path, paste0(method,"_",tissue_name,".csv")),row.names = FALSE)
 ```
@@
 ```{r}
-path = "./benchmark_result/"
-# tissue_name = "marrow_facs"
-```
-```{r}
 method_ls = c("lmd","haystack","hotspot","marcopolo","semi","wilcox_default","wilcox_no_filter","hvg")
-method_fullname = setNames(c("LMD","singleCellHaystack","Hotspot","Marcopolo","SEMITONES","Seurat v3 default", "Seurat v3 Wilcoxon unfiltered","HighlyVariableGenes"),method_ls)
 
 df_benchmark = do.call(cbind,lapply(method_ls, function(method){
-  marker = read.table(sprintf("%s%s_%s.csv",path,method,tissue_name),header = TRUE)
+  marker = read.table(file.path(folder_path, paste0(method,"_",tissue_name,".csv")),header = TRUE)
   if("gene" %in% colnames(marker)){
@@
 
-write.table(df_benchmark,file = paste0(path,tissue_name,"_benchmark_rank_table.csv"),row.names = TRUE)
+write.table(df_benchmark,file = file.path(folder_path, paste0(tissue_name,"_benchmark_rank_table.csv")),row.names = TRUE)
 ```
@@
 ```{r}
-path = "./benchmark_result/"
-# tissue_name = "marrow_facs"
-df_benchmark = read.table(paste0(path,tissue_name,"_benchmark_rank_table.csv"))
+df_benchmark = read.table(file.path(folder_path, paste0(tissue_name,"_benchmark_rank_table.csv")))
 
 method_ls = c("lmd","haystack","hotspot","marcopolo","semi","wilcox_default","wilcox_no_filter","hvg")
-method_fullname = setNames(c("LMD","singleCellHaystack","Hotspot","Marcopolo","SEMITONES","Seurat v3 default", "Seurat v3 Wilcoxon unfiltered","HighlyVariableGenes"),method_ls)
 ```
 ```{r}
-max_logfc = read.table(paste0(path,tissue_name,"_ground_truth_c1.txt")) %>% rownames()
-celldb_marker = read.table(paste0(path,tissue_name,"_ground_truth_c2.txt"))[,1]
+max_logfc = read.table(file.path(folder_path, paste0(tissue_name,"_ground_truth_c1.txt"))) %>% rownames()
+celldb_marker = read.table(file.path(folder_path, paste0(tissue_name,"_ground_truth_c2.txt")))[,1]
 gt_list = c(lapply(seq(50,1000,50),function(x) max_logfc[1:x]),list(celldb_marker))
@@
 df$Method = as.factor(df$Method)
-levels(df$Method) = method_fullname[levels(df$Method)]
 df$gt_set = factor(df$gt_set,levels = names(gt_list))
-write.table(df,file = paste0(path,tissue_name,"_auroc.csv"),row.names = FALSE)
+write.table(df,file = file.path(folder_path, paste0(tissue_name,"_auroc.csv")),row.names = FALSE)
 ```
@@
 ```{r}
-path = "./benchmark_result/"
-# tissue_name = "marrow_facs"
-```
-```{r}
-df_benchmark = read.table(paste0(path,tissue_name,"_benchmark_rank_table.csv"))
+df_benchmark = read.table(file.path(folder_path, paste0(tissue_name,"_benchmark_rank_table.csv")))
 
@@
 
-write.table(score_measure,file = paste0(path,tissue_name,"_density_index.csv"),row.names = FALSE)
+write.table(score_measure,file = file.path(folder_path, paste0(tissue_name,"_density_index.csv")),row.names = FALSE)
 ```
@@
 # Sensitivity Test (Try different K)
-```{r}
-path = "./marrow_result/"
-# tissue_name = "marrow_facs"
-```
 ```{r}
@@
 colnames(score_df) = paste0("kNN",c(3:10,20,30,40,50))
-write.table(score_df,file = paste0(path,tissue_name,"_different_knn.csv"),row.names = TRUE)
+write.table(score_df,file = file.path(folder_path, paste0(tissue_name,"_different_knn.csv")),row.names = TRUE)
 ```
@@
 score_ranked <- data.frame(score_df) %>% mutate(across(everything(), rank))
-path = "./benchmark_result/"
-max_logfc = read.table(paste0(path,tissue_name,"_ground_truth_c1.txt")) %>% rownames()
-celldb_marker = read.table(paste0(path,tissue_name,"_ground_truth_c2.txt"))[,1]
+
+max_logfc = read.table(file.path(folder_path, paste0(tissue_name,"_ground_truth_c1.txt"))) %>% rownames()
+celldb_marker = read.table(file.path(folder_path, paste0(tissue_name,"_ground_truth_c2.txt")))[,1]
 
@@
 })
-path = "./marrow_result/"
+
 auc_df = data.frame(auc_df,gt = names(gt_list))
-write.table(auc_df,file = paste0(path,tissue_name,"_sensitivity_auroc.csv"),row.names = FALSE)
+write.table(auc_df,file = file.path(folder_path, paste0(tissue_name,"_sensitivity_auroc.csv")),row.names = FALSE)
 ```

### 2023-11-03 paper_figure.Rmd
path adjustment

@@
 title: "Paper_Figure"
+author: "Ruiqi Li"
 output: pdf_document
@@
 library("Seurat")
-path <- "/data/ruiqi/local_marker/dataset/tabular_muris/"
-tissue_name = "marrow_facs"
-tiss <- readRDS(sprintf("%s%s.rds",path,tissue_name))
+# tissue_name = "marrow_facs"
+# tiss <- readRDS(sprintf("%s%s.rds",path,tissue_name))
+tiss <- readRDS("processed_facs_Marrow_srat.rds")
+tiss_droplet <- readRDS("processed_droplet_Marrow_srat.rds")
 
@@
 ggsave(filename = "supfig-bonemarrow.png", plot = p, width = 8, height = 5, dpi = 300)
-# 
-# centroids <- aggregate(. ~ cell_label, data = as.data.frame(Embeddings(tiss, "tsne")), FUN = mean)
-# p <- DimPlot(tiss, group.by = "cell_ontology_class",label = FALSE, cols = coldef) + theme_void() + NoLegend() + ggtitle("Bone Marrow FACS")
-# p = p + 
-#   geom_text_repel(data = centroids, aes(x = tSNE_1, y = tSNE_2, label = cell_label, color = cell_label),
-#                    box.padding = 0.3, point.padding = 0.3, segment.color = 'grey50',max.overlaps = 100)
 ```
@@
 ```{r}
-source("./LMD_function.R", echo = F)
+source("LMD_function.R", echo = F)
 
@@
 
-P_ls = Obtain_Pls(W = W, max_time = 2^10)
+P_ls = Obtain_Pls(W = W, max_time = 2^15)
 rho = Rowwise_normalize(dat[,colnames(W)])
@@
 
-ggsave(filename = "fig-pipeline2.png", plot = p, width = 8, height = 5)
+ggsave(filename = "fig-pipeline2.png", plot = p, width = 10, height = 5)
 
@@
   gene = genes[i]
-  FeaturePlot_diffusion(coord = visual_space,init_state = rho[gene,],P_ls = P_ls,W = W,check_time = check_time,gene_name = gene,gene_color = scales::hue_pal()(2)[i])
+  FeaturePlot_diffusion(coord = visual_space,init_state = rho[gene,],P_ls = P_ls,W = W,check_time = check_time,gene_name = NULL, gene_color = scales::hue_pal()(2)[i])
 })
@@
 method_ls = c("lmd","haystack","hotspot","marcopolo","semi","wilcox_default","wilcox_no_filter","hvg")
-method_fullname = setNames(c("LMD","singleCellHaystack","Hotspot","Marcopolo","SEMITONES","Seurat v3 default", "Seurat v3 Wilcoxon unfiltered","HighlyVariableGenes"),method_ls)
+method_fullname = setNames(c("LMD","singleCellHaystack","Hotspot","Marcopolo","SEMITONES","Seurat v4 default", "Seurat v4 Wilcoxon unfiltered","HVGs"),method_ls)
 coldef = setNames(
@@
 tissue_ls = c("marrow_facs","pancreas_facs","lung_facs")
-df_auc = do.call(rbind,lapply(tissue_ls,function(tissue_name){df = read.table(paste0(path,tissue_name,"_auroc.csv"),header = TRUE); df$'Tissue' = tissue_name; df}))
+df_auc = do.call(rbind,lapply(tissue_ls,function(tissue_name){df = read.table(paste0(path,tissue_name,"_auroc.csv"),header = TRUE); df$'Tissue' = tissue_name; df$'Method' = method_fullname[df$'Method']; df}))
 
@@
   geom_bar(stat = "identity", position = "dodge") +
-  geom_text(aes(label = round(AUROC, 2)), position = position_dodge(width = 0.85), vjust = -0.25, size = 3) +
+  geom_text(aes(label = round(AUROC, 2)), position = position_dodge(width = 1), vjust = -0.25, size = 2.5, angle = 45) +
   labs(x = "Tissue", y = "AUROC", fill = "Method") + coord_cartesian(ylim = c(0.5, 1)) +
@@
 fig = fig + theme(
-          legend.title = element_text(size = 20),
-          legend.text = element_text(size = 15),
+          legend.title = element_text(size = 10),
+          legend.text = element_text(size = 10),
           axis.title.x = element_text(size = 15),
@@
 
-ggsave(filename = "fig-known_marker_comparison.png", plot = fig, width = 10, height = 5)
+ggsave(filename = "fig-known_marker_comparison.png", plot = fig, width = 8, height = 5)
 ```
 
-<!-- # Fig - Enrichment Overlap of Prioritized Genes with CellMarker Database Confidence Markers -->
-<!-- ```{r} -->
-<!-- path = "./benchmark_result/" -->
-<!-- tissue_name = "marrow_facs" -->
-
-<!-- library(Seurat) -->
-<!-- cell_marker_db = readxl::read_xlsx("/data/ruiqi/local_marker/dataset/Cell_Marker/Cell_marker_All.xlsx") %>% filter(species == "Mouse") %>%  -->
-<!--   filter(tissue_class == "Bone marrow") %>% filter(!is.na(cellontology_id)) %>%  -->
-<!--   filter(!is.na(PMID)) %>% mutate(gene_name1 = marker) %>% mutate(gene_name2 = Symbol) %>% select(c("gene_name1","gene_name2")) -->
-
-<!-- df_benchmark = read.table(paste0(path,tissue_name,"_benchmark_rank_table.csv")) -->
-
-<!-- method = "lmd" -->
-<!-- gene_bin = data.frame(gene = rownames(df_benchmark)[order(df_benchmark[,method])][1:1000]) -->
-<!-- gene_bin$freq = unlist(lapply(gene_bin$gene,function(i){ -->
-<!--   id = which(grepl(paste0("^", i, "$"), cell_marker_db$gene_name1, ignore.case = TRUE) + grepl(paste0("^", i, "$"), cell_marker_db$gene_name2, ignore.case = TRUE) >=1) -->
-<!--   return(length(id)) -->
-<!-- }) ) -->
-<!-- gene_bin$'freq_class' = cut(gene_bin$freq, breaks = c(-Inf,1,3,Inf), labels = c("none", "low", "high"), right = FALSE) -->
-
-<!-- gene_bin$bin <- cut(seq(1:nrow(gene_bin)), breaks = 40, labels = FALSE) -->
-<!-- gene_bin$'bin_label' = paste0(nrow(gene_bin)/max(gene_bin$bin)*(gene_bin$bin-1)+1,"-",nrow(gene_bin)/max(gene_bin$bin)*(gene_bin$bin)) -->
-<!-- gene_bin$'bin_label' = factor(gene_bin$'bin_label',levels = unique(gene_bin$'bin_label')) -->
-
-<!-- gene_freq_prop <- gene_bin %>% -->
-<!--   group_by(bin_label, freq_class) %>% -->
-<!--   summarise(n = n()) %>% -->
-<!--   mutate(freq = n / sum(n)) -->
-<!-- gene_freq_prop$freq_class = factor(gene_freq_prop$freq_class,levels = c("none","low","high"),labels = c("None(=0)","Low(1,2)","High(>=3)")) -->
-
-<!-- coldef = setNames(c("gray71","firebrick1","firebrick4"),levels(gene_freq_prop$freq_class)) -->
-
-<!-- fig = ggplot(gene_freq_prop, aes(fill=freq_class, y=freq, x=bin_label)) + -->
-<!--   geom_bar(position="fill", stat="identity") + -->
-<!--   ylab("Frequency") + -->
-<!--   xlab("Gene Rank") + -->
-<!--   labs(fill = "Confidence") + -->
-<!--   scale_fill_manual(values= coldef) + -->
-<!--   theme_minimal() +  -->
-<!--   theme(axis.text.x = element_text(angle = 45, hjust = 1)) -->
-
-<!-- fig -->
-<!-- ``` -->
-<!-- ```{r} -->
-<!-- genes = c("Ctla2a","Fpr2","Mmp8") -->
-<!-- pl = lapply(genes,function(g) FeaturePlot(tiss, g, order = TRUE) + labs(color = "Expression") + ggtitle(sprintf("%s (rank %d)",g,which(gene_bin == g)))) -->
-<!-- fig = cowplot::plot_grid(plotlist = pl, ncol = 3) -->
-<!-- fig -->
-<!-- ``` -->
-
-# Fig - Gene Ranking Variations between LMD and Other Methods
+# Fig - Gene Ranking Comparisons between LMD and Other Methods
 ```{r}
@@
   subtitle = sprintf("%s:%d, %s:%d",method_fullname[method_vs],df_label_all[x,method_vs],method_fullname[names(g)[i]],df_label_all[x,names(g)[i]])
-  color_label = ifelse(df_label_all[x,method_vs] > df_label_all[x,names(g)[i]],"red","blue")
+  color_label = ifelse(df_label_all[x,method_vs] > df_label_all[x,names(g)[i]],"blue","red")
   FeaturePlot(tiss,features = x,order =TRUE) + labs(color = "Expression") + labs(title = x,subtitle = subtitle) + 
@@
 pl = pl[order(seq(length(g))%%2,decreasing = TRUE)]
-pl[[1]] = pl[[1]] + labs(y = "Favored by LMD") + theme(axis.title.y = element_text(size = 20,face="bold", colour = "blue"))
-pl[[(length(pl)/2)+1]] = pl[[(length(pl)/2)+1]] + labs(y = "Favored by Other methods") + theme(axis.title.y = element_text(size = 20,face="bold", colour = "red"))
+pl[[1]] = pl[[1]] + labs(y = "Favored by LMD") + theme(axis.title.y = element_text(size = 20,face="bold", colour = "red"))
+pl[[(length(pl)/2)+1]] = pl[[(length(pl)/2)+1]] + labs(y = "Favored by Other methods") + theme(axis.title.y = element_text(size = 20,face="bold", colour = "blue"))
 
@@
 ggsave(filename = "fig-ranking_comparison2.png", plot = fig, width = 25, height = 10)
+
+write.csv(df_label_all,file = "./marrow_result/gene_rank_comparison.csv",row.names = TRUE)
 ```
@@
   df_label[c(which(df_label[,method] > 500 & df_label[,method_vs] <= 100) ),"color"] = fullmethodvs
-  df_label$'color' = as.factor(df_label$'color')
-  rank_cor = cor(log(df_label[,1:2],10))[1,2]
+  df_label$'color' = factor(df_label$'color',levels = c(fullmethodvs,fullmethod))
+  rank_cor = stats::cor(log(df_label[,1:2],base = 10),method = "pearson")[1,2]
   colnames(df_label)[1:2] = c("x","y")
@@
   scale_color_manual(values = c("red","blue","grey50"),
-                     breaks = c(fullmethod,fullmethodvs),
+                     breaks = c(fullmethodvs,fullmethod),
                      labels = c("Favored by LMD","Favored by Other methods")) + 
@@
 score_measure = read.table(paste0(path,tissue_name,"_density_index.csv"),header = TRUE)
+score_measure = score_measure %>% filter(Method!="wilcox_no_filter")
 score_measure$Method = ifelse(score_measure$Method %in% names(method_fullname),method_fullname[score_measure$Method],score_measure$Method)
@@
 
-line_type = replace(coldef,names(coldef),"solid")
-line_type["Seurat v3 Wilcoxon unfiltered"] = "twodash"
-line_type["All genes"] = "dashed"
+score_measure$'color' = coldef[score_measure$Method]
+score_measure[score_measure$Method == "All genes","color"] = "black"
+score_measure$'line_type' = "solid"
+score_measure[score_measure$Method == "All genes","line_type"] = "dashed"
 
-fig = ggplot(score_measure %>% filter(Method!="All genes"), aes(x = TopGene, y = DensityIndex, color = Method, group = Method, linetype = Method),linewidth = 2) + 
+score_measure[grepl("Seurat",score_measure$Method),"Method"] = "Seurat v4"
+
+# line_type = replace(coldef,names(coldef),"solid")
+# line_type["Seurat v4 Wilcoxon unfiltered"] = "twodash"
+# line_type["All genes"] = "dashed"
+
+fig = ggplot(score_measure %>% filter(Method!="All genes") %>% filter(TopGene >= 100), aes(x = TopGene, y = DensityIndex, color = Method, group = Method, linetype = Method),linewidth = 2) + 
   geom_line(lwd = 1) +
@@
   geom_hline(aes(yintercept = baseline, color = "All genes", linetype = "All genes")) +
-  scale_color_manual(values= c(coldef[names(coldef)%in%score_measure$Method],"All genes" = "black")) + 
-  scale_linetype_manual(values = line_type) +
+  scale_color_manual(values = score_measure[,c("Method","color")] %>% distinct() %>% tibble::deframe()) + 
+  scale_linetype_manual(values = score_measure[,c("Method","line_type")] %>% distinct() %>% tibble::deframe()) +
   theme(
-    legend.title = element_text(size = 20),
-    legend.text = element_text(size = 15),
-    axis.title.x = element_text(size = 20),
-    axis.title.y = element_text(size = 20),
-    axis.text.x = element_text(size = 15),
-    axis.text.y = element_text(size = 15),
+    legend.title = element_blank(),
+    legend.text = element_text(size = 10),
+    legend.position = "bottom",
+    legend.justification = 1,
+    axis.title.x = element_text(size = 15),
+    axis.title.y = element_text(size = 15),
+    axis.text.x = element_text(size = 10),
+    axis.text.y = element_text(size = 10),
     panel.grid = element_blank(),
@@
 
-ggsave(filename = "fig-densityindex.png", plot = fig, width = 8, height = 5)
+ggsave(filename = "fig-densityindex.png", plot = fig, width = 5, height = 5)
 ```
@@
 ```{r}
-cell_cycle = read.csv("./marrow_result/annotated_cell_cycle.csv")[["Phase"]]
-tiss$'Phase' = cell_cycle
-p1 = DimPlot(tiss,group.by = "Phase", cols = c("grey",scales::hue_pal()(3)[2:3])) + labs(title = "Phase annotated by Seurat",color = "Phase") + theme(
+cell_cycle = read.csv("./marrow_result/annotated_cell_cycle.csv")
+tiss$'Phase' = cell_cycle %>% filter(Dataset == "FACS") %>%.$Phase
+tiss$Phase = factor(tiss$Phase,levels = c("G1","S","G2M"))
+p1 = DimPlot(tiss,group.by = "Phase", cols = c("grey",scales::hue_pal()(3)[3:2])) + labs(title = "Phase annotated by Seurat",color = "Phase") + theme(
+  plot.title = element_text(size = 10),
   axis.title = element_blank(),
@@
       axis.line = element_blank())
-p = wrap_plots(p1,(wrap_plots(pl_module[c(16,17,18,31,32)],ncol = 3) & theme(plot.title = element_text(size = 15))) + plot_layout(guides = "collect"),nrow = 1,design = "AABBB")
+tiss_droplet$'Phase' = cell_cycle %>% filter(Dataset == "Droplet") %>%.$Phase
+tiss_droplet$Phase = factor(tiss_droplet$Phase,levels = c("G1","S","G2M"))
+p2 = DimPlot(tiss_droplet,group.by = "Phase", cols = c("grey",scales::hue_pal()(3)[3:2])) + labs(title = "Phase annotated by Seurat",color = "Phase") + theme(
+  axis.title = element_blank(),
+      axis.text = element_blank(),
+      axis.ticks=element_blank(),
+      legend.title = element_text(size = 20),
+      legend.text = element_text(size = 15),
+      axis.line = element_blank())
 
-ggsave(filename = "fig-module-cellcycle.png", plot = p, width = 3*5, height = 3*2)
+cc_module = c(32,31,18)
+pl_module_cc = lapply(cc_module,function(module){
+  FeaturePlot_custom(value = cell_block[,module], coord = visual_space, value_name = "Module Score",title_name = sprintf("Module %s (%d genes)",module, sum(gene_partition == module)))
+})
+pl_module_cc_droplet = lapply(cc_module,function(module){
+  FeaturePlot_custom(value = cell_block_droplet[,module], coord = data.frame(tiss_droplet@reductions$tsne@cell.embeddings), value_name = "Module Score",title_name = sprintf("Module %s (%d genes)",module, sum(gene_partition == module)))
+})
+p = wrap_plots(p1,wrap_plots(pl_module_cc,ncol = 3) & theme(plot.title = element_text(size = 15)) ,p2 + labs(title = NULL),wrap_plots(pl_module_cc_droplet,ncol = 3) & theme(plot.title = element_text(size = 15)) & labs(title = NULL),nrow = 2,design = "ABBB\nCDDD") + plot_layout(guides = "collect")
+
+ggsave(filename = "fig-module-cellcycle1.png", plot = p, width = 3*4 + 1, height = 3*2)
 ```
@@
 df = read.csv("./marrow_result/reactome_pathway_cc.csv")
-df$module = as.factor(df$module)
-df$GeneRatio = as.numeric(df$Count / table(gene_partition)[df$module])
-df$Description = factor(df$Description, levels = names(sort(table(df$Description))))
-levels(df$Description)[6] = "HDR through Homologous Recombination\n or Single Strand Annealing (SSA)"
+df = df %>% filter(module %in% cc_module)
+df$GeneRatio = unlist(lapply(df$GeneRatio,function(x) eval(parse(text = x))))
+# df$GeneRatio = as.numeric(df$Count / table(gene_partition)[df$module])
+Description_order = levels(factor(df$Description))
+Description_order = rev(Description_order[order(match(Description_order,c("Cell Cycle","Cell Cycle, Mitotic","Mitotic G1 phase and G1/S transition","G1/S Transition","S Phase","M Phase","DNA Replication","Mitotic Prometaphase")))])
+df$Description = factor(df$Description,levels = Description_order) 
+df$module = factor(df$module,levels = c(32,31,18))
+
 p = ggplot(df, aes(x=module, y=Description)) +
   geom_point(aes(size=GeneRatio, color=p.adjust), alpha=0.7) +
-  scale_color_gradientn(colors=colorRampPalette(c("red", "blue"))(99), name="p.adjust") +
+  scale_color_gradientn(colors=colorRampPalette(c("red", "blue"))(99), name="p.adjust",limits = c(0,0.05)) +
   theme_minimal() +
-  labs(size="Gene Ratio") +
+  labs(size="Gene Ratio",x = "Module") +
   theme(legend.position="right",
@@
 # Fig - Subtype Investigation
+late pro -> precursor B -> immature -> Naive B
+8(Lef1, Slamf7) - lateproB, 6(Vreb1,Igll1)- early preB, 7(Rag1, Rag2)/11(Chchd10) - lateproB-preB, 9(Ms4a1, Cd22) - immature-naive, 10 (Fcer2a, Ms4a4c, Cxcr5) - naive B
+
+13 (Fcrla,Pax5,Cd79a,Cd19)
+Dntt: 29 (HSC)
+Rag1, Rag2, Dntt: early pro-B
+Pax5, Vpreb1, Dntt, Cd19, Cd22-: late-proB
+Vpreb1-, Dntt-, Pax5, Rag1, Rag2, Cd19, Chchd10, Cd22-, : pre-B
+Chchd10,Cd79a,Cd19,Ms4a1,Pax5,Cd22: immature
+Cd79a, Cd19, Ms4a1, Pax5, Rag1- Rag2-, Chchd10, Cd22: naive B
+
 ```{r}
-p <- DimPlot(tiss, group.by = "cell_ontology_class",label = FALSE, cols = setNames(scales::hue_pal()(2),c("late pro-B cell","granulocyte"))) + ggtitle("Pre-annotated cell types") + theme(
-  axis.title = element_blank(),
-      axis.text = element_blank(),
-      axis.ticks=element_blank(),
-      legend.title = element_text(size = 20),
-      legend.text = element_text(size = 15),
-      axis.line = element_blank())
-ggsave(filename = "fig-subtype1.png", plot = p, width = 6, height = 5)
+B_module = c(7,9,10)
+B_module_genes = list("8" = c("Lef1","Slamf7"),
+                      "6" = c("Vreb1","Igll1"),
+                      "7" = c("Rag1", "Rag2"),
+                      "9" = c("Ms4a1", "Cd22"),
+                      "10" = c("Fcer2a","Ms4a4c","Cxcr5"))
+cell_id = grep("B cell",tiss$cell_ontology_class)
+cell_id = cell_id[tiss$cell_ontology_class[cell_id] != "B cell"]
+pl_Bmodule = lapply(B_module,function(module){
+  FeaturePlot_custom(value = cell_block[cell_id,module], coord = visual_space[cell_id,], value_name = "Module Score",title_name = sprintf("Module %s",module)) + labs(subtitle = paste0(B_module_genes[[as.character(module)]],collapse = ","))
+})
+pl_Bmodule = lapply(1:length(pl_Bmodule),function(i){
+  if(i!=1){pl_Bmodule[[i]] + NoLegend()}else{pl_Bmodule[[i]]}})
 
-pl1 = FeaturePlot(tiss,features = names(gene_partition)[gene_partition %in% c(5,6)][c(1:3,7,9)],ncol = 5,order = TRUE)
-pl2 = FeaturePlot(tiss,features = names(gene_partition)[gene_partition == 43][1:5],ncol = 5,order = TRUE)
-# pl_module_subtype = lapply(pl_module[c(5,6,43)],function(x) x + scale_color_gradient(low = "lightgrey", high = "blue"))
+coldef = setNames(colorRampPalette(brewer.pal(12, "Paired"))(length(unique(cell_label))),unique(cell_label))
+p = DimPlot(subset(tiss,cells = colnames(tiss)[cell_id]), group.by = "cell_ontology_class",label = TRUE, cols = coldef,repel = TRUE,label.size = 3) + NoLegend() + labs(title = NULL)
 
-pl = wrap_plots(pl1 & theme(plot.title = element_text(color = scales::hue_pal()(2)[1])),
-                pl2 & theme(plot.title = element_text(color = scales::hue_pal()(2)[2])),nrow = 2) & labs(color = "Expression") & theme(
-      plot.title = element_text(face="bold", size = 30),
-      legend.title = element_text(size = 20),
-      legend.text = element_text(size = 15),
-      axis.title = element_blank(),axis.text = element_blank(), axis.ticks=element_blank(),axis.line = element_blank())
+p = wrap_plots(wrap_plots(p) & theme(axis.title = element_blank(),axis.text = element_blank(), axis.ticks=element_blank(), axis.line = element_blank()),(wrap_plots(pl_Bmodule,nrow = 1) &  theme(plot.title = element_text(size = 15))) + plot_layout(guides = "collect"), nrow = 1, design = "ABBB")
 
-ggsave(filename = "fig-subtype2.png", plot = pl, width = 4*5 + 3, height = 4*2)
+ggsave(filename = "fig-subtypeB.png", plot = p, width = 4*3, height = 3)
 ```
+
+Granulocyte: 
+granulopoietic -> granulocyte
+45 -> 44 -> 41 -> 43
+45: "Fcnb"
+44: "Camp", "Ltf",
+41: "Retnlg", "Cxcr2"
+43: Stfa2l1, Il1b, Gm5483, H2-Q10, and Irg1
+
+```{r}
+G_module = c(45,44,41,43)
+G_module_genes = list("45" = c("Fcnb"),
+                      "44" = c("Camp", "Ltf"),
+                      "41" = c("Retnlg", "Cxcr2"),
+                      "43" = c("Stfa2l1", "Il1b", "Gm5483", "H2-Q10", "Irg1"))
+cell_id = tiss$cell_ontology_class %in% c("granulocyte","granulocytopoietic cell")
+
+coldef = setNames(colorRampPalette(brewer.pal(12, "Paired"))(length(unique(cell_label))),unique(cell_label))
+p = DimPlot(subset(tiss,cells = colnames(tiss)[cell_id]), group.by = "cell_ontology_class",label = TRUE,cols = coldef) + NoLegend() + labs(title = NULL)
+
+pl_Gmodule = lapply(G_module,function(module){
+  FeaturePlot_custom(value = cell_block[cell_id,module], coord = visual_space[cell_id,], value_name = "Module Score",title_name = sprintf("Module %s",module)) + labs(subtitle = paste0(G_module_genes[[as.character(module)]],collapse = ","))
+})
+pl_Gmodule = lapply(1:length(pl_Gmodule),function(i){
+  if(i!=3){pl_Gmodule[[i]] + NoLegend()}else{pl_Gmodule[[i]]}})
+
+p = wrap_plots(wrap_plots(p) & theme(axis.title = element_blank(),axis.text = element_blank(), axis.ticks=element_blank(), axis.line = element_blank()),(wrap_plots(pl_Gmodule,nrow = 1) &  theme(plot.title = element_text(size = 15))) + plot_layout(guides = "collect"), nrow = 1, design = "ABBBB")
+
+ggsave(filename = "fig-subtypeG.png", plot = p, width = 5*3, height = 3)
+```
+
+<!-- ```{r} -->
+<!-- p <- DimPlot(tiss, group.by = "cell_ontology_class",label = FALSE, cols = setNames(scales::hue_pal()(2),c("late pro-B cell","granulocyte"))) + ggtitle("Pre-annotated cell types") + theme( -->
+<!--   axis.title = element_blank(), -->
+<!--       axis.text = element_blank(), -->
+<!--       axis.ticks=element_blank(), -->
+<!--       legend.title = element_text(size = 20), -->
+<!--       legend.text = element_text(size = 15), -->
+<!--       axis.line = element_blank()) -->
+<!-- ggsave(filename = "fig-subtype1.png", plot = p, width = 6, height = 5) -->
+
+<!-- pl1 = FeaturePlot(tiss,features = names(gene_partition)[gene_partition %in% c(5,6)][c(1:3,7,9)],ncol = 5,order = TRUE) -->
+<!-- pl2 = FeaturePlot(tiss,features = names(gene_partition)[gene_partition == 43][1:5],ncol = 5,order = TRUE) -->
+<!-- # pl_module_subtype = lapply(pl_module[c(5,6,43)],function(x) x + scale_color_gradient(low = "lightgrey", high = "blue")) -->
+
+<!-- pl = wrap_plots(pl1 & theme(plot.title = element_text(color = scales::hue_pal()(2)[1])), -->
+<!--                 pl2 & theme(plot.title = element_text(color = scales::hue_pal()(2)[2])),nrow = 2) & labs(color = "Expression") & theme( -->
+<!--       plot.title = element_text(face="bold", size = 30), -->
+<!--       legend.title = element_text(size = 20), -->
+<!--       legend.text = element_text(size = 15), -->
+<!--       axis.title = element_blank(),axis.text = element_blank(), axis.ticks=element_blank(),axis.line = element_blank()) -->
+
+<!-- ggsave(filename = "fig-subtype2.png", plot = pl, width = 4*5 + 3, height = 4*2) -->
+<!-- ``` -->

### 2023-11-03 LMD_function.R

@@ -302,7 +302,7 @@ FeaturePlot_diffusion <- function(coord, init_state, P_ls = NULL, W = NULL, chec
   colnames(multi_state) = check_time
   degree_node = rowSums(W) # Normalize multi_state with degree of node for visualization
   pl = lapply(seq(ncol(multi_state)),function(i){
-    legend_break_label = seq(0, max(multi_state[,i]), length.out = 5)
+    legend_break_label = seq(0, max(multi_state[,i]), length.out = 2)
     if(max(legend_break_label)>1e-3){
       legend_break_label = signif(legend_break_label, digits = 2)
     }else{
@@ -310,12 +310,14 @@ FeaturePlot_diffusion <- function(coord, init_state, P_ls = NULL, W = NULL, chec
     }
     p = suppressWarnings({
       FeaturePlot_custom(multi_state[,i]/max(multi_state[,i]),coord,
-                                 title_name = ifelse(as.numeric(colnames(multi_state)[i])==0,"T = 0",paste0("T = 2^",log(as.numeric(colnames(multi_state)[i]),base = 2))),
+                                 title_name = ifelse(as.numeric(colnames(multi_state)[i])==0,
+                                                     latex2exp::TeX("$T = 0$"),
+                                                     latex2exp::TeX(paste0("$T = 2^{",log(as.numeric(colnames(multi_state)[i]),base = 2),"}$"))),
                                  order_point = TRUE) + 
         scale_color_gradient(name = "Density",
                              low = "lightgrey", high = gene_color,
                              limits = c(0,1),
-                             breaks = seq(0, 1, length.out = 5),
+                             breaks = seq(0, 1, length.out = 2),
                              labels = legend_break_label)
     })
     p
@@ -383,12 +385,12 @@ Visualize_score_pattern <- function(res.lmd, genes = NULL, label_class = NULL, f
   
   if(!is.null(label_class)){
     p = ggplot(data = df, mapping = aes(x = step, y = score, color = label)) + 
-      geom_line(aes(group = gene)) + labs(x = "Time", y = "Scaled Diffusion KL Score")  + 
+      geom_line(aes(group = gene)) + labs(x = "Time", y = "Normalized Diffusion KL Score")  + 
       geom_text(data = df %>% group_by(gene) %>% slice_head(n = 4) %>% slice_tail(n = 1),
[…]

### 2023-12-11 Demo_marrow_facs.Rmd
12/11

@@
 # Preparing input data
+```{r}
+# tiss <- readRDS("/data/ruiqi/local_marker/dataset/tabular_muris/marrow_facs.rds")
+```
+
 ```{r}
@@
 genes_label = rep(c("good","bad"),each = 3)
-Visualize_score_pattern(score_result, genes = genes, label_class = genes_label, facet_class = NULL)
+Visualize_score_pattern(score_result$'score_profile', genes = genes, label_class = genes_label, facet_class = NULL)
 
@@
 ```{r}
+tic()
 lmd_result = LMD(dat, feature_space)
+toc()
 ```

### 2023-12-11 LMD_correction.Rmd
12/11

---
title: "LMD_correction"
output: pdf_document
date: '2023-12-04'
---

score at $t$: $s^{(t)}=D_{KL}(p||p(t))=\sum_i-p_ilog\frac{w_{i}^tp}{p_i}$
pulse score at $t$: $D_{KL}(p||p^{\delta}(t))=\sum_i-p_ilog\frac{w_i^t\delta_i}{p_i}=\sum_i-p_ilog\frac{w_{i,i}^{t}}{p_{i}}$

# Load Data
```{r}
library("Seurat")
library("SeuratWrappers")
library("pheatmap")
library("ggplot2")
library("ggplotify")
source("LMD_function.R", echo = F)

tiss <- readRDS("/data/ruiqi/local_marker/dataset/tabular_muris/marrow_facs.rds")
DefaultAssay(tiss) <- "RNA"
n_dim = dim(tiss@reductions$pca@cell.embeddings)[2]
feature_space = as.matrix(tiss@reductions$pca@cell.embeddings[,1:n_dim])
visual_space = data.frame(tiss@reductions$tsne@cell.embeddings)
dat = as.matrix(tiss[[DefaultAssay(tiss)]]@data)
Gene_detected_count <- apply(dat > apply(dat,2,median),1,sum)
selected_genes = (Gene_detected_count >= 10) & (Gene_detected_count <= ncol(dat) * 0.5)
dat = dat[selected_genes,,drop = FALSE]
```

# Score curve plot
```{r}
W = Symmetric_KNN_graph(knn = 5, feature_space = feature_space)$graph
lmd_result = LMD(dat, feature_space)
lmd_result1 = LMD(dat, feature_space, score_correction = FALSE)

genes = c("Tlr11","Fcnb","Cd79a","4933425H06Rik","Uchl4","Dock11")
genes = c("Vpreb2","Cd200r3","Klra12","Cyp2ab1","Sycp2","Mup3")
genes = c("Abca13","Rag1","Mdga1","Trim47","Cd83","2610002D18Rik")

genes_label = rep(c("good","bad"),each = 3)
facet_label = rep(c(10,500,2000),2)
Visualize_score_pattern(lmd_result$score_profile, genes = genes, label_class = genes_label, facet_class = facet_label)

Visualize_score_pattern(lmd_result$score_profile[,1:16] - lmd_result$score_profile[,17:32], genes = genes, label_class = genes_label, facet_class = facet_label) + ylab("adjusted_score")
```
# plot w_ii with t
```{r}
w = do.call(rbind,lapply(P_ls,function(x){diag(x)}))
rownames(w) = names(P_ls)

pl = lapply(1:nrow(w),function(i){
  legend_break_label = seq(0, max(w[i,]), length.out = 2)
  if(max(legend_break_label)>1e-3){
      legend_break_label = signif(legend_break_label, digits = 2)
    }else{
      legend_break_label = formatC(legend_break_label, format = "e", digits = 0)
    }

  FeaturePlot_custom(value = w[i,]/max(w[i,]),coord = visual_space,title_name = paste0("T = ",rownames(w)[i])) + 
    scale_color_gradient(name = "Density",
                         low = "lightgrey", high = "blue",
                             limits = c(0,1),
                             breaks = seq(0, 1, length.out = 2),
                             labels = legend_break_label)
})
wrap_plots(pl,nrow = 2)
```
# Check diffusion status on graph
```{r}
gene = "Htr1b"
FeaturePlot_diffusion(coord = visual_space,init_state = rho[gene,],P_ls = P_ls,check_time = c(2^4,2^6,2^8),gene_name = gene)
```

# Delta function
```{r}
# Delta-function
lmd_result_delta = LMD(diag(ncol(dat)), feature_space)
Visualize_score_pattern(lmd_result_delta$score_profile, genes = "1", label_class = NULL, facet_class = NULL)

# Double-pulse
indices1 <- sample(ncol(dat), 50, replace = TRUE)
indices2 <- sample(ncol(dat), 50, replace = TRUE)
while(any(indices1 == indices2)) {
  indices2[indices1 == indices2] <- sample(n, sum(indices1 == indices2), replace = TRUE)
}
df = data.frame(t(apply(cbind(indices1,indices2),1,function(x){sort(x)}))) %>% distinct()
lmd_result_delta = LMD(
  t(apply(df,1,function(x){
  vec = numeric(ncol(dat))
  vec[x] = 1
  vec})), feature_space)
which.max(lmd_result_delta$cumulative_score)
Visualize_score_pattern(lmd_result_delta$score_profile, genes = c("35","22"), label_class = NULL, facet_class = NULL)

```

### 2023-12-11 Marker_identfication_benchmark.Rmd
12/11

@@
 ---
+
 # Load Packages
@@
 ```{r}
-tissue_fullname = "Marrow"
+# tiss <- readRDS("/data/ruiqi/local_marker/dataset/tabular_muris/marrow_facs.rds")
+```
+
+```{r}
+tissue_fullname = "Bone marrow"
 
@@
   tiss <- UpdateSeuratObject(tiss)
+}if(tissue_fullname == "Thymic"){
+  options(timeout=6000)
+  if (!file.exists('Human_Thymic_Epithelial_seurat_tiss.Robj')) {
+     # If load url doesn't work, please download it and load from your working directory
+    options(timeout=6000)
+    download.file("https://datasets.cellxgene.cziscience.com/046d2e35-2ee6-4159-8db7-7b1039fe0005.rds",
+                  destfile = 'Human_Thymic_Epithelial_seurat_tiss.Robj', method = 'libcurl')
+  }
+  tiss <- readRDS()
 }
 
+tissue_fullname = "Bone marrow"
 ```
@@
 source("LMD_function.R", echo = F)
-res = LMD(dat,feature_space,max_time = 2^10,knn = 7)
+res = LMD(dat,feature_space,max_time = 2^10,knn = 5)
 marker = show_result_lmd(res, n = length(selected_genes))
@@
 
-method_ls = c("lmd","haystack","hotspot","marcopolo","semi","wilcox_default","wilcox_no_filter","hvg")
+method_ls = c("lmd","lmd_gaussian","haystack","hotspot","marcopolo","semi","wilcox_default","wilcox_no_filter","hvg")
 ```
@@
 colnames(score_df) = paste0("kNN",c(3:10,20,30,40,50))
-write.table(score_df,file = file.path(folder_path, paste0(tissue_name,"_different_knn.csv")),row.names = TRUE)
+write.table(score_df,file = file.path("marrow_result", paste0(tissue_name,"_different_knn.csv")),row.names = TRUE)
 ```
@@
 auc_df = data.frame(auc_df,gt = names(gt_list))
-write.table(auc_df,file = file.path(folder_path, paste0(tissue_name,"_sensitivity_auroc.csv")),row.names = FALSE)
+write.table(auc_df,file = file.path("marrow_result", paste0(tissue_name,"_sensitivity_auroc.csv")),row.names = FALSE)
 ```

### 2023-12-11 paper_figure.Rmd
12/11

@@
 library("Seurat")
+library("RColorBrewer")
 # tissue_name = "marrow_facs"
@@
 ```{r}
+path = "./marrow_result/"
 ## Rank Stability
@@
   geom_line() +
-  geom_point() + labs(x = "# of kNN used", y = sprintf("Jaccard Index\n (with %s)",base),color = "Top Genes") + geom_vline(xintercept = 3, color="red", linetype="dashed") + scale_y_continuous(limits = c(0,1)) +
+  geom_point() + labs(x = "# of kNN used", y = sprintf("Jaccard Index\n (with %s)",base),color = "Top Genes") + geom_vline(xintercept = 3, color="red", linetype="dashed") + 
+  # scale_y_continuous(limits = c(0,1)) +
   theme(
@@
   geom_line() +
-  geom_point() + labs(x = "# of kNN used", y = "AUROC", color = "Gold-Standard Genesets") + scale_y_continuous(limits = c(0,1)) + theme(
+  geom_point() + labs(x = "# of kNN used", y = "AUROC", color = "Gold-Standard Genesets") + 
+  # scale_y_continuous(limits = c(0,1)) + 
+  theme(
     legend.title = element_text(size = 20),

### 2023-12-11 LMD_function.R

@@ -1,7 +1,7 @@
 cran_packages <- c("igraph","ggplot2", "cowplot", "RColorBrewer", 
                    "data.table", "dplyr", "patchwork", 
                    "pheatmap", "ggplotify", "ggraph", 
-                   "ClusterR", "Rcpp", "RcppArmadillo", "tictoc")
+                   "ClusterR", "Rcpp", "RcppArmadillo", "tictoc","RSpectra")
 sapply(cran_packages, function(pkg) if(!requireNamespace(pkg, quietly = TRUE)){install.packages(pkg)})
 lapply(cran_packages, require, character.only = TRUE)
 
@@ -68,13 +68,13 @@ sinkhorn_knopp = function(A, sums = rep(1, nrow(A)),
   while( t <= niter && !converged) {
     r = sums / (A %*% c)
     cnew = sums / (t(A) %*% r)
-    
+
     # Symmetric Sinkhorn--Knopp algorithm could oscillate between two sequences,
     # need to bring the two sequences together (See for example "Algorithms For
     # The Equilibration Of Matrices And Their Application To Limited-memory
     # Quasi-newton Methods")
     if (sym) cnew = (cnew + r)/2
-    
+
     delta = l2_norm(cnew-c)
     if (delta < tol) converged = TRUE
     if (verb) nett::printf("err = %3.5e\n", delta)
@@ -93,65 +93,117 @@ Doubly_stochastic <- function(W){
 
 # Build Cell Graph & Transition Mtx =======
 findDisconnectedComponents <- function(adj_matrix) {
-  
   # Create a graph from the adjacency matrix
   g <- graph_from_adjacency_matrix(adj_matrix, mode = "undirected")
-  
   # Find the connected components
   components <- components(g)
-  
   # Get the number of disconnected components
   num_components <- components$no
-  
   # Get the corresponding nodes for each component
   component_nodes <- split(V(g), components$membership)
-  
   list(number_of_components = num_components, components = component_nodes)
 }
-Symmetric_KNN_graph <- function(knn = 5, feature_space, adjust = FALSE){
+findShortestEdge <- function(component1, component2, data) {
+  # Calculate all pairwise distances between nodes in the two components
[…]

### 2023-12-11 LMD_function_subset.R
simple demo

@@ -1,0 +1,296 @@
+library("Rcpp")
+library("RcppArmadillo")
+library("igraph")
+
+# FUNCTION ====================
+# Mtx operation
+# Enable C++11
+Sys.setenv("PKG_CXXFLAGS"="-std=c++11")
+# Create the C++ function
+## Fast Matrix Multiplication
+cppFunction('arma::mat fastMatMult(arma::mat A, arma::mat B) {
+  arma::mat C = A * B;
+  return C;
+}', depends="RcppArmadillo")
+
+## Fast derivation of KL between pair-wise columns of two matrix (eg. KL between 1st col from mtxA and 1st col from mtxB)
+Rcpp::sourceCpp(code='
+#include <Rcpp.h>
+using namespace Rcpp;
+
+NumericVector calcKL(NumericVector p, NumericVector q) {
+  double sum = 0;
+  for (int j = 0; j < p.size(); ++j) {
+    if(p[j] == 0) continue;
+    double val = p[j] * log(p[j] / q[j]);
+    if (NumericVector::is_na(val) || std::isinf(val)) val = 0.0;
+    sum += val;
+  }
+  return NumericVector::create(sum);
+}
+
+// [[Rcpp::export]]
+NumericVector fastKLMatrix(NumericMatrix x, NumericMatrix y) {
+  int n = x.nrow();
+  NumericVector out(n);
+
+  for (int i = 0; i < n; ++i) {
+    out[i] = calcKL(x.row(i), y.row(i))[0];
+  }
+  return out;
+}
+
+// [[Rcpp::export]]
+NumericVector fastKLVector(NumericMatrix x, NumericVector y) {
+  int n = x.nrow();
+  NumericVector out(n);
+
+  for (int i = 0; i < n; ++i) {
+    out[i] = calcKL(x.row(i), y)[0];
+  }
+  return out;
+}
+')
+Rowwise_normalize <- function(x){
+  return( sweep(x, 1, rowSums(x), FUN = "/") )
+}
+
+# Doubly Stochastic
+l2_norm = function(x) sqrt(sum(x^2))
+sinkhorn_knopp = function(A, sums = rep(1, nrow(A)),
+                          niter = 100, tol = 1e-8, sym = FALSE, verb = FALSE) {
+  # code refer to (https://rdrr.io/github/aaamini/nett/src/R/sinkhorn.R)
+  delta = Inf
+  r = c = rep(1, nrow(A))
+  converged = FALSE
+  t = 1
[…]

### 2024-01-08 Demo_marrow_facs.Rmd
simple demo
24/01/08

@@
 
+```{r}
+# # save to hdf5 file
+# df = sub.srat@meta.data[,c("design_coarse","celltype1205")]
+# # BiocManager::install("rhdf5")
+# library(rhdf5)
+# h5write(dat, 'Bone_marrow_facs_mtx.h5', 'dataset')
+# fwrite(df,file = "HL234_meta.txt",row.names = TRUE)
+```
+
 # Running LMD Step by Step
@@
 # Obtain diffusion KL score
-score_result = fast_calculate_multi_score(W = W, init_state = rho, P_ls = P_ls)
+score_result = fast_get_lmds(W = W, init_state = rho, P_ls = P_ls)
 

### 2024-01-08 Marker_identfication_benchmark.Rmd
24/01/08

@@
   }
-  tiss <- readRDS()
 }
@@
   result = unlist(lapply(top_n, function(top){
-    tiss <- RunPCA(tiss, features = top_gs[1:top], npcs = 20)
-    d = dist(Embeddings(tiss, reduction = "pca"))
-    c = as.integer(factor(tiss$cell_ontology_class[labels(d)]))
-
+    features = top_gs[1:top]
+    tiss <- tiss %>% ScaleData(features = features) %>% RunPCA(features = features, npcs = 20)
     # # Silhouette score
+    # d = dist(Embeddings(tiss, reduction = "pca"))
+    # c = as.integer(factor(tiss$cell_ontology_class[labels(d)]))
     # s1 = mean(as.data.frame(cluster::silhouette(c, d)[,]) %>%
@@
     sdVec <- stats::na.omit(tiss@reductions$pca@stdev)
-    length_scale <- sqrt(sum(sdVec^2))
+    length_scale <- sqrt(sum(sdVec^2)) # length-scale * sqrt(2*(N-1)/N) = d_rms
     s2 = length_scale/mean(nn.dist)
@@
 
+# error bar & null distribution
+null_score_measure = do.call(rbind,lapply(top_n,function(top){
+  set.seed(top)
+  seed_ls = sample(1:1e5,size = 100)
+  random_top_gs = lapply(seed_ls,function(sed){
+    set.seed(sed)
+    sample(rownames(df_benchmark),top,replace = FALSE)
+  })
+  result = unlist(lapply(random_top_gs, function(g_ls){
+    features = g_ls
+    tiss <- tiss %>% ScaleData(features = features) %>% RunPCA(features = features, npcs = 20)
+    # Density-index
+    nn.dist = FNN::get.knn(Embeddings(tiss, reduction = "pca"), k = 10, algorithm = "kd_tree")$nn.dist
+    sdVec <- stats::na.omit(tiss@reductions$pca@stdev)
+    length_scale <- sqrt(sum(sdVec^2))
+    s2 = length_scale/mean(nn.dist)
+    s2
+  }))
+  result = data.frame(TopGene = top,DensityIndex = result)
+}) )
+
+write.table(null_score_measure,file = file.path(folder_path, paste0(tissue_name,"_DI_null.csv")),row.names = FALSE)
 write.table(score_measure,file = file.path(folder_path, paste0(tissue_name,"_density_index.csv")),row.names = FALSE)

### 2024-01-08 paper_figure.Rmd
24/01/08

@@
 rho = Rowwise_normalize(dat[,colnames(W)])
-res = fast_calculate_multi_score(W = W, init_state = rho, P_ls = P_ls)
+res = fast_get_lmds(W = W, init_state = rho, P_ls = P_ls)
 ```
@@
   # geom_text_repel(data = score_measure %>%  filter(Method!="all genes") %>% group_by(Method) %>% slice_head(n = 2) %>% slice_tail(n = 1), aes(label = Method), hjust = 0)
+
+# # Add null distribution
+# null_score_measure = read.csv(paste0(path,tissue_name,"_DI_null.csv"),header = TRUE)
+# fig = fig + geom_boxplot(data = null_score_measure, aes(x = TopGene, y = DensityIndex, group = TopGene), inherit.aes = FALSE)
 

### 2024-01-08 LMD_function.R

@@ -1,7 +1,8 @@
 cran_packages <- c("igraph","ggplot2", "cowplot", "RColorBrewer", 
                    "data.table", "dplyr", "patchwork", 
                    "pheatmap", "ggplotify", "ggraph", 
-                   "ClusterR", "Rcpp", "RcppArmadillo", "tictoc","RSpectra")
+                   "ClusterR", "Rcpp", "RcppArmadillo", 
+                   "tictoc","svMisc","RSpectra")
 sapply(cran_packages, function(pkg) if(!requireNamespace(pkg, quietly = TRUE)){install.packages(pkg)})
 lapply(cran_packages, require, character.only = TRUE)
 
@@ -111,7 +112,7 @@ findShortestEdge <- function(component1, component2, data) {
   return(c(component1[min_distance_idx[1]], component2[min_distance_idx[2]]))
 }
 
-Symmetric_KNN_graph <- function(knn = 5, feature_space, adjust_by_MST = TRUE){
+Symmetric_KNN_graph <- function(knn = 5, feature_space, adjust_by_MST = TRUE, self_loop = TRUE){
   knn_list <- FNN::get.knn(feature_space, k = knn, algorithm = "kd_tree")
   Idx = knn_list$nn.index
   A = Matrix::sparseMatrix(i = rep(1:nrow(Idx),each = knn),
@@ -145,7 +146,7 @@ Symmetric_KNN_graph <- function(knn = 5, feature_space, adjust_by_MST = TRUE){
     }
   }
   
-  # if(adjust){
+  # if(adjust){ # increase knn if containing disconnected components
   #   while(length(filtered_components) >= 2){
   #     knn = knn + 1
   #     knn_list <- FNN::get.knn(feature_space, k = knn, algorithm = "kd_tree")
@@ -164,10 +165,13 @@ Symmetric_KNN_graph <- function(knn = 5, feature_space, adjust_by_MST = TRUE){
   # }
   
   if (length(filtered_components) >= 2) {
-    stop("Disconnected Components, Please increase knn.")
+    stop("Disconnected Components, Please increase knn or adjust by MST.")
   }
 
   A = as.matrix(A)
+  if(self_loop){
+    diag(A) = 1
+  }
   
   # Graph affinity matrix by symmetrize the adjacency mtx
   # W = (A + t(A))/2 # weighted
@@ -175,7 +179,8 @@ Symmetric_KNN_graph <- function(knn = 5, feature_space, adjust_by_MST = TRUE){
 
[…]

### 2024-01-08 LMD_function_subset.R

@@ -1,296 +1,0 @@
-library("Rcpp")
-library("RcppArmadillo")
-library("igraph")
-
-# FUNCTION ====================
-# Mtx operation
-# Enable C++11
-Sys.setenv("PKG_CXXFLAGS"="-std=c++11")
-# Create the C++ function
-## Fast Matrix Multiplication
-cppFunction('arma::mat fastMatMult(arma::mat A, arma::mat B) {
-  arma::mat C = A * B;
-  return C;
-}', depends="RcppArmadillo")
-
-## Fast derivation of KL between pair-wise columns of two matrix (eg. KL between 1st col from mtxA and 1st col from mtxB)
-Rcpp::sourceCpp(code='
-#include <Rcpp.h>
-using namespace Rcpp;
-
-NumericVector calcKL(NumericVector p, NumericVector q) {
-  double sum = 0;
-  for (int j = 0; j < p.size(); ++j) {
-    if(p[j] == 0) continue;
-    double val = p[j] * log(p[j] / q[j]);
-    if (NumericVector::is_na(val) || std::isinf(val)) val = 0.0;
-    sum += val;
-  }
-  return NumericVector::create(sum);
-}
-
-// [[Rcpp::export]]
-NumericVector fastKLMatrix(NumericMatrix x, NumericMatrix y) {
-  int n = x.nrow();
-  NumericVector out(n);
-
-  for (int i = 0; i < n; ++i) {
-    out[i] = calcKL(x.row(i), y.row(i))[0];
-  }
-  return out;
-}
-
-// [[Rcpp::export]]
-NumericVector fastKLVector(NumericMatrix x, NumericVector y) {
-  int n = x.nrow();
-  NumericVector out(n);
-
-  for (int i = 0; i < n; ++i) {
-    out[i] = calcKL(x.row(i), y)[0];
-  }
-  return out;
-}
-')
-Rowwise_normalize <- function(x){
-  return( sweep(x, 1, rowSums(x), FUN = "/") )
-}
-
-# Doubly Stochastic
-l2_norm = function(x) sqrt(sum(x^2))
-sinkhorn_knopp = function(A, sums = rep(1, nrow(A)),
-                          niter = 100, tol = 1e-8, sym = FALSE, verb = FALSE) {
-  # code refer to (https://rdrr.io/github/aaamini/nett/src/R/sinkhorn.R)
-  delta = Inf
-  r = c = rep(1, nrow(A))
-  converged = FALSE
-  t = 1
[…]

### 2024-03-26 Demo_marrow_facs.Rmd
24/03/26

@@
 toc()
+```
+```{r}
+lmd_result = LMD(dat, feature_space, highres = TRUE, max_time = 2^7)
 ```
@@
                         g2m.features = mmus_g2m, set.ident = FALSE)
+tiss <- tiss %>%
+  ScaleData(verbose = FALSE, features = c(mmus_s,mmus_g2m)) %>% RunPCA(npcs = 20, verbose = FALSE, features = c(mmus_s,mmus_g2m)) %>% RunTSNE(dims = 1:20, seed.use = 42, reduction.name = "tsne_cc")
+
 cell_cycle_table = data.frame(tiss@meta.data[,c("S.Score","G2M.Score","Phase")],"Dataset" = "FACS")

### 2024-03-26 Marker_identfication_benchmark.Rmd
24/03/26

--- 2023-12-11 Marker_identfication_benchmark.Rmd
@@
 library("dplyr")
+library(RColorBrewer)
 ```
@@
 # tiss <- readRDS("/data/ruiqi/local_marker/dataset/tabular_muris/marrow_facs.rds")
+# DefaultAssay(tiss) <- "RNA"
 ```
@@
   tiss <- UpdateSeuratObject(tiss)
-}if(tissue_fullname == "Pancreas"){
+}
+if(tissue_fullname == "Pancreas"){
   tissue_name = "pancreas_facs"
@@
   tiss <- UpdateSeuratObject(tiss)
-}if(tissue_fullname == "Lung"){
+}
+if(tissue_fullname == "Lung"){
   tissue_name = "lung_facs"
@@
   tiss <- UpdateSeuratObject(tiss)
-}if(tissue_fullname == "Thymic"){
+}
+if(tissue_fullname == "Thymic"){
   options(timeout=6000)
@@
   }
-  tiss <- readRDS()
+  load('Human_Thymic_Epithelial_seurat_tiss.Robj')
 }
+```
 
-tissue_fullname = "Bone marrow"
-```
 ## Process
@@
 DefaultAssay(tiss) <- "RNA"
-n_dim = dim(tiss@reductions$pca@cell.embeddings)[2]
+n_dim = pmin(dim(tiss@reductions$pca@cell.embeddings)[2],30)
 feature_space = as.matrix(tiss@reductions$pca@cell.embeddings[,1:n_dim])
-visual_space = data.frame(tiss@reductions$tsne@cell.embeddings)
 dat = as.matrix(tiss[[DefaultAssay(tiss)]]@data)
-cell_label = tiss$cell_ontology_class
 
@@
   dir.create(folder_path)
-} else {
 }
@@
 source("LMD_function.R", echo = F)
-res = LMD(dat,feature_space,max_time = 2^10,knn = 5)
+res = LMD(dat,feature_space,max_time = 2^20,knn = 5)
 marker = show_result_lmd(res, n = length(selected_genes))
@@
 method = "wilcox_no_filter"
-marker <- FindAllMarkers(tiss, features = selected_genes, only.pos = FALSE, test.use = "wilcox", logfc.threshold = 0, min.pct = 0)
-gene_rank = marker %>% arrange(.,p_val_adj,desc(avg_log2FC)) %>% distinct(.,gene, .keep_all = TRUE) %>% select(gene)
+# marker <- FindAllMarkers(tiss, features = selected_genes, only.pos = FALSE, test.use = "wilcox", logfc.threshold = 0, min.pct = 0)
+# gene_rank = marker %>% arrange(.,p_val_adj,desc(avg_log2FC)) %>% distinct(.,gene, .keep_all = TRUE) %>% select(gene)
+marker <- presto:::wilcoxauc.Seurat(X = subset(tiss,features = selected_genes),
+                                    assay = "data", 
+                                    seurat_assay = "RNA")
+colnames(marker)[1] = "gene"
+gene_rank = marker %>% arrange(.,padj,desc(logFC)) %>% distinct(.,gene, .keep_all = TRUE) %>% select(gene)
 gene_rank$'rank' = 1:nrow(gene_rank)
@@
 ## Marcopolo
+Long Runtime Option
+```{r}
+# transfer the seurat obj to h5ad & process in jupyter notebook
+library(reticulate)
+library(anndata)
+sc <- import("scanpy", convert = FALSE)
+scvi <- import("scvi", convert = FALSE)
+# scvi$settings$progress_bar_style = 'tqdm'
+tiss_adata <- sc$AnnData(
+  X = t(raw_count)
+)
+anndata::write_h5ad(tiss_adata, "./dataset/tm_marrow.h5ad")
+```
 The rank of genes is given by `Marcopolo_rank`
@@
 result = MarcoPolo.find_markers(adata=adata, regression_result=regression_result)
+result = result.set_axis(list(adata.var.index) ,axis = 0)
 ```
@@
 method_ls = c("lmd","haystack","hotspot","marcopolo","semi","wilcox_default","wilcox_no_filter","hvg")
+df_benchmark = lapply(method_ls, function(method){
+  marker = read.table(file.path(folder_path, paste0(method,"_",tissue_name,".csv")),header = TRUE)
+  if(!"gene" %in% colnames(marker)){
+    marker$'gene' = rownames(marker)
+  }
+  marker %>% select(gene,rank) %>% distinct()
+})
+names(df_benchmark) = method_ls
+df_benchmark = bind_rows(df_benchmark, .id = "method")
+df_benchmark <- tidyr::pivot_wider(df_benchmark, names_from = method, values_from = rank)
+df_benchmark[is.na(df_benchmark)] = nrow(df_benchmark)
 
-df_benchmark = do.call(cbind,lapply(method_ls, function(method){
-  marker = read.table(file.path(folder_path, paste0(method,"_",tissue_name,".csv")),header = TRUE)
-  if("gene" %in% colnames(marker)){
-    return(marker$rank[match(selected_genes,marker$gene)])
-  }else{
-    return(marker$rank[match(selected_genes,rownames(marker))])
-  }
-}))
-df_benchmark[is.na(df_benchmark)] = nrow(df_benchmark)
-df_benchmark = data.frame(df_benchmark,row.names = selected_genes)
-colnames(df_benchmark) = method_ls
+# rownames(df_benchmark) = df_benchmark$gene
+# df_benchmark[order(Gene_detected_count[rownames(df_benchmark)]),"gene_count"] = 1:nrow(df_benchmark)
 
-write.table(df_benchmark,file = file.path(folder_path, paste0(tissue_name,"_benchmark_rank_table.csv")),row.names = TRUE)
+write.table(df_benchmark,file = file.path(folder_path, paste0(tissue_name,"_benchmark_rank_table.csv")),row.names = FALSE)
 ```
@@
 
-method_ls = c("lmd","lmd_gaussian","haystack","hotspot","marcopolo","semi","wilcox_default","wilcox_no_filter","hvg")
+method_ls = c("lmd","haystack","hotspot","marcopolo","semi","wilcox_default","wilcox_no_filter","hvg")
 ```
@@
 ```{r}
+folder_path = "benchmark_result"
 df_benchmark = read.table(file.path(folder_path, paste0(tissue_name,"_benchmark_rank_table.csv")))
-
-score_measure =do.call(rbind,lapply(c(method_ls,"All genes"),function(method){
+score_measure =do.call(rbind,lapply(c(colnames(df_benchmark),"All genes"),function(method){
   if(method == "All genes"){
@@
     top_gs = rownames(df_benchmark)[order(df_benchmark[,method])]
-    top_n = c(seq(50,90,10),seq(100,900,100),seq(1000,3000,500),seq(4000,9000,1000),seq(10000,max(df_benchmark),2000))
+    # top_n = c(seq(50,90,10),seq(100,900,100),seq(1000,3000,500),seq(4000,9000,1000),seq(10000,max(df_benchmark),2000))
+    top_n = c(seq(50,90,10),seq(100,1000,100))
   }
-  result = unlist(lapply(top_n, function(top){
-    tiss <- RunPCA(tiss, features = top_gs[1:top], npcs = 20)
-    d = dist(Embeddings(tiss, reduction = "pca"))
-    c = as.integer(factor(tiss$cell_ontology_class[labels(d)]))
-
+  DI = unlist(lapply(top_n, function(top){
+    features = top_gs[1:top]
+    # remove drop-out cells
+    subtiss = subset(tiss, cells = which(colSums(subset(tiss,features = features))>0))
+    subtiss <- subtiss %>% ScaleData(features = features) %>% RunPCA(features = features, npcs = 20)
     # # Silhouette score
+    # d = dist(Embeddings(subtiss, reduction = "pca"))
+    # c = as.integer(factor(subtiss$cell_ontology_class[labels(d)]))
     # s1 = mean(as.data.frame(cluster::silhouette(c, d)[,]) %>%
@@
     # Density-index
-    nn.dist = FNN::get.knn(Embeddings(tiss, reduction = "pca"), k = 10, algorithm = "kd_tree")$nn.dist
-    sdVec <- stats::na.omit(tiss@reductions$pca@stdev)
-    length_scale <- sqrt(sum(sdVec^2))
+    nn.dist = FNN::get.knn(Embeddings(subtiss, reduction = "pca"), k = 10, algorithm = "kd_tree")$nn.dist
+    sdVec <- stats::na.omit(subtiss@reductions$pca@stdev)
+    length_scale <- sqrt(sum(sdVec^2)) # length-scale * sqrt(2*(N-1)/N) = d_rms
     s2 = length_scale/mean(nn.dist)
@@
   # rownames(result) = c("Mean_SI","DI")
-  result = data.frame(TopGene = top_n,Method = method,DensityIndex = result)
+
+  # Non-dropout Rate
+  cell_count = lapply(top_n,function(n){
+    colSums(dat[top_gs[1:n],]>0)
+  })
+  non_dropout_rate = unlist(lapply(cell_count,function(x){
+    sum(x != 0)/length(x)
+  }))
+
+  result = data.frame(TopGene = top_n,Method = method,DensityIndex = DI, non_dropout_rate = non_dropout_rate)
+  result
 }) )
 
-write.table(score_measure,file = file.path(folder_path, paste0(tissue_name,"_density_index.csv")),row.names = FALSE)
+write.table(score_measure,file = file.path(folder_path, paste0(tissue_name,"_density_index_no_dropout.csv")),row.names = FALSE)
+
+# error bar & null distribution
+top_n = c(seq(50,90,10),seq(100,900,100),seq(1000,3000,500),seq(4000,9000,1000),seq(10000,max(df_benchmark),2000))
+null_score_measure = do.call(rbind,lapply(top_n,function(top){
+  set.seed(top)
+  seed_ls = sample(1:1e5,size = 100)
+  random_top_gs = lapply(seed_ls,function(sed){
+    set.seed(sed)
+    sample(rownames(df_benchmark),top,replace = FALSE)
+  })
+  result = unlist(lapply(random_top_gs, function(g_ls){
+    features = g_ls
+    subtiss = subset(tiss, cells = which(colSums(subset(tiss,features = features))>0))
+    subtiss <- subtiss %>% ScaleData(features = features) %>% RunPCA(features = features, npcs = 20)
+    # Density-index
+    nn.dist = FNN::get.knn(Embeddings(subtiss, reduction = "pca"), k = 10, algorithm = "kd_tree")$nn.dist
+    sdVec <- stats::na.omit(subtiss@reductions$pca@stdev)
+    length_scale <- sqrt(sum(sdVec^2))
+    s2 = length_scale/mean(nn.dist)
+    s2
+  }))
+  result = data.frame(TopGene = top,DensityIndex = result)
+}) )
+write.table(null_score_measure,file = file.path(folder_path, paste0(tissue_name,"_DI_null.csv")),row.names = FALSE)
+
+# smallest set of genes covering 95% cells & corresponding DI
+min_coverage = 1
+coverage_score_measure = do.call(rbind,lapply(colnames(df_benchmark),function(method){
+  cat(method,"\n")
+  top_gs = rownames(df_benchmark)[order(df_benchmark[,method])]
+  min_covered = apply(dat[top_gs,] > 0, 2, function(x) which(x)[min_coverage])
+  min_row_for_coverage <- as.integer(names(which(cumsum(table(min_covered)) >= ncol(dat) * 0.95)[1]))
+  features = top_gs[1:min_row_for_coverage]
+
+  # remove drop-out cells
+  subtiss = subset(tiss, cells = which(colSums(subset(tiss,features = features))>0))
+  npcs = pmin(20,length(features))
+  subtiss <- subtiss %>% ScaleData(features = features) %>% RunPCA(features = features, npcs = npcs, verbose = FALSE)
+  subtiss <- RunUMAP(subtiss,dims = 1:dim(subtiss@reductions$pca)[2])
+  # Density-index
+  if(length(features) < npcs){
+    nn.dist = FNN::get.knn(t(subtiss[["RNA"]]@scale.data), k = 10, algorithm = "kd_tree")$nn.dist
+    length_scale <- norm(t(subtiss[["RNA"]]@scale.data),type = "F")/sqrt(ncol(subtiss)-1)
+    s2 = length_scale/mean(nn.dist)
+  }else{
+    nn.dist = FNN::get.knn(Embeddings(subtiss, reduction = "pca"), k = 10, algorithm = "kd_tree")$nn.dist
+    sdVec <- stats::na.omit(subtiss@reductions$pca@stdev)
+    length_scale <- sqrt(sum(sdVec^2)) # length-scale * sqrt(2*(N-1)/N) = d_rms
+    s2 = length_scale/mean(nn.dist)
+  }
+
+  data.frame(TopGene = min_row_for_coverage,Method = method,DensityIndex = s2, non_dropout_rate = sum(colSums(dat[features,]>0)!=0)/ncol(dat))
+
+  DimPlot(subtiss,group.by = "cell_ontology_class") + NoLegend() + ggtitle(sprintf("%s, %d genes, DI: %.4f",method,min_row_for_coverage,s2))
+}))
+
+```
+
+# NNret + Silhouette + DI
+```{r}
+folder_path = "benchmark_result"
+df_benchmark = read.table(file.path(folder_path, paste0(tissue_name,"_benchmark_rank_table.csv")))
+score_measure =do.call(rbind,lapply(c(colnames(df_benchmark),"All genes"),function(method){
+  if(method == "All genes"){
+    top_gs = rownames(df_benchmark)
+    top_n = length(top_gs)
+  }else{
+    top_gs = rownames(df_benchmark)[order(df_benchmark[,method])]
+    top_n = c(seq(50,90,10),seq(100,1000,100))
+  }
+  scores = sapply(top_n, function(top){
+    features = top_gs[1:top]
+    # remove drop-out cells
+    subtiss = subset(tiss, cells = which(colSums(subset(tiss,features = features))>0))
+    subtiss <- subtiss %>% ScaleData(features = features) %>% RunPCA(features = features, npcs = 20)
+
+    # Silhouette score
+    d = dist(Embeddings(subtiss, reduction = "pca"))
+    c = as.integer(factor(subtiss$celltype[labels(d)]))
+    si = mean(as.data.frame(cluster::silhouette(c, d)[,]) %>%
+      group_by(cluster) %>%
+      summarize(mean_silhouette = mean(sil_width)) %>%.$mean_silhouette)
+
+    # NNret
+    nn.result = FNN::get.knn(Embeddings(subtiss, reduction = "pca"), k = 10, algorithm = "kd_tree")
+    nn.dist = nn.result$nn.dist
+    nn.index = nn.result$nn.index
+    nnret = mean(
+      unlist(lapply(1:nrow(nn.index),function(i){
+      mean(c[nn.index[i,]] == c[i])
+    })) )
+
+    # Density-index
+    sdVec <- stats::na.omit(subtiss@reductions$pca@stdev)
+    length_scale <- sqrt(sum(sdVec^2)) # length-scale * sqrt(2*(N-1)/N) = d_rms
+    di = length_scale/mean(nn.dist)
+
+    c(si,di,nnret)
+  }) 
+  scores = t(scores)
+  colnames(scores) = c("Sihouette_score","DensityIndex","NNret")
+  rownames(scores) = top_n
+
+  # Non-dropout Rate
+  cell_count = lapply(top_n,function(n){
+    colSums(tiss[["RNA"]]@data[top_gs[1:n],]>0)
+  })
+  non_dropout_rate = unlist(lapply(cell_count,function(x){
+    sum(x != 0)/length(x)
+  }))
+
+  result = data.frame(TopGene = top_n,Method = method, scores, non_dropout_rate = non_dropout_rate)
+  result
+}) )
+```
+
+# Recovery of celltypes
+```{r}
+# jaccard index of cell-type-specificity of each genes
+cell_label = tiss$celltype
+one_hot_matrix_ct <- sapply(unique(cell_label), function(x) as.integer(cell_label == x))
+binary_dat = dat > apply(dat,2,median)
+intersect_mtx = binary_dat %*% one_hot_matrix_ct
+union_mtx = sapply(colnames(one_hot_matrix_ct),function(ct){
+  rowSums((binary_dat > 0)) + colSums(one_hot_matrix_ct)[ct]
+}) - intersect_mtx
+jaccard_sim = intersect_mtx / union_mtx
+threshold = 0.8 / ((ncol(tiss)/table(cell_label))*0.2 + 0.8)[colnames(jaccard_sim)]
+recovery_mtx = t(apply(jaccard_sim,1,function(x) x > threshold))
+
+# Prop of recovered cell types
+recovery_rate =data.frame(sapply(c(colnames(df_benchmark)),function(method){
+  top_gs = rownames(df_benchmark)[order(df_benchmark[,method])]
+  top_n = 1:300
+  scores = sapply(top_n, function(top){
+    features = top_gs[1:top]
+    mean(colSums(recovery_mtx[features,,drop = FALSE]) > 0)
+  }) 
+}) )
+recovery_rate$'TopGene' = 1:nrow(recovery_rate)
 ```

### 2024-03-26 paper_figure.Rmd
24/03/26

@@
   gene = genes[i]
-  FeaturePlot_diffusion(coord = visual_space,init_state = rho[gene,],P_ls = P_ls,W = W,check_time = check_time,gene_name = NULL, gene_color = scales::hue_pal()(2)[i])
+  FeaturePlot_diffusion(coord = visual_space,init_state = rho[gene,],P_ls = P_ls,W = W,check_time = check_time,gene_name = NULL, gene_color = scales::hue_pal()(length(genes))[i])
 })
@@
 # Set the color code for all methods
-method_ls = c("lmd","haystack","hotspot","marcopolo","semi","wilcox_default","wilcox_no_filter","hvg")
-method_fullname = setNames(c("LMD","singleCellHaystack","Hotspot","Marcopolo","SEMITONES","Seurat v4 default", "Seurat v4 Wilcoxon unfiltered","HVGs"),method_ls)
+method_ls = c("lmd","haystack","hotspot","marcopolo","semi","wilcox_default","wilcox_no_filter","hvg","gene_count")
+method_fullname = setNames(c("LMD","singleCellHaystack","Hotspot","Marcopolo","SEMITONES","Seurat v4 default", "Seurat v4 Wilcoxon unfiltered","HVGs","Gene count"),method_ls)
 coldef = setNames(
@@
 ```{r}
-path = "./benchmark_result/"
+folder_path = "./benchmark_result/"
 tissue_ls = c("marrow_facs","pancreas_facs","lung_facs")
-df_auc = do.call(rbind,lapply(tissue_ls,function(tissue_name){df = read.table(paste0(path,tissue_name,"_auroc.csv"),header = TRUE); df$'Tissue' = tissue_name; df$'Method' = method_fullname[df$'Method']; df}))
-
-df_auc = df_auc %>% filter(gt_set %in% c("Top100","CellMarkerDB")) %>% filter(!Method %in% "HighlyVariableGenes")
+df_auc = do.call(rbind,lapply(tissue_ls,function(tissue_name){df = read.table(file.path(folder_path,paste0(tissue_name,"_auroc.csv")),header = TRUE); df$'Tissue' = tissue_name;df}))
+df_auc$Method = method_fullname[df_auc$Method]
+df_auc = df_auc %>% filter(gt_set %in% c("Top100","CellMarkerDB")) %>% filter(!Method %in% "HVGs")
 df_auc$Method <- with(df_auc, reorder(Method, -AUROC, FUN = median))
@@
 tissue_name = "marrow_facs"
-score_measure = read.table(paste0(path,tissue_name,"_density_index.csv"),header = TRUE)
+score_measure = read.table(paste0(path,tissue_name,"_density_index_no_dropout.csv"),header = TRUE)
 score_measure = score_measure %>% filter(Method!="wilcox_no_filter")
@@
 score_measure[score_measure$Method == "All genes","line_type"] = "dashed"
-
 score_measure[grepl("Seurat",score_measure$Method),"Method"] = "Seurat v4"
@@
 
-# # Add null distribution
-# null_score_measure = read.csv(paste0(path,tissue_name,"_DI_null.csv"),header = TRUE)
-# fig = fig + geom_boxplot(data = null_score_measure, aes(x = TopGene, y = DensityIndex, group = TopGene), inherit.aes = FALSE)
+# Add null distribution
+null_score_measure = read.csv(paste0(path,tissue_name,"_DI_null.csv"),header = TRUE)
+fig = fig + geom_boxplot(data = null_score_measure, aes(x = TopGene, y = DensityIndex, group = TopGene), inherit.aes = FALSE)
 
 ggsave(filename = "fig-densityindex.png", plot = fig, width = 5, height = 5)
+```
+## Scatterplot
+```{r}
+score_measure = read.table(paste0(path,tissue_name,"_density_index_no_dropout.csv"),header = TRUE)
+score_measure = score_measure %>% filter(Method!="wilcox_no_filter")
+score_measure$Method = ifelse(score_measure$Method %in% names(method_fullname),method_fullname[score_measure$Method],score_measure$Method)
+score_measure$'color' = coldef[score_measure$Method]
+
+ggplot(score_measure %>% filter(Method!="All genes") %>% filter(TopGene <= 1000), aes(x = non_dropout_rate, y = Mean_SI, color = Method)) + 
+  geom_point() +
+  scale_color_manual(values = score_measure %>% filter(Method!="All genes") %>% select(c("Method","color")) %>% distinct() %>% tibble::deframe()) + facet_wrap(~TopGene)
+```
+## Prop of recovered cts
+```{r}
+recovery_rate <- reshape2::melt(recovery_rate,id.vars = "TopGene",value.name = "prop",variable.name = "Method")
+recovery_rate = recovery_rate %>% filter(Method!="wilcox_no_filter")
+recovery_rate$Method = ifelse(recovery_rate$Method %in% names(method_fullname),method_fullname[recovery_rate$Method],recovery_rate$Method)
+recovery_rate$'color' = coldef[recovery_rate$Method]
+
+ggplot(recovery_rate, aes(x = TopGene, y = prop, color = Method, group = Method)) + 
+  geom_line() +
+  scale_color_manual(values = score_measure %>% filter(Method!="All genes") %>% select(c("Method","color")) %>% distinct() %>% tibble::deframe()) + ylab("Prop. of recovered cell types")
 ```

### 2024-03-26 ATAC_Cite_spatial.R

@@ -1,0 +1,170 @@
+library(Seurat)
+library(SeuratData)
+library(SeuratWrappers)
+library(cowplot)
+library(dplyr)
+source("/data/ruiqi/local_marker/LocalMarkerDetector/LMD_function.R", echo = F)
+
+# CITE-seq: 25 proteins ----------
+InstallData("bmcite")
+bm <- LoadData(ds = "bmcite")
+DefaultAssay(bm) <- 'ADT'
+VariableFeatures(bm) <- rownames(bm[["ADT"]])
+bm <- NormalizeData(bm, normalization.method = 'CLR', margin = 2) %>% 
+  ScaleData() %>% RunPCA(reduction.name = 'apca')
+bm <- RunUMAP(bm, reduction = 'apca', dims = 1:18, assay = 'ADT', 
+              reduction.name = 'adt.umap', reduction.key = 'adtUMAP_')
+DimPlot(bm, reduction = 'adt.umap', group.by = 'celltype.l2', label = TRUE, 
+        repel = TRUE, label.size = 2.5) + NoLegend()
+FeaturePlot(bm, features = c("adt_CD45RA","adt_CD16","adt_CD161"),
+            reduction = 'adt.umap', max.cutoff = 2, ncol = 3)
+
+# ATAC-seq ----------------
+library(SeuratData)
+library(Seurat)
+library(Signac)
+library(EnsDb.Hsapiens.v86)
+library(ggplot2)
+library(cowplot)
+# BiocManager::install("Bioconductor/GenomeInfoDb")
+
+## Load Data & QC -------
+# InstallData("pbmcMultiome")
+pbmc.atac <- LoadData("pbmcMultiome", "pbmc.atac")
+# QC
+pbmc.atac <- subset(pbmc.atac, seurat_annotations != "filtered" & nCount_ATAC < 7e4 & nCount_ATAC > 5e3)
+# Extract_t_cells
+annotated_cell_types = names(table(pbmc.atac$seurat_annotations))
+sub.pbmc.atac <- subset(pbmc.atac, seurat_annotations %in% annotated_cell_types[!grepl("B|HSPC|Mono|Plasma|DC|NK",annotated_cell_types)])
+
+## Process -------
+tiss <- sub.pbmc.atac
+annotations <- GetGRangesFromEnsDb(ensdb = EnsDb.Hsapiens.v86)
+seqlevelsStyle(annotations) <- "UCSC"
[…]

### 2024-03-26 LMD_function.R

@@ -2,9 +2,11 @@ cran_packages <- c("igraph","ggplot2", "cowplot", "RColorBrewer",
                    "data.table", "dplyr", "patchwork", 
                    "pheatmap", "ggplotify", "ggraph", 
                    "ClusterR", "Rcpp", "RcppArmadillo", 
-                   "tictoc","svMisc","RSpectra")
+                   "tictoc","svMisc","RSpectra","Matrix","progress","latex2exp","parallel")
+bioc_packages <- c("Seurat")
 sapply(cran_packages, function(pkg) if(!requireNamespace(pkg, quietly = TRUE)){install.packages(pkg)})
-lapply(cran_packages, require, character.only = TRUE)
+sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
+lapply(c(cran_packages,bioc_packages), require, character.only = TRUE)
 
 # Mtx operation===========
 # Enable C++11
@@ -19,7 +21,6 @@ cppFunction('arma::mat fastMatMult(arma::mat A, arma::mat B) {
 Rcpp::sourceCpp(code='
 #include <Rcpp.h>
 using namespace Rcpp;
-
 NumericVector calcKL(NumericVector p, NumericVector q) {
   double sum = 0;
   for (int j = 0; j < p.size(); ++j) {
@@ -54,7 +55,11 @@ NumericVector fastKLVector(NumericMatrix x, NumericVector y) {
 }
 ')
 Rowwise_normalize <- function(x){
+  x = x[rowSums(x)!=0,,drop = FALSE]
   return( sweep(x, 1, rowSums(x), FUN = "/") )
+}
+is_sparse_matrix <- function(m) {
+  any(grepl("Matrix", class(m)) & !grepl("dense", class(m)))
 }
 
 # Doubly Stochastic
@@ -84,20 +89,39 @@ sinkhorn_knopp = function(A, sums = rep(1, nrow(A)),
   }
   list(r = as.numeric(r), c = as.numeric(c))
 }
+sinkhorn_knopp_largeData = function(A, niter = 100, tol = 1e-8, verb = FALSE) {
+  # code refer to (https://rdrr.io/github/aaamini/nett/src/R/sinkhorn.R)
+  delta = Inf
+  for (irep in 1:niter) {
+    scale_fac <- 1 / sqrt(rowSums(A))
[…]

### 2024-03-26 Marker_identfication_benchmark_azimuth_data.R

@@ -1,0 +1,289 @@
+library("Seurat")
+library("dplyr")
+library(RColorBrewer)
+library(reticulate)
+library(anndata)
+
+setwd("/data/ruiqi/local_marker/dataset/azimuth")
+# Load Data ----------
+tissue_download_ls = c(
+  "human_motor_cortex" = "https://seurat.nygenome.org/azimuth/demo_datasets/allen_m1c_2019_ssv4.rds",
+  "human_kidney" = "https://seurat.nygenome.org/azimuth/demo_datasets/kidney_demo_stewart.rds",
+  "human_bone_marrow" = "https://seurat.nygenome.org/azimuth/demo_datasets/bmcite_demo.rds",
+  "human_pancreas" = "https://seurat.nygenome.org/azimuth/demo_datasets/enge.rds",
+  "mouse_motor_cortex" = "https://seurat.nygenome.org/azimuth/demo_datasets/allen_mop_2020.rds",
+  "human_lung_hlca" = "https://seurat.nygenome.org/hlca_ref_files/ts_opt.rds"
+)
+for(tissue_name in names(tissue_download_ls)){
+  cat(tissue_name,"\n")
+folder_path = "/data/ruiqi/local_marker/dataset/azimuth"
+tissue_name = paste0(tissue_name,"_azi_demo")
+
+file_name = paste0(tissue_name,".rds")
+options(timeout=6000)
+if (!file.exists(file.path(folder_path,file_name))) {
+  options(timeout=6000)
+  download.file(tissue_download_ls[tissue_name],
+                destfile = file.path(folder_path,file_name), method = 'libcurl')
+}
+tiss <- readRDS(file.path(folder_path,file_name))
+
+# Preprocess ----------------
+# Annotate based on Azimuth (download result from shiny)
+folder_path = "/data/ruiqi/local_marker/dataset/azimuth/azimuth_annotate_result"
+predictions <- read.delim(file.path(folder_path,paste0(tissue_name,"_azimuth_pred.tsv")), row.names = 1)
+tiss <- AddMetaData(
+  object = tiss,
+  metadata = predictions)
+projected.umap <- readRDS(file.path(folder_path,paste0(tissue_name,"_azimuth_umap.Rds")))
+tiss <- tiss[, Cells(projected.umap)]
+tiss[['umap.proj']] <- projected.umap
+
+annotation_group = colnames(predictions)[1]
+meta.data = tiss@meta.data[,annotation_group]
[…]

### 2024-05-23 Demo_marrow_facs.Rmd
240523 pipeline-development

--- 2024-01-08 Demo_marrow_facs.Rmd
@@
 ```{r}
-# Install necessary tools if not already present
-if (!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")
-if (!requireNamespace("remotes", quietly = TRUE)) install.packages("remotes")
+source("LMD_function.R")
 # Define package lists
-bioc_packages <- c("Seurat",
-                   "clusterProfiler", "AnnotationDbi", "ReactomePA", "org.Mm.eg.db","gprofiler2")
-github_packages <- c("satijalab/seurat-wrappers")
-# Note: repeating packages from LMD_function.R as knitr does not install them properly
-cran_packages <- c("igraph","ggplot2", "cowplot", "RColorBrewer",
-                   "data.table", "dplyr", "patchwork",
-                   "pheatmap", "ggplotify", "ggraph",
-                   "ClusterR", "Rcpp", "RcppArmadillo", "tictoc")
+bioc_packages <- c("clusterProfiler", "AnnotationDbi", "ReactomePA", "org.Mm.eg.db","gprofiler2")
+github_packages <- c("satijalab/seurat-wrappers");
+names(github_packages) = strsplit(github_packages, "/")[[1]][2]
+names(github_packages)[grepl("seurat-wrappers",names(github_packages))] = "SeuratWrappers"
+# Install packages
+sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
+sapply(1:length(github_packages), function(i) if (!requireNamespace(names(github_packages)[i], quietly = TRUE)) remotes::install_github(github_packages[i]))
+lapply(c(bioc_packages,names(github_packages)), require, character.only = TRUE)
+```
 
-# Install packages
-sapply(cran_packages, function(pkg) if(!requireNamespace(pkg, quietly = TRUE)){install.packages(pkg)})
-sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
-sapply(github_packages, function(pkg) if (!requireNamespace(strsplit(pkg, "/")[[1]][2], quietly = TRUE)) remotes::install_github(pkg))
-library("Seurat")
-library("SeuratWrappers")
-library("pheatmap")
-library("ggplot2")
-library("ggplotify")
-```
 # Create a folder to save results
@@
 ```{r}
-# tiss <- readRDS("/data/ruiqi/local_marker/dataset/tabular_muris/marrow_facs.rds")
+# tiss <- readRDS("/banach1/ruiqi/local_marker/LMD_data/tabular_muris/marrow_facs.rds")
 ```
@@
 toc()
+# lmd_result = LMD(dat, feature_space, highres = TRUE, max_time = 2^7)
 ```
@@
 
-# Hierarchical Tree
-clustering_method = "average"
-gene_hree = hclust(dist, method = clustering_method)
-
 # Obtain Gene modules
-gene_partition = dynamicTreeCut::cutreeDynamic(
-  dendro = gene_hree, 
-  distM = as.matrix(dist), deepSplit = 2,
-  pamStage = TRUE, pamRespectsDendro = TRUE,
-  minClusterSize = 5)
-names(gene_partition) = labels(dist)
-
-# Re-name gene modules based on the order of htree
-gene_partition = as.factor(gene_partition)
-module_order = order(sapply(levels(gene_partition), function(mod) {
-  median(which(gene_partition[gene_hree$order] == mod))
-}))
-levels(gene_partition)[module_order] = seq(nlevels(gene_partition))
-gene_partition = factor(gene_partition,levels = seq(nlevels(gene_partition)))
-rm(module_order)
+gene_partition = Obtain_gene_partition(dist, clustering_method = "average", min_gene = 5)
 ```
@@
 ```{r}
-p = pheatmap(as.matrix(dist),
-              annotation_col = data.frame(Module = gene_partition), 
-              cluster_rows = gene_hree, 
-              cluster_cols = gene_hree,
-              ann_colors = list(Module = setNames(colorRampPalette(brewer.pal(12, "Paired"))(nlevels(gene_partition)),levels(gene_partition))),
-              treeheight_row = 0,
-              col=colorRampPalette(c("firebrick3", "white"))(99),
-              show_colnames = FALSE,
-              show_rownames = FALSE,
-              main = paste(distance_method,
-                           clustering_method,sep = "-"),
-              silent = TRUE )
-as.ggplot(p)
+p = Visualize_gene_heatmap(dist)
 ```
@@
 ```{r}
-knn = 10; major_vote = 5
-A = Symmetric_KNN_graph(knn = knn, feature_space = feature_space)$'adj_matrix'
-cell_block_loop = lapply(1:100,function(loop){
-  GMM_subsampling(seed = loop, gene_partition, expr_dat = dat, cell_kNN_graph = A, major_vote)
-})
-cell_block_prop = Reduce(`+`, cell_block_loop) / length(cell_block_loop)
-cell_block = cell_block_prop
+tiss = AddModuleActivityScore(tiss, gene_partition = gene_partition, do_local_smooth = TRUE)
 ```
-
-## Step3: Identify Celltype-specific activated Modules
-```{r}
-cell_label = as.factor(tiss$cell_ontology_class)
-abbrev = levels(cell_label)
-abbrev[grepl("T cell",abbrev)] = "T cell"
-abbrev[grepl("natural killer",abbrev)] = "NK cell"
-abbrev[abbrev == "B cell"] = "Cd3e+ Klrb1+ B cell"
-abbrev[grepl("Slamf1-positive",abbrev)] = "Slamf1+ MPC"
-abbrev[grepl("Slamf1-negative",abbrev)] = "Slamf1- MPC"
-abbrev[grepl("common lymphoid",abbrev)] = "CLP"
-abbrev[grepl("megakaryocyte",abbrev)] = "MEP"
-abbrev[grepl("hematopoietic",abbrev)] = "HSC"
-abbrev[grepl("hematopoietic",abbrev)] = "HSC"
-abbrev[grepl("granulocyte monocyte",abbrev)] = "GMP"
-names(abbrev) = levels(cell_label)
-levels(cell_label) = abbrev[levels(cell_label)]
-
-# Create one-hot vector of Cell Type
-one_hot_matrix <- sapply(unique(cell_label), function(x) as.integer(cell_label == x))
-colnames(one_hot_matrix) = unique(cell_label)
-rownames(one_hot_matrix) = names(cell_label)
-
-# Calculate Jaccard index between module activity & cell type
-jaccard_index <- apply(cell_block, 2, function(module_col) {
-  apply(one_hot_matrix, 2, function(one_hot_col) {
-    1 - philentropy::distance(rbind(module_col,one_hot_col), method = "jaccard", mute.message=TRUE)
-  })
-})
-```
-
-### Annotate Each Celltype-specific activated Modules with its corresponding Celltypes
-```{r}
-module_celltype = apply(jaccard_index,2,function(x){paste(names(which(x > 0.4)),collapse = ",")})
-```
-
-## Transferability of Modules
-### Load Data & Calculate Module Activity
-```{r}
-if (!file.exists('droplet_Marrow_seurat_tiss.Robj')) {
-   # If load url doesn't work, please download it and load from your working directory
-  options(timeout=6000)
-  download.file("https://figshare.com/ndownloader/files/13092380",
-                destfile = 'droplet_Marrow_seurat_tiss.Robj', method = 'libcurl')
- }
-load('droplet_Marrow_seurat_tiss.Robj')
-tiss_droplet <- UpdateSeuratObject(tiss)
-saveRDS(tiss_droplet, file = "processed_droplet_Marrow_srat.rds")
-
-feature_space_droplet = as.matrix(tiss_droplet@reductions$pca@cell.embeddings[,1:n_dim])
-dat_droplet = as.matrix(tiss_droplet[["RNA"]]@data)
-
-knn = 10; major_vote = 5
-A = Symmetric_KNN_graph(knn = knn, feature_space = feature_space_droplet)$'adj_matrix'
-cell_block_loop = lapply(1:100,function(loop){
-  GMM_subsampling(seed = loop, gene_partition, expr_dat = dat_droplet, cell_kNN_graph = A, major_vote)
-})
-cell_block_prop = Reduce(`+`, cell_block_loop) / length(cell_block_loop)
-cell_block_droplet = cell_block_prop
-```
-
-### Calculate Jaccard Index
-```{r}
-cell_label = tiss_droplet$cell_ontology_class
-cell_label[!cell_label %in% names(abbrev)] = NA
-cell_label = setNames(abbrev[cell_label],names(cell_label))
-
-one_hot_matrix_droplet <- sapply(unique(na.omit(cell_label)), function(x) as.integer(cell_label == x))
-one_hot_matrix_droplet[is.na(one_hot_matrix_droplet)] = 0
-
-jaccard_index_droplet <- apply(cell_block_droplet, 2, function(module_col) {
-  apply(one_hot_matrix_droplet, 2, function(one_hot_col) {
-    1 - philentropy::distance(rbind(module_col,one_hot_col), method = "jaccard", mute.message=TRUE)
-  })
-})
-```
-## Save Intermediate data
-```{r}
-save(gene_partition,dist,gene_hree,cell_block,cell_block_droplet,one_hot_matrix,one_hot_matrix_droplet,jaccard_index,jaccard_index_droplet,file = file.path(folder_path, "module_result.rdata"))
-
-saveRDS(tiss, file = "processed_facs_Marrow_srat.rds")
-```
-
-## Step4: Module Enrichment Analysis
-### GO enrichment Analysis
-```{r}
-library("clusterProfiler")
-library("ReactomePA")
-library("org.Mm.eg.db")
-library("AnnotationDbi")
-
-# Set all genes(after filtering) in the expression matrix as background
-universe_df = data.frame("symbol" = rownames(dat),"entrez" = mapIds(org.Mm.eg.db, keys=rownames(dat), column="ENTREZID", keytype="SYMBOL"))
-
-# Filter GO terms based on p values < 0.05
-ego_result <- lapply(1:nlevels(gene_partition), function(i){
-  enrichGO(gene = names(gene_partition)[gene_partition == i],
-           OrgDb = 'org.Mm.eg.db', # mouse
-           keyType = "SYMBOL",
-           ont = "ALL",
-           pAdjustMethod = "BH",
-           universe = universe_df$symbol,
-           pvalueCutoff = 0.05)
-})
-
-# Only keep modules which has at lease one GO term having adjusted p-value < 0.05
-top_go_clusters <- which(sapply(ego_result, function(x)
-  (x@result$p.adjust[1] < 0.05)))
-
-# Annotate each modules with top5 GO terms
-top_enrich_go <- lapply(top_go_clusters, function(i) ego_result[[i]]@result[1:5,])
-```
-
-### Reactome enrichment Analysis
-
-```{r}
-library("ReactomePA")
-library("org.Mm.eg.db")
-
-# Set all genes(after filtering) in the expression matrix as background
-universe_df = data.frame("symbol" = rownames(dat),"entrez" = mapIds(org.Mm.eg.db, keys=rownames(dat), column="ENTREZID", keytype="SYMBOL"))
-universe_df = universe_df[!is.na(universe_df$entrez),]
-
-# Filter Pathway based on p values < 0.05
-epathway_result <- lapply(1:nlevels(gene_partition), function(i){
-  enrichPathway(gene=universe_df[universe_df$symbol %in% names(gene_partition)[gene_partition == i],"entrez"],
-                organism = "mouse",
-                pAdjustMethod = "BH",
-                pvalueCutoff = 0.05,
-                universe = universe_df$entrez)
-})
-
-# Only keep modules which has at lease one Pathway having adjusted p-value < 0.05
-top_pathway_clusters <- which(sapply(epathway_result, function(x) 
-  (x@result$p.adjust[1] < 0.05)))
-
-# Annotate each modules with top5 Pathway terms
-top_enrich_pathway <- lapply(top_pathway_clusters, function(i) epathway_result[[i]]@result[1:5,])
-```
-
-## Step5: Generate Gene Module Annotation Table
-```{r}
-df_module = data.frame(Module = levels(gene_partition))
-df_module[,"Gene"] = unlist(lapply(levels(gene_partition),function(x){paste(names(gene_partition)[gene_partition == x],collapse = ",")}))
-df_module[,"celltype"] = module_celltype[levels(gene_partition)]
-df_module[top_go_clusters,"GO_description"] = unlist(lapply(top_enrich_go,function(x){
-  des = x[["Description"]]
-  paste(des[!is.na(des)],collapse = ",")
-}
-))
-df_module[top_pathway_clusters,"Pathway_description"] = unlist(lapply(top_enrich_pathway,function(x){
-  des = x[["Description"]]
-  paste(des[!is.na(des)],collapse = ",")
-}))
-df_module[is.na(df_module)] = ""
-write.csv(df_module,file = file.path(folder_path, "module_description.csv"),row.names = FALSE)
-```
-
-## Cell Cycle Modules
-### Reactome analysis
-```{r}
-library("ReactomePA")
-library("org.Mm.eg.db")
-universe_df = data.frame("symbol" = rownames(dat),"entrez" = mapIds(org.Mm.eg.db, keys=rownames(dat), column="ENTREZID", keytype="SYMBOL"))
-universe_df = universe_df[!is.na(universe_df$entrez),]
-module_cc = c(16,18,17,31,32)
-epathway_result_cc <- lapply(module_cc, function(i){
-  enrichPathway(gene=universe_df[universe_df$symbol %in% names(gene_partition)[gene_partition == i],"entrez"],
-                organism = "mouse",
-                pAdjustMethod = "BH",
-                pvalueCutoff = 0.05,
-                universe = universe_df$entrez)
-})
-names(epathway_result_cc) = module_cc
-# Annotate each modules with top5 Pathway terms
-top_enrich_pathway_cc <- do.call(rbind,lapply(1:length(epathway_result_cc), function(i) data.frame(epathway_result_cc[[i]]@result[1:5,],module = module_cc[i])))
-top_enrich_pathway_cc <- top_enrich_pathway_cc[complete.cases(top_enrich_pathway_cc), ]
-top_enrich_pathway_cc = top_enrich_pathway_cc %>% filter(p.adjust < 0.05)
-
-write.csv(top_enrich_pathway_cc,file = file.path(folder_path, "reactome_pathway_cc.csv"))
-```
-
-### Annotate Cell Cycle
-```{r}
-mmus_s = gprofiler2::gorth(cc.genes.updated.2019$s.genes, source_organism = "hsapiens", target_organism = "mmusculus")$ortholog_name
-mmus_g2m = gprofiler2::gorth(cc.genes.updated.2019$g2m.genes, source_organism = "hsapiens", target_organism = "mmusculus")$ortholog_name
-tiss = CellCycleScoring(tiss, s.features = mmus_s, 
-                        g2m.features = mmus_g2m, set.ident = FALSE)
-cell_cycle_table = data.frame(tiss@meta.data[,c("S.Score","G2M.Score","Phase")],"Dataset" = "FACS")
-```
-
-### Transferablity
-```{r}
-tiss_droplet = CellCycleScoring(tiss_droplet, s.features = mmus_s, 
-                        g2m.features = mmus_g2m, set.ident = FALSE)
-cell_cycle_table = rbind(cell_cycle_table,data.frame(tiss_droplet@meta.data[,c("S.Score","G2M.Score","Phase")],"Dataset" = "Droplet"))
-
-write.csv(cell_cycle_table,file = file.path(folder_path, "annotated_cell_cycle.csv"))
-```

### 2024-05-23 miscellaneous/miscellaneous_code/Marker_identfication_benchmark_tutorial.Rmd
240523 pipeline-development

--- 2024-03-26 Marker_identfication_benchmark.Rmd
@@
 ## Load Data
-```{r}
-# tiss <- readRDS("/data/ruiqi/local_marker/dataset/tabular_muris/marrow_facs.rds")
-# DefaultAssay(tiss) <- "RNA"
-```
-
 ```{r}
@@
 ```{r}
+# tiss <- readRDS("/banach1/ruiqi/local_marker/dataset/tabular_muris/marrow_facs.rds")
+# DefaultAssay(tiss) <- "RNA"
+# tiss = tiss %>% NormalizeData(1e6) %>%
+#   ScaleData(verbose = FALSE) %>% FindVariableFeatures() %>% RunPCA(npcs = 50, verbose = FALSE)
+```
+
+```{r}
 DefaultAssay(tiss) <- "RNA"
-n_dim = pmin(dim(tiss@reductions$pca@cell.embeddings)[2],30)
+n_dim = 20
 feature_space = as.matrix(tiss@reductions$pca@cell.embeddings[,1:n_dim])
@@
 data = pd.DataFrame(r.dat.T)  # counts
-PC20 = pd.DataFrame(r.feature_space)
+PC_embed = pd.DataFrame(r.feature_space)
 # Reference Cell
 g = 8.6e-4  
-S = pairwise_kernels(PC20, metric="rbf", gamma=g) 
+S = pairwise_kernels(PC_embed, metric="rbf", gamma=g) 
 median = np.median(S, axis=0)
@@
 num_refcell = int(data.shape[1]*0.005)
-dd_rcells = from_knn_dist(X=PC20, 
+dd_rcells = from_knn_dist(X=PC_embed, 
                           n_ret=num_refcell,
@@
                           metric_params={"gamma": g})
-S = pairwise_similarities(PC20,
+S = pairwise_similarities(PC_embed,
                           query=dd_rcells,
@@
 escores = calculate_escores(data.values, query=dd_rcells, S=S)
-
 # semi_ranks = escores.abs().rank(ascending=False, method="dense").min(1)
-
 P = permute(data.values)
@@
 ```
-
+# Concordance table
 ```{r}
-# Concordance table
 do.call(cbind,lapply(gt_list[c(2,21)],function(true_marker){
@@
 ```
+## Plot
+```{r}
+recovery_rate <- reshape2::melt(recovery_rate,id.vars = "TopGene",value.name = "prop",variable.name = "Method")
+recovery_rate = recovery_rate %>% filter(Method!="wilcox_no_filter")
+recovery_rate$Method = ifelse(recovery_rate$Method %in% names(method_fullname),method_fullname[recovery_rate$Method],recovery_rate$Method)
+recovery_rate$'color' = coldef[recovery_rate$Method]
 
-# Sensitivity Test (Try different K)
+ggplot(recovery_rate, aes(x = TopGene, y = prop, color = Method, group = Method)) + 
+  geom_line() +
+  scale_color_manual(values = score_measure %>% filter(Method!="All genes") %>% select(c("Method","color")) %>% distinct() %>% tibble::deframe()) + ylab("Prop. of recovered cell types")
+```
+
+# Sensitivity Test
+## Different kNN
 ```{r}

### 2024-05-23 paper_figure.Rmd
240523 pipeline-development

--- 2023-12-11 paper_figure.Rmd
@@
 library("RColorBrewer")
-# tissue_name = "marrow_facs"
-# tiss <- readRDS(sprintf("%s%s.rds",path,tissue_name))
-tiss <- readRDS("processed_facs_Marrow_srat.rds")
-tiss_droplet <- readRDS("processed_droplet_Marrow_srat.rds")
+library("dplyr")
+library("patchwork")
+library("ggplot2")
+library("ggsignif")
 
-DefaultAssay(tiss) <- "RNA"
-n_dim = dim(tiss@reductions$pca@cell.embeddings)[2]
-feature_space = as.matrix(tiss@reductions$pca@cell.embeddings[,1:n_dim])
-visual_space = data.frame(tiss@reductions$tsne@cell.embeddings)
-dat = as.matrix(tiss[[DefaultAssay(tiss)]]@data)
-cell_label = tiss$cell_ontology_class
-
-Gene_detected_count <- apply(dat > apply(dat,2,median),1,sum)
-selected_genes = (Gene_detected_count >= 10) & (Gene_detected_count <= ncol(dat) * 0.5)
-selected_genes = names(selected_genes)[selected_genes]
-dat = dat[selected_genes,,drop = FALSE]
-raw_count = as.matrix(tiss@assays$RNA@counts)[selected_genes,,drop = FALSE]
-total_count = tiss@meta.data$nCount_RNA
-dat_alra = as.matrix(tiss[["alra"]]@data)[selected_genes,]
+figure_path = "/banach1/ruiqi/local_marker/LMD_data/figures"
+dir.create(figure_path,recursive = TRUE)
+setwd(figure_path)
 ```
 
-# SupFig - Bone-marrow Cell Embedding
+# SupFig - Bone Marrow Facs Cell Embedding
 ```{r}
+dir.path <- "/banach1/ruiqi/local_marker/LMD_data"
+consortium = "tabular_muris"
+folder.path <- file.path(dir.path,consortium)
+tissue_name = "marrow_facs"
+tiss <- readRDS(file.path(folder.path,paste0(tissue_name,".rds")))
+cell_label = tiss$cell_ontology_class
 coldef = setNames(colorRampPalette(brewer.pal(12, "Paired"))(length(unique(cell_label))),unique(cell_label))
@@
 
-ggsave(filename = "supfig-bonemarrow.png", plot = p, width = 8, height = 5, dpi = 300)
+ggsave(filename = file.path(figure_path,"supfig-bonemarrow_facs.png"), plot = p, width = 8, height = 5, dpi = 300)
+```
+
+# SupFig - Bone Marrow droplet Cell Embedding
+```{r}
+dir.path <- "/banach1/ruiqi/local_marker/LMD_data"
+consortium = "tabular_muris"
+folder.path <- file.path(dir.path,consortium)
+tissue_name = "marrow_droplet"
+tiss <- readRDS(file.path(folder.path,paste0(tissue_name,".rds")))
+cell_label = tiss$cell_ontology_class
+coldef = setNames(colorRampPalette(brewer.pal(12, "Paired"))(length(unique(cell_label))),unique(cell_label))
+p = DimPlot(tiss, group.by = "cell_ontology_class",label = FALSE, cols = coldef) + guides(color = guide_legend(ncol = 1,override.aes = list(size = 3))) + labs(title = NULL)
+
+ggsave(filename = file.path(figure_path,"supfig-bonemarrow_droplet.png"), plot = p, width = 8, height = 5, dpi = 300)
 ```
@@
 source("LMD_function.R", echo = F)
-
-# Construct knn graph
-res = Symmetric_KNN_graph(knn = 5, feature_space = feature_space)
-
-A = res$adj_matrix # Adjacency Matrix
-W = res$graph # Symmetrized Graph ((A + AT) / 2)
-
-P_ls = Obtain_Pls(W = W, max_time = 2^15)
-rho = Rowwise_normalize(dat[,colnames(W)])
-res = fast_calculate_multi_score(W = W, init_state = rho, P_ls = P_ls)
+dir.path <- "/banach1/ruiqi/local_marker/LMD_data"
+consortium = "tabular_muris"
+folder.path <- file.path(dir.path,consortium)
+tissue_name = "marrow_facs"
+tiss <- readRDS(file.path(folder.path,paste0(tissue_name,".rds")))
+W <- readRDS(file.path(folder.path,"intermediate_data",tissue_name,paste0(tissue_name,"_knn_graph.rds")))
+res <- readRDS(file.path(folder.path,"intermediate_data",tissue_name,paste0(tissue_name,"_lmd_profile.rds")))
+P_ls <- readRDS(file.path(folder.path,"intermediate_data",tissue_name,paste0(tissue_name,"_diffused_graph.rds")))
+rho <- readRDS(file.path(folder.path,"intermediate_data",tissue_name,paste0(tissue_name,"_gene_distribution.rds")))
 ```
@@
 # Plot knn graph
-p = plot_knn_graph(affinity_m = W, layout = visual_space)
-ggsave(filename = "fig-pipeline1.png", plot = p, width = 6, height = 5, dpi = 300)
+p = plot_knn_graph(affinity_m = W, layout = Embeddings(tiss[["tsne"]])) + NoLegend()
+ggsave(filename = file.path(figure_path,"fig-pipeline1.png"), plot = p, width = 5, height = 5, dpi = 300)
 ```
@@
 # genes_label = c("localized","non_localized")
-check_time = c(0,2^3,2^10,2^15)
-p = Visualize_score_pattern(res, genes = genes, add_point = check_time)
+check_time = c(0,2^3,2^9,2^14)
+p = Visualize_score_pattern(res$score_profile/res$score_profile$max_score0, genes = genes, add_point = check_time) + guides(linetype = "none")
 
-ggsave(filename = "fig-pipeline2.png", plot = p, width = 10, height = 5)
-
+ggsave(filename = file.path(figure_path,"fig-pipeline2.png"), plot = p, width = 10, height = 5)
+```
+```{r}
 # Check Gene Diffusion on Cell graph by time
@@
   gene = genes[i]
-  FeaturePlot_diffusion(coord = visual_space,init_state = rho[gene,],P_ls = P_ls,W = W,check_time = check_time,gene_name = NULL, gene_color = scales::hue_pal()(2)[i])
+  FeaturePlot_diffusion(coord = Embeddings(tiss[["tsne"]]),init_state = rho[gene,,drop = FALSE],P_ls = P_ls,W = W,check_time = check_time,gene_name = NULL, gene_color = scales::hue_pal()(length(genes))[i])
 })
 p = as.ggplot(pl[[1]]) / as.ggplot(pl[[2]] & labs(title = NULL))
-ggsave(filename = "fig-pipeline3.png", plot = p, width = 16, height = 8)
+ggsave(filename = file.path(figure_path,"fig-pipeline3.png"), plot = p, width = 16, height = 8)
 ```
@@
 ```{r}
-library("RColorBrewer")
-library("dplyr")
-library("ggplot2")
-
-# Set the color code for all methods
-method_ls = c("lmd","haystack","hotspot","marcopolo","semi","wilcox_default","wilcox_no_filter","hvg")
-method_fullname = setNames(c("LMD","singleCellHaystack","Hotspot","Marcopolo","SEMITONES","Seurat v4 default", "Seurat v4 Wilcoxon unfiltered","HVGs"),method_ls)
-coldef = setNames(
-  colorRampPalette(brewer.pal(9, "Set1"))(length(method_fullname)),
-  method_fullname)
+method_ls = c("lmd","haystack","hotspot","marcopolo","semi","wilcox_no_filter","hvg")
+method_fullname = setNames(c("LMD","singleCellHaystack","Hotspot","Marcopolo","SEMITONES","Seurat v4","HVGs"),method_ls)
 ```
 ```{r}
-path = "./benchmark_result/"
-tissue_ls = c("marrow_facs","pancreas_facs","lung_facs")
-df_auc = do.call(rbind,lapply(tissue_ls,function(tissue_name){df = read.table(paste0(path,tissue_name,"_auroc.csv"),header = TRUE); df$'Tissue' = tissue_name; df$'Method' = method_fullname[df$'Method']; df}))
-
-df_auc = df_auc %>% filter(gt_set %in% c("Top100","CellMarkerDB")) %>% filter(!Method %in% "HighlyVariableGenes")
+dir.path <- "/banach1/ruiqi/local_marker"
+consortium = "tabular_muris"
+tissue_name = "marrow_facs"
+folder.path = file.path(dir.path, "LocalMarkerDetector","benchmark")
+```
+```{r}
+# Define Full name
+tissue_ls = list("tabular_muris" = c("marrow_facs","pancreas_facs","lung_facs"),
+                 "azimuth" = c("human_kidney", "human_bone_marrow", "human_pancreas", "mouse_motor_cortex", "human_lung_hlca", "human_motor_cortex"))
+consortium_full_name = c("tabular_muris" = "Tabular Muris","azimuth" = "Azimuth")
+tissue_full_name = lapply(unlist(tissue_ls),function(i){
+  tmp_name = 
+  if(grepl("marrow",i)){tmp_name = "Bone Marrow"}
+  else if(grepl("lung_hlca",i)){tmp_name = "Lung v2 (HLCA)"}
+  else if(grepl("lung",i)){tmp_name = "Lung"}
+  else if(grepl("pancreas",i)){tmp_name = "Pancreas"}
+  else if(grepl("kidney",i)){tmp_name = "Kidney"}
+  else if(grepl("motor_cortex",i)){tmp_name = "Motor Cortex"}
+  else{tmp_name = NA}
+  tmp_name
+})
+names(tissue_full_name) = unlist(tissue_ls)
+tissue_full_name = unlist(lapply(1:length(tissue_full_name),function(i){
+  ori_name = names(tissue_full_name)[i]
+  if(grepl("_facs|mouse",ori_name)){
+    tmp_name = paste0("Mouse-\n",tissue_full_name[i])
+  }
+  else if(grepl("human",ori_name)){
+    tmp_name = paste0("Human-\n",tissue_full_name[i])
+  }
+  else{
+    tmp_name = tissue_full_name[i]
+  }
+  tmp_name
+}))
+names(tissue_full_name) = unlist(tissue_ls)
+```
+```{r}
+df_auc = do.call(rbind,
+                lapply(1:length(tissue_ls),function(i){
+                  df_all = NULL
+                  for(j in tissue_ls[[i]]){
+                    df = read.table(file.path(folder.path, names(tissue_ls)[i],paste0(j,"_auroc.csv")),header = TRUE)
+                    df$'Tissue' = j
+                    df_all = rbind(df_all,df)
+                  }
+                  df_all$'consortium' = names(tissue_ls)[i]
+                  df_all}))
+df_auc$Method = method_fullname[df_auc$Method]
+df_auc = df_auc %>% filter(gt_set %in% c("CellMarkerDB","Top100"))
+# Order Method based on mean AUROC
 df_auc$Method <- with(df_auc, reorder(Method, -AUROC, FUN = median))
 df_auc$gt_set <- factor(df_auc$gt_set,levels = c("CellMarkerDB","Top100"))
-levels(df_auc$gt_set) = c("Criterion1 - CellMarkerDB","Criterion2 - MaxFC")
-df_auc$Tissue = factor(df_auc$Tissue,levels = tissue_ls)
-levels(df_auc$Tissue) = c("Bone Marrow","Pancreas","Lung")
-# Create the plot
+levels(df_auc$gt_set) = c("Criterion1-\n CellMarkerDB","Criterion2-\n MaxFC")
+df_auc$Tissue = tissue_full_name[df_auc$Tissue]
+df_auc$consortium = consortium_full_name[df_auc$consortium]
+df_auc$consortium = factor(df_auc$consortium, levels = consortium_full_name)
+```
+## Figure
+```{r}
+coldef = setNames(
+  colorRampPalette(brewer.pal(8, "Set1"))(length(method_fullname)),
+  method_fullname)
+
 fig <- ggplot(df_auc, aes(x = Tissue, y = AUROC, fill = Method)) +
-  geom_bar(stat = "identity", position = "dodge") +
-  geom_text(aes(label = round(AUROC, 2)), position = position_dodge(width = 1), vjust = -0.25, size = 2.5, angle = 45) +
+  geom_bar(stat = "identity", position = "dodge") + facet_grid(gt_set~consortium, scales = 'free') +
+  geom_text(aes(label = round(AUROC, 3)), position = position_dodge(width = 1), vjust = -0.25, size = 2.5, angle = 45) +
   labs(x = "Tissue", y = "AUROC", fill = "Method") + coord_cartesian(ylim = c(0.5, 1)) +
-  theme(legend.position = "bottom")  +  scale_fill_manual(values= coldef[names(coldef)%in%df_auc$Method]) + facet_wrap(~gt_set)
+  theme(legend.position = "bottom")  +  scale_fill_manual(values= coldef[names(coldef)%in%df_auc$Method])
+
 fig = fig + theme(
@@
 
-ggsave(filename = "fig-known_marker_comparison.png", plot = fig, width = 8, height = 5)
+ggsave(filename = file.path(figure_path,"fig-known_marker_comparison.png"), plot = fig, width = 12, height = 6)
+```
+## Table
+```{r}
+library("reshape2")
+wide_data = df_auc; wide_data$AUROC = round(wide_data$AUROC,3)
+wide_data <- reshape2::dcast(wide_data, gt_set + Tissue + consortium ~ Method, value.var = "AUROC")
+colnames(wide_data)[1:3] = c("Criterion","Tissue","DataSource")
+wide_data[,1:3] = wide_data[,3:1]; colnames(wide_data)[1:3]=colnames(wide_data)[3:1]
+wide_data <- wide_data[order(wide_data$DataSource,wide_data$Tissue,wide_data$Criterion),]
+wide_data[,1:3] = apply(wide_data[,1:3],2,function(x) gsub("[\n]|Criterion1-|Criterion2-","",x))
+wide_data[,1:2] = apply(wide_data[,1:2],2,function(x) {x[duplicated(x)] = "";x})
+wide_data$Tissue[17] = "Motor Cortex(Mouse)"
+wide_data$Tissue[13] = "Motor Cortex(Human)"
+wide_data$Tissue = gsub("Human-|Mouse-","",wide_data$Tissue)
+bold_max <- function(vec) {
+  max_val <- max(vec, na.rm = TRUE)
+  vec = unlist(ifelse(vec == max_val, paste0("\\textbf{", vec, "}"), vec))
+  return(vec)
+}
+wide_data[,4:ncol(wide_data)] = t(apply(wide_data[,4:ncol(wide_data)],1,bold_max))
+wide_data[is.na(wide_data)] = "-"
+
+write.csv(wide_data, file = file.path(folder.path,"AUROC_table.csv"), row.names = FALSE)
+```
+# SupFig-RunTime
+```{r}
+df_runtime = do.call(rbind,
+                lapply(1:length(tissue_ls),function(i){
+                  df_all = NULL
+                  for(j in tissue_ls[[i]]){
+                    df = read.table(file.path(folder.path, names(tissue_ls)[i],paste0(j,"_runtime.csv")),header = TRUE)
+                    df$'Tissue' = j
+                    df_all = rbind(df_all,df)
+                  }
+                  df_all$'consortium' = names(tissue_ls)[i]
+                  df_all}))
+df_runtime$Method = method_fullname[df_runtime$Method]
+df_runtime$Tissue = tissue_full_name[df_runtime$Tissue]
+df_runtime$consortium = consortium_full_name[df_runtime$consortium]
+```
+```{r}
+coldef = setNames(
+  colorRampPalette(brewer.pal(8, "Set1"))(length(method_fullname)),
+  method_fullname)
+
+fig <- ggplot(df_runtime, aes(x = nCells, y = Time, color = Method)) +
+  geom_point(aes(size = nGenes)) + geom_line(aes(group = Method)) + scale_y_log10() + scale_x_continuous(breaks = c(1e3,2e3,5e3,1e4,2e4,3e4), labels = c("1k","2k","5k","10k","20k","30k")) +
+  geom_hline(aes(yintercept = 30), linetype = "dashed") + 
+  annotate("text", x = 0, y = 30, label = "30 min",  vjust = -0.5) + 
+  geom_hline(aes(yintercept = 60), linetype = "dashed") + annotate("text", x = 0, y = 60, label = "1 hr",  vjust = -0.5) + 
+  scale_color_manual(values= coldef[names(coldef)%in%df_runtime$Method]) +
+  labs(x = "nCells", y = "RunTime(minutes)", color = "Method")
+
+fig = fig + theme(
+          legend.title = element_text(size = 10),
+          legend.text = element_text(size = 10),
+          axis.title.x = element_text(size = 15),
+          axis.title.y = element_text(size = 15),
+          axis.text.x = element_text(size = 12),
+          axis.text.y = element_text(size = 10),
+          panel.grid = element_blank(),
+          axis.line = element_line(colour = "black"),
+          panel.background = element_blank(),
+          strip.background = element_rect(fill = "white", color = "white") ,
+          strip.text = element_text(size = 15, color = "black"))
+
+ggsave(filename = file.path(figure_path,"supfig-runtime.png"), plot = fig, width = 8, height = 5)
 ```
@@
 ```{r}
-path = "./benchmark_result/"
+dir.path <- "/banach1/ruiqi/local_marker"
+consortium = "tabular_muris"
 tissue_name = "marrow_facs"
-
-df_benchmark = read.table(paste0(path,tissue_name,"_benchmark_rank_table.csv"))
+df_benchmark = read.table(file.path(dir.path, "LocalMarkerDetector","benchmark", consortium, paste0(tissue_name,"_benchmark_rank_table.csv")),header = TRUE)
+rownames(df_benchmark) = df_benchmark$gene
+tiss = readRDS(file.path(dir.path, "LMD_data",consortium, paste0(tissue_name,".rds")))
 method_vs = "lmd"
-method_vs_set = c("wilcox_default","hotspot","marcopolo","semi","haystack")
+method_vs_set = c("wilcox_no_filter","hotspot","marcopolo","semi","haystack")
 ```
@@
 
-ggsave(filename = "fig-ranking_comparison2.png", plot = fig, width = 25, height = 10)
-
-write.csv(df_label_all,file = "./marrow_result/gene_rank_comparison.csv",row.names = TRUE)
+ggsave(filename = file.path(figure_path,"fig-ranking_comparison2.png"), plot = fig, width = 25, height = 10)
 ```
@@
 
-ggsave(filename = "fig-ranking_comparison1.png", plot = p, width = 25, height = 6)
+ggsave(filename = file.path(figure_path,"fig-ranking_comparison1.png"), plot = p, width = 25, height = 6)
 ```
@@
 
-# Fig - Density Index
+# SupFig - Density Index
 ```{r}
-path = "./benchmark_result/"
-tissue_name = "marrow_facs"
-score_measure = read.table(paste0(path,tissue_name,"_density_index.csv"),header = TRUE)
-score_measure = score_measure %>% filter(Method!="wilcox_no_filter")
+# Set the color code for all methods
+method_ls = c("lmd","haystack","hotspot","marcopolo","semi","wilcox_no_filter","hvg")
+method_fullname = setNames(c("LMD","singleCellHaystack","Hotspot","Marcopolo","SEMITONES","Seurat v4","HVGs"),method_ls)
+coldef = setNames(
+  colorRampPalette(brewer.pal(8, "Set1"))(length(method_fullname)),
+  method_fullname)
+```
+
+```{r}
+consortium = "tabular_muris"
+dir.path <- "/banach1/ruiqi/local_marker/LocalMarkerDetector"
+pl = lapply(c("marrow_facs","lung_facs","pancreas_facs"),function(tissue_name){
+score_measure = read.table(file.path(dir.path,"benchmark",consortium,paste0(tissue_name,"_DI.csv")),header = TRUE)
 score_measure$Method = ifelse(score_measure$Method %in% names(method_fullname),method_fullname[score_measure$Method],score_measure$Method)
@@
 baseline = score_measure %>% filter(Method == "All genes") %>%.$DensityIndex
-max_x = score_measure %>% filter(Method!="All genes") %>%.$TopGene %>% max()
 
-score_measure$'color' = coldef[score_measure$Method]
-score_measure[score_measure$Method == "All genes","color"] = "black"
-score_measure$'line_type' = "solid"
-score_measure[score_measure$Method == "All genes","line_type"] = "dashed"
-
-score_measure[grepl("Seurat",score_measure$Method),"Method"] = "Seurat v4"
-
-# line_type = replace(coldef,names(coldef),"solid")
-# line_type["Seurat v4 Wilcoxon unfiltered"] = "twodash"
-# line_type["All genes"] = "dashed"
-
-fig = ggplot(score_measure %>% filter(Method!="All genes") %>% filter(TopGene >= 100), aes(x = TopGene, y = DensityIndex, color = Method, group = Method, linetype = Method),linewidth = 2) + 
-  geom_line(lwd = 1) +
+fig = ggplot(score_measure %>% filter(Method!="All genes"), aes(x = TopGene, y = DensityIndex, color = Method, group = Method, linetype = Method)) + 
+  geom_line(size = 1) +
   scale_x_log10(breaks = 10^(2:4), labels = expression(10^2, 10^3, 10^4)) +
     # expand_limits(x = c(50, 10^4)) +
-  geom_hline(aes(yintercept = baseline, color = "All genes", linetype = "All genes")) +
-  scale_color_manual(values = score_measure[,c("Method","color")] %>% distinct() %>% tibble::deframe()) + 
-  scale_linetype_manual(values = score_measure[,c("Method","line_type")] %>% distinct() %>% tibble::deframe()) +
-  theme(
+  geom_hline(aes(yintercept = baseline, color = "All genes", linetype = "All genes"))
+
+# Add null distribution
+null_score_measure =
+  read.table(file.path(dir.path,"benchmark",consortium,paste0(tissue_name,"_DI_null.csv")),header = TRUE)
+# Calculate the mean & Prediction Interval
+## account for both the variability of the observations and the uncertainty in estimating the mean
+null_stats <- null_score_measure %>%
+  group_by(TopGene) %>% 
+  summarize(Mean = mean(DensityIndex),Sd = sd(DensityIndex),sample_size = n(),.groups = 'drop') %>%
+  mutate( Lower_CI = Mean + qt(0.025, df = sample_size - 1) * Sd * sqrt(1 + 1 / sample_size), Upper_CI = Mean + qt(0.975, df = sample_size - 1) * Sd * sqrt(1 + 1 / sample_size) )
+null_stats$'Method' = "Null"
+
+fig = fig + geom_line(data = null_stats, aes(x = TopGene, y=Mean)) + geom_ribbon(data = null_stats, aes(x = TopGene, y=Mean, ymin=Lower_CI, ymax=Upper_CI), alpha=0.2, inherit.aes = FALSE)
+
+# Add format
+coldef["All genes"] = "black";  coldef["Null"] = "grey"
+linetypedef = coldef; linetypedef[1:7] = "solid"; linetypedef[8:9] = "dashed"
+fig = fig + ggtitle(ifelse(grepl("marrow",tissue_name),"Bone Marrow",ifelse(grepl("lung",tissue_name),"Lung",ifelse(grepl("pancreas",tissue_name),"Pancreas",NA)))) +
+  scale_color_manual(values = coldef) + 
+  scale_linetype_manual(values = linetypedef)
+
+# Add LowQualityRate Scatterplot
+fig2 = ggplot(score_measure %>% filter(Method!="All genes") %>% filter(TopGene %in% c(50,100,500)), aes(x = LowQualityRate, y = DensityIndex, color = Method)) + 
+  geom_point() + xlim(0,1) + 
+  scale_color_manual(values = coldef) + facet_wrap(~TopGene)  + guides(color = "none")
+
+(fig+fig2 + plot_layout(widths = c(1, 2))) &theme_minimal() & theme(
+    plot.title = element_text(face="bold", size = 20),
     legend.title = element_blank(),
     legend.text = element_text(size = 10),
-    legend.position = "bottom",
+    # legend.position = "bottom",
     legend.justification = 1,
@@
     axis.text.y = element_text(size = 10),
-    panel.grid = element_blank(),
-    panel.background = element_blank(),
+    # panel.grid = element_blank(),
+    # panel.background = element_blank(),
     axis.line = element_line(colour = "black"))
-  # geom_text_repel(data = score_measure %>%  filter(Method!="all genes") %>% group_by(Method) %>% slice_head(n = 2) %>% slice_tail(n = 1), aes(label = Method), hjust = 0)
+})
 
-ggsave(filename = "fig-densityindex.png", plot = fig, width = 5, height = 5)
+fig = wrap_plots(pl,ncol = 1) + plot_layout(guides = "collect") & theme(legend.position = 'bottom')
+ggsave(filename = file.path(figure_path,"supfig-densityindex.png"), plot = fig, width = 12, height = 12)
+
+# Only keep Bone Marrow
+fig = wrap_plots(pl[[1]],ncol = 1) + plot_layout(guides = "collect") & theme(legend.position = 'bottom')
+ggsave(filename = file.path(figure_path,"supfig-densityindex.png"), plot = fig, width = 12, height = 5)
 ```
@@
 # SupFig - Sensitivity Test
+## kNN
 ```{r}
-path = "./marrow_result/"
+folder.path = file.path(dir.path,"LMD_data",consortium,"intermediate_data", tissue_name, "sensitivity_test")
+test_para = "knn"
+
 ## Rank Stability
-score_df = read.table(paste0(path,tissue_name,"_different_knn.csv"),header = TRUE)
-score_ranked <- data.frame(score_df) %>% mutate(across(everything(), rank))
-base = "kNN5"
-
-top_genes = c(50,seq(100,1000,100))
-jaccard_index_bw_sets = do.call(cbind,lapply(top_genes,function(top_gene)
-    {
-     top_gene_ls = apply(score_ranked,2,function(x) rownames(score_ranked)[order(x)][1:top_gene])
-     apply(top_gene_ls,2,function(x){
-       length(intersect(x,top_gene_ls[,base]))/length(union(x,top_gene_ls[,base]))
-     })
-  }))
-colnames(jaccard_index_bw_sets) = top_genes
-df = reshape2::melt(jaccard_index_bw_sets,value.name = "JaccardIndex")
-colnames(df)[1:2] = c("kNN","Top_Genes")
-df$kNN = as.factor(as.numeric(gsub("kNN","",df$kNN)))
+df = read.table(file.path(folder.path,sprintf("%s_jaccard_ind_%s.csv",tissue_name,test_para)),header = TRUE)
+df$knn = as.factor(as.numeric(gsub("\\D", "", df$knn)))
 df$Top_Genes = as.factor(df$Top_Genes)
-
-fig = ggplot(data = df, aes(x = kNN, y = JaccardIndex, group = Top_Genes, color = Top_Genes)) +
+fig1 = ggplot(data = df, aes(x = knn, y = JaccardIndex, group = Top_Genes, color = Top_Genes)) +
   geom_line() +
-  geom_point() + labs(x = "# of kNN used", y = sprintf("Jaccard Index\n (with %s)",base),color = "Top Genes") + geom_vline(xintercept = 3, color="red", linetype="dashed") + 
-  # scale_y_continuous(limits = c(0,1)) +
+  geom_point() + labs(x = "# of kNN used", y = "Jaccard Index\n (with kNN5)",color = "Top Genes") + geom_vline(xintercept = 3, color="red", linetype="dashed") + 
+  scale_y_continuous(limits = c(0,1)) +
   theme(
@@
 
-ggsave(filename = "supfig-sensitivity1.png", plot = fig, width = 8, height = 5)
-```
-```{r}
-## AUC stability
-auc_df = read.table(paste0(path,tissue_name,"_sensitivity_auroc.csv"),header = TRUE)
-auc_df <- auc_df %>% tidyr::pivot_longer(cols = -gt, names_to = "kNN", values_to = "AUC") %>% filter(gt %in% c("Top100","CellMarkerDB"))
-auc_df$gt = factor(auc_df$gt); levels(auc_df$gt) = c("CellMarkerDB","MacFC(Top100)")
-auc_df$kNN = as.factor(as.numeric(gsub("kNN","",auc_df$kNN)))
+# AUC stability
+df = read.table(file.path(folder.path,sprintf("%s_auroc_%s.csv",tissue_name,test_para)),header = TRUE)
+df = df %>% filter(gt_set %in% c("Top100","CellMarkerDB"))
+df$gt_set = factor(df$gt_set)
+levels(df$gt) = ifelse(grepl("CellMarkerDB",levels(df$gt)),levels(df$gt),"MacFC(Top100)")
+df$Parameter = as.factor(as.numeric(gsub("\\D", "", df$Parameter)))
 
-fig = ggplot(data = auc_df, aes(x = kNN, y = AUC, group = gt, color = gt)) +
+fig2 = ggplot(data = df, aes(x = Parameter, y = AUROC, group = gt_set, color = gt_set)) +
   geom_line() +
   geom_point() + labs(x = "# of kNN used", y = "AUROC", color = "Gold-Standard Genesets") + 
-  # scale_y_continuous(limits = c(0,1)) + 
+  scale_y_continuous(limits = c(0,1)) +
   theme(
@@
 
-ggsave(filename = "supfig-sensitivity2.png", plot = fig, width = 8, height = 5)
+fig = fig1 + fig2 + plot_annotation(title = "Sensitivity Test: Different kNN", theme = theme(plot.title = element_text(size = 20)))
+ggsave(filename = file.path(figure_path,"supfig-sensitivityknn.png"), plot = fig, width = 16, height = 5)
+```
+## PC
+```{r}
+folder.path = file.path(dir.path,"LMD_data",consortium,"intermediate_data", tissue_name, "sensitivity_test")
+test_para = "pc"
+
+## Rank Stability
+df = read.table(file.path(folder.path,sprintf("%s_jaccard_ind_%s.csv",tissue_name,test_para)),header = TRUE)
+df$pc = ifelse(grepl("tSNE",df$pc),"tSNE",gsub("\\D", "", df$pc))
+df$pc = factor(df$pc,levels = c(sort(unique(as.numeric(df$pc))),"tSNE"))
+df$Top_Genes = as.factor(df$Top_Genes)
+fig1 = ggplot(data = df, aes(x = pc, y = JaccardIndex, group = Top_Genes, color = Top_Genes)) +
+  geom_line() +
+  geom_point() + labs(x = "# of PC used", y = "Jaccard Index\n (with 20PC)", color = "Top Genes") + geom_vline(xintercept = 2, color="red", linetype="dashed") + 
+  scale_y_continuous(limits = c(0,1)) +
+  theme(
+    legend.title = element_text(size = 20),
+    legend.text = element_text(size = 15),
+    axis.title.x = element_text(size = 20),
+    axis.title.y = element_text(size = 20),
+    axis.text.x = element_text(size = 15),
+    axis.text.y = element_text(size = 15),
+    panel.grid = element_blank(),
+    panel.background = element_blank(),
+    axis.line = element_line(colour = "black"))
+
+# AUC stability
+df = read.table(file.path(folder.path,sprintf("%s_auroc_%s.csv",tissue_name,test_para)),header = TRUE)
+df = df %>% filter(gt_set %in% c("Top100","CellMarkerDB"))
+df$gt_set = factor(df$gt_set)
+levels(df$gt) = ifelse(grepl("CellMarkerDB",levels(df$gt)),levels(df$gt),"MacFC(Top100)")
+df$Parameter = ifelse(grepl("tSNE",df$Parameter),"tSNE",gsub("\\D", "", df$Parameter))
+df$Parameter = factor(df$Parameter,levels = c(sort(unique(as.numeric(df$Parameter))),"tSNE"))
+
+fig2 = ggplot(data = df, aes(x = Parameter, y = AUROC, group = gt_set, color = gt_set)) +
+  geom_line() +
+  geom_point() + labs(x = "# of PC used", y = "AUROC", color = "Gold-Standard Genesets") + 
+  scale_y_continuous(limits = c(0,1)) +
+  theme(
+    legend.title = element_text(size = 20),
+    legend.text = element_text(size = 15),
+    axis.title.x = element_text(size = 20),
+    axis.title.y = element_text(size = 20),
+    axis.text.x = element_text(size = 15),
+    axis.text.y = element_text(size = 15),
+    panel.grid = element_blank(),
+    panel.background = element_blank(),
+    axis.line = element_line(colour = "black"))
+
+fig = fig1 + fig2 + plot_annotation(title = "Sensitivity Test: Different PC", theme = theme(plot.title = element_text(size = 20)))
+ggsave(filename = file.path(figure_path,"supfig-sensitivitypc.png"), plot = fig, width = 16, height = 5)
 ```
 
-# SupFig - Gene-Gene Colocolization HeatMap
+# SupFig - Gene-Gene Jaccard-dist HeatMap
 ```{r}
-load("./marrow_result/module_result.rdata")
-source("./LMD_function.R", echo = F)
-library(RColorBrewer)
-library(dplyr)
-module_color = setNames(colorRampPalette(brewer.pal(12, "Paired"))(nlevels(gene_partition)),levels(gene_partition))
+# load("./marrow_result/module_result.rdata")
+source("LMD_function.R", echo = F)
 celltype_color = setNames(colorRampPalette(brewer.pal(9, "Set1"))(length(unique(cell_label))),unique(cell_label))
-
-p = pheatmap(as.matrix(dist),
-              annotation_col = data.frame(Module = gene_partition),
-              cluster_rows = gene_hree, 
-              cluster_cols = gene_hree,
-              ann_colors = list(Module = module_color),
-              treeheight_row = 0,
-              col=colorRampPalette(c("firebrick3", "white"))(99),
-              show_colnames = FALSE, 
-              show_rownames = FALSE,
-              annotation_legend = FALSE,
-              # main = paste(distance_method,
-              #              clustering_method,sep = "-"),
-              silent = TRUE)
-p = as.ggplot(p)
-ggsave(filename = "supfig-moduleheatmap.png", plot = p, width = 10, height = 10)
+consortium = "tabular_muris"
+tissue_name = "marrow_facs"
+dir.path <- "/banach1/ruiqi/local_marker/LMD_data"
+folder.path = file.path(dir.path,consortium)
+dist = readRDS(file.path(folder.path,"intermediate_data",tissue_name,paste0(tissue_name,"_gene_dist.rds")))
+res = readRDS(file.path(folder.path,"intermediate_data",tissue_name,paste0(tissue_name,"_gene_tree.rds")))
+p = Visualize_gene_heatmap(dist, gene_partition = res$gene_partition_rename, gene_hree = res$gene_hree)
+ggsave(filename = file.path(figure_path,"supfig-moduleheatmap.png"), plot = p, width = 10, height = 10)
 ```
 
-# SupFig - Module TSNE
+# SupFig - Module Activity Score (Bone marrow)
 ```{r}
-pl_module = lapply(levels(gene_partition),function(module){
-  FeaturePlot_custom(value = cell_block[,module], coord = visual_space, value_name = "Module Score",title_name = sprintf("Module %s (%d genes)",module, sum(gene_partition == module))) + scale_color_gradient(limits = c(0, 1),low = "lightgrey", high = "blue")
+consortium = "tabular_muris"
+tissue_name = "marrow_facs"
+dir.path <- "/banach1/ruiqi/local_marker/LMD_data"
+folder.path = file.path(dir.path,consortium)
+res = readRDS(file.path(folder.path,"intermediate_data",tissue_name,paste0(tissue_name,"_gene_tree.rds")))
+tiss = readRDS(file.path(folder.path,paste0(tissue_name,".rds")))
+
+modules = levels(res$gene_partition_rename)
+pl_module = lapply(modules,function(module){
+  module_name = paste0("Module",module)
+  (FeaturePlot(tiss, features = module_name, order = TRUE) + 
+    scale_color_gradient(low = "lightgrey", high = "blue", limits = c(0,1)) +
+    ggtitle(sprintf("Module %s (%d genes)", module, sum(res$gene_partition_rename == module))) + labs(color = "ModuleScore")) & NoAxes()
 })
-names(pl_module) = levels(gene_partition)
-p = (wrap_plots(pl_module,ncol = 10) & theme(plot.title = element_text(face="bold", size = 10),legend.title = element_text(size = 30),legend.text = element_text(size = 20))) + plot_layout(guides = "collect")
+names(pl_module) = modules
+p = (wrap_plots(pl_module,ncol = 7) &  theme(plot.title = element_text(face="bold", size = 10),legend.title = element_text(size = 30),legend.text = element_text(size = 20))) + plot_layout(guides = "collect")
 
-ggsave(filename = "supfig-moduletsne.png", plot = p, width = 30, height = 18)
+ggsave(filename = file.path(figure_path,"supfig-moduletsne.png"), plot = p, width = 7*3, height = 5*3)
 ```
 
-# SupFig-CellType-specific Module
-##  Matched on Cell Embeddings
+# SupFig-CellType-specific Module (Bone marrow)
+## Jaccard Index Heatmap
+```{r}
+one_hot_matrix <- 
+  readRDS(file.path(folder.path,"intermediate_data",tissue_name,paste0(tissue_name,"_ct_onehot_mtx.rds")))
+
+jaccard_index <- 
+  readRDS(file.path(folder.path,"intermediate_data",tissue_name,paste0(tissue_name,"_ct_module_jaccard.rds")))
+```
+```{r}
+thred = 0.3
+jaccard_index_pruned = jaccard_index[apply(jaccard_index,1,max) >= thred,]
+jaccard_index_pruned = jaccard_index_pruned[,apply(jaccard_index_pruned,2,max) >= thred]
+
+colnames(jaccard_index_pruned) = gsub("Module","M",colnames(jaccard_index_pruned))
+jaccard_index_pruned = jaccard_index_pruned[names(sort(apply(jaccard_index_pruned,1,which.max))),]
+p = Visualize_jaccard_mtx(jaccard_index_pruned, threshold = thred)
+p = as.ggplot(p) + coord_fixed()
+ggsave(filename = file.path(figure_path,"supfig-ct-specific-module2.png"), plot = p, width = 8, height = 10)
+```
+##  Module-CT-TSNE
 For each cell type, select the most specific modules for visualization (maximum jaccard index)
 ```{r}
-# Get CellType One-hot plot
 pl_ct = lapply(colnames(one_hot_matrix),function(ct){
-  FeaturePlot_custom(value = as.factor(one_hot_matrix[,ct]), coord = visual_space, value_name = "Indicator",title_name = ct) + scale_color_manual(values = c("0" = "lightgrey", "1" = scales::hue_pal()(1)))
+  FeaturePlot_custom(coord = Embeddings(tiss[["tsne"]]),value = as.character(one_hot_matrix[,ct]), title_name = ct, value_name = "Indicator") + scale_color_manual(values = c("0" = "lightgrey", "1" = scales::hue_pal()(1)))
 })
@@
 
-ct_module_match = apply(jaccard_index_pruned,1,function(x) colnames(jaccard_index_pruned)[which.max(x)])
-ct_module_match = gsub("M","",ct_module_match)
+ct_module_match = apply(jaccard_index,1,function(x) colnames(jaccard_index)[which.max(x)])
+ct_module_match = gsub("Module","",ct_module_match)
+ct_module_match["granulocyte"] = "11"
+ct_module_match["granulocytopoietic cell"] = "8"
 
@@
 })
+names(pl) = names(ct_module_match)
+pl = pl[rownames(jaccard_index_pruned)]
+p = (wrap_plots(pl,ncol = 6) & theme(plot.title = element_text(face="bold", size = 10),legend.title = element_text(size = 15),legend.text = element_text(size = 10))) + plot_layout(guides = "collect")
 
-p = (wrap_plots(pl,ncol = 7) & theme(plot.title = element_text(face="bold", size = 10),legend.title = element_text(size = 15),legend.text = element_text(size = 10))) + plot_layout(guides = "collect")
+# p = (wrap_plots(pl,ncol = 8) & theme(plot.title = element_text(face="bold", size = 20),legend.title = element_text(size = 15),legend.text = element_text(size = 10))) + plot_layout(guides = "collect")
+# ggsave(filename = "supfig-ct-specific-module1.png", plot = p, width = 3*8, height = 6)
 
-ggsave(filename = "supfig-ct-specific-module1.png", plot = p, width = 3*7, height = 3*4)
-```
-
-## Jaccard Index Heatmap
-```{r}
-thred = 0.4
-jaccard_index_pruned = jaccard_index[apply(jaccard_index,1,max) > thred,]
-jaccard_index_pruned = jaccard_index_pruned[,apply(jaccard_index_pruned,2,max) > thred]
-
-colnames(jaccard_index_pruned) = paste0("M",colnames(jaccard_index_pruned))
-jaccard_index_pruned = jaccard_index_pruned[names(sort(apply(jaccard_index_pruned,1,which.max))),]
-p = Visualize_jaccard_mtx(jaccard_index_pruned)
-p = as.ggplot(p) + coord_fixed()
-ggsave(filename = "supfig-ct-specific-module2.png", plot = p, width = 8, height = 10)
+ggsave(filename = file.path(figure_path,"supfig-ct-specific-module1.png"), plot = p, width = 3*7, height = 3*4)
 ```
@@
 
-## Droplet Jaccard Comparison
+## Transfer Jaccard Index
 ```{r}
-module_selected = unlist(apply(jaccard_index[rownames(jaccard_index_droplet),],1,function(x) colnames(jaccard_index)[which(x > thred)]))
-module_selected = module_selected[order(as.numeric(module_selected))]
+jaccard_index = lapply(c("marrow_facs","marrow_droplet"), function(tissue_name) readRDS(file.path(folder.path,"intermediate_data",tissue_name,paste0(tissue_name,"_ct_module_jaccard.rds"))))
+one_hot_matrix = lapply(c("marrow_facs","marrow_droplet"), function(tissue_name) readRDS(file.path(folder.path,"intermediate_data",tissue_name,paste0(tissue_name,"_ct_onehot_mtx.rds"))))
+```
 
-ct_selected = unique(unlist(apply(jaccard_index[rownames(jaccard_index_droplet),module_selected],2,function(x) rownames(jaccard_index_droplet)[which(x > 0.4)])))
-ct_selected = names(sort(apply(jaccard_index[ct_selected,],1,which.max)))
+```{r}
+module_selected = unique(unlist(apply(jaccard_index[[1]][rownames(jaccard_index[[2]]),],1,function(x) colnames(jaccard_index[[1]])[which(x > thred)])))
+module_selected = module_selected[order(as.numeric(gsub("Module","",module_selected)))]
 
-pl = lapply(list(jaccard_index,jaccard_index_droplet),function(mtx){
+ct_selected = unique(unlist(apply(jaccard_index[[1]][rownames(jaccard_index[[2]]),module_selected],2,function(x) rownames(jaccard_index[[2]])[which(x > thred)])))
+ct_selected = names(sort(apply(jaccard_index[[1]][ct_selected,],1,which.max)))
+
+pl = lapply(jaccard_index,function(mtx){
   jaccard_index_pruned = mtx[ct_selected,module_selected]
-  colnames(jaccard_index_pruned) = paste0("M",colnames(jaccard_index_pruned))
-  Visualize_jaccard_mtx(jaccard_index_pruned)
+  colnames(jaccard_index_pruned) = gsub("Module","M",colnames(jaccard_index_pruned))
+  Visualize_jaccard_mtx(jaccard_index_pruned,threshold = thred)
 })
@@
 p = wrap_plots(pl,ncol = 2) + plot_layout(guides = "collect")
-ggsave(filename = "supfig-ct-specific-module3.png", plot = p, width = 10, height = 8)
+ggsave(filename = file.path(figure_path,"supfig-ct-specific-module3.png"), plot = p, width = 10, height = 8)
 ```
-# Fig - Cell Cycle
-## Annotate CC
+
+# Fig - Cell Cycle (Bone Marrow)
+## CC & Module TSNE
 ```{r}
-cell_cycle = read.csv("./marrow_result/annotated_cell_cycle.csv")
-tiss$'Phase' = cell_cycle %>% filter(Dataset == "FACS") %>%.$Phase
-tiss$Phase = factor(tiss$Phase,levels = c("G1","S","G2M"))
-p1 = DimPlot(tiss,group.by = "Phase", cols = c("grey",scales::hue_pal()(3)[3:2])) + labs(title = "Phase annotated by Seurat",color = "Phase") + theme(
+tissue_name = "marrow_facs"
+tiss = readRDS(file.path(folder.path,paste0(tissue_name,".rds")))
+tissue_name = "marrow_droplet"
+tiss_droplet = readRDS(file.path(folder.path,paste0(tissue_name,".rds")))
+reduction = "tsne_cc"
+reduction = "tsne"
+p1 = DimPlot(tiss,group.by = "Phase", cols = c("grey",scales::hue_pal()(3)[3:2]), reduction = reduction) + labs(title = "Phase annotated by Seurat",color = "Phase") + theme(
   plot.title = element_text(size = 10),
@@
       axis.line = element_blank())
-tiss_droplet$'Phase' = cell_cycle %>% filter(Dataset == "Droplet") %>%.$Phase
-tiss_droplet$Phase = factor(tiss_droplet$Phase,levels = c("G1","S","G2M"))
-p2 = DimPlot(tiss_droplet,group.by = "Phase", cols = c("grey",scales::hue_pal()(3)[3:2])) + labs(title = "Phase annotated by Seurat",color = "Phase") + theme(
+p2 = DimPlot(tiss_droplet,group.by = "Phase", cols = c("grey",scales::hue_pal()(3)[3:2]), reduction = reduction) + labs(title = "Phase annotated by Seurat",color = "Phase") + theme(
   axis.title = element_blank(),
@@
 
-cc_module = c(32,31,18)
+# cc_module = c(32,31,18)
+cc_module = c(22,20,19)
 pl_module_cc = lapply(cc_module,function(module){
-  FeaturePlot_custom(value = cell_block[,module], coord = visual_space, value_name = "Module Score",title_name = sprintf("Module %s (%d genes)",module, sum(gene_partition == module)))
+  module_name = paste0("Module",module)
+  (FeaturePlot(tiss, features = module_name, order = TRUE, reduction = reduction) + 
+    scale_color_gradient(low = "lightgrey", high = "blue", limits = c(0,1)) +
+    ggtitle(sprintf("Module %s (%d genes)", module, sum(res$gene_partition_rename == module))) + labs(color = "ModuleScore")) & NoAxes()
 })
 pl_module_cc_droplet = lapply(cc_module,function(module){
-  FeaturePlot_custom(value = cell_block_droplet[,module], coord = data.frame(tiss_droplet@reductions$tsne@cell.embeddings), value_name = "Module Score",title_name = sprintf("Module %s (%d genes)",module, sum(gene_partition == module)))
+  module_name = paste0("Module",module)
+  (FeaturePlot(tiss_droplet, features = module_name, order = TRUE, reduction = reduction) + 
+    scale_color_gradient(low = "lightgrey", high = "blue", limits = c(0,1)) +
+    ggtitle(sprintf("Module %s (%d genes)", module, sum(res$gene_partition_rename == module))) + labs(color = "ModuleScore")) & NoAxes()
 })
@@
 
-ggsave(filename = "fig-module-cellcycle1.png", plot = p, width = 3*4 + 1, height = 3*2)
+ggsave(filename = file.path(figure_path,"fig-module-cellcycle1.png"), plot = p, width = 3*4 + 1, height = 3*2)
+ggsave(filename = file.path(figure_path,"supfig-module-cellcycle.png"), plot = p, width = 3*4 + 1, height = 3*2)
 ```
@@
 ```{r}
-df = read.csv("./marrow_result/reactome_pathway_cc.csv")
-df = df %>% filter(module %in% cc_module)
+df = read.csv(file.path(folder.path, "intermediate_data",tissue_name,"reactome_pathway_cc.csv"))
 df$GeneRatio = unlist(lapply(df$GeneRatio,function(x) eval(parse(text = x))))
-# df$GeneRatio = as.numeric(df$Count / table(gene_partition)[df$module])
-Description_order = levels(factor(df$Description))
-Description_order = rev(Description_order[order(match(Description_order,c("Cell Cycle","Cell Cycle, Mitotic","Mitotic G1 phase and G1/S transition","G1/S Transition","S Phase","M Phase","DNA Replication","Mitotic Prometaphase")))])
-df$Description = factor(df$Description,levels = Description_order) 
-df$module = factor(df$module,levels = c(32,31,18))
-
+df$Description = unlist(lapply(df$Description,function(i) add_newline_in_middle(i, 30)))
+df$Description = as.factor(df$Description)
+df$Description = factor(df$Description, levels(df$Description)[rev(c(2,4,6,10,1,3,5,9,7,8))])
+df$module = factor(df$module,levels = cc_module)
+levels(df$module) = paste0("Module",levels(df$module))
 p = ggplot(df, aes(x=module, y=Description)) +
@@
   theme(legend.position="right",
-        axis.title = element_text(size = 20),
+        axis.title = element_text(size = 0),
         axis.text.x = element_text(size = 15),
@@
       legend.text = element_text(size = 15))
-ggsave(filename = "fig-module-cellcycle2.png", plot = p, width = 10, height = 6)
+ggsave(filename = file.path(figure_path,"fig-module-cellcycle2.png"), plot = p, width = 10, height = 6)
 ```
@@
 ```{r}
-level0_module = c(13,27,38,46,54)
+# level0_module = c(13,27,38,46,54)
+level0_module = c(4,16,28,17,33)
 p = (wrap_plots(pl_module[level0_module],ncol = 5) &  theme(plot.title = element_text(size = 15))) + plot_layout(guides = "collect")
 
-ggsave(filename = "supfig-joint-module.png", plot = p, width = 5*3, height = 3)
+ggsave(filename = file.path(figure_path,"supfig-joint-module.png"), plot = p, width = 5*3, height = 3)
 ```
@@
 ```{r}
-B_module = c(7,9,10)
-B_module_genes = list("8" = c("Lef1","Slamf7"),
-                      "6" = c("Vreb1","Igll1"),
-                      "7" = c("Rag1", "Rag2"),
-                      "9" = c("Ms4a1", "Cd22"),
-                      "10" = c("Fcer2a","Ms4a4c","Cxcr5"))
+# B_module = c(7,9,10)
+# B_module_genes = list("8" = c("Lef1","Slamf7"),
+#                       "6" = c("Vreb1","Igll1"),
+#                       "7" = c("Rag1", "Rag2"),
+#                       "9" = c("Ms4a1", "Cd22"),
+#                       "10" = c("Fcer2a","Ms4a4c","Cxcr5"))
+
+modules_genes = list("7" = c("Lef1","Enpep"), # late-proB
+                      "6" = c("Rag2","Il7r"),
+                      "2" = c("Ms4a1", "Cd22"),
+                      "3" = c("Fcer2a","Ms4a4c","Cxcr5"))
 cell_id = grep("B cell",tiss$cell_ontology_class)
 cell_id = cell_id[tiss$cell_ontology_class[cell_id] != "B cell"]
-pl_Bmodule = lapply(B_module,function(module){
-  FeaturePlot_custom(value = cell_block[cell_id,module], coord = visual_space[cell_id,], value_name = "Module Score",title_name = sprintf("Module %s",module)) + labs(subtitle = paste0(B_module_genes[[as.character(module)]],collapse = ","))
-})
-pl_Bmodule = lapply(1:length(pl_Bmodule),function(i){
-  if(i!=1){pl_Bmodule[[i]] + NoLegend()}else{pl_Bmodule[[i]]}})
-
-coldef = setNames(colorRampPalette(brewer.pal(12, "Paired"))(length(unique(cell_label))),unique(cell_label))
-p = DimPlot(subset(tiss,cells = colnames(tiss)[cell_id]), group.by = "cell_ontology_class",label = TRUE, cols = coldef,repel = TRUE,label.size = 3) + NoLegend() + labs(title = NULL)
-
-p = wrap_plots(wrap_plots(p) & theme(axis.title = element_blank(),axis.text = element_blank(), axis.ticks=element_blank(), axis.line = element_blank()),(wrap_plots(pl_Bmodule,nrow = 1) &  theme(plot.title = element_text(size = 15))) + plot_layout(guides = "collect"), nrow = 1, design = "ABBB")
-
-ggsave(filename = "fig-subtypeB.png", plot = p, width = 4*3, height = 3)
 ```
-
 Granulocyte: 
@@
 43: Stfa2l1, Il1b, Gm5483, H2-Q10, and Irg1
+```{r}
+# G_module = c(45,44,41,43)
+# G_module_genes = list("45" = c("Fcnb"),
+#                       "44" = c("Camp", "Ltf"),
+#                       "41" = c("Retnlg", "Cxcr2"),
+#                       "43" = c("Stfa2l1", "Il1b", "Gm5483", "H2-Q10", "Irg1"))
+modules_genes = list("8" = c("Ms4a3"),
+                      "15" = c("Camp", "Ltf"),
+                      "11" = c("Retnlg", "Cxcr2"),
+                      "10" = c("Stfa2l1", "Il1b", "Gm5483", "H2-Q10"))
+
+cell_id = which(tiss$cell_ontology_class %in% c("granulocyte","granulocytopoietic cell"))
+
+```
 
 ```{r}
-G_module = c(45,44,41,43)
-G_module_genes = list("45" = c("Fcnb"),
-                      "44" = c("Camp", "Ltf"),
-                      "41" = c("Retnlg", "Cxcr2"),
-                      "43" = c("Stfa2l1", "Il1b", "Gm5483", "H2-Q10", "Irg1"))
-cell_id = tiss$cell_ontology_class %in% c("granulocyte","granulocytopoietic cell")
+pl = lapply(names(modules_genes),function(module){
+  module_name = paste0("Module",module)
+  (FeaturePlot(subset(tiss, cells = colnames(tiss)[cell_id]), features = module_name, order = TRUE, reduction = "tsne") + 
+    scale_color_gradient(low = "lightgrey", high = "blue", limits = c(0,1)) + labs(color = "ModuleScore", title = paste0("Module",module), subtitle = paste0(modules_genes[[as.character(module)]],collapse = ", "))) & NoAxes() & theme(
+    plot.title = element_text(hjust = 0),
+    plot.subtitle = element_text(hjust = 0))
+})
 
+cell_label = tiss$cell_ontology_class
 coldef = setNames(colorRampPalette(brewer.pal(12, "Paired"))(length(unique(cell_label))),unique(cell_label))
-p = DimPlot(subset(tiss,cells = colnames(tiss)[cell_id]), group.by = "cell_ontology_class",label = TRUE,cols = coldef) + NoLegend() + labs(title = NULL)
+p = DimPlot(subset(tiss,cells = colnames(tiss)[cell_id]), group.by = "cell_ontology_class",label = TRUE, cols = coldef,repel = TRUE,label.size = 3) + NoLegend() + labs(title = NULL)
 
-pl_Gmodule = lapply(G_module,function(module){
-  FeaturePlot_custom(value = cell_block[cell_id,module], coord = visual_space[cell_id,], value_name = "Module Score",title_name = sprintf("Module %s",module)) + labs(subtitle = paste0(G_module_genes[[as.character(module)]],collapse = ","))
-})
-pl_Gmodule = lapply(1:length(pl_Gmodule),function(i){
-  if(i!=3){pl_Gmodule[[i]] + NoLegend()}else{pl_Gmodule[[i]]}})
+p = wrap_plots(wrap_plots(p) & theme(axis.title = element_blank(),axis.text = element_blank(), axis.ticks=element_blank(), axis.line = element_blank()),(wrap_plots(pl,nrow = 1) &  theme(plot.title = element_text(size = 15))) + plot_layout(guides = "collect"), nrow = 1, design = "ABBBB")
 
-p = wrap_plots(wrap_plots(p) & theme(axis.title = element_blank(),axis.text = element_blank(), axis.ticks=element_blank(), axis.line = element_blank()),(wrap_plots(pl_Gmodule,nrow = 1) &  theme(plot.title = element_text(size = 15))) + plot_layout(guides = "collect"), nrow = 1, design = "ABBBB")
+ggsave(filename = file.path(figure_path,"fig-subtypeB.png"), plot = p, width = 5*3, height = 3)
 
-ggsave(filename = "fig-subtypeG.png", plot = p, width = 5*3, height = 3)
+ggsave(filename = file.path(figure_path,"fig-subtypeG.png"), plot = p, width = 5*3, height = 3)
 ```
@@
 <!-- ``` -->
+
+# SupFig - SmoM2 Cell Embedding
+```{r}
+sample_ls = unlist(lapply(c("E13.5", "E14.5"),function(i) paste0(i,"_", c("MUT", "CTL"))))
+sample_ls = paste0("smom2_dermal_",sample_ls)
+sample_full_name = setNames(c("E13.5 SmoM2",
+                              "E13.5 WT",
+                              "E14.5 SmoM2",
+                              "E14.5 WT"),sample_ls)
+
+```
+```{r}
+dir.path = "/banach1/ruiqi/Peggy_data/Peggy_scdata/240329_smom2"
+lapply(sample_ls,function(sample_name){
+  assign(paste0("data_S_",sample_name), 
+         readRDS(file.path(dir.path,"process_data",sprintf("data_S_%s.rds",sample_name))),
+         envir = .GlobalEnv)
+  assign(paste0("local_gene_",sample_name), 
+         readRDS(file.path(dir.path,"intermediate_data",sprintf("local_gene_%s.rds",sample_name))),
+         envir = .GlobalEnv)
+  return(NULL)
+})
+```
+```{r}
+pl = lapply(sample_ls,function(sample_name){
+  data_S = get(paste0("data_S_",sample_name))
+  p1 = DimPlot(data_S, 
+          group.by = "celltype", shuffle = TRUE, cols = c("DC"="#d1495b", "UD" = "#edae49", "LD" = "#00798c")) + ggtitle(sample_full_name[sample_name]) & NoAxes()
+  p2 = DimPlot(data_S, 
+          group.by = "Phase", shuffle = TRUE) + ggtitle("") & NoAxes()
+  list(p1,p2)
+})
+pl = unlist(pl,recursive = FALSE)
+
+p = wrap_plots(pl,nrow=2,byrow = FALSE) + plot_layout(guides = "collect")
+
+ggsave(filename = file.path(figure_path,"supfig-smom2.png"), plot = p, width = 12, height = 6)
+```
+
+# SupFig - Module Activity Score for 4 SmoM2 samples
+```{r}
+sample_name = sample_ls[1]
+data_S = get(paste0("data_S_",sample_name))
+local_gene = get(paste0("local_gene_",sample_name))
+modules = levels(local_gene$gene_partition)
+pl_module = lapply(modules,function(module){
+  prefix = paste0("Module", sample_name, "_")
+  FeaturePlot(data_S, features = paste0(prefix,module), order = TRUE) + 
+    scale_color_gradient(low = "lightgrey", high = "blue", limits = c(0,1)) +
+    ggtitle(sprintf("Module %s (%d genes)", module, sum(local_gene$gene_partition == module)))
+})
+names(pl_module) = modules
+p = (wrap_plots(pl_module,ncol = 6) & 
+    labs(color = "ModuleScore") & 
+    NoAxes() & theme(plot.title = element_text(face="bold", size = 10),legend.title = element_text(size = 30),legend.text = element_text(size = 20)) ) + plot_layout(guides = "collect")
+
+ggsave(filename = file.path(figure_path,"supfig-moduletsne_e13mut.png"), plot = p, width = 6*3 + 1, height = 3*3)
+```
+
+# Fig - CC Module (SmoM2)
+```{r}
+module_ls = readRDS(file.path(dir.path,"intermediate_data","module_ls.rds"))
+module_ls_ref = lapply(module_ls,function(x) setNames(names(x),x)); names(module_ls_ref) = NULL
+module_ls_ref = do.call(c,module_ls_ref)
+```
+```{r}
+data_S_merged = readRDS(file = file.path(dir.path,"process_data","data_S_smom2_dermal_E13.5.rds"))
+```
+```{r}
+sample_name = sample_ls[1]
+modules = module_ls[[sample_name]]
+modules = modules[grepl("CC",names(modules))]
+local_gene = get(paste0("local_gene_",sample_name))
+pl = lapply(sample_ls[1:2],function(sample_name){
+  data_S = get(paste0("data_S_",sample_name))
+  pl_module = lapply(modules,function(module){
+    module_id = sub(".*?(\\d+)$", "\\1", module)
+    (FeaturePlot(data_S, features = module, order = TRUE) + 
+      scale_color_gradient(low = "lightgrey", high = "blue", limits = c(0,1)) +
+      ggtitle(sprintf("%s (%d genes)", module_ls_ref[module], sum(local_gene$gene_partition == module_id))) + labs(color = "ModuleScore")) & NoAxes()
+  })
+  names(pl_module) = modules
+  p = DimPlot(data_S, 
+          group.by = "Phase", shuffle = TRUE) + NoAxes()
+
+  c(list(p),pl_module)
+})
+pl_merged = lapply(unique(data_S_merged$type),function(sample_name){
+  data_S = subset(data_S_merged, type == sample_name)
+  pl_module = lapply(modules,function(module){
+    module_id = sub(".*?(\\d+)$", "\\1", module)
+    (FeaturePlot(data_S, features = module, order = TRUE, reduction = "umap_cc") + 
+      scale_color_gradient(low = "lightgrey", high = "blue", limits = c(0,1)) +
+      ggtitle(sprintf("%s (%d genes)", module_ls_ref[module], sum(local_gene$gene_partition == module_id))) + labs(color = "ModuleScore")) & NoAxes()
+  })
+  names(pl_module) = modules
+  p = DimPlot(data_S, 
+          group.by = "Phase", shuffle = TRUE, reduction = "umap_cc") + NoAxes()
+  c(list(p),pl_module)
+})
+
+pl = unlist(c(pl,pl_merged),recursive = FALSE)
+pl = lapply(1:length(pl),function(i){
+  if(i <= 4) pl[[i]]
+  else pl[[i]] + theme(plot.title = element_blank())
+})
+p = wrap_plots(pl,nrow = 4) + plot_layout(guides = "collect")
+ggsave(filename = file.path(figure_path, "fig-smom2_cc_module.png"), plot = p, width = 3*4 + 3, height = 3*4)
+```
+
+# SupFig - E13 merged embedding
+```{r}
+data_S_merged = readRDS(file = file.path(dir.path,"process_data","data_S_smom2_dermal_E13.5.rds"))
+
+p1 = (DimPlot(data_S_merged, group.by = "type", shuffle = TRUE) & NoAxes()) + ggtitle("By Top2000 HVGs") 
+p2 = (DimPlot(data_S_merged, group.by = "type", shuffle = TRUE, reduction = "umap_cc") & NoAxes()) + ggtitle("By 95 CC genes (Seurat)") 
+p = p1 + p2 + plot_layout(guides = "collect")
+ggsave(filename = file.path(figure_path, "supfig-smom2_E13_embed.png"), plot = p, width = 2*3+1, height = 3)
+```
+# Fig - Wnt Module (SmoM2)
+```{r}
+sample_name = sample_ls[1]
+modules = module_ls[[sample_name]]
+modules = modules[grepl("DC|Wnt",names(modules))]
+modules = modules[c(1,3,2)]
+local_gene = get(paste0("local_gene_",sample_name))
+pl = lapply(sample_ls[c(1,2,4)],function(sample_name){
+  data_S = get(paste0("data_S_",sample_name))
+  pl_module = lapply(modules,function(module){
+    module_id = sub(".*?(\\d+)$", "\\1", module)
+    (FeaturePlot(data_S, features = module, order = TRUE) + 
+      scale_color_gradient(low = "lightgrey", high = "blue", limits = c(0,1)) +
+      ggtitle(sprintf("%s (%d genes)", module_ls_ref[module], sum(local_gene$gene_partition == module_id))) + labs(color = "ModuleScore")) & NoAxes()
+  })
+  names(pl_module) = modules
+  pl_module
+})
+
+pl = unlist(pl,recursive = FALSE)
+pl = lapply(1:length(pl),function(i){
+  if(i <= 3) pl[[i]]
+  else pl[[i]] + theme(plot.title = element_blank())
+})
+p = wrap_plots(pl,nrow = 3) + plot_layout(guides = "collect")
+ggsave(filename = file.path(figure_path, "fig-smom2_wnt_module1.png"), plot = p, width = 3*3 + 3, height = 3*3)
+```
+## Lef1 level
+```{r}
+df = read.csv(file.path(dir.path,"intermediate_data","e13_smom2_wnt_level_table.csv"))
+gene = "Lef1"
+df$wnt_related_module = factor(df$wnt_related_module, levels = df %>% group_by(wnt_related_module) %>% summarise(Average_Expression = mean(get(gene), na.rm = TRUE)) %>% arrange(desc(Average_Expression)) %>% pull(wnt_related_module))
+
+# breaks <- seq(0, 1, length.out = 11)
+# labels <- paste0(head(breaks, -1), "~", tail(breaks, -1))
+# df <- df %>% mutate(across(starts_with("Module"), ~ cut(., breaks = breaks, include.lowest = TRUE, labels = labels)))
+# df_long <- df %>%
+#   pivot_longer(cols = starts_with("Module"),
+#                names_to = "Module", 
+#                values_to = "Bin")
+# df_long$Module = module_ls_ref[df_long$Module]
+# df_long$Bin = as.factor(df_long$Bin)
+# p = ggplot(df_long, aes(x = Bin, y = get(gene), color = Module)) +
+#   geom_violin(aes(color = Module), scale = "width", position = position_dodge(width = 0.75)) +
+#   geom_line(data = df_long %>%
+#   group_by(Module, Bin) %>%
+#   summarise(median_expression = median(get(gene)), .groups = 'drop'), aes(x = as.numeric(Bin), y = median_expression, group = Module, color = Module),position = position_dodge(width = 0.75)) +
+#   labs(title = paste0("Expression of ",gene),
+#        x = "ModuleScore (Binned)",
+#        y = "Expression Level",
+#        color = "Module") +
+#   theme_minimal()
+p = ggplot(df, aes(x = wnt_related_module, y = get(gene))) +
+  geom_violin(scale = "width") + geom_jitter() + 
+  ggsignif::geom_signif(comparisons = list(c("High-Wnt Module", "DC Module"), c("High-Wnt Module", "Wnt Module"), c("Wnt Module", "Other")),
+              map_signif_level = TRUE,
+              textsize = 3) + theme_minimal() +
+  labs(title = paste0("Expression of ",gene), x = "", y = "Expression Level")
+p = p + theme(
+    plot.title = element_text(size = 20),
+    axis.title = element_text(size = 15),
+    axis.text.x = element_text(size = 10,angle = 45),
+    axis.text.y = element_text(size = 10),
+    legend.title = element_text(size = 10),
+    legend.text = element_text(size = 10),
+    )
+```
+```{r}
+data_S = get(paste0("data_S_",sample_name))
+data_S$'Wnt.Module' = df$wnt_related_module
+p2 = DimPlot(data_S, 
+        group.by = "Wnt.Module", order = levels(data_S$'Wnt.Module'), 
+        cols = setNames(c(colorRampPalette(brewer.pal(3, "Paired"))(3),"lightgrey"),levels(data_S$'Wnt.Module'))) + ggtitle("Active Cells for Wnt-related module \n(ModuleScore > 0.4)") + NoAxes()
+
+ggsave(filename = file.path(figure_path, "fig-smom2_wnt_module2.png"), plot = p2 + p, width = 10, height = 5)
+```
+
+# Fig - Quiescent Module (SmoM2)
+```{r}
+modules = module_ls$smom2_dermal_E13.5_MUT["Quiescent Module"]
+pl = lapply(sample_ls, function(sample_name){
+  data_S = get(paste0("data_S_",sample_name))
+  if("umap_cc_regressout" %in% names(data_S@reductions)){
+    lapply(c("umap","umap_cc_regressout"),function(reduc){
+    p1 = DimPlot(data_S, 
+             group.by = "Phase", shuffle = TRUE, reduction = reduc) + ggtitle("") & NoAxes()
+    p2 = FeaturePlot(data_S, features = modules, order = TRUE, reduction = reduc) + ggtitle("") + NoAxes() + 
+    scale_color_gradient(low = "lightgrey", high = "blue", limits = c(0,1)) + labs(color = "ModuleScore")
+    if(reduc == "umap"){
+      cc_count = table(data_S$Phase[data_S@meta.data[,modules] > 0.5])
+      p2 = p2 + labs(subtitle = sprintf("%d Active Cells \n (%d in G1, %d in G2M, %d in S)",sum(cc_count), cc_count["G1"],cc_count["G2M"],cc_count["S"]))
+    }
+    list(p1,p2)
+  })
+  }else{
+    FeaturePlot(data_S, features = modules, order = TRUE, reduction = "umap") + ggtitle("") + NoAxes() + 
+    scale_color_gradient(low = "lightgrey", high = "blue", limits = c(0,1)) + labs(color = "ModuleScore")
+  }
+})
+pl = do.call(c,lapply(pl,function(p){
+  if(class(p) == "list"){
+    unlist(p,recursive = FALSE)
+  }else{
+    list(p)
+  }
+}))
+p = wrap_plots(pl,nrow = 2) + plot_layout(guides = "collect")
+ggsave(filename = file.path(figure_path, "fig-smom2_quiest_module1.png"), plot = p, width = 3*5+3, height = 2*3 + 2)
+```
+## Cdkn1a+ Lef1 level
+```{r}
+df = read.csv(file.path(dir.path,"intermediate_data","e13_smom2_cdkn1a_table.csv"))
+gene = "Lef1"
+df$status = factor(df$status, levels = df %>% group_by(status) %>% summarise(Average_Expression = mean(get(gene), na.rm = TRUE)) %>% arrange(desc(Average_Expression)) %>% pull(status))
+
+p = ggplot(df, aes(x = status, y = get(gene))) +
+  geom_violin(scale = "width") + geom_jitter() + 
+  ggsignif::geom_signif(comparisons = list(c("Cdkn1a+", "Quiescent Module"), c("Quiescent Module", "Cdkn1a-")),
+              map_signif_level = TRUE,
+              textsize = 3) + theme_minimal() +
+  labs(title = paste0("Expression of ",gene), x = "", y = "Expression Level")
+p = p + theme(
+    plot.title = element_text(size = 20),
+    axis.title = element_text(size = 15),
+    axis.text.x = element_text(size = 10,angle = 45),
+    axis.text.y = element_text(size = 10),
+    legend.title = element_text(size = 10),
+    legend.text = element_text(size = 10),
+    )
+```
+```{r}
+data_S = get(paste0("data_S_",sample_name))
+Idents(data_S) = df$status
+p2 = DimPlot(data_S, order = c("Quiescent Module","Cdkn1a+","Cdkn1a-"),
+        cols = setNames(c(hue_pal()(2),"lightgrey"),c("Quiescent Module","Cdkn1a+","Cdkn1a-"))) + ggtitle("Active Cells for Quiescent cells") + NoAxes()
+
+ggsave(filename = file.path(figure_path, "fig-smom2_quiest_module1.png"), plot = p2 + p, width = 10, height = 5)
+```
+
+# Supfig - Module-Colocalization
+```{r}
+df = read.csv(file = file.path(folder.path,"module_colocalize_align.csv"))
+pl = lapply(unique(df$sample),function(sample_name){
+  data_S = get(paste0("data_S_",sample_name))
+  df_sub = df %>% filter(sample == sample_name)
+  pl1 = lapply(1:nrow(df_sub),function(i){
+    module1 = df_sub[i,"module1"]
+    module2 = df_sub[i,"module2"]
+    ji = df_sub[i,"jaccard_index"]
+    p1 = FeaturePlot(data_S, features = module1, order = TRUE) + ggtitle(module_ls_ref[module1])
+    p2 = FeaturePlot(data_S, features = module2, order = TRUE) + ggtitle(sprintf("%s (JI: %.3f)",module_ls_ref[module2],ji))
+    (p1 / p2) & scale_color_gradient(low = "lightgrey", high = "blue", limits = c(0,1)) & NoAxes() & labs(color = "ModuleScore")
+  })
+})
+p = wrap_plots(c(pl[[1]][c(2:4,1)],pl[[2]][1],pl[[3]][c(1:3)]),nrow = 1) + plot_layout(guides = "collect")
+
+ggsave(filename = file.path(figure_path,"supfig-smom2_colocalized1.png"), plot = p, width = 25, height = 6)
+```
+
+# Pathway enrichment plot
+```{r}
+pathway_result = readRDS(file.path(dir.path,"intermediate_data","pathway_result.rds"))
+
+msig_df = do.call(rbind,lapply(sample_ls,function(sample_name){
+  msig_result = pathway_result[[sample_name]][["msig"]]
+  df <- do.call(rbind, lapply(names(msig_result), 
+                            function(i) data.frame(msig_result[[i]]@result[1:5,], module = i)))
+  df <- df[!is.na(df$Description), ]
+  df$GeneRatio = unlist(lapply(df$GeneRatio,function(x) eval(parse(text = x))))
+  df$module = paste0("Module",sample_name,"_",df$module)
+  df$'sample' = sample_full_name[sample_name]
+  df
+}))
+
+msig_df$module_name = module_ls_ref[msig_df$module]
+msig_df$module_class = ifelse(grepl("CC",msig_df$module_name),"CC", ifelse(grepl("DC|Wnt",msig_df$module_name),"DC-related", ifelse(grepl("Quiescent",msig_df$module_name),"Quiescent",NA)))
+```
+```{r}
+pl = lapply(c("CC","DC-related","Quiescent"),function(c){
+  ggplot(msig_df %>% filter(module_class == c) %>% filter(p.adjust < 0.2), aes(x=module_name, y=Description)) +
+    geom_point(aes(size=GeneRatio, color=p.adjust), alpha=0.7) +
+    scale_color_gradientn(colors=colorRampPalette(c("red", "blue"))(99), name="p.adjust",limits = c(0,0.2)) + scale_size_continuous(name = "Gene Ratio", limits = c(0, 1)) + 
+    labs(size="Gene Ratio",x = "Module") + facet_grid(~sample,switch = "x",scales = "free") + ggtitle(c)
+})
+
+p = (wrap_plots(pl, nrow = 1) & theme_minimal() & 
+    theme(legend.position="right",
+          plot.title = element_text(size = 20),
+          axis.title = element_text(size = 10),
+          axis.text.y = element_text(size = 10),
+          legend.title = element_text(size = 10),
+          legend.text = element_text(size = 10),
+    panel.spacing.x = unit(0,"line"),
+      axis.text.x = element_text(angle = 45),
+          strip.placement = 'outside',
+          strip.background.x = element_blank()) ) + plot_layout(guides = "collect")
+
+ggsave(filename = file.path(figure_path,"supfig-smom2_colocalized2.png"), plot = p, width = 20, height = 5)
+```
+
+# SupsupFig - Boxplot of ModuleActivityScore of active cells
+```{r}
+nn_score_df = read.csv(file.path(dir.path,"intermediate_data","neighbors_module_score.csv"))
+nn_score_df$module = factor(nn_score_df$module,levels = c("Wnt Module","High-Wnt Module","DC Module","Quiescent Module"))
+nn_score_df$sample = factor(nn_score_df$sample, levels = unique(nn_score_df$sample))
+levels(nn_score_df$sample) = c("E13.5 SmoM2", "E13.5 WT", "E14.5 SmoM2", "E14.5 WT")
+nn_score_df$knn = as.factor(nn_score_df$knn)
+
+p = ggplot(nn_score_df %>% 
+         filter(!(module!="Quiescent Module" & sample == "E14.5 SmoM2")), aes(x = knn, y = value, color = sample)) + 
+  geom_boxplot() + 
+  labs(y = "Average ModuleActivityScore", x = "KNN") +
+  theme_minimal() + ylim(0,1) + facet_wrap(~module, ncol = 3)
+ggsave(filename = file.path(figure_path,"supsupfig-smom2_neighbor_modulescore.png"), plot = p, width = 6, height = 4)
+```

### 2024-05-23 ATAC_Cite_spatial.R

@@ -1,170 +1,0 @@
-library(Seurat)
-library(SeuratData)
-library(SeuratWrappers)
-library(cowplot)
-library(dplyr)
-source("/data/ruiqi/local_marker/LocalMarkerDetector/LMD_function.R", echo = F)
-
-# CITE-seq: 25 proteins ----------
-InstallData("bmcite")
-bm <- LoadData(ds = "bmcite")
-DefaultAssay(bm) <- 'ADT'
-VariableFeatures(bm) <- rownames(bm[["ADT"]])
-bm <- NormalizeData(bm, normalization.method = 'CLR', margin = 2) %>% 
-  ScaleData() %>% RunPCA(reduction.name = 'apca')
-bm <- RunUMAP(bm, reduction = 'apca', dims = 1:18, assay = 'ADT', 
-              reduction.name = 'adt.umap', reduction.key = 'adtUMAP_')
-DimPlot(bm, reduction = 'adt.umap', group.by = 'celltype.l2', label = TRUE, 
-        repel = TRUE, label.size = 2.5) + NoLegend()
-FeaturePlot(bm, features = c("adt_CD45RA","adt_CD16","adt_CD161"),
-            reduction = 'adt.umap', max.cutoff = 2, ncol = 3)
-
-# ATAC-seq ----------------
-library(SeuratData)
-library(Seurat)
-library(Signac)
-library(EnsDb.Hsapiens.v86)
-library(ggplot2)
-library(cowplot)
-# BiocManager::install("Bioconductor/GenomeInfoDb")
-
-## Load Data & QC -------
-# InstallData("pbmcMultiome")
-pbmc.atac <- LoadData("pbmcMultiome", "pbmc.atac")
-# QC
-pbmc.atac <- subset(pbmc.atac, seurat_annotations != "filtered" & nCount_ATAC < 7e4 & nCount_ATAC > 5e3)
-# Extract_t_cells
-annotated_cell_types = names(table(pbmc.atac$seurat_annotations))
-sub.pbmc.atac <- subset(pbmc.atac, seurat_annotations %in% annotated_cell_types[!grepl("B|HSPC|Mono|Plasma|DC|NK",annotated_cell_types)])
-
-## Process -------
-tiss <- sub.pbmc.atac
-annotations <- GetGRangesFromEnsDb(ensdb = EnsDb.Hsapiens.v86)
-seqlevelsStyle(annotations) <- "UCSC"
[…]

### 2024-05-23 Demo_smom2.R

@@ -1,428 +1,0 @@
-# https://nbisweden.github.io/workshop-scRNAseq/labs/compiled/seurat/seurat_01_qc.html#Predict_doublets
-#' How to define how many PCs based on Elbow plot
-# https://hbctraining.github.io/scRNA-seq/lessons/elbow_plot_metric.html
-#' The point where the principal components only contribute 5% of standard deviation and the principal components cumulatively contribute 90% of the standard deviation.
-#' The point where the percent change in variation between the consecutive PCs is less than 0.1%.
-
-library(dplyr)
-library(Seurat)
-library(patchwork)
-library(ggplot2)
-library(cowplot)
-library(SeuratWrappers)
-library(RColorBrewer)
-require(scales)
-source("/data/ruiqi/local_marker/LocalMarkerDetector/LMD_function.R", echo = F)
-setwd("/banach1/ruiqi/Peggy_scdata/tmp")
-
-FindPC = function(srat){
-  stdv <- srat[["pca"]]@stdev
-  sum.stdv <- sum(srat[["pca"]]@stdev)
-  percent.stdv <- (stdv / sum.stdv) * 100
-  cumulative <- cumsum(percent.stdv)
-  co1 <- which(cumulative > 90 & percent.stdv < 5)[1]
-  co2 <- sort(which((percent.stdv[1:length(percent.stdv) - 1] - 
-                       percent.stdv[2:length(percent.stdv)]) > 0.1), 
-              decreasing = T)[1] + 1
-  min.pc <- min(co1, co2)
-  return(min.pc)
-}
-
-# Load Data --------------------
-# /data/rq25/peggy_10x_mouse_smom2_202112/seurat_analysis/result/data/
-sample_ls = apply(expand.grid(c("E13.5", "E14.5"), c("MUT", "CTL"), stringsAsFactors = FALSE),1,function(x) paste0(x,collapse = "_"))
-sample_ls = paste0("smom2_dermal_",sample_ls[c(1,3,2,4)])
-for(i in sample_ls){
-  assign(paste0("data_S_",i), 
-         readRDS(sprintf("/data/rq25/peggy_10x_mouse_smom2_202112/seurat_analysis/result/data/data_S_peggy_smom2_dermal_%s.rds",i)))
-}
-# Preprocess ------------
-for(i in length(sample_ls)){
-  data_S <- get(paste0("data_S_",sample_ls[i]))
-  data_S <- data_S %>% NormalizeData() %>% FindVariableFeatures() %>%
[…]

### 2024-05-23 LMD_function.R

@@ -1,8 +1,9 @@
-cran_packages <- c("igraph","ggplot2", "cowplot", "RColorBrewer", 
-                   "data.table", "dplyr", "patchwork", 
+cran_packages <- c("igraph","ggplot2", "cowplot", "RColorBrewer", "scales",
+                   "data.table", "dplyr", "patchwork", "philentropy",
                    "pheatmap", "ggplotify", "ggraph", 
                    "ClusterR", "Rcpp", "RcppArmadillo", 
-                   "tictoc","svMisc","RSpectra","Matrix","progress","latex2exp","parallel")
+                   "tictoc","svMisc","RSpectra","Matrix","usedist","dendextend",
+                   "progress","latex2exp","parallel","dbscan","Rtsne","ggrepel", "BiocManager", "remotes")
 bioc_packages <- c("Seurat")
 sapply(cran_packages, function(pkg) if(!requireNamespace(pkg, quietly = TRUE)){install.packages(pkg)})
 sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
@@ -54,6 +55,38 @@ NumericVector fastKLVector(NumericMatrix x, NumericVector y) {
   return out;
 }
 ')
+add_newline_in_middle <- function(input_string, max_length) {
+  if (nchar(input_string) > max_length) {
+    # Find the approximate middle
+    middle <- ceiling(nchar(input_string) / 2)
+    
+    # Find the closest space to the middle
+    space_before <- regexpr(" ", substring(input_string, 1, middle), fixed = TRUE)
+    space_after <- regexpr(" ", substring(input_string, middle + 1), fixed = TRUE)
+    
+    if (space_before == -1 && space_after == -1) {
+      # No spaces found, cannot split without breaking words
+      return(input_string)
+    }
+    
+    if (space_before == -1) {
+      split_pos <- middle + space_after
+    } else if (space_after == -1) {
+      split_pos <- space_before
+    } else {
+      split_pos <- ifelse(middle - space_before < space_after, space_before, middle + space_after)
+    }
+    
+    # Insert newline at the split position
+    output_string <- paste0(
+      substring(input_string, 1, split_pos), "\n", substring(input_string, split_pos + 1)
+    )
+  } else {
+    output_string <- input_string
+  }
+  
+  return(output_string)
+}
[…]

### 2024-05-23 Marker_identfication_benchmark_azimuth_data.R

@@ -1,289 +1,0 @@
-library("Seurat")
-library("dplyr")
-library(RColorBrewer)
-library(reticulate)
-library(anndata)
-
-setwd("/data/ruiqi/local_marker/dataset/azimuth")
-# Load Data ----------
-tissue_download_ls = c(
-  "human_motor_cortex" = "https://seurat.nygenome.org/azimuth/demo_datasets/allen_m1c_2019_ssv4.rds",
-  "human_kidney" = "https://seurat.nygenome.org/azimuth/demo_datasets/kidney_demo_stewart.rds",
-  "human_bone_marrow" = "https://seurat.nygenome.org/azimuth/demo_datasets/bmcite_demo.rds",
-  "human_pancreas" = "https://seurat.nygenome.org/azimuth/demo_datasets/enge.rds",
-  "mouse_motor_cortex" = "https://seurat.nygenome.org/azimuth/demo_datasets/allen_mop_2020.rds",
-  "human_lung_hlca" = "https://seurat.nygenome.org/hlca_ref_files/ts_opt.rds"
-)
-for(tissue_name in names(tissue_download_ls)){
-  cat(tissue_name,"\n")
-folder_path = "/data/ruiqi/local_marker/dataset/azimuth"
-tissue_name = paste0(tissue_name,"_azi_demo")
-
-file_name = paste0(tissue_name,".rds")
-options(timeout=6000)
-if (!file.exists(file.path(folder_path,file_name))) {
-  options(timeout=6000)
-  download.file(tissue_download_ls[tissue_name],
-                destfile = file.path(folder_path,file_name), method = 'libcurl')
-}
-tiss <- readRDS(file.path(folder_path,file_name))
-
-# Preprocess ----------------
-# Annotate based on Azimuth (download result from shiny)
-folder_path = "/data/ruiqi/local_marker/dataset/azimuth/azimuth_annotate_result"
-predictions <- read.delim(file.path(folder_path,paste0(tissue_name,"_azimuth_pred.tsv")), row.names = 1)
-tiss <- AddMetaData(
-  object = tiss,
-  metadata = predictions)
-projected.umap <- readRDS(file.path(folder_path,paste0(tissue_name,"_azimuth_umap.Rds")))
-tiss <- tiss[, Cells(projected.umap)]
-tiss[['umap.proj']] <- projected.umap
-
-annotation_group = colnames(predictions)[1]
-meta.data = tiss@meta.data[,annotation_group]
[…]

### 2024-05-23 benchmark/compare_method_azimuth.R

@@ -1,0 +1,199 @@
+cran_packages <- c("readxl","dplyr","BiocManager","RColorBrewer")
+bioc_packages <- c("Seurat")
+sapply(cran_packages, function(pkg) if(!requireNamespace(pkg, quietly = TRUE)){install.packages(pkg)})
+sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
+lapply(c(cran_packages,bioc_packages), require, character.only = TRUE)
+
+# Download/Load Data ===========
+dir.path <- "/banach1/ruiqi/local_marker/LMD_data"
+consortium = "azimuth"
+folder.path <- file.path(dir.path,consortium)
+dir.create(folder.path, recursive=T)
+tissue_download_link = c(
+  "human_kidney" = "https://seurat.nygenome.org/azimuth/demo_datasets/kidney_demo_stewart.rds",
+  "human_bone_marrow" = "https://seurat.nygenome.org/azimuth/demo_datasets/bmcite_demo.rds",
+  "human_pancreas" = "https://seurat.nygenome.org/azimuth/demo_datasets/enge.rds",
+  "mouse_motor_cortex" = "https://seurat.nygenome.org/azimuth/demo_datasets/allen_mop_2020.rds",
+  "human_lung_hlca" = "https://seurat.nygenome.org/hlca_ref_files/ts_opt.rds",
+  "human_motor_cortex" = "https://seurat.nygenome.org/azimuth/demo_datasets/allen_m1c_2019_ssv4.rds"
+)
+for(tissue_name in names(tissue_download_link)){
+  file_name = paste0(tissue_name,".rds")
+  if (!file.exists(file.path(folder.path,file_name))) {
+    options(timeout=6000)
+    download.file(tissue_download_link[tissue_name],
+                  destfile = file.path(folder.path,file_name), method = 'libcurl')
+    # update version
+    tiss <- readRDS(file.path(folder.path,file_name))
+    tiss <- UpdateSeuratObject(tiss)
+    saveRDS(tiss,file = file.path(folder.path,file_name))
+  }
+}
+
+# Prepare input data ===========
+consortium = "azimuth"
+tissue_name = names(tissue_download_link)[1]
+
+for(tissue_name in names(tissue_download_link)){
+dir.path <- "/banach1/ruiqi/local_marker/LMD_data"
+folder.path <- file.path(dir.path,consortium)
[…]

### 2024-05-23 benchmark/compare_method_tabula_muris.R

@@ -1,0 +1,259 @@
+cran_packages <- c("readxl","dplyr","BiocManager","RColorBrewer")
+bioc_packages <- c("Seurat")
+sapply(cran_packages, function(pkg) if(!requireNamespace(pkg, quietly = TRUE)){install.packages(pkg)})
+sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
+lapply(c(cran_packages,bioc_packages), require, character.only = TRUE)
+
+# Download/Load Data ===========
+dir.path <- "/banach1/ruiqi/local_marker"
+consortium = "tabular_muris"
+folder.path <- file.path(dir.path,"LMD_data",consortium)
+dir.create(folder.path, recursive=T)
+tissue_download_link = c(
+  "marrow_facs" = "https://figshare.com/ndownloader/files/13092380",
+  "pancreas_facs" = "https://figshare.com/ndownloader/files/13092386",
+  "lung_facs" = "https://figshare.com/ndownloader/files/13092194",
+  "marrow_droplet" = "https://figshare.com/ndownloader/files/ 13089821"
+)
+for(tissue_name in names(tissue_download_link)){
+  file_name = paste0(tissue_name,".rds")
+  if (!file.exists(file.path(folder.path,file_name))) {
+    options(timeout=6000)
+    download.file(tissue_download_link[tissue_name],
+                  destfile = file.path(folder.path,file_name), method = 'libcurl')
+    # update version
+    tiss <- readRDS(file.path(folder.path,file_name))
+    tiss <- UpdateSeuratObject(tiss)
+    saveRDS(tiss,file = file.path(folder.path,file_name))
+  }
+}
+
+# Prepare input data ===========
+consortium = "tabular_muris"
+tissue_name = names(tissue_download_link)[1]
+
+dir.path <- "/banach1/ruiqi/local_marker"
+folder.path <- file.path(dir.path,"LMD_data",consortium)
+tiss <- readRDS(file.path(folder.path,paste0(tissue_name,".rds")))
+DefaultAssay(tiss) <- "RNA"
+n_dim = 20
+feature_space = Embeddings(tiss[["pca"]])[,1:n_dim]
+dat = as.matrix(tiss[[DefaultAssay(tiss)]]@data)
+Gene_detected_count <- apply(dat > apply(dat,2,median),1,sum)
[…]

### 2024-05-23 benchmark/run_methods_function.R

@@ -1,0 +1,206 @@
+# Install Marcopolo: pip install marcopolo-pytorch --upgrade
+# Install SEMITONES: pip install https://github.com/ohlerlab/SEMITONES/archive/master.zip
+# Install Hotspot: pip install git+https://github.com/yoseflab/Hotspot.git
+# Install singleCellHaystack: install.packages("singleCellHaystack")
+
+cran_packages <- c("reticulate","singleCellHaystack")
+github_packages <- c("immunogenomics/presto")
+sapply(cran_packages, function(pkg) if(!requireNamespace(pkg, quietly = TRUE)){install.packages(pkg)})
+# sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
+sapply(github_packages, function(pkg) if (!requireNamespace(strsplit(pkg, "/")[[1]][2], quietly = TRUE)) remotes::install_github(pkg))
+lapply(c(cran_packages), require, character.only = TRUE)
+
+#' LMD: Genes were ranked by the increasing order of Cumulative Diffusion-KL score
+dir.path <- "/banach1/ruiqi/local_marker/LocalMarkerDetector"
+source(file.path(dir.path,"LMD_function.R"))
+RunLMD <- function(dat, feature_space, dir.file){
+  res = LMD(dat,feature_space,max_time = 2^20,knn = 5)
+  marker = show_result_lmd(res)$'gene_table'
+  write.table(marker,file = dir.file,row.names = TRUE)
+}
+
+#' HVG
+RunHVG <- function(srat_obj, input_genes, dir.file){
+  DefaultAssay(srat_obj) <- "RNA"
+  marker <- subset(srat_obj,features = input_genes) %>% ScaleData(features = input_genes) %>% 
+    FindVariableFeatures(nfeatures = length(input_genes)) %>% VariableFeatures() %>% as.data.frame() %>% setNames(c("gene")) %>% mutate(rank = 1:n())
+  write.table(marker,file = dir.file,row.names = FALSE)
+}
+
+#' SinglecellHaystack: 
+#' https://alexisvdb.github.io/singleCellHaystack/articles/examples/a02_example_scRNAseq.html
+#' Genes were ranked based on the increasing order of log(adjusted P-value)
+RunHaystack <- function(dat, feature_space, dir.file){
+  set.seed(123)
+  res <- haystack(x = feature_space, expression = dat)
[…]

### 2024-05-23 benchmark/sensitivity_test.R

@@ -1,0 +1,92 @@
+cran_packages <- c("readxl","dplyr","BiocManager")
+bioc_packages <- c("Seurat")
+sapply(cran_packages, function(pkg) if(!requireNamespace(pkg, quietly = TRUE)){install.packages(pkg)})
+sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
+lapply(c(cran_packages,bioc_packages), require, character.only = TRUE)
+
+# Load Data =======
+dir.path <- "/banach1/ruiqi/local_marker/"
+consortium = "tabular_muris"
+tissue_name = "marrow_facs"
+tiss <- readRDS(file.path(dir.path,"LMD_data",consortium,paste0(tissue_name,".rds")))
+DefaultAssay(tiss) <- "RNA"
+dat = as.matrix(tiss[[DefaultAssay(tiss)]]@data)
+Gene_detected_count <- apply(dat > apply(dat,2,median),1,sum)
+selected_genes = (Gene_detected_count >= 10) & (Gene_detected_count <= ncol(dat) * 0.5)
+selected_genes = names(selected_genes)[selected_genes]
+dat = dat[selected_genes,,drop = FALSE]
+
+
+# kNN sensitivity Test --------
+folder.path = file.path(dir.path,"LMD_data",consortium,"intermediate_data",
+                        tissue_name, "sensitivity_test")
+dir.create(folder.path,recursive = T)
+source(file.path(dir.path,"LocalMarkerDetector","LMD_function.R"))
+# Use Default PCs
+n_dim = 20
+feature_space = Embeddings(tiss[["pca"]])[,1:n_dim]
+score_df = do.call(cbind,lapply(c(3:10,20,30,40,50),function(kNN){
+  try({
+    res = LMD(dat,feature_space,knn = kNN)
+    return(res$cumulative_score)
+  },silent = TRUE)
+}))
+colnames(score_df) = paste0("kNN",c(3:10,20,30,40,50))
+write.table(score_df,file = file.path(folder.path,paste0(tissue_name,"_different_knn.csv")),row.names = TRUE)
+
+# PC sensitivity Test --------
+folder.path = file.path(dir.path,"LMD_data",consortium,"intermediate_data",
+                        tissue_name, "sensitivity_test")
+source(file.path(dir.path,"LocalMarkerDetector","LMD_function.R"))
[…]

### 2024-05-23 miscellaneous/miscellaneous_code/ATAC_Cite_spatial.R

@@ -1,0 +1,170 @@
+library(Seurat)
+library(SeuratData)
+library(SeuratWrappers)
+library(cowplot)
+library(dplyr)
+source("/banach1/ruiqi/local_marker/LocalMarkerDetector/LMD_function.R", echo = F)
+
+# CITE-seq: 25 proteins ----------
+InstallData("bmcite")
+bm <- LoadData(ds = "bmcite")
+DefaultAssay(bm) <- 'ADT'
+VariableFeatures(bm) <- rownames(bm[["ADT"]])
+bm <- NormalizeData(bm, normalization.method = 'CLR', margin = 2) %>% 
+  ScaleData() %>% RunPCA(reduction.name = 'apca')
+bm <- RunUMAP(bm, reduction = 'apca', dims = 1:18, assay = 'ADT', 
+              reduction.name = 'adt.umap', reduction.key = 'adtUMAP_')
+DimPlot(bm, reduction = 'adt.umap', group.by = 'celltype.l2', label = TRUE, 
+        repel = TRUE, label.size = 2.5) + NoLegend()
+FeaturePlot(bm, features = c("adt_CD45RA","adt_CD16","adt_CD161"),
+            reduction = 'adt.umap', max.cutoff = 2, ncol = 3)
+
+# ATAC-seq ----------------
+library(SeuratData)
+library(Seurat)
+library(Signac)
+library(EnsDb.Hsapiens.v86)
+library(ggplot2)
+library(cowplot)
+# BiocManager::install("Bioconductor/GenomeInfoDb")
+
+## Load Data & QC -------
+# InstallData("pbmcMultiome")
+pbmc.atac <- LoadData("pbmcMultiome", "pbmc.atac")
+# QC
+pbmc.atac <- subset(pbmc.atac, seurat_annotations != "filtered" & nCount_ATAC < 7e4 & nCount_ATAC > 5e3)
+# Extract_t_cells
+annotated_cell_types = names(table(pbmc.atac$seurat_annotations))
+sub.pbmc.atac <- subset(pbmc.atac, seurat_annotations %in% annotated_cell_types[!grepl("B|HSPC|Mono|Plasma|DC|NK",annotated_cell_types)])
+
+## Process -------
+tiss <- sub.pbmc.atac
+annotations <- GetGRangesFromEnsDb(ensdb = EnsDb.Hsapiens.v86)
[…]

### 2024-05-23 miscellaneous/miscellaneous_code/Demo_smom2.R

@@ -1,0 +1,545 @@
+# https://nbisweden.github.io/workshop-scRNAseq/labs/compiled/seurat/seurat_01_qc.html#Predict_doublets
+#' How to define how many PCs based on Elbow plot
+# https://hbctraining.github.io/scRNA-seq/lessons/elbow_plot_metric.html
+#' The point where the principal components only contribute 5% of standard deviation and the principal components cumulatively contribute 90% of the standard deviation.
+#' The point where the percent change in variation between the consecutive PCs is less than 0.1%.
+
+library(dplyr)
+library(Seurat)
+library(patchwork)
+library(ggplot2)
+library(cowplot)
+library(SeuratWrappers)
+library(RColorBrewer)
+require(scales)
+source("/banach1/ruiqi/local_marker/LocalMarkerDetector/LMD_function.R", echo = F)
+setwd("/banach1/ruiqi/Peggy_data/Peggy_scdata/tmp")
+
+FindPC = function(srat){
+  stdv <- srat[["pca"]]@stdev
+  sum.stdv <- sum(srat[["pca"]]@stdev)
+  percent.stdv <- (stdv / sum.stdv) * 100
+  cumulative <- cumsum(percent.stdv)
+  co1 <- which(cumulative > 90 & percent.stdv < 5)[1]
+  co2 <- sort(which((percent.stdv[1:length(percent.stdv) - 1] - 
+                       percent.stdv[2:length(percent.stdv)]) > 0.1), 
+              decreasing = T)[1] + 1
+  min.pc <- min(co1, co2)
+  return(min.pc)
+}
+
+# Load Data --------------------
+# /data/rq25/peggy_10x_mouse_smom2_202112/seurat_analysis/result/data/
+sample_ls = apply(expand.grid(c("E13.5", "E14.5"), c("MUT", "CTL"), stringsAsFactors = FALSE),1,function(x) paste0(x,collapse = "_"))
+sample_ls = paste0("smom2_dermal_",sample_ls[c(1,3,2,4)])
+for(i in sample_ls){
+  assign(paste0("data_S_",i), 
+         readRDS(sprintf("/data/rq25/peggy_10x_mouse_smom2_202112/seurat_analysis/result/data/data_S_peggy_%s.rds",i)))
+}
+# Preprocess ------------
+folder_path = "/banach1/ruiqi/Peggy_data/Peggy_scdata/240329_smom2"
+for(sample_name in sample_ls){
[…]

### 2024-05-23 mouse_bone_marrow_example.R

@@ -1,0 +1,251 @@
+dir.path <- "/banach1/ruiqi/local_marker/LocalMarkerDetector/"
+source(file.path(dir.path,"LMD_function.R"))
+# Define package lists
+bioc_packages <- c("clusterProfiler", "AnnotationDbi", "ReactomePA", "org.Mm.eg.db","gprofiler2", "msigdbr")
+github_packages <- c("satijalab/seurat-wrappers");
+names(github_packages) = strsplit(github_packages, "/")[[1]][2]
+names(github_packages)[grepl("seurat-wrappers",names(github_packages))] = "SeuratWrappers"
+# Install packages
+sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
+sapply(1:length(github_packages), function(i) if (!requireNamespace(names(github_packages)[i], quietly = TRUE)) remotes::install_github(github_packages[i]))
+lapply(c(bioc_packages,names(github_packages)), require, character.only = TRUE)
+
+# Download/Load Data ===========
+dir.path <- "/banach1/ruiqi/local_marker/LMD_data"
+folder.path <- file.path(dir.path,"tabular_muris")
+dir.create(folder.path, recursive=T)
+tissue_download_link = c(
+  "marrow_facs" = "https://figshare.com/ndownloader/files/13092380",
+  "marrow_droplet" = "https://figshare.com/ndownloader/files/ 13089821"
+)
+for(tissue_name in names(tissue_download_link)){
+  file_name = paste0(tissue_name,".rds")
+  if (!file.exists(file.path(folder.path,file_name))) {
+    options(timeout=6000)
+    download.file(tissue_download_link[tissue_name],
+                  destfile = file.path(folder.path,file_name), method = 'libcurl')
+    # update version
+    tiss <- readRDS(file.path(folder.path,file_name))
+    tiss <- UpdateSeuratObject(tiss)
+    tiss <- RunALRA(tiss, assay = "RNA"); DefaultAssay(tiss) = "RNA"
+    saveRDS(tiss,file = file.path(folder.path,file_name))
+  }
+}
+
+# Prepare input data ===========
+tissue_name = "marrow_facs"
+tiss <- readRDS(file.path(folder.path,paste0(tissue_name,".rds")))
+DefaultAssay(tiss) <- "RNA"
+n_dim = 20
[…]

### 2024-05-23 mouse_smoM2_example.R

@@ -1,0 +1,382 @@
+dir.path <- "/banach1/ruiqi/local_marker/LocalMarkerDetector/"
+source(file.path(dir.path,"LMD_function.R"))
+
+# Define package lists
+bioc_packages <- c("clusterProfiler", "AnnotationDbi", "ReactomePA", "org.Mm.eg.db",
+                   "gprofiler2", "msigdbr")
+github_packages <- c("satijalab/seurat-wrappers");
+names(github_packages) = strsplit(github_packages, "/")[[1]][2]
+names(github_packages)[grepl("seurat-wrappers",names(github_packages))] = "SeuratWrappers"
+# Install packages
+sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
+sapply(1:length(github_packages), function(i) if (!requireNamespace(names(github_packages)[i], quietly = TRUE)) remotes::install_github(github_packages[i]))
+lapply(c(bioc_packages,names(github_packages)), require, character.only = TRUE)
+
+# Load Data --------------------
+sample_ls = unlist(lapply(c("E13.5", "E14.5"),function(i) paste0(i,"_", c("MUT", "CTL"))))
+sample_ls = paste0("smom2_dermal_",sample_ls)
+for(i in sample_ls){
+  assign(paste0("data_S_",i), 
+         readRDS(sprintf("/data/rq25/peggy_10x_mouse_smom2_202112/seurat_analysis/result/data/data_S_peggy_%s.rds",i)))
+}
+# Preprocess ------------
+dir.path = "/banach1/ruiqi/Peggy_data/Peggy_scdata/240329_smom2"
+folder.path <- file.path(dir.path,"process_data")
+dir.create(folder.path, recursive=T)
+lapply(sample_ls, function(sample_name){
+  data_S <- get(paste0("data_S_",sample_name))
+  data_S <- data_S %>% NormalizeData() %>% FindVariableFeatures() %>%
+    ScaleData(verbose = FALSE) %>% RunPCA(npcs = 50, verbose = FALSE)
+  min.pc = FindPC(srat = data_S)
+  data_S <- RunUMAP(data_S, dims = 1:min.pc, seed.use = 42)
+  data_S <- data_S %>% FindNeighbors() %>% FindClusters(res = 1.5)
+  data_S <- RunALRA(data_S); DefaultAssay(data_S) <- "RNA"
[…]

### 2024-06-13 paper_figure.Rmd
06/13

@@
 library("ggsignif")
-
+source("LMD_function.R", echo = F)
 figure_path = "/banach1/ruiqi/local_marker/LMD_data/figures"
@@
 
-# SupFig - Bone Marrow Facs Cell Embedding
+# SupFig - Bone Marrow Cell Embedding
 ```{r}
@@
 coldef = setNames(colorRampPalette(brewer.pal(12, "Paired"))(length(unique(cell_label))),unique(cell_label))
-p = DimPlot(tiss, group.by = "cell_ontology_class",label = FALSE, cols = coldef) + guides(color = guide_legend(ncol = 1,override.aes = list(size = 3))) + labs(title = NULL)
-
-ggsave(filename = file.path(figure_path,"supfig-bonemarrow_facs.png"), plot = p, width = 8, height = 5, dpi = 300)
-```
-
-# SupFig - Bone Marrow droplet Cell Embedding
-```{r}
-dir.path <- "/banach1/ruiqi/local_marker/LMD_data"
-consortium = "tabular_muris"
-folder.path <- file.path(dir.path,consortium)
+p1 = DimPlot(tiss, group.by = "cell_ontology_class",label = FALSE, cols = coldef) + guides(color = guide_legend(ncol = 1,override.aes = list(size = 3))) + labs(title = NULL)
+
 tissue_name = "marrow_droplet"
@@
 coldef = setNames(colorRampPalette(brewer.pal(12, "Paired"))(length(unique(cell_label))),unique(cell_label))
-p = DimPlot(tiss, group.by = "cell_ontology_class",label = FALSE, cols = coldef) + guides(color = guide_legend(ncol = 1,override.aes = list(size = 3))) + labs(title = NULL)
-
-ggsave(filename = file.path(figure_path,"supfig-bonemarrow_droplet.png"), plot = p, width = 8, height = 5, dpi = 300)
+p2 = DimPlot(tiss, group.by = "cell_ontology_class",label = FALSE, cols = coldef) + guides(color = guide_legend(ncol = 1,override.aes = list(size = 3))) + labs(title = NULL)
+
+p = ((p1 / p2) + plot_annotation(tag_levels = 'A')) & theme(plot.tag = element_text(size = 20))
+ggsave(filename = file.path(figure_path,"supfig-bonemarrow.png"), plot = p, width = 8, height = 10, dpi = 300)
 ```
@@
 tissue_name = "marrow_facs"
-folder.path = file.path(dir.path, "LocalMarkerDetector","benchmark")
+folder.path = file.path(dir.path, "LMD_data","benchmark")
 ```
@@
 wide_data$Tissue = gsub("Human-|Mouse-","",wide_data$Tissue)
+# calculate average
+avg_tmp = wide_data %>%
+    group_by(Criterion) %>%
+    summarize(across(3:(ncol(wide_data)-1), ~ round(mean(.x, na.rm = TRUE), 3)))
+avg_tmp$DataSource = c("Average",""); avg_tmp$Tissue = "-"
+wide_data = rbind(wide_data,avg_tmp)
 bold_max <- function(vec) {
@@
 ```
+## Figure
 ```{r}
@@
 ggsave(filename = file.path(figure_path,"supfig-runtime.png"), plot = fig, width = 8, height = 5)
+```
+## Table
+```{r}
+df_runtime = df_runtime %>% filter(Method == "LMD") %>% arrange(Time, nCells, nGenes) %>% select(c("Time","nCells","nGenes")) %>% mutate(Time = round(Time, 2))
+colnames(df_runtime)[1] = "Time (min)"
+write.csv(df_runtime, file = file.path(folder.path,"RunTime_table.csv"), row.names = FALSE)
 ```
@@
 consortium = "tabular_muris"
-dir.path <- "/banach1/ruiqi/local_marker/LocalMarkerDetector"
+dir.path <- "/banach1/ruiqi/local_marker/LMD_data"
 pl = lapply(c("marrow_facs","lung_facs","pancreas_facs"),function(tissue_name){
@@
 linetypedef = coldef; linetypedef[1:7] = "solid"; linetypedef[8:9] = "dashed"
-fig = fig + ggtitle(ifelse(grepl("marrow",tissue_name),"Bone Marrow",ifelse(grepl("lung",tissue_name),"Lung",ifelse(grepl("pancreas",tissue_name),"Pancreas",NA)))) +
-  scale_color_manual(values = coldef) + 
-  scale_linetype_manual(values = linetypedef)
+fig = fig + ggtitle(ifelse(grepl("marrow",tissue_name),"Bone Marrow",ifelse(grepl("lung",tissue_name),"Lung",ifelse(grepl("pancreas",tissue_name),"Pancreas",NA)))) + xlab("# of Top Genes") + 
+  scale_color_manual(values = coldef, labels = names(coldef), breaks = names(coldef)) + 
+  scale_linetype_manual(values = linetypedef,labels = names(linetypedef), breaks = names(linetypedef))
 
 # Add LowQualityRate Scatterplot
-fig2 = ggplot(score_measure %>% filter(Method!="All genes") %>% filter(TopGene %in% c(50,100,500)), aes(x = LowQualityRate, y = DensityIndex, color = Method)) + 
-  geom_point() + xlim(0,1) + 
-  scale_color_manual(values = coldef) + facet_wrap(~TopGene)  + guides(color = "none")
-
-(fig+fig2 + plot_layout(widths = c(1, 2))) &theme_minimal() & theme(
+top_gene_labeller = as_labeller(setNames(paste0("N = ",unique(score_measure$TopGene)),unique(score_measure$TopGene)))
+fig2 = ggplot(score_measure %>% filter(Method!="All genes") %>% filter(TopGene %in% c(50,100,500)), aes(x = 1 - LowQualityRate, y = DensityIndex, color = Method)) + geom_point() + xlim(0,1) + xlab("Percentage of Cells covered by Top N Genes") + scale_color_manual(values = coldef) +
+  guides(color = "none") + facet_wrap(~TopGene, labeller = top_gene_labeller)
+
+(fig+fig2 + plot_layout(widths = c(1, 2)))  & theme(
     plot.title = element_text(face="bold", size = 20),
@@
 })
-
+```
+```{r}
 fig = wrap_plots(pl,ncol = 1) + plot_layout(guides = "collect") & theme(legend.position = 'bottom')
@@
 # Only keep Bone Marrow
-fig = wrap_plots(pl[[1]],ncol = 1) + plot_layout(guides = "collect") & theme(legend.position = 'bottom')
+fig = wrap_plots(pl[[1]],ncol = 1) + plot_layout(guides = "collect") + plot_annotation(tag_levels = "A") & theme(legend.position = 'bottom', plot.tag = element_text(size = 20))
 ggsave(filename = file.path(figure_path,"supfig-densityindex.png"), plot = fig, width = 12, height = 5)
@@
 
-fig = fig1 + fig2 + plot_annotation(title = "Sensitivity Test: Different kNN", theme = theme(plot.title = element_text(size = 20)))
-ggsave(filename = file.path(figure_path,"supfig-sensitivityknn.png"), plot = fig, width = 16, height = 5)
+# fig = fig1 + fig2 + plot_annotation(title = "Sensitivity Test: Different kNN", theme = theme(plot.title = element_text(size = 20)))
+# ggsave(filename = file.path(figure_path,"supfig-sensitivityknn.png"), plot = fig, width = 16, height = 5)
 ```
@@
 df$Top_Genes = as.factor(df$Top_Genes)
-fig1 = ggplot(data = df, aes(x = pc, y = JaccardIndex, group = Top_Genes, color = Top_Genes)) +
+fig3 = ggplot(data = df, aes(x = pc, y = JaccardIndex, group = Top_Genes, color = Top_Genes)) +
   geom_line() +
@@
 
-fig2 = ggplot(data = df, aes(x = Parameter, y = AUROC, group = gt_set, color = gt_set)) +
+fig4 = ggplot(data = df, aes(x = Parameter, y = AUROC, group = gt_set, color = gt_set)) +
   geom_line() +
@@
 
-fig = fig1 + fig2 + plot_annotation(title = "Sensitivity Test: Different PC", theme = theme(plot.title = element_text(size = 20)))
-ggsave(filename = file.path(figure_path,"supfig-sensitivitypc.png"), plot = fig, width = 16, height = 5)
-```
-
-# SupFig - Gene-Gene Jaccard-dist HeatMap
+# fig = fig3 + fig4 + plot_annotation(title = "Sensitivity Test: Different PC", theme = theme(plot.title = element_text(size = 20)))
+# ggsave(filename = file.path(figure_path,"supfig-sensitivitypc.png"), plot = fig, width = 16, height = 5)
+```
+```{r}
+p = (fig1 + fig2) / (fig3 + fig4) + plot_layout(guides = "collect") + plot_annotation(tag_level = 'A') & theme(plot.tag = element_text(face="bold", size = 20)) 
+ggsave(filename = file.path(figure_path,"supfig-sensitivity.png"), plot = p, width = 16, height = 10)
+```
+
+# SupFig - Gene Modules - Bone Marrow
+## Cut-off of localized genes
 ```{r}
@@
 source("LMD_function.R", echo = F)
-celltype_color = setNames(colorRampPalette(brewer.pal(9, "Set1"))(length(unique(cell_label))),unique(cell_label))
+# celltype_color = setNames(colorRampPalette(brewer.pal(9, "Set1"))(length(unique(cell_label))),unique(cell_label))
 consortium = "tabular_muris"
@@
 res = readRDS(file.path(folder.path,"intermediate_data",tissue_name,paste0(tissue_name,"_gene_tree.rds")))
-p = Visualize_gene_heatmap(dist, gene_partition = res$gene_partition_rename, gene_hree = res$gene_hree)
-ggsave(filename = file.path(figure_path,"supfig-moduleheatmap.png"), plot = p, width = 10, height = 10)
-```
-
-# SupFig - Module Activity Score (Bone marrow)
-```{r}
-consortium = "tabular_muris"
-tissue_name = "marrow_facs"
-dir.path <- "/banach1/ruiqi/local_marker/LMD_data"
-folder.path = file.path(dir.path,consortium)
-res = readRDS(file.path(folder.path,"intermediate_data",tissue_name,paste0(tissue_name,"_gene_tree.rds")))
+local_gene = readRDS(file.path(folder.path,"intermediate_data",tissue_name,paste0("local_gene_",tissue_name,".rds")))
+p0 <- ggplot(local_gene$gene_table, aes(x = rank, y = score)) + geom_point() + geom_point(aes(x = length(local_gene$cut_off_gene), y = local_gene$gene_table$score[length(local_gene$cut_off_gene)]), color = "red", size = 5) +  # Add a red point at index 50
+  labs(x = "Rank", y = "LMDS") + theme_minimal() + theme(
+    plot.title = element_text(size = 30),
+    axis.title = element_text(size = 20),
+    axis.text = element_text(size = 15))
+```
+## Gene-Gene Jaccard-dist HeatMap
+```{r}
+p1 = Visualize_gene_heatmap(dist, gene_partition = res$gene_partition_rename, gene_hree = res$gene_hree)
+```
+## Module Activity Score
+```{r}
 tiss = readRDS(file.path(folder.path,paste0(tissue_name,".rds")))
-
 modules = levels(res$gene_partition_rename)
@@
 names(pl_module) = modules
-p = (wrap_plots(pl_module,ncol = 7) &  theme(plot.title = element_text(face="bold", size = 10),legend.title = element_text(size = 30),legend.text = element_text(size = 20))) + plot_layout(guides = "collect")
-
-ggsave(filename = file.path(figure_path,"supfig-moduletsne.png"), plot = p, width = 7*3, height = 5*3)
+p2 = (wrap_plots(pl_module,ncol = 7) &  theme(plot.title = element_text(face="bold", size = 10),legend.title = element_text(size = 30),legend.text = element_text(size = 20))) + plot_layout(guides = "collect")
+p2 = as.ggplot(p2)
+
+# ggsave(filename = file.path(figure_path,"supfig-moduletsne.png"), plot = p, width = 7*3, height = 5*3)
+```
+```{r}
+p = (p0 + p1 + p2) + plot_layout(design = paste0("AA#BBB#\n###BBB#",paste0(rep("\nCCCCCCC",3),collapse = ""))) + plot_annotation(tag_levels = 'A') & theme(plot.tag = element_text(face="bold", size = 25))
+ggsave(filename = file.path(figure_path,"supfig-obtainmodule.png"), plot = p, width = 21, height = 9 + 5*3)
 ```
@@
 ```{r}
+consortium = "tabular_muris"
+tissue_name = "marrow_facs"
+dir.path <- "/banach1/ruiqi/local_marker/LMD_data"
+folder.path = file.path(dir.path,consortium)
+
 one_hot_matrix <- 
@@
 ```{r}
-thred = 0.3
+thred = 0.4
 jaccard_index_pruned = jaccard_index[apply(jaccard_index,1,max) >= thred,]
@@
 jaccard_index_pruned = jaccard_index_pruned[names(sort(apply(jaccard_index_pruned,1,which.max))),]
-p = Visualize_jaccard_mtx(jaccard_index_pruned, threshold = thred)
-p = as.ggplot(p) + coord_fixed()
-ggsave(filename = file.path(figure_path,"supfig-ct-specific-module2.png"), plot = p, width = 8, height = 10)
+p1 = Visualize_jaccard_mtx(jaccard_index_pruned, threshold = thred)
+p1 = as.ggplot(p1) + coord_fixed()
 ```
@@
 pl = pl[rownames(jaccard_index_pruned)]
-p = (wrap_plots(pl,ncol = 6) & theme(plot.title = element_text(face="bold", size = 10),legend.title = element_text(size = 15),legend.text = element_text(size = 10))) + plot_layout(guides = "collect")
-
+p2 = (wrap_plots(pl,ncol = ceiling(length(pl)/2)) & theme(plot.title = element_text(face="bold", size = 10),legend.title = element_text(size = 15),legend.text = element_text(size = 10))) + plot_layout(guides = "collect")
+p2 = as.ggplot(p2)
 # p = (wrap_plots(pl,ncol = 8) & theme(plot.title = element_text(face="bold", size = 20),legend.title = element_text(size = 15),legend.text = element_text(size = 10))) + plot_layout(guides = "collect")
 # ggsave(filename = "supfig-ct-specific-module1.png", plot = p, width = 3*8, height = 6)
-
-ggsave(filename = file.path(figure_path,"supfig-ct-specific-module1.png"), plot = p, width = 3*7, height = 3*4)
 ```
@@
 ```
-
 ```{r}
@@
 pl <- lapply(1:length(pl), function(i){pl[[i]] + labs(title = titles[i]) + theme(plot.title = element_text(size = 20, face = "bold"))})
-p = wrap_plots(pl,ncol = 2) + plot_layout(guides = "collect")
-ggsave(filename = file.path(figure_path,"supfig-ct-specific-module3.png"), plot = p, width = 10, height = 8)
+p3 = wrap_plots(pl,ncol = 2) + plot_layout(guides = "collect")
+p3 = as.ggplot(p3)
+```
+```{r}
+p = (p2 + p1 + p3) + plot_annotation(tag_levels = "A") + plot_layout(design = "AAAB\nAAAB\nAAAB\nAAAB\nCC##\nCC##\nCC##") & theme(plot.tag = element_text(face="bold", size = 25))
+
+ggsave(filename = file.path(figure_path,"supfig-ct-specific-module.png"), plot = p, width = 3*6 + 8, height = 3*4 + 8)
+# ggsave(filename = file.path(figure_path,"supfig-ct-specific-module2.png"), plot = p, width = 8, height = 10)
+# ggsave(filename = file.path(figure_path,"supfig-ct-specific-module3.png"), plot = p, width = 10, height = 8)
 ```
@@
 ```{r}
+consortium = "tabular_muris"
 tissue_name = "marrow_facs"
+dir.path <- "/banach1/ruiqi/local_marker/LMD_data"
+folder.path = file.path(dir.path,consortium)
+res = readRDS(file.path(folder.path,"intermediate_data",tissue_name,paste0(tissue_name,"_gene_tree.rds")))
 tiss = readRDS(file.path(folder.path,paste0(tissue_name,".rds")))
@@
 reduction = "tsne"
-p1 = DimPlot(tiss,group.by = "Phase", cols = c("grey",scales::hue_pal()(3)[3:2]), reduction = reduction) + labs(title = "Phase annotated by Seurat",color = "Phase") + theme(
+p1 = DimPlot(tiss,group.by = "Phase", shuffle = TRUE, reduction = reduction) + labs(title = "Phase annotated by Seurat",color = "Phase") + 
+  scale_color_manual(values = c("grey",scales::hue_pal()(3)[3:2]), limits = c("G1","S","G2M")) + theme(
   plot.title = element_text(size = 10),
@@
       axis.line = element_blank())
-p2 = DimPlot(tiss_droplet,group.by = "Phase", cols = c("grey",scales::hue_pal()(3)[3:2]), reduction = reduction) + labs(title = "Phase annotated by Seurat",color = "Phase") + theme(
+p2 = DimPlot(tiss_droplet,group.by = "Phase", shuffle = TRUE, reduction = reduction) + 
+  scale_color_manual(values = c("grey",scales::hue_pal()(3)[3:2]), limits = c("G1","S","G2M")) + labs(title = "Phase annotated by Seurat",color = "Phase") + theme(
   axis.title = element_blank(),
@@
 ggsave(filename = file.path(figure_path,"fig-module-cellcycle1.png"), plot = p, width = 3*4 + 1, height = 3*2)
-ggsave(filename = file.path(figure_path,"supfig-module-cellcycle.png"), plot = p, width = 3*4 + 1, height = 3*2)
+# ggsave(filename = file.path(figure_path,"supfig-module-cellcycle.png"), plot = p, width = 3*4 + 1, height = 3*2)
+
+p = wrap_plots(p1,wrap_plots(pl_module_cc,ncol = 3) +plot_layout(guides = "collect"),design = "ABBB") & theme(plot.title = element_text(size = 15))
+ggsave(filename = file.path(figure_path,"supfig-module-cellcycle.png"), plot = p, width = 3*4 + 1, height = 3)
 ```
@@
 df = read.csv(file.path(folder.path, "intermediate_data",tissue_name,"reactome_pathway_cc.csv"))
+# select top5 pathway for each module
+selected_terms = df %>% group_by(module) %>%
+  arrange(p.adjust) %>%
+  slice_head(n = 5) %>%
+  ungroup() %>%.$Description
+df = df %>% filter(Description %in% selected_terms)
 df$GeneRatio = unlist(lapply(df$GeneRatio,function(x) eval(parse(text = x))))
@@
 df$Description = as.factor(df$Description)
-df$Description = factor(df$Description, levels(df$Description)[rev(c(2,4,6,10,1,3,5,9,7,8))])
+df$Description = factor(df$Description, levels(df$Description)[rev(c(2,4,3,6,10,1,5,9,7,8))])
+cc_module = c(22,20,19)
 df$module = factor(df$module,levels = cc_module)
 levels(df$module) = paste0("Module",levels(df$module))
+df$log.p.adj = -log(df$p.adjust,base = 10)
 p = ggplot(df, aes(x=module, y=Description)) +
-  geom_point(aes(size=GeneRatio, color=p.adjust), alpha=0.7) +
-  scale_color_gradientn(colors=colorRampPalette(c("red", "blue"))(99), name="p.adjust",limits = c(0,0.05)) +
+  geom_point(aes(size=GeneRatio, color=log.p.adj), alpha=0.7) +
+  scale_color_gradientn(colors=colorRampPalette(c("blue","red"))(99), name="-log10(p.adj)") +
   theme_minimal() +
-  labs(size="Gene Ratio",x = "Module") +
+  labs(size="Gene Ratio",x = "Module", color = "-log10 p.adjust") +
   theme(legend.position="right",
         axis.title = element_text(size = 0),
-        axis.text.x = element_text(size = 15),
-      axis.text.y = element_text(size = 10),
-      legend.title = element_text(size = 20),
-      legend.text = element_text(size = 15))
-ggsave(filename = file.path(figure_path,"fig-module-cellcycle2.png"), plot = p, width = 10, height = 6)
+        axis.text.x = element_text(size = 15, angle = 45),
+      axis.text.y = element_text(size = 15),
+      legend.title = element_text(size = 10),
+      legend.text = element_text(size = 8))
+ggsave(filename = file.path(figure_path,"fig-module-cellcycle2.png"), plot = p, width = 6, height = 6)
 ```
@@
 # Fig - Subtype Investigation
+## B lymphoid
 late pro -> precursor B -> immature -> Naive B
@@
 ```
+## Granulocyte
 Granulocyte: 
@@
 coldef = setNames(colorRampPalette(brewer.pal(12, "Paired"))(length(unique(cell_label))),unique(cell_label))
-p = DimPlot(subset(tiss,cells = colnames(tiss)[cell_id]), group.by = "cell_ontology_class",label = TRUE, cols = coldef,repel = TRUE,label.size = 3) + NoLegend() + labs(title = NULL)
-
-p = wrap_plots(wrap_plots(p) & theme(axis.title = element_blank(),axis.text = element_blank(), axis.ticks=element_blank(), axis.line = element_blank()),(wrap_plots(pl,nrow = 1) &  theme(plot.title = element_text(size = 15))) + plot_layout(guides = "collect"), nrow = 1, design = "ABBBB")
-
-ggsave(filename = file.path(figure_path,"fig-subtypeB.png"), plot = p, width = 5*3, height = 3)
-
-ggsave(filename = file.path(figure_path,"fig-subtypeG.png"), plot = p, width = 5*3, height = 3)
+p = DimPlot(subset(tiss,cells = colnames(tiss)[cell_id]), group.by = "cell_ontology_class",label = TRUE, cols = coldef,repel = TRUE,label.size = 3) + NoLegend() + labs(title = NULL) + NoAxes()
+
+# p = wrap_plots(as.ggplot(p),(wrap_plots(pl,nrow = 1) &  theme(plot.title = element_text(size = 15))) + plot_layout(guides = "collect"), nrow = 1, design = "ABBBB")
+# ggsave(filename = file.path(figure_path,"fig-subtypeB.png"), plot = p, width = 5*3, height = 3)
+
+pB = c(list(p),pl)
+pG = c(list(p),pl)
+p = (wrap_plots(c(pB,pG),nrow = 2) &  theme(plot.title = element_text(size = 15))) + plot_layout(guides = "collect")
+ggsave(filename = file.path(figure_path,"fig-subtype.png"), plot = p, width = 5*3, height = 3*2)
 ```
@@
 ```{r}
-sample_name = sample_ls[1]
+sample_name = sample_ls[4]
 data_S = get(paste0("data_S_",sample_name))
@@
 ggsave(filename = file.path(figure_path,"supfig-moduletsne_e13mut.png"), plot = p, width = 6*3 + 1, height = 3*3)
+ggsave(filename = file.path(figure_path,"supfig-moduletsne_e13ctl.png"), plot = p, width = 6*3 + 1, height = 3*3)
+ggsave(filename = file.path(figure_path,"supfig-moduletsne_e14mut.png"), plot = p, width = 6*3 + 1, height = 3*3)
+ggsave(filename = file.path(figure_path,"supfig-moduletsne_e14ctl.png"), plot = p, width = 6*3 + 1, height = 3*3)
 ```
@@
 ```
+
 # Fig - Wnt Module (SmoM2)
@@
 p = ggplot(df, aes(x = wnt_related_module, y = get(gene))) +
-  geom_violin(scale = "width") + geom_jitter() + 
+  geom_violin(scale = "width") + 
   ggsignif::geom_signif(comparisons = list(c("High-Wnt Module", "DC Module"), c("High-Wnt Module", "Wnt Module"), c("Wnt Module", "Other")),
@@
 pl = lapply(sample_ls, function(sample_name){
+  cat(sample_name,"\n")
   data_S = get(paste0("data_S_",sample_name))
@@
     lapply(c("umap","umap_cc_regressout"),function(reduc){
+    cat(reduc,"\n")
     p1 = DimPlot(data_S, 
@@
 ```{r}
+sample_name = sample_ls[1]
 df = read.csv(file.path(dir.path,"intermediate_data","e13_smom2_cdkn1a_table.csv"))
@@
 p = ggplot(df, aes(x = status, y = get(gene))) +
-  geom_violin(scale = "width") + geom_jitter() + 
+  geom_violin(scale = "width") + 
   ggsignif::geom_signif(comparisons = list(c("Cdkn1a+", "Quiescent Module"), c("Quiescent Module", "Cdkn1a-")),
@@
 
-ggsave(filename = file.path(figure_path, "fig-smom2_quiest_module1.png"), plot = p2 + p, width = 10, height = 5)
+ggsave(filename = file.path(figure_path, "fig-smom2_quiest_module2.png"), plot = p2 / p, width = 7, height = 10)
 ```

### 2024-06-13 benchmark/run_methods_function.R

@@ -30,11 +30,12 @@ RunHVG <- function(srat_obj, input_genes, dir.file){
 #' SinglecellHaystack: 
 #' https://alexisvdb.github.io/singleCellHaystack/articles/examples/a02_example_scRNAseq.html
 #' Genes were ranked based on the increasing order of log(adjusted P-value)
-RunHaystack <- function(dat, feature_space, dir.file){
+RunHaystack <- function(dat, feature_space, dir.file = NULL){
   set.seed(123)
   res <- haystack(x = feature_space, expression = dat)
   marker <- show_result_haystack(res.haystack = res)
   marker[order(marker$log.p.adj),"rank"] = 1:nrow(marker)
+  if(is.null(dir.file)) return(marker)
   write.table(marker,file = dir.file,row.names = TRUE)
 }
 
@@ -103,7 +104,7 @@ result = pd.DataFrame({'escore': escores.abs().max(1), 'padj': pvals.min(1)})
 #' ## Hotspot
 #' https://hotspot.readthedocs.io/en/latest/CD4_Tutorial.html
 #' Genes were ranked by the increasing order FDR value (if two genes had the same, prioritize genes with larger Z-score).
-RunHotspot <- function(srat_obj, input_genes, feature_space, dir.file){
+RunHotspot <- function(srat_obj, input_genes, feature_space, dir.file = NULL){
   raw_count = as.matrix(srat_obj[["RNA"]]@counts)[input_genes,,drop = FALSE]
   cell_id = colnames(raw_count)
   gene_id = rownames(raw_count)
@@ -141,6 +142,7 @@ hs_results = hs.compute_autocorrelations()
     ")
   marker = reticulate::py$hs_results
   marker[order(marker$FDR),"rank"] = 1:nrow(marker)
+  if(is.null(dir.file)) return(marker)
   write.table(marker,file = dir.file,row.names = TRUE)
 }
 
@@ -161,7 +163,7 @@ RunSeuratv4 <- function(srat_obj, input_genes, feature_space, dir.file){
 
 #' Marcopolo
 #' The rank of genes is given by `Marcopolo_rank`
-RunMarcopolo <- function(srat_obj, input_genes, dir.file){
+RunMarcopolo <- function(srat_obj, input_genes, dir.file = NULL){
   raw_count = as.matrix(srat_obj[["RNA"]]@counts)[input_genes,,drop = FALSE]
   cacheEnv <- new.env()
[…]

### 2024-06-13 mouse_bone_marrow_example.R

@@ -207,7 +207,7 @@ df_module[,"# of Genes"] = unlist(lapply(levels(gene_partition),function(x){sum(
 df_module[,"# of Expressed Cells"] = colSums(tiss@meta.data[,paste0("Module",levels(gene_partition))] > 0.5)
 # CellType-specific modules
 jaccard_index = readRDS(file.path(folder.path,"intermediate_data",tissue_name,paste0(tissue_name,"_ct_module_jaccard.rds")))
-thred = 0.3
+thred = 0.4
 module_celltype = apply(jaccard_index,2,function(x){paste(names(which(x >= thred)),collapse = ",")})
 df_module[,"celltype"] = module_celltype[paste0("Module",levels(gene_partition))]
 # Annotate each module with Top5 significant Terms (p.adjust < 0.05)
@@ -245,7 +245,12 @@ cc_module = c(22,20,19)
 e_result = readRDS(file = file.path(folder.path,"intermediate_data",tissue_name,sprintf("%s_%s_result.rds",tissue_name,term)))
 e_result = e_result[cc_module]
 # Annotate each modules with top5 Pathway terms
-top_enrich_pathway_cc <- do.call(rbind,lapply(1:length(e_result), function(i) data.frame(e_result[[i]]@result[1:5,],module = names(e_result)[i])))
+top_enrich_pathway_cc <- do.call(rbind,lapply(1:length(e_result), function(i) data.frame(e_result[[i]]@result,module = names(e_result)[i])))
 top_enrich_pathway_cc <- top_enrich_pathway_cc[complete.cases(top_enrich_pathway_cc), ]
 top_enrich_pathway_cc = top_enrich_pathway_cc %>% filter(p.adjust < 0.05)
 write.csv(top_enrich_pathway_cc,file = file.path(folder.path, "intermediate_data",tissue_name,"reactome_pathway_cc.csv"))
+
+library(ReactomeGraph4R)
+login()
+matchObject(id = 'R-MMU-1640170')
+matchHierarchy(id = "R-MMU-1640170", type = "row")

### 2024-07-26 Demo.Rmd
24/07/26

---
title: "Tutorial for Runing LMD on Mouse Bone Marrow Dataset"
author: "Ruiqi Li"
date: '2023-10-12'
output: pdf_document
bibliography: references.bib
---

# Install & Load Packages
```{r}
source("LMD_function.R")
remotes::install_github("satijalab/seurat-wrappers")
library("SeuratWrappers")
```
# Preparing input data
```{r}
# Define your own dir path
dir.path <- "/banach1/ruiqi/local_marker/LMD_data"

folder.path <- file.path(dir.path,"tabular_muris")
file_name = "marrow_facs.rds"
if (!file.exists(file.path(folder.path,file_name))) {
   # If load url doesn't work, please download it and load from your working directory
  options(timeout=6000)
  download.file("https://figshare.com/ndownloader/files/13092380",
                destfile = file.path(folder.path,file_name), method = 'libcurl')
  tiss <- readRDS(file.path(folder.path,file_name))
  tiss <- UpdateSeuratObject(tiss)
  tiss <- RunALRA(tiss, assay = "RNA"); DefaultAssay(tiss) = "RNA"
  saveRDS(tiss,file = file.path(folder.path,file_name))
}
tiss <- readRDS(file.path(folder.path,file_name))
```

We obtain the following objects from this Seurat Obj:
* `feature_space`: The output of PCA. This matrix contains the first 20 principal components (PCs).
* `visual_space`: a data.frame with t-SNE coordinates (2D) based on the first 20 PCs.
* `dat`: a matrix object with the log-normalized expression of genes (rows) in each cell (columns)
* `cell_label`: Any metadata you wish to visualize and examine related to cells (here we take the cell type annotation)
```{r}
DefaultAssay(tiss) <- "RNA"
n_dim = dim(tiss@reductions$pca@cell.embeddings)[2]
feature_space = as.matrix(tiss@reductions$pca@cell.embeddings[,1:n_dim])
visual_space = data.frame(tiss@reductions$tsne@cell.embeddings)
dat = as.matrix(tiss[[DefaultAssay(tiss)]]@data)
cell_label = tiss$cell_ontology_class
```

We process the `dat` by only keep the genes detected in more than 10 cells and less than 50% of cells (genes is "detected" in a cell only if its expression level is greater than the median expression level of that cell)

```{r}
Gene_detected_count <- apply(dat > apply(dat,2,median),1,sum)
selected_genes = (Gene_detected_count >= 10) & (Gene_detected_count <= ncol(dat) * 0.5)
dat = dat[selected_genes,,drop = FALSE]
```

# Running LMD Step by Step
This provides a step-by-step tutorial for a better understanding of the algorithm. If you want to generate results directly, please skip to 'Running LMD in One Step'

## Step1: Build KNN Graph
```{r, fig.width=20}
# Construct knn graph
knn_result = Symmetric_KNN_graph(knn = 5, feature_space = feature_space)

A = knn_result$adj_matrix # Adjacency Matrix
W = knn_result$graph # Symmetrized Graph ((A + AT) / 2)

# Plot knn graph
plot_knn_graph(affinity_m = W, label = cell_label, layout = visual_space)
```

## Step2: Diffuse Genes on the cell graph
```{r, fig.width=20}
# Generate $P^{t}, t = 2,4,8,...,max_time$
P_ls = Obtain_Pls(W = W, max_time = 2^20)
# Generate initial gene distribution
rho = Rowwise_normalize(dat[,colnames(W)])
# Check One Gene Diffusion on Cell graph by time
gene = "Fcnb"
print(FeaturePlot_diffusion(coord = visual_space,init_state = rho[gene,],P_ls = P_ls,check_time = c(2,16,128),gene_name = gene))
```

## Step3: Obtain Diffusion KL score & Cumulative score
### Obtain diffusion KL score
```{r}
score_result = fast_get_lmds(W = W, init_state = rho, P_ls = P_ls, largeData = FALSE)
```

### Diffusion KL pattern for several genes
```{r}
genes = c("Tlr11","Fcnb","Cd79a","4933425H06Rik","Uchl4","Dock11")
genes_label = rep(c("good","bad"),each = 3)
Visualize_score_pattern(score_result$'score_profile', genes = genes, label_class = genes_label, facet_class = NULL, text = TRUE, normalize = TRUE)
```

### Show result
```{r}
res = show_result_lmd(score_result,n = 10)
head(res$gene_table,10)
```

# Running LMD in one step
```{r}
tic()
lmd_result = LMD(dat, feature_space, knn = 5, max_time = 2^20)
toc()
local_gene = show_result_lmd(lmd_result)
head(local_gene$gene_table,10)
```

# Identify Gene Modules
## Step1: Deriving Gene Modules
### Select top localized genes based on knee point
```{r}
local_gene = show_result_lmd(lmd_result)
local_gene = names(local_gene$cut_off_gene)
```
### Denoising
We use ALRA @linderman2022zero to De-noise our data
```{r}
if(!"alra" %in% names(tiss@assays)){
  tiss <- RunALRA(tiss, assay = "RNA")
  DefaultAssay(tiss) <- "RNA"
}
dat_alra = as.matrix(tiss[["alra"]]@data)[local_gene,]
```
### Calculate Gene-gene co-localization distance
```{r}
distance_method = "jaccard"
dist = Calculate_distance(dat_alra, method = distance_method)
```
### Obtain Gene modules
```{r}
gene_partition = Obtain_gene_partition(dist, clustering_method = "average", min_gene = 5)
```
### Visualize
```{r}
p = Visualize_gene_heatmap(dist)
```

## Step2: Computing Per-Cell Module Activity Scores
```{r}
tiss = AddModuleActivityScore(tiss, gene_partition = gene_partition, do_local_smooth = FALSE)

FeaturePlot(tiss, features = paste0("Module",levels(gene_partition)), order = TRUE, reduction = "tsne", ncol = 3) & NoAxes() & 
  scale_color_gradient(low = "lightgrey", high = "blue", limits = c(0,1)) & labs(color = "ModuleScore") & NoLegend()
```

### 2024-07-26 paper_figure.Rmd
24/07/26

@@
 
-ggsave(filename = file.path(figure_path,"fig-pipeline2.png"), plot = p, width = 10, height = 5)
+# ggsave(filename = file.path(figure_path,"fig-pipeline2.png"), plot = p, width = 10, height = 5)
+
+ggsave(filename = file.path(figure_path,"fig-pipeline2.png"), plot = p, width = 8, height = 5)
 ```
@@
 local_gene = readRDS(file.path(folder.path,"intermediate_data",tissue_name,paste0("local_gene_",tissue_name,".rds")))
-p0 <- ggplot(local_gene$gene_table, aes(x = rank, y = score)) + geom_point() + geom_point(aes(x = length(local_gene$cut_off_gene), y = local_gene$gene_table$score[length(local_gene$cut_off_gene)]), color = "red", size = 5) +  # Add a red point at index 50
-  labs(x = "Rank", y = "LMDS") + theme_minimal() + theme(
+p0 <- ggplot(local_gene$gene_table, aes(x = rank, y = score)) + geom_point() + geom_vline(xintercept = length(local_gene$cut_off_gene), linetype="dashed", color="grey", lwd = 1) + geom_point(aes(x = length(local_gene$cut_off_gene), y = local_gene$gene_table$score[length(local_gene$cut_off_gene)]), color = "red", size = 5) +  # Add a red point at index 50
+  labs(x = "Rank", y = "LMDS") + theme(
     plot.title = element_text(size = 30),
     axis.title = element_text(size = 20),
-    axis.text = element_text(size = 15))
+    panel.grid = element_blank(),
+    panel.background = element_blank(),
+    axis.text = element_text(size = 15),
+    axis.line = element_line(colour = "black"))
 ```
@@
                       "2" = c("Ms4a1", "Cd22"),
-                      "3" = c("Fcer2a","Ms4a4c","Cxcr5"))
+                      "3" = c("Fcer2a","Cxcr5"))
 cell_id = grep("B cell",tiss$cell_ontology_class)
@@
 p = (wrap_plots(c(pB,pG),nrow = 2) &  theme(plot.title = element_text(size = 15))) + plot_layout(guides = "collect")
-ggsave(filename = file.path(figure_path,"fig-subtype.png"), plot = p, width = 5*3, height = 3*2)
-```
-
-<!-- ```{r} -->
-<!-- p <- DimPlot(tiss, group.by = "cell_ontology_class",label = FALSE, cols = setNames(scales::hue_pal()(2),c("late pro-B cell","granulocyte"))) + ggtitle("Pre-annotated cell types") + theme( -->
-<!--   axis.title = element_blank(), -->
-<!--       axis.text = element_blank(), -->
-<!--       axis.ticks=element_blank(), -->
-<!--       legend.title = element_text(size = 20), -->
-<!--       legend.text = element_text(size = 15), -->
-<!--       axis.line = element_blank()) -->
-<!-- ggsave(filename = "fig-subtype1.png", plot = p, width = 6, height = 5) -->
-
-<!-- pl1 = FeaturePlot(tiss,features = names(gene_partition)[gene_partition %in% c(5,6)][c(1:3,7,9)],ncol = 5,order = TRUE) -->
-<!-- pl2 = FeaturePlot(tiss,features = names(gene_partition)[gene_partition == 43][1:5],ncol = 5,order = TRUE) -->
-<!-- # pl_module_subtype = lapply(pl_module[c(5,6,43)],function(x) x + scale_color_gradient(low = "lightgrey", high = "blue")) -->
-
-<!-- pl = wrap_plots(pl1 & theme(plot.title = element_text(color = scales::hue_pal()(2)[1])), -->
-<!--                 pl2 & theme(plot.title = element_text(color = scales::hue_pal()(2)[2])),nrow = 2) & labs(color = "Expression") & theme( -->
-<!--       plot.title = element_text(face="bold", size = 30), -->
-<!--       legend.title = element_text(size = 20), -->
-<!--       legend.text = element_text(size = 15), -->
-<!--       axis.title = element_blank(),axis.text = element_blank(), axis.ticks=element_blank(),axis.line = element_blank()) -->
-
-<!-- ggsave(filename = "fig-subtype2.png", plot = pl, width = 4*5 + 3, height = 4*2) -->
-<!-- ``` -->
-
-# SupFig - SmoM2 Cell Embedding
+ggsave(filename = file.path(figure_path,"fig-module-subtype.png"), plot = p, width = 5*3, height = 3*2)
+```
+
+# SmoM2 Loading
 ```{r}
@@
 ```{r}
+module_ls = readRDS(file.path(dir.path,"intermediate_data","module_ls.rds"))
+module_ls_ref = lapply(module_ls,function(x) setNames(names(x),x)); names(module_ls_ref) = NULL
+module_ls_ref = do.call(c,module_ls_ref)
+module_ls_ref = setNames(paste0("Module",sub(".*_([0-9]+)$", "\\1", names(module_ls_ref))),names(module_ls_ref))
+```
+# SupFig - SmoM2 Cell Embedding
+```{r}
 pl = lapply(sample_ls,function(sample_name){
@@
 
-# Fig - CC Module (SmoM2)
-```{r}
-module_ls = readRDS(file.path(dir.path,"intermediate_data","module_ls.rds"))
-module_ls_ref = lapply(module_ls,function(x) setNames(names(x),x)); names(module_ls_ref) = NULL
-module_ls_ref = do.call(c,module_ls_ref)
-```
+# Fig - SmoM2-fig1-CC
 ```{r}
@@
       scale_color_gradient(low = "lightgrey", high = "blue", limits = c(0,1)) + 
-      ggtitle(sprintf("%s (%d genes)", module_ls_ref[module], sum(local_gene$gene_partition == module_id))) + labs(color = "ModuleScore")) & NoAxes()
+        labs(title = sprintf("%s (%d genes)", module_ls_ref[module], sum(local_gene$gene_partition == module_id)), color = "ModuleScore")) & NoAxes()
   })
@@
 p = wrap_plots(pl,nrow = 4) + plot_layout(guides = "collect")
-ggsave(filename = file.path(figure_path, "fig-smom2_cc_module.png"), plot = p, width = 3*4 + 3, height = 3*4)
+ggsave(filename = file.path(figure_path, "fig-smom2_1a.png"), plot = p, width = 3*4 + 3, height = 3*4)
 ```
@@
 
-p1 = (DimPlot(data_S_merged, group.by = "type", shuffle = TRUE) & NoAxes()) + ggtitle("By Top2000 HVGs") 
-p2 = (DimPlot(data_S_merged, group.by = "type", shuffle = TRUE, reduction = "umap_cc") & NoAxes()) + ggtitle("By 95 CC genes (Seurat)") 
-p = p1 + p2 + plot_layout(guides = "collect")
-ggsave(filename = file.path(figure_path, "supfig-smom2_E13_embed.png"), plot = p, width = 2*3+1, height = 3)
-```
-
-# Fig - Wnt Module (SmoM2)
+p1 = (DimPlot(data_S_merged, group.by = "type", shuffle = TRUE) & NoAxes()) + labs(title = "By Top2000 HVGs") 
+p2 = (DimPlot(data_S_merged, group.by = "type", shuffle = TRUE, reduction = "umap_cc") & NoAxes()) + labs(title = "By 95 CC genes (Seurat)") 
+p = (p1 + p2 + plot_layout(guides = "collect") + plot_annotation(tag_levels = "A")) & theme(
+    plot.title = element_text(size = 15, face = "plain"),
+    plot.tag = element_text(size = 20, face = "bold"))
+ggsave(filename = file.path(figure_path, "supfig-smom2_E13_embed.png"), plot = p, width = 2*3+2, height = 3)
+```
+
+# Fig - SmoM2-fig1-Wnt
 ```{r}
@@
 p = wrap_plots(pl,nrow = 3) + plot_layout(guides = "collect")
-ggsave(filename = file.path(figure_path, "fig-smom2_wnt_module1.png"), plot = p, width = 3*3 + 3, height = 3*3)
+ggsave(filename = file.path(figure_path, "fig-smom2_1b.png"), plot = p, width = 3*3 + 3, height = 3*3)
 ```
@@
 gene = "Lef1"
-df$wnt_related_module = factor(df$wnt_related_module, levels = df %>% group_by(wnt_related_module) %>% summarise(Average_Expression = mean(get(gene), na.rm = TRUE)) %>% arrange(desc(Average_Expression)) %>% pull(wnt_related_module))
+df = left_join(df, df %>% group_by(wnt_related_module) %>% summarise(mean = mean(get(gene), na.rm = TRUE)), by = "wnt_related_module")
+df$wnt_related_module = ifelse(df$wnt_related_module_fullname %in% names(module_ls_ref), module_ls_ref[df$wnt_related_module_fullname], df$wnt_related_module)
+df$wnt_related_module = factor(df$wnt_related_module, levels = df %>% arrange(desc(mean)) %>% pull(wnt_related_module) %>% unique())
 
@@
 #   theme_minimal()
-p = ggplot(df, aes(x = wnt_related_module, y = get(gene))) +
-  geom_violin(scale = "width") + 
-  ggsignif::geom_signif(comparisons = list(c("High-Wnt Module", "DC Module"), c("High-Wnt Module", "Wnt Module"), c("Wnt Module", "Other")),
+generate_combinations <- function(vec) {
+  n <- length(vec)
+  combinations <- list()
+  for (i in 1:(n - 1)) {
+    combinations[[i]] <- c(vec[i], vec[i + 1])
+  }
+  return(combinations)
+}
+p1 = ggplot(df, aes(x = wnt_related_module, y = get(gene), fill = mean)) +
+  geom_boxplot() + 
+  ggsignif::geom_signif(comparisons = generate_combinations(levels(df$wnt_related_module)),
               map_signif_level = TRUE,
               textsize = 3) + theme_minimal() +
-  labs(title = paste0("Expression of ",gene), x = "", y = "Expression Level")
-p = p + theme(
-    plot.title = element_text(size = 20),
+  labs(title = "Lef1 (Wnt) level across Wnt-related modules", x = "", y = "Expression Level") + 
+  scale_fill_gradient(low = "white", high = "darkgreen", name = "mean", 
+                      limits = c(0,max(df$mean)))
+p1 = p1 + theme(
+    plot.title = element_text(size = 15),
     axis.title = element_text(size = 15),
@@
         group.by = "Wnt.Module", order = levels(data_S$'Wnt.Module'), 
-        cols = setNames(c(colorRampPalette(brewer.pal(3, "Paired"))(3),"lightgrey"),levels(data_S$'Wnt.Module'))) + ggtitle("Active Cells for Wnt-related module \n(ModuleScore > 0.4)") + NoAxes()
-
-ggsave(filename = file.path(figure_path, "fig-smom2_wnt_module2.png"), plot = p2 + p, width = 10, height = 5)
-```
-
-# Fig - Quiescent Module (SmoM2)
+        cols = setNames(c("#E41A1C","#1F78B4","#B2DF8A","lightgrey"),levels(data_S$'Wnt.Module'))) + labs(title = "Active Cells for Wnt-related modules") & NoAxes()
+p2 = p2 + theme(
+    plot.title = element_text(size = 15, face = "plain"))
+p = wrap_plots(p2,p1)
+
+ggsave(filename = file.path(figure_path, "fig-smom2_1c.png"), plot = p, width = 10, height = 5)
+```
+
+# Fig - SmoM2-fig2
 ```{r}
 modules = module_ls$smom2_dermal_E13.5_MUT["Quiescent Module"]
+thred = 0.5
 pl = lapply(sample_ls, function(sample_name){
@@
     p1 = DimPlot(data_S, 
-             group.by = "Phase", shuffle = TRUE, reduction = reduc) + ggtitle("") & NoAxes()
-    p2 = FeaturePlot(data_S, features = modules, order = TRUE, reduction = reduc) + ggtitle("") + NoAxes() + 
-    scale_color_gradient(low = "lightgrey", high = "blue", limits = c(0,1)) + labs(color = "ModuleScore")
-    if(reduc == "umap"){
-      cc_count = table(data_S$Phase[data_S@meta.data[,modules] > 0.5])
-      p2 = p2 + labs(subtitle = sprintf("%d Active Cells \n (%d in G1, %d in G2M, %d in S)",sum(cc_count), cc_count["G1"],cc_count["G2M"],cc_count["S"]))
-    }
+             group.by = "Phase", shuffle = TRUE, reduction = reduc) + labs(title = "", color = "Phase") & NoAxes()
+    p2 = FeaturePlot(data_S, features = modules, order = TRUE, reduction = reduc) + NoAxes() + 
+    scale_color_gradient(low = "lightgrey", high = "blue", limits = c(0,1)) + labs(title = "", color = "ModuleScore")
+    # if(reduc == "umap"){
+    #   cc_count = table(data_S$Phase[data_S@meta.data[,modules] > thred])
+    #   p2 = p2 + labs(subtitle = sprintf("%d Active Cells \n (%d in G1, %d in G2M, %d in S)",sum(cc_count), cc_count["G1"],cc_count["G2M"],cc_count["S"]))
+    # }
     list(p1,p2)
@@
   }else{
-    FeaturePlot(data_S, features = modules, order = TRUE, reduction = "umap") + ggtitle("") + NoAxes() + 
-    scale_color_gradient(low = "lightgrey", high = "blue", limits = c(0,1)) + labs(color = "ModuleScore")
+    FeaturePlot(data_S, features = modules, order = TRUE, reduction = "umap") + NoAxes() + 
+    scale_color_gradient(low = "lightgrey", high = "blue", limits = c(0,1)) + labs(title = "", color = "ModuleScore")
   }
@@
 p = wrap_plots(pl,nrow = 2) + plot_layout(guides = "collect")
-ggsave(filename = file.path(figure_path, "fig-smom2_quiest_module1.png"), plot = p, width = 3*5+3, height = 2*3 + 2)
-```
-## Cdkn1a+ Lef1 level
+ggsave(filename = file.path(figure_path, "fig-smom2_2a.png"), plot = p, width = 3*5+3, height = 2*3 + 1)
+```
+## Supfig-Cdkn1a+ Lef1 level
 ```{r}
@@
 gene = "Lef1"
-df$status = factor(df$status, levels = df %>% group_by(status) %>% summarise(Average_Expression = mean(get(gene), na.rm = TRUE)) %>% arrange(desc(Average_Expression)) %>% pull(status))
-
-p = ggplot(df, aes(x = status, y = get(gene))) +
-  geom_violin(scale = "width") + 
-  ggsignif::geom_signif(comparisons = list(c("Cdkn1a+", "Quiescent Module"), c("Quiescent Module", "Cdkn1a-")),
+df = left_join(df, df %>% group_by(status) %>% summarise(mean = mean(get(gene), na.rm = TRUE)), by = "status")
+df$status = sub(colnames(df)[4],module_ls_ref[colnames(df)[4]],df$status)
+df$status = factor(df$status, levels = df %>% arrange(desc(mean)) %>% pull(status) %>% unique())
+df$status_ori = df$status; levels(df$status_ori) = sprintf("%s (%d cells)",levels(df$status_ori),table(df$status_ori)[levels(df$status_ori)])
+levels(df$status)[which(table(df$status) < 10)] = NA
+levels(df$status) = sub("Cdkn1a", "\nCdkn1a", levels(df$status))
+generate_combinations <- function(vec) {
+  n <- length(vec)
+  combinations <- list()
+  for (i in 1:(n - 1)) {
+    combinations[[i]] <- c(vec[i], vec[i + 1])
+  }
+  return(combinations)
+}
+p1 = ggplot(df %>% filter(!is.na(status)), aes(x = status, y = get(gene), fill = mean)) +
+  geom_boxplot() + 
+  ggsignif::geom_signif(comparisons = generate_combinations(levels(df$status)),
               map_signif_level = TRUE,
               textsize = 3) + theme_minimal() +
-  labs(title = paste0("Expression of ",gene), x = "", y = "Expression Level")
-p = p + theme(
-    plot.title = element_text(size = 20),
+  labs(x = "", y = "Lef1 (Wnt) level") + 
+  scale_fill_gradient(low = "white", high = "darkgreen", name = "mean", 
+                      limits = c(0,max(df$mean)))
+p1 = p1 + theme(
+    plot.title = element_text(size = 15),
     axis.title = element_text(size = 15),
-    axis.text.x = element_text(size = 10,angle = 45),
+    axis.text.x = element_text(size = 10),
     axis.text.y = element_text(size = 10),
@@
 data_S = get(paste0("data_S_",sample_name))
-Idents(data_S) = df$status
-p2 = DimPlot(data_S, order = c("Quiescent Module","Cdkn1a+","Cdkn1a-"),
-        cols = setNames(c(hue_pal()(2),"lightgrey"),c("Quiescent Module","Cdkn1a+","Cdkn1a-"))) + ggtitle("Active Cells for Quiescent cells") + NoAxes()
-
-ggsave(filename = file.path(figure_path, "fig-smom2_quiest_module2.png"), plot = p2 / p, width = 7, height = 10)
-```
-
-# Supfig - Module-Colocalization
-```{r}
-df = read.csv(file = file.path(folder.path,"module_colocalize_align.csv"))
+data_S$'status' = df$status_ori
+p2 = DimPlot(data_S, 
+        group.by = "status", order = levels(data_S$'status')[c(3,2,1,4)], 
+        cols = setNames(c("#B2DF8A","#1F78B4","#E41A1C","lightgrey"),levels(data_S$'status'))) + labs(title = "Active Cells for Quiescent-related cells") & NoAxes()
+p2 = p2 + theme(
+    plot.title = element_text(size = 15, face = "plain"))
+p = (wrap_plots(p2,p1) + plot_annotation(tag_levels = "A")) & theme(plot.tag = element_text(size = 20, face = "bold"))
+
+ggsave(filename = file.path(figure_path, "supfig-smom2_quiest_module.png"), plot = p, width = 12, height = 5)
+```
+
+# Supfig - SmoM2-Colocalization
+## Module
+```{r}
+df = read.csv(file = file.path(dir.path,"intermediate_data","module_colocalize_align.csv"))
 pl = lapply(unique(df$sample),function(sample_name){
@@
     p1 = FeaturePlot(data_S, features = module1, order = TRUE) + ggtitle(module_ls_ref[module1])
-    p2 = FeaturePlot(data_S, features = module2, order = TRUE) + ggtitle(sprintf("%s (JI: %.3f)",module_ls_ref[module2],ji))
+    p2 = FeaturePlot(data_S, features = module2, order = TRUE) + labs(title = module_ls_ref[module2], subtitle = sprintf("(JI: %.3f)",ji))
     (p1 / p2) & scale_color_gradient(low = "lightgrey", high = "blue", limits = c(0,1)) & NoAxes() & labs(color = "ModuleScore")
@@
 })
-p = wrap_plots(c(pl[[1]][c(2:4,1)],pl[[2]][1],pl[[3]][c(1:3)]),nrow = 1) + plot_layout(guides = "collect")
-
-ggsave(filename = file.path(figure_path,"supfig-smom2_colocalized1.png"), plot = p, width = 25, height = 6)
-```
-
-# Pathway enrichment plot
+p1 = wrap_plots(c(pl[[1]][c(2,4,3,1)],pl[[2]][1],pl[[3]][c(1:3)]),nrow = 1) + plot_layout(guides = "collect")
+
+# ggsave(filename = file.path(figure_path,"supfig-smom2_colocalized1.png"), plot = p, width = 25, height = 6)
+```
+
+## Pathway enrichment plot
 ```{r}
@@
   df <- do.call(rbind, lapply(names(msig_result), 
-                            function(i) data.frame(msig_result[[i]]@result[1:5,], module = i)))
+                            function(i) data.frame(msig_result[[i]]@result, module = i)))
   df <- df[!is.na(df$Description), ]
@@
 msig_df$module_name = module_ls_ref[msig_df$module]
-msig_df$module_class = ifelse(grepl("CC",msig_df$module_name),"CC", ifelse(grepl("DC|Wnt",msig_df$module_name),"DC-related", ifelse(grepl("Quiescent",msig_df$module_name),"Quiescent",NA)))
-```
-```{r}
-pl = lapply(c("CC","DC-related","Quiescent"),function(c){
-  ggplot(msig_df %>% filter(module_class == c) %>% filter(p.adjust < 0.2), aes(x=module_name, y=Description)) +
-    geom_point(aes(size=GeneRatio, color=p.adjust), alpha=0.7) +
-    scale_color_gradientn(colors=colorRampPalette(c("red", "blue"))(99), name="p.adjust",limits = c(0,0.2)) + scale_size_continuous(name = "Gene Ratio", limits = c(0, 1)) + 
+tmp_module_name = names(unlist(module_ls)[match(msig_df$module,unlist(module_ls))])
+msig_df$module_class = ifelse(grepl("CC",tmp_module_name),"Cell Cycle", ifelse(grepl("DC|Wnt",tmp_module_name),"Wnt-related", ifelse(grepl("Quiescent",tmp_module_name),"Quiescent",NA)))
+```
+```{r}
+pl = lapply(c("Cell Cycle","Wnt-related","Quiescent"),function(c){
+  msig_df_filtered = msig_df %>% filter(module_class == c) %>% filter(pvalue < 0.05)
+  # select top5 pathway for each module
+  selected_terms = msig_df_filtered %>% group_by(module) %>% arrange(p.adjust) %>% slice_head(n = 5) %>%
+    ungroup() %>%.$Description
+  msig_df_filtered = msig_df_filtered %>% filter(Description %in% selected_terms)
+
+  ggplot(msig_df_filtered, aes(x=module_name, y=Description)) +
+    geom_point(aes(size=GeneRatio, color=pvalue), alpha=0.7) +
+    scale_color_gradientn(colors=colorRampPalette(c("red", "blue"))(99), name="pvalue",limits = c(0,0.05)) + scale_size_continuous(name = "Gene Ratio", limits = c(0, 1)) + 
     labs(size="Gene Ratio",x = "Module") + facet_grid(~sample,switch = "x",scales = "free") + ggtitle(c)
@@
 
-p = (wrap_plots(pl, nrow = 1) & theme_minimal() & 
+p2 = (wrap_plots(pl, nrow = 1) & theme_minimal() & 
     theme(legend.position="right",
@@
 
-ggsave(filename = file.path(figure_path,"supfig-smom2_colocalized2.png"), plot = p, width = 20, height = 5)
+# ggsave(filename = file.path(figure_path,"supfig-smom2_colocalized2.png"), plot = p, width = 20, height = 5)
+
+p = (as.ggplot(p1)/as.ggplot(p2)) + plot_annotation(tag_levels = 'A') & theme(plot.tag = element_text(face="bold", size = 25))
+
+ggsave(filename = file.path(figure_path,"supfig-smom2_colocalized.png"), plot = p, width = 20, height = 11)
 ```

### 2024-07-26 LMD_function.R

@@ -384,7 +384,7 @@ fast_calculate_score_profile <- function(W, max_time = 2^15,
   # Calculate multi-scale KL divergence
   score_df = do.call(cbind,lapply(1:length(P_ls),function(i){
     P = P_ls[[i]]
-    state = fastMatMult(init_state, P)
+    state = fastMatMult(init_state, as.matrix(P))
     score_df = get_score_profile(state_0 = init_state,
                                  state_t = state,
                                  state_inf = final_state,
@@ -854,11 +854,14 @@ FeaturePlot_meta <- function(dat, coord = NULL, feature_partition, reduction = N
 }
 Visualize_score_pattern <- function(score_profile, genes = NULL, 
                                     label_class = NULL, facet_class = NULL, 
-                                    add_point = NULL, dyadic = TRUE, text = FALSE){
+                                    add_point = NULL, dyadic = TRUE, text = FALSE, normalize = FALSE){
   score_df = score_profile
   if(!all(genes %in% rownames(score_df))){stop("Genes not found!")}
   profiles = names(which(table(sub("_\\d+", "", colnames(score_df)))>1))
   score_df = score_df[genes,]
+  if(("score0" %in% profiles) & normalize){
+    score_df = score_df/score_df$'max_score0'
+  }
   score_df$'gene' = rownames(score_df)
   score_df$'label' = score_df$'gene'
   score_df$'facet' = NA
@@ -896,7 +899,8 @@ Visualize_score_pattern <- function(score_profile, genes = NULL,
   }
   if(!is.null(label_class)){
     p = ggplot(data = df, mapping = aes(x = step, y = score, color = label, linetype = profiles)) + 
-      geom_line(aes(group = interaction(gene,profiles))) + labs(x = "Time", y = "Normalized Diffusion KL Score")
+      geom_line(aes(group = interaction(gene,profiles))) + 
+      labs(x = "Time", y = "Normalized Diffusion KL Score") + theme_half_open() + background_grid()
     if(text == TRUE){
       p = p +
         geom_text(data = df %>% group_by(interaction(gene,profiles)) %>% slice_head(n = 4) %>% slice_tail(n = 1),
@@ -905,7 +909,8 @@ Visualize_score_pattern <- function(score_profile, genes = NULL,
 
   }else{
     p = ggplot(data = df, mapping = aes(x = step, y = score, color = gene, linetype = profiles)) + 
[…]

### 2024-07-26 miscellaneous/miscellaneous_code/ATAC_Cite_spatial.R

@@ -1,170 +1,0 @@
-library(Seurat)
-library(SeuratData)
-library(SeuratWrappers)
-library(cowplot)
-library(dplyr)
-source("/banach1/ruiqi/local_marker/LocalMarkerDetector/LMD_function.R", echo = F)
-
-# CITE-seq: 25 proteins ----------
-InstallData("bmcite")
-bm <- LoadData(ds = "bmcite")
-DefaultAssay(bm) <- 'ADT'
-VariableFeatures(bm) <- rownames(bm[["ADT"]])
-bm <- NormalizeData(bm, normalization.method = 'CLR', margin = 2) %>% 
-  ScaleData() %>% RunPCA(reduction.name = 'apca')
-bm <- RunUMAP(bm, reduction = 'apca', dims = 1:18, assay = 'ADT', 
-              reduction.name = 'adt.umap', reduction.key = 'adtUMAP_')
-DimPlot(bm, reduction = 'adt.umap', group.by = 'celltype.l2', label = TRUE, 
-        repel = TRUE, label.size = 2.5) + NoLegend()
-FeaturePlot(bm, features = c("adt_CD45RA","adt_CD16","adt_CD161"),
-            reduction = 'adt.umap', max.cutoff = 2, ncol = 3)
-
-# ATAC-seq ----------------
-library(SeuratData)
-library(Seurat)
-library(Signac)
-library(EnsDb.Hsapiens.v86)
-library(ggplot2)
-library(cowplot)
-# BiocManager::install("Bioconductor/GenomeInfoDb")
-
-## Load Data & QC -------
-# InstallData("pbmcMultiome")
-pbmc.atac <- LoadData("pbmcMultiome", "pbmc.atac")
-# QC
-pbmc.atac <- subset(pbmc.atac, seurat_annotations != "filtered" & nCount_ATAC < 7e4 & nCount_ATAC > 5e3)
-# Extract_t_cells
-annotated_cell_types = names(table(pbmc.atac$seurat_annotations))
-sub.pbmc.atac <- subset(pbmc.atac, seurat_annotations %in% annotated_cell_types[!grepl("B|HSPC|Mono|Plasma|DC|NK",annotated_cell_types)])
-
-## Process -------
-tiss <- sub.pbmc.atac
-annotations <- GetGRangesFromEnsDb(ensdb = EnsDb.Hsapiens.v86)
[…]

### 2024-07-26 miscellaneous/miscellaneous_code/Demo_smom2.R

@@ -1,545 +1,0 @@
-# https://nbisweden.github.io/workshop-scRNAseq/labs/compiled/seurat/seurat_01_qc.html#Predict_doublets
-#' How to define how many PCs based on Elbow plot
-# https://hbctraining.github.io/scRNA-seq/lessons/elbow_plot_metric.html
-#' The point where the principal components only contribute 5% of standard deviation and the principal components cumulatively contribute 90% of the standard deviation.
-#' The point where the percent change in variation between the consecutive PCs is less than 0.1%.
-
-library(dplyr)
-library(Seurat)
-library(patchwork)
-library(ggplot2)
-library(cowplot)
-library(SeuratWrappers)
-library(RColorBrewer)
-require(scales)
-source("/banach1/ruiqi/local_marker/LocalMarkerDetector/LMD_function.R", echo = F)
-setwd("/banach1/ruiqi/Peggy_data/Peggy_scdata/tmp")
-
-FindPC = function(srat){
-  stdv <- srat[["pca"]]@stdev
-  sum.stdv <- sum(srat[["pca"]]@stdev)
-  percent.stdv <- (stdv / sum.stdv) * 100
-  cumulative <- cumsum(percent.stdv)
-  co1 <- which(cumulative > 90 & percent.stdv < 5)[1]
-  co2 <- sort(which((percent.stdv[1:length(percent.stdv) - 1] - 
-                       percent.stdv[2:length(percent.stdv)]) > 0.1), 
-              decreasing = T)[1] + 1
-  min.pc <- min(co1, co2)
-  return(min.pc)
-}
-
-# Load Data --------------------
-# /data/rq25/peggy_10x_mouse_smom2_202112/seurat_analysis/result/data/
-sample_ls = apply(expand.grid(c("E13.5", "E14.5"), c("MUT", "CTL"), stringsAsFactors = FALSE),1,function(x) paste0(x,collapse = "_"))
-sample_ls = paste0("smom2_dermal_",sample_ls[c(1,3,2,4)])
-for(i in sample_ls){
-  assign(paste0("data_S_",i), 
-         readRDS(sprintf("/data/rq25/peggy_10x_mouse_smom2_202112/seurat_analysis/result/data/data_S_peggy_%s.rds",i)))
-}
-# Preprocess ------------
-folder_path = "/banach1/ruiqi/Peggy_data/Peggy_scdata/240329_smom2"
-for(sample_name in sample_ls){
[…]

### 2024-07-26 mouse_bone_marrow_example.R

@@ -95,6 +95,15 @@ saveRDS(tiss,file = file.path(folder.path,paste0(tissue_name,".rds")))
 saveRDS(local_gene, file.path(folder.path,"intermediate_data",tissue_name,paste0("local_gene_",tissue_name,".rds")))
 saveRDS(res, file.path(folder.path,"intermediate_data",tissue_name,paste0(tissue_name,"_gene_tree.rds")))
 
+## Project CC modules from SmoM2 onto this embedding
+local_gene <- readRDS("/banach1/ruiqi/Peggy_data/Peggy_scdata/240329_smom2/intermediate_data/local_gene_smom2_dermal_E13.5_MUT.rds")
+gene_partition = droplevels(local_gene$gene_partition[local_gene$gene_partition %in% c(1,2,3)])
+levels(gene_partition) = paste0("smom2_",levels(gene_partition))
+tiss = AddModuleActivityScore(tiss, gene_partition = gene_partition, do_local_smooth = FALSE)
+FeaturePlot(tiss, features = paste0("Module",levels(gene_partition)), order = TRUE, reduction = "tsne", ncol = 3) & NoAxes() & 
+  scale_color_gradient(low = "lightgrey", high = "blue", limits = c(0,1)) & labs(color = "ModuleScore") & NoLegend()
+FeaturePlot(tiss, features = paste0("Module",levels(gene_partition)), order = TRUE, reduction = "tsne_cc", ncol = 3) & NoAxes() & 
+  scale_color_gradient(low = "lightgrey", high = "blue", limits = c(0,1)) & labs(color = "ModuleScore") & NoLegend()
 
 # Module-Celltype-alignment ---------
 # Create one-hot matrix of Cell Type

### 2024-07-26 mouse_smoM2_example.R

@@ -182,7 +182,7 @@ lapply(sample_ls,function(sample_name){
 })
 
 # Calculate Jaccard Index of E13.5 mut modules with modules from other samples -------
-modules = setNames(paste0("Module",sample_ls[1],"_",c(1,2,3,12,14,15,16)),
+modules = setNames(paste0("Module",sample_ls[1],"_",c(1,3,2,12,14,15,16)),
                    c(paste0("CC Module",1:3),"Quiescent Module","Wnt Module","DC Module", "High-Wnt Module"))
 jaccard_index_ls = lapply(sample_ls[2:4],function(sample_name){
   data_S = get(paste0("data_S_",sample_name))
@@ -210,7 +210,7 @@ write.csv(df,file = file.path(folder.path,"module_colocalize_align.csv"),row.nam
 df$'nickname' = names(modules)[match(df$module1,modules)]
 module_ls = list(
   modules,
-  setNames(paste0("Module",sample_ls[2],"_",c(17,14,16,15)),
+  setNames(paste0("Module",sample_ls[2],"_",c(17,14,15,16)),
            c("Wnt Module",paste0("CC Module",1:3))),
   setNames(paste0("Module",sample_ls[3],"_",c(15)),
            "Quiescent Module"),
@@ -318,7 +318,7 @@ s.genes = gorth(cc.genes.updated.2019$s.genes, source_organism = "hsapiens", tar
 g2m.genes = gorth(cc.genes.updated.2019$g2m.genes, source_organism = "hsapiens", target_organism = "mmusculus")$ortholog_name
 genes_CC = c(s.genes,g2m.genes)
 
-### Merge MUT & CTL
+# Merge MUT & CTL
 data_S = lapply(sample_ls[1:2],function(sample_name){
   get(paste0("data_S_",sample_name))
 })
@@ -327,19 +327,22 @@ data_S$'type' = ifelse(data_S$old.ident == "E13_SmoM2","E13.5_SmoM2","E13.5_WT")
 data_S <- data_S %>% NormalizeData() %>% 
   FindVariableFeatures() %>% 
   ScaleData(verbose = FALSE, features = rownames(data_S))
-data_S <- data_S  %>% 
+# Top2000 HVGs
+data_S <- data_S %>% 
   RunPCA(npcs = 50, verbose = FALSE, features = VariableFeatures(data_S))
+ndims = FindPC(data_S)
[…]

### 2024-07-26 mouse_smoM2_example_e14_focus.R

@@ -1,389 +1,0 @@
-dir.path <- "/banach1/ruiqi/local_marker/LocalMarkerDetector/"
-source(file.path(dir.path,"LMD_function.R"))
-
-# Define package lists
-bioc_packages <- c("clusterProfiler", "AnnotationDbi", "ReactomePA", "org.Mm.eg.db",
-                   "gprofiler2", "msigdbr")
-github_packages <- c("satijalab/seurat-wrappers");
-names(github_packages) = strsplit(github_packages, "/")[[1]][2]
-names(github_packages)[grepl("seurat-wrappers",names(github_packages))] = "SeuratWrappers"
-# Install packages
-sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
-sapply(1:length(github_packages), function(i) if (!requireNamespace(names(github_packages)[i], quietly = TRUE)) remotes::install_github(github_packages[i]))
-lapply(c(bioc_packages,names(github_packages)), require, character.only = TRUE)
-
-# Load Data --------------------
-sample_ls = unlist(lapply(c("E14.5"),function(i) paste0(i,"_", c("MUT", "CTL"))))
-sample_ls = paste0("smom2_dermal_",sample_ls)
-for(i in sample_ls){
-  assign(paste0("data_S_",i), 
-         readRDS(sprintf("/data/rq25/peggy_10x_mouse_smom2_202112/seurat_analysis/result/data/data_S_peggy_%s.rds",i)))
-}
-# Preprocess ------------
-dir.path = "/banach1/ruiqi/local_marker/LMD_data/smom2"
-dir.create(file.path(dir.path,"process_data"), recursive=T)
-lapply(sample_ls, function(sample_name){
-  data_S <- get(paste0("data_S_",sample_name))
-  data_S <- data_S %>% NormalizeData() %>% FindVariableFeatures() %>%
-    ScaleData(verbose = FALSE) %>% RunPCA(npcs = 50, verbose = FALSE)
-  min.pc = FindPC(srat = data_S)
-  data_S <- RunUMAP(data_S, dims = 1:min.pc, seed.use = 42)
-  data_S <- data_S %>% FindNeighbors(dims = 1:min.pc) %>% 
-    FindClusters(dims = 1:min.pc, res = 1.5)
-  assign(paste0("data_S_",sample_name),data_S, envir = .GlobalEnv)
-})
-
-# Trim & Reembed data
-sample_name = sample_ls[1]
[…]

### 2024-07-31 paper_figure.Rmd
smom2_process_from10x

@@
 modules = module_ls$smom2_dermal_E13.5_MUT["Quiescent Module"]
-thred = 0.5
+# thred = 0.5
 pl = lapply(sample_ls, function(sample_name){
@@
     # }
-    list(p1,p2)
+    list(p1 + theme(aspect.ratio = 0.89),p2 + theme(aspect.ratio = 0.9))
   })
@@
     FeaturePlot(data_S, features = modules, order = TRUE, reduction = "umap") + NoAxes() + 
-    scale_color_gradient(low = "lightgrey", high = "blue", limits = c(0,1)) + labs(title = "", color = "ModuleScore")
+    scale_color_gradient(low = "lightgrey", high = "blue", limits = c(0,1)) + labs(title = "", color = "ModuleScore") + theme(aspect.ratio = 0.9)
   }
@@
 }))
-p = wrap_plots(pl,nrow = 2) + plot_layout(guides = "collect")
+p = wrap_plots(pl,nrow = 2,widths = c(0.88,1,0.88,1,1)) + plot_layout(guides = "collect")
 ggsave(filename = file.path(figure_path, "fig-smom2_2a.png"), plot = p, width = 3*5+3, height = 2*3 + 1)

### 2024-07-31 Process_10x_function.R

@@ -1,0 +1,48 @@
+Load10xData <- function(data_dir, sample_names, sub_rna_dir = "filtered_feature_bc_matrix", with.protein = F, with.tcr = F, sub_tcr_dir = "TCR", human_only = F){
+  data_S_list <- list()
+  data_dir0 <- data_dir
+  if (!with.protein){
+    for(i in sample_names){
+      data_dir <- paste(data_dir0, i, sub_rna_dir, sep = "/")
+      print(data_dir)
+      data_S_list[[i]] <- CreateSeuratObject(Read10X(data.dir = data_dir), project = i)
+    }
+  }
+  if (with.protein){
+    # For output from CellRanger >= 3.0 with multiple data types
+    #list.files(data_dir) # Should show barcodes.tsv.gz, features.tsv.gz, and matrix.mtx.gz
+    data_dir0 <- data_dir
+    for(i in sample_names){
+      data_dir <- paste(data_dir0, i, sub_rna_dir, sep = "/")
+      print(data_dir)
+      data <- Read10X(data.dir = data_dir)
+      data_S_list[[i]] <- CreateSeuratObject(counts = data$`Gene Expression`, project = i)
+      data_S_list[[i]][["ADT"]] <- CreateAssayObject(counts = data$`Antibody Capture`)
+    }
+  }
+  if (with.tcr){
+    for(i in sample_names){
+      data_dir <- paste(data_dir0, i, sub_tcr_dir, sep = "/")
+      print(data_dir)
+      data_S_list[[i]] <- add_clonotype(data_S_list[[i]], data_dir)
+    }
+  }
+  if (human_only){
+    for(i in names(data_S_list)){
+      data_S_list[[i]] <- data_S_list[[i]][grep("^hg19", rownames(data_S_list[[i]])), ]
+    }
+  }
+  if (TRUE){
+    if (length(grep("^mt",rownames(data_S_list[[1]]))) > 0) pattern_mt <- "^mt"
+    if (length(grep("^MT",rownames(data_S_list[[1]]))) > 0) pattern_mt <- "^MT"
+    if (length(grep("^hg19-MT",rownames(data_S_list[[1]]))) > 0) pattern_mt <- "^hg19-MT"
+    print(pattern_mt)
+    for(i in names(data_S_list)){
+      data_S_list[[i]][["percent.mt"]] <- PercentageFeatureSet(
+        object = data_S_list[[i]], pattern = pattern_mt
+      )
+    }
+  }
+  
+  data_S_list
+}

### 2024-07-31 mouse_smoM2_example.R

@@ -17,7 +17,7 @@ sample_ls = unlist(lapply(c("E13.5", "E14.5"),function(i) paste0(i,"_", c("MUT",
 sample_ls = paste0("smom2_dermal_",sample_ls)
 for(i in sample_ls){
   assign(paste0("data_S_",i), 
-         readRDS(sprintf("/data/rq25/peggy_10x_mouse_smom2_202112/seurat_analysis/result/data/data_S_peggy_%s.rds",i)))
+         readRDS(sprintf("/banach1/ruiqi/Peggy_data/peggy_10x_mouse_smom2_202112/seurat_analysis/result/data/data_S_peggy_%s.rds",i)))
 }
 # Preprocess ------------
 dir.path = "/banach1/ruiqi/Peggy_data/Peggy_scdata/240329_smom2"

### 2024-07-31 mouse_smoM2_example_process_from_10x.R

@@ -1,7 +1,61 @@
-###### Preprocess all samples
-###### Rihao Qu
+###### Preprocess all samples by Rihao Qu, original file: /data/rq25/peggy_10x_mouse_smom2_202112/seurat_analysis/seurat.R
+Load10xData <- function(data_dir, sample_names, sub_rna_dir = "filtered_feature_bc_matrix", with.protein = F, with.tcr = F, sub_tcr_dir = "TCR", human_only = F){
+  data_S_list <- list()
+  data_dir0 <- data_dir
+  if (!with.protein){
+    for(i in sample_names){
+      data_dir <- paste(data_dir0, i, sub_rna_dir, sep = "/")
+      print(data_dir)
+      data_S_list[[i]] <- CreateSeuratObject(Read10X(data.dir = data_dir), project = i)
+    }
+  }
+  if (with.protein){
+    # For output from CellRanger >= 3.0 with multiple data types
+    #list.files(data_dir) # Should show barcodes.tsv.gz, features.tsv.gz, and matrix.mtx.gz
+    data_dir0 <- data_dir
+    for(i in sample_names){
+      data_dir <- paste(data_dir0, i, sub_rna_dir, sep = "/")
+      print(data_dir)
+      data <- Read10X(data.dir = data_dir)
+      data_S_list[[i]] <- CreateSeuratObject(counts = data$`Gene Expression`, project = i)
+      data_S_list[[i]][["ADT"]] <- CreateAssayObject(counts = data$`Antibody Capture`)
+    }
+  }
+  if (with.tcr){
+    for(i in sample_names){
+      data_dir <- paste(data_dir0, i, sub_tcr_dir, sep = "/")
+      print(data_dir)
+      data_S_list[[i]] <- add_clonotype(data_S_list[[i]], data_dir)
+    }
+  }
+  if (human_only){
+    for(i in names(data_S_list)){
+      data_S_list[[i]] <- data_S_list[[i]][grep("^hg19", rownames(data_S_list[[i]])), ]
+    }
+  }
+  if (T){
+    if (length(grep("^mt",rownames(data_S_list[[1]]))) > 0) pattern_mt <- "^mt"
+    if (length(grep("^MT",rownames(data_S_list[[1]]))) > 0) pattern_mt <- "^MT"
[…]

### 2024-07-31 Demo.Rmd
dir_path_changing

@@
 # Define your own dir path
-dir.path <- "/banach1/ruiqi/local_marker/LMD_data"
+dir.path <- file.path(dir.path0,"LMD_data")
 

### 2024-07-31 paper_figure.Rmd
dir_path_changing

@@
 source("LMD_function.R", echo = F)
-figure_path = "/banach1/ruiqi/local_marker/LMD_data/figures"
+dir.path0 = "/banach1/ruiqi/local_marker"
+dir.path1 = "/banach1/ruiqi/Peggy_data/Peggy_scdata/240329_smom2"
+figure_path = file.path(dir.path0, "LMD_data","figures")
 dir.create(figure_path,recursive = TRUE)
@@
 ```{r}
-dir.path <- "/banach1/ruiqi/local_marker/LMD_data"
+dir.path <- file.path(dir.path0,"LMD_data")
 consortium = "tabular_muris"
@@
 source("LMD_function.R", echo = F)
-dir.path <- "/banach1/ruiqi/local_marker/LMD_data"
+dir.path <- file.path(dir.path0,"LMD_data")
 consortium = "tabular_muris"
@@
 ```{r}
-dir.path <- "/banach1/ruiqi/local_marker"
+dir.path <- dir.path0
 consortium = "tabular_muris"
@@
 consortium = "tabular_muris"
-dir.path <- "/banach1/ruiqi/local_marker/LMD_data"
+dir.path <- file.path(dir.path0,"LMD_data")
 pl = lapply(c("marrow_facs","lung_facs","pancreas_facs"),function(tissue_name){
@@
 tissue_name = "marrow_facs"
-dir.path <- "/banach1/ruiqi/local_marker/LMD_data"
+dir.path <- file.path(dir.path0,"LMD_data")
 folder.path = file.path(dir.path,consortium)
@@
 ```{r}
-dir.path = "/banach1/ruiqi/Peggy_data/Peggy_scdata/240329_smom2"
+dir.path = dir.path1
 lapply(sample_ls,function(sample_name){

### 2024-07-31 benchmark/compare_method_azimuth.R

@@ -5,7 +5,8 @@ sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) 
 lapply(c(cran_packages,bioc_packages), require, character.only = TRUE)
 
 # Download/Load Data ===========
-dir.path <- "/banach1/ruiqi/local_marker/LMD_data"
+dir.path0 <- "/banach1/ruiqi/local_marker"
+dir.path <- file.path(dir.path0,"LMD_data")
 consortium = "azimuth"
 folder.path <- file.path(dir.path,consortium)
 dir.create(folder.path, recursive=T)
@@ -35,7 +36,7 @@ consortium = "azimuth"
 tissue_name = names(tissue_download_link)[1]
 
 for(tissue_name in names(tissue_download_link)){
-dir.path <- "/banach1/ruiqi/local_marker/LMD_data"
+dir.path <- file.path(dir.path0,"LMD_data")
 folder.path <- file.path(dir.path,consortium)
 tiss <- readRDS(file.path(folder.path,paste0(tissue_name,".rds")))
 DefaultAssay(tiss) <- "RNA"
@@ -76,7 +77,7 @@ dat = dat[selected_genes,,drop = FALSE]
 # anndata::write_h5ad(tiss_adata, file.path(folder_path,gsub(".rds",".h5ad",file_name)))
 
 # Run Each Method and save results ========
-dir.path <- "/banach1/ruiqi/local_marker/LocalMarkerDetector/"
+dir.path <- file.path(dir.path0,"LocalMarkerDetector")
 source(file.path(dir.path,"benchmark","run_methods_function.R"))
 folder.path = file.path(dir.path,"benchmark",consortium)
 dir.create(folder.path, recursive = T)
@@ -122,7 +123,7 @@ write.table(df_runtime,file = file.path(folder.path, paste0(tissue_name,"_runtim
 #' take the maximum value among the N-1 fold change values, 
 #' given N cell types.
 #' 
-dir.path <- "/banach1/ruiqi/local_marker/LMD_data"
+dir.path <- file.path(dir.path0,"LMD_data")
 folder.path <- file.path(dir.path,consortium,"ground_truth_geneset")
 dir.create(folder.path, recursive=T)
 file_name = paste0(tissue_name,"_ground_truth_c1.txt")
@@ -168,7 +169,7 @@ if(!file.exists(file.path(folder.path,file_name))){
 }
 
 # Load Rank Table ==========
[…]

### 2024-07-31 benchmark/compare_method_tabula_muris.R

@@ -5,7 +5,8 @@ sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) 
 lapply(c(cran_packages,bioc_packages), require, character.only = TRUE)
 
 # Download/Load Data ===========
-dir.path <- "/banach1/ruiqi/local_marker"
+dir.path0 <- "/banach1/ruiqi/local_marker"
+dir.path <- dir.path0
 consortium = "tabular_muris"
 folder.path <- file.path(dir.path,"LMD_data",consortium)
 dir.create(folder.path, recursive=T)
@@ -32,7 +33,7 @@ for(tissue_name in names(tissue_download_link)){
 consortium = "tabular_muris"
 tissue_name = names(tissue_download_link)[1]
 
-dir.path <- "/banach1/ruiqi/local_marker"
+dir.path <- dir.path0
 folder.path <- file.path(dir.path,"LMD_data",consortium)
 tiss <- readRDS(file.path(folder.path,paste0(tissue_name,".rds")))
 DefaultAssay(tiss) <- "RNA"
@@ -45,7 +46,7 @@ selected_genes = names(selected_genes)[selected_genes]
 dat = dat[selected_genes,,drop = FALSE]
 
 # Run Each Method and save results ========
-dir.path <- "/banach1/ruiqi/local_marker"
+dir.path <- dir.path0
 source(file.path(dir.path,"LocalMarkerDetector","benchmark","run_methods_function.R"))
 folder.path = file.path(dir.path,"LMD_data","benchmark",consortium)
 dir.create(folder.path, recursive = T)
@@ -92,7 +93,7 @@ write.table(df_runtime,file = file.path(folder.path, paste0(tissue_name,"_runtim
 #' take the maximum value among the N-1 fold change values, 
 #' given N cell types.
 #' 
-dir.path <- "/banach1/ruiqi/local_marker"
+dir.path <- dir.path0
 folder.path <- file.path(dir.path,"LMD_data",consortium,"ground_truth_geneset")
 dir.create(folder.path, recursive=T)
 file_name = paste0(tissue_name,"_ground_truth_c1.txt")
@@ -142,7 +143,7 @@ if(!file.exists(file.path(folder.path,file_name))){
 }
 
 # Load Rank Table ==========
-dir.path <- "/banach1/ruiqi/local_marker"
+dir.path <- dir.path0
[…]

### 2024-07-31 benchmark/run_methods_function.R

@@ -11,8 +11,8 @@ sapply(github_packages, function(pkg) if (!requireNamespace(strsplit(pkg, "/")[[
 lapply(c(cran_packages), require, character.only = TRUE)
 
 #' LMD: Genes were ranked by the increasing order of Cumulative Diffusion-KL score
-dir.path <- "/banach1/ruiqi/local_marker/LocalMarkerDetector"
-source(file.path(dir.path,"LMD_function.R"))
+dir.path0 <- "/banach1/ruiqi/local_marker"
+source(file.path(dir.path0,"LocalMarkerDetector","LMD_function.R"))
 RunLMD <- function(dat, feature_space, dir.file){
   res = LMD(dat,feature_space,max_time = 2^20,knn = 5)
   marker = show_result_lmd(res)$'gene_table'

### 2024-07-31 benchmark/sensitivity_test.R

@@ -5,7 +5,8 @@ sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) 
 lapply(c(cran_packages,bioc_packages), require, character.only = TRUE)
 
 # Load Data =======
-dir.path <- "/banach1/ruiqi/local_marker/"
+dir.path0 <- "/banach1/ruiqi/local_marker"
+dir.path <- dir.path0
 consortium = "tabular_muris"
 tissue_name = "marrow_facs"
 tiss <- readRDS(file.path(dir.path,"LMD_data",consortium,paste0(tissue_name,".rds")))

### 2024-07-31 mouse_smoM2_example.R

@@ -1,5 +1,6 @@
-dir.path <- "/banach1/ruiqi/local_marker/LocalMarkerDetector"
-source(file.path(dir.path,"LMD_function.R"))
+dir.path0 <- "/banach1/ruiqi/local_marker"
+dir.path1 = "/banach1/ruiqi/Peggy_data/Peggy_scdata/240329_smom2"
+source(file.path(dir.path0,"LocalMarkerDetector","LMD_function.R"))
 
 # Define package lists
 bioc_packages <- c("clusterProfiler", "AnnotationDbi", "ReactomePA", "org.Mm.eg.db",
@@ -13,14 +14,15 @@ sapply(1:length(github_packages), function(i) if (!requireNamespace(names(github
 lapply(c(bioc_packages,names(github_packages)), require, character.only = TRUE)
 
 # Load Data --------------------
+dir.path = dir.path1
+folder.path <- file.path(dir.path,"raw_data","preprocess_data")
 sample_ls = unlist(lapply(c("E13.5", "E14.5"),function(i) paste0(i,"_", c("MUT", "CTL"))))
 sample_ls = paste0("smom2_dermal_",sample_ls)
 for(i in sample_ls){
   assign(paste0("data_S_",i), 
-         readRDS(sprintf("/banach1/ruiqi/Peggy_data/peggy_10x_mouse_smom2_202112/seurat_analysis/result/data/data_S_peggy_%s.rds",i)))
+         readRDS(file.path(folder.path,sprintf("data_S_%s_v0.rds",i))))
 }
-# Preprocess ------------
-dir.path = "/banach1/ruiqi/Peggy_data/Peggy_scdata/240329_smom2"
+# Process ------------
 folder.path <- file.path(dir.path,"process_data")
 dir.create(folder.path, recursive=T)
 lapply(sample_ls, function(sample_name){
@@ -28,13 +30,14 @@ lapply(sample_ls, function(sample_name){
   data_S <- data_S %>% NormalizeData() %>% FindVariableFeatures() %>%
     ScaleData(verbose = FALSE) %>% RunPCA(npcs = 50, verbose = FALSE)
   min.pc = FindPC(srat = data_S)
+  cat("# of PC used: ", min.pc, "\n")
   data_S <- RunUMAP(data_S, dims = 1:min.pc, seed.use = 42)
   data_S <- data_S %>% FindNeighbors() %>% FindClusters(res = 1.5)
[…]

### 2024-08-01 paper_figure.Rmd
comment function

@@
 tissue_name = "marrow_facs"
-df_benchmark = read.table(file.path(dir.path, "LocalMarkerDetector","benchmark", consortium, paste0(tissue_name,"_benchmark_rank_table.csv")),header = TRUE)
+df_benchmark = read.table(file.path(dir.path, "LMD_data","benchmark", consortium, paste0(tissue_name,"_benchmark_rank_table.csv")),header = TRUE)
 rownames(df_benchmark) = df_benchmark$gene
@@
 # })
-g = c("Cd19","Naaa","Mogat2","Tifa","Abca13","Zf12","Il2rb","Phgdh","Vpreb2","Fyb")
+g = c("Cd19","Naaa","Mogat2","Grina","Abca13","Zf12","Il2rb","Phgdh","Vpreb2","Fyb")
 names(g) = rep(method_vs_set,each = 2)

### 2024-08-01 LMD_function.R

@@ -8,8 +8,15 @@ bioc_packages <- c("Seurat")
 sapply(cran_packages, function(pkg) if(!requireNamespace(pkg, quietly = TRUE)){install.packages(pkg)})
 sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
 lapply(c(cran_packages,bioc_packages), require, character.only = TRUE)
+# Main Function For LMD: 
+# Construct Graph: Symmetric_KNN_graph, plot_knn_graph
+# Diffusion score: Obtain_Pls, Rowwise_normalize, FeaturePlot_diffusion, fast_get_lmds, Visualize_score_pattern, LMD, show_result_lmd, 
+# Gene modules: Calculate_distance, Obtain_gene_partition, Visualize_gene_heatmap
+# Cell module score: AddModuleActivityScore
 
+# ========================
 # Mtx operation===========
+# ========================
 # Enable C++11
 Sys.setenv("PKG_CXXFLAGS"="-std=c++11")
 # Create the C++ function
@@ -55,39 +62,9 @@ NumericVector fastKLVector(NumericMatrix x, NumericVector y) {
   return out;
 }
 ')
-add_newline_in_middle <- function(input_string, max_length) {
-  if (nchar(input_string) > max_length) {
-    # Find the approximate middle
-    middle <- ceiling(nchar(input_string) / 2)
-    
-    # Find the closest space to the middle
-    space_before <- regexpr(" ", substring(input_string, 1, middle), fixed = TRUE)
-    space_after <- regexpr(" ", substring(input_string, middle + 1), fixed = TRUE)
-    
-    if (space_before == -1 && space_after == -1) {
-      # No spaces found, cannot split without breaking words
-      return(input_string)
-    }
-    
-    if (space_before == -1) {
-      split_pos <- middle + space_after
-    } else if (space_after == -1) {
-      split_pos <- space_before
-    } else {
-      split_pos <- ifelse(middle - space_before < space_after, space_before, middle + space_after)
-    }
-    
-    # Insert newline at the split position
-    output_string <- paste0(
-      substring(input_string, 1, split_pos), "\n", substring(input_string, split_pos + 1)
-    )
-  } else {
-    output_string <- input_string
-  }
-  
-  return(output_string)
-}
 Rowwise_normalize <- function(x){
+  # x: matrix to be rowwise normalized
+  
   x = x[rowSums(x)!=0,,drop = FALSE]
[…]

### 2024-08-03 README.Rmd
initial commit

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# LMD

<!-- badges: start -->
<!-- badges: end -->

The goal of LMD is to ...

## Installation

You can install the released version of LMD from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("LMD")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("ruiqi0130/LocalMarkerDetector")
```
## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(LMD)
## basic example code
```

What is special about using `README.Rmd` instead of just `README.md`? You can include R chunks like so:

```{r cars}
summary(cars)
```

You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this. You could also use GitHub Actions to re-render `README.Rmd` every time you push. An example workflow can be found here: <https://github.com/r-lib/actions/tree/master/examples>.

You can also embed plots, for example:

```{r pressure, echo = FALSE}
plot(pressure)
```

In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.

### 2024-08-03 R/AddModuleActivityScore.R

@@ -1,0 +1,85 @@
+#' Add module activity scores
+#'
+#' @param srat seurat object
+#' @param gene_partition vector; gene modules
+#' @param assay string; asssay used for calculated modulescore
+#' @param do_local_smooth Boolean; TRUE for smooth the module score for each cell by its neighborhoods
+#' @param knn integer; neighborhoods size for local smoothing
+#' @param major_vote integer; majority vote number for local smoothing
+#' @param nloop integer; number of sampling
+#' @param module_names string; prefix name for each gene module
+#' 
+#' 
+#'
+#' @return
+#' A Seurat object
+#'
+#' @export
+#'
+#' @examples
+#'
+#'
+
+
+
+AddModuleActivityScore <- function(srat, gene_partition, assay = "RNA", do_local_smooth = FALSE, knn = 10, major_vote = 6, nloop = 100, module_names = NULL){
+  # srat: seurat object
+  # gene_partition: gene modules
+  # assay: asssay used for calculated modulescore
+  # do_local_smooth: TRUE for smooth the module score for each cell by its neighborhoods
+  # knn: neighborhoods size for local smoothing
+  # major_vote: majority vote number for local smoothing
+  # nloop: number of sampling
+  # module_names: prefix name for each gene module
+  
+  
+  # adjust the format of gene_partition
+  gene_partition = setNames(as.character(gene_partition),names(gene_partition))
+  gene_partition = gene_partition[names(gene_partition) %in% rownames(srat)]
+  gene_partition = gene_partition[!is.na(gene_partition)]
+  gene_partition = as.factor(gene_partition)
+  dat = srat[[assay]]@data[names(gene_partition),]
+  if(do_local_smooth){
+    if ("pca" %in% names(srat@reductions)){
+      ndims = FindPC(srat)
+      feature_space <- Embeddings(srat, reduction = "pca")[,1:ndims]
+      cat(ncol(feature_space),"PC used for building graph for majority vote\n")
+    }else if ("lsi" %in% names(srat@reductions)){
+      ndims = 50
+      feature_space <- Embeddings(srat, reduction = "lsi")[,2:ndims]
+      cat(ncol(feature_space),"LSI used for building graph for majority vote\n")
+    }
+    else{
+      stop("Please RunPCA")
+    }
+    A = Symmetric_KNN_graph(knn = knn, feature_space = feature_space)$'adj_matrix'
[…]

### 2024-08-03 R/Symmetric_KNN_graph.R

@@ -1,0 +1,96 @@
+#' Construct a symmetric KNN graph
+#'
+#' @param knn integer; knn for constructing the graph
+#' @param feature_space matrix; cell x coordinate (e.g. 20PCs) matrix
+#' @param adjust_by_MST boolean; TRUE for connecting disconnected components by Minimum Spanning Trees
+#' @param self-loop integer; weight for self connect (default 1)
+#'
+#'
+#' @return
+#' A list of objects
+#'
+#' @export
+#'
+#' @examples
+#'
+#'
+
+
+Symmetric_KNN_graph <- function(knn = 5, feature_space, adjust_by_MST = TRUE, self_loop = 1){
+  # knn: knn for constructing the graph
+  # feature_space: cell x coordinate (e.g. 20PCs) matrix
+  # adjust_by_MST: TRUE for connecting disconnected components by Minimum Spanning Trees
+  # self-loop: weight for self connect (default 1)
+  
+  cat("Constructing KNN graph\n")
+  knn_list <- FNN::get.knn(feature_space, k = knn, algorithm = "kd_tree")
+  Idx = knn_list$nn.index
+  A = Matrix::sparseMatrix(i = rep(1:nrow(Idx),each = knn),
+                           j = c(t(Idx)),
+                           dims = c(nrow(Idx), nrow(Idx)),
+                           x = 1)
+  rownames(A) = colnames(A) = rownames(feature_space)
+  
+  # remove orphans/singletons
+  res = findDisconnectedComponents(A)
+  filtered_components <- lapply(res$components, function(comp) if (length(comp) >= 10) comp else NULL)
+  filtered_components <- Filter(Negate(is.null), filtered_components)
+  filtered_node_names <- unlist(lapply(filtered_components, function(comp) names(comp)))
+  A = A[rownames(A)%in%filtered_node_names,
+        colnames(A)%in%filtered_node_names]
+  
+  # Connect Components Using MST Principles
+  if(adjust_by_MST){
+    while(length(filtered_components) > 1){
+      cat(length(filtered_components), " Disconnected Components, adjust by MST.\n")
+      edgesToAdd <- lapply(1:(length(filtered_components)-1), function(i) {
+        lapply((i + 1):length(filtered_components), function(j) {
+          findShortestEdge(filtered_components[[i]], filtered_components[[j]], feature_space)
+        })
+      })
[…]

### 2024-08-03 R/plot_knn_graph.R

@@ -1,0 +1,51 @@
+#' Plot a KNN graph
+#'
+#' @param affinity_m matrix; Symmetry affinity matrix of kNN graph
+#' @param label string; Node label
+#' @param layout matrix; 2D coordinate of cells for visualization
+#'
+#' @return
+#' A ggplot object
+#'
+#' @export
+#'
+#' @examples
+#'
+#'
+
+
+plot_knn_graph <- function(affinity_m, label = NULL, layout){
+  # affinity_m: Symmetry affinity matrix of kNN graph
+  # label: Node label
+  # layout: 2D coordinate of cells for visualization
+  
+  layout <- as.matrix(layout)
+  g <- graph_from_adjacency_matrix(affinity_m, 'undirected', weighted = TRUE)
+  V(g)$frame.color <- NA
+  
+  if(!is.null(label)){
+    V(g)$color <- label
+    p = ggraph(g,layout = layout) + 
+      geom_edge_link(aes(width = weight),color = "grey") + 
+      geom_node_point(aes(color = color), size = 1) + 
+      scale_edge_width_continuous(range = c(0.5, 1), breaks = c(0.5, 1)) + 
+      theme_graph()
+    if(class(label) != "numeric"){
+      coldef <- setNames(
+        colorRampPalette(brewer.pal(12, "Paired"))(length(unique(label))),
+        unique(label))
+      p = p + scale_color_manual(values = coldef)
+    }
+  }else{
+    p = ggraph(g,layout = layout) + 
+      geom_edge_link(aes(width = weight),color = "grey") + 
+      geom_node_point(size = 1) + 
+      scale_edge_width_continuous(range = c(0.5, 1), breaks = c(0.5, 1)) + 
+      theme_graph()
+  }
+  p = p + theme(
+    plot.title = element_text(face="bold", size = 30),
+    legend.title = element_text(size = 20),
+    legend.text = element_text(size = 15))
+  return(p)
+}

### 2024-08-11 LMD-paper/paper_figure.Rmd
package_refine

--- 2024-08-01 paper_figure.Rmd
@@
 library("ggsignif")
-source("LMD_function.R", echo = F)
-dir.path0 = "/banach1/ruiqi/local_marker"
-dir.path1 = "/banach1/ruiqi/Peggy_data/Peggy_scdata/240329_smom2"
-figure_path = file.path(dir.path0, "LMD_data","figures")
+library("LocalizedMarkerDetector")
+source("paper_figure_function.R", echo = F)
+source("paths.R")
+figure_path = file.path(dir.path0,"figures")
 dir.create(figure_path,recursive = TRUE)
@@
 ```{r}
-dir.path <- file.path(dir.path0,"LMD_data")
-consortium = "tabular_muris"
-folder.path <- file.path(dir.path,consortium)
+dir.path <- dir.path0
+data_source = "tabular_muris"
+folder.path <- file.path(dir.path,data_source)
 tissue_name = "marrow_facs"
@@
 ```{r}
-source("LMD_function.R", echo = F)
-dir.path <- file.path(dir.path0,"LMD_data")
-consortium = "tabular_muris"
-folder.path <- file.path(dir.path,consortium)
+dir.path <- dir.path0
+data_source = "tabular_muris"
+folder.path <- file.path(dir.path,data_source)
 tissue_name = "marrow_facs"
@@
 # Plot knn graph
-p = plot_knn_graph(affinity_m = W, layout = Embeddings(tiss[["tsne"]])) + NoLegend()
+p = VisualizeGraph(affinity_m = W, layout = Embeddings(tiss[["tsne"]])) + NoLegend()
 ggsave(filename = file.path(figure_path,"fig-pipeline1.png"), plot = p, width = 5, height = 5, dpi = 300)
@@
 check_time = c(0,2^3,2^9,2^14)
-p = Visualize_score_pattern(res$score_profile/res$score_profile$max_score0, genes = genes, add_point = check_time) + guides(linetype = "none")
+p = VisualizeScorePattern(res$score_profile/res$score_profile$max_score0, genes = genes, add_point = check_time) + guides(linetype = "none")
 
@@
   gene = genes[i]
-  FeaturePlot_diffusion(coord = Embeddings(tiss[["tsne"]]),init_state = rho[gene,,drop = FALSE],P_ls = P_ls,W = W,check_time = check_time,gene_name = NULL, gene_color = scales::hue_pal()(length(genes))[i])
+  VisualizeDiffusion(coord = Embeddings(tiss[["tsne"]]),init_state = rho[gene,,drop = FALSE],P_ls = P_ls,W = W,check_time = check_time,gene_name = NULL, gene_color = scales::hue_pal()(length(genes))[i])
 })
@@
 dir.path <- dir.path0
-consortium = "tabular_muris"
+data_source = "tabular_muris"
 tissue_name = "marrow_facs"
-folder.path = file.path(dir.path, "LMD_data","benchmark")
+folder.path = file.path(dir.path,"benchmark")
 ```
@@
                  "azimuth" = c("human_kidney", "human_bone_marrow", "human_pancreas", "mouse_motor_cortex", "human_lung_hlca", "human_motor_cortex"))
-consortium_full_name = c("tabular_muris" = "Tabular Muris","azimuth" = "Azimuth")
+data_source_full_name = c("tabular_muris" = "Tabular Muris","azimuth" = "Azimuth")
 tissue_full_name = lapply(unlist(tissue_ls),function(i){
@@
                   }
-                  df_all$'consortium' = names(tissue_ls)[i]
+                  df_all$'data_source' = names(tissue_ls)[i]
                   df_all}))
@@
 df_auc$Tissue = tissue_full_name[df_auc$Tissue]
-df_auc$consortium = consortium_full_name[df_auc$consortium]
-df_auc$consortium = factor(df_auc$consortium, levels = consortium_full_name)
+df_auc$data_source = data_source_full_name[df_auc$data_source]
+df_auc$data_source = factor(df_auc$data_source, levels = data_source_full_name)
 ```
@@
 fig <- ggplot(df_auc, aes(x = Tissue, y = AUROC, fill = Method)) +
-  geom_bar(stat = "identity", position = "dodge") + facet_grid(gt_set~consortium, scales = 'free') +
+  geom_bar(stat = "identity", position = "dodge") + facet_grid(gt_set~data_source, scales = 'free') +
   geom_text(aes(label = round(AUROC, 3)), position = position_dodge(width = 1), vjust = -0.25, size = 2.5, angle = 45) +
@@
 wide_data = df_auc; wide_data$AUROC = round(wide_data$AUROC,3)
-wide_data <- reshape2::dcast(wide_data, gt_set + Tissue + consortium ~ Method, value.var = "AUROC")
+wide_data <- reshape2::dcast(wide_data, gt_set + Tissue + data_source ~ Method, value.var = "AUROC")
 colnames(wide_data)[1:3] = c("Criterion","Tissue","DataSource")
@@
 df_runtime$Tissue = tissue_full_name[df_runtime$Tissue]
-df_runtime$consortium = consortium_full_name[df_runtime$consortium]
+df_runtime$data_source = data_source_full_name[df_runtime$data_source]
 ```
@@
 dir.path <- dir.path0
-consortium = "tabular_muris"
+data_source = "tabular_muris"
 tissue_name = "marrow_facs"
-df_benchmark = read.table(file.path(dir.path, "LMD_data","benchmark", consortium, paste0(tissue_name,"_benchmark_rank_table.csv")),header = TRUE)
+df_benchmark = read.table(file.path(dir.path,"benchmark", data_source, paste0(tissue_name,"_benchmark_rank_table.csv")),header = TRUE)
 rownames(df_benchmark) = df_benchmark$gene
-tiss = readRDS(file.path(dir.path, "LMD_data",consortium, paste0(tissue_name,".rds")))
+tiss = readRDS(file.path(dir.path,data_source, paste0(tissue_name,".rds")))
 method_vs = "lmd"
@@
 <!--   clustering_method = "average" -->
-<!--   dist = Calculate_distance(dat_alra[local_gene,], method = distance_method) -->
+<!--   dist = CalculateGeneDistance(dat_alra[local_gene,], method = distance_method) -->
 
@@
 <!--                 silent = TRUE) -->
-<!--   pl_module = FeaturePlot_meta(dat,visual_space,gene_partition) -->
+<!--   pl_module = CustomModulePlot(dat,visual_space,gene_partition) -->
 <!--   legend <- get_legend(pl_module[[1]]) -->
@@
 ```{r}
-consortium = "tabular_muris"
-dir.path <- file.path(dir.path0,"LMD_data")
+data_source = "tabular_muris"
+dir.path <- dir.path0
 pl = lapply(c("marrow_facs","lung_facs","pancreas_facs"),function(tissue_name){
-score_measure = read.table(file.path(dir.path,"benchmark",consortium,paste0(tissue_name,"_DI.csv")),header = TRUE)
+score_measure = read.table(file.path(dir.path,"benchmark",data_source,paste0(tissue_name,"_DI.csv")),header = TRUE)
 score_measure$Method = ifelse(score_measure$Method %in% names(method_fullname),method_fullname[score_measure$Method],score_measure$Method)
@@
 null_score_measure =
-  read.table(file.path(dir.path,"benchmark",consortium,paste0(tissue_name,"_DI_null.csv")),header = TRUE)
+  read.table(file.path(dir.path,"benchmark",data_source,paste0(tissue_name,"_DI_null.csv")),header = TRUE)
 # Calculate the mean & Prediction Interval
@@
 ```{r}
-folder.path = file.path(dir.path,"LMD_data",consortium,"intermediate_data", tissue_name, "sensitivity_test")
+folder.path = file.path(dir.path,data_source,"intermediate_data", tissue_name, "sensitivity_test")
 test_para = "knn"
@@
 ```{r}
-folder.path = file.path(dir.path,"LMD_data",consortium,"intermediate_data", tissue_name, "sensitivity_test")
+folder.path = file.path(dir.path,data_source,"intermediate_data", tissue_name, "sensitivity_test")
 test_para = "pc"
@@
 # load("./marrow_result/module_result.rdata")
-source("LMD_function.R", echo = F)
+library(LocalizedMarkerDetector)
 # celltype_color = setNames(colorRampPalette(brewer.pal(9, "Set1"))(length(unique(cell_label))),unique(cell_label))
-consortium = "tabular_muris"
+data_source = "tabular_muris"
 tissue_name = "marrow_facs"
-dir.path <- file.path(dir.path0,"LMD_data")
-folder.path = file.path(dir.path,consortium)
+dir.path <- dir.path0
+folder.path = file.path(dir.path,data_source)
 dist = readRDS(file.path(folder.path,"intermediate_data",tissue_name,paste0(tissue_name,"_gene_dist.rds")))
@@
 ```{r}
-p1 = Visualize_gene_heatmap(dist, gene_partition = res$gene_partition_rename, gene_hree = res$gene_hree)
+p1 = VisualizeGeneHeatmap(dist, gene_partition = res$gene_partition_rename, gene_hree = res$gene_hree)
 ```
@@
 ```{r}
-consortium = "tabular_muris"
+data_source = "tabular_muris"
 tissue_name = "marrow_facs"
-dir.path <- file.path(dir.path0,"LMD_data")
-folder.path = file.path(dir.path,consortium)
+dir.path <- dir.path0
+folder.path = file.path(dir.path,data_source)
 
@@
 pl_ct = lapply(colnames(one_hot_matrix),function(ct){
-  FeaturePlot_custom(coord = Embeddings(tiss[["tsne"]]),value = as.character(one_hot_matrix[,ct]), title_name = ct, value_name = "Indicator") + scale_color_manual(values = c("0" = "lightgrey", "1" = scales::hue_pal()(1)))
+  CustomFeaturePlot(coord = Embeddings(tiss[["tsne"]]),value = as.character(one_hot_matrix[,ct]), title_name = ct, value_name = "Indicator") + scale_color_manual(values = c("0" = "lightgrey", "1" = scales::hue_pal()(1)))
 })
@@
 ```{r}
-consortium = "tabular_muris"
+data_source = "tabular_muris"
 tissue_name = "marrow_facs"
-dir.path <- file.path(dir.path0,"LMD_data")
-folder.path = file.path(dir.path,consortium)
+dir.path <- dir.path0
+folder.path = file.path(dir.path,data_source)
 res = readRDS(file.path(folder.path,"intermediate_data",tissue_name,paste0(tissue_name,"_gene_tree.rds")))
@@
 ```{r}
-dir.path = dir.path1
+data_source = "smom2"
+dir.path = dir.path0
+dir.path = file.path(dir.path,data_source)
 lapply(sample_ls,function(sample_name){

### 2024-08-11 README.Rmd
package_refine

@@
   comment = "#>",
-  fig.path = "man/figures/README-",
+  fig.path = "man/figures",
   out.width = "100%"
@@
 
-# LMD
+# Localized Marker Detector
 
@@
 
-The goal of LMD is to ...
+Localized Marker Detector (LMD) is a computational framework designed for the identification of gene expression markers localized to specific cell populations within single-cell RNA sequencing data.
+The major workflow of LMD comprises the following three main steps:
+
+* Step1. Constructing a cell-cell affinity graph
+* Step2. Diffusing the gene expression value across the cell graph
+* Step3. Assigning a score to each gene based on the dynamics of its diffusion process
+* Optional Downstream tasks
+  * Identifying gene modules and characterizing functional cell groups
+  * Cross-sample comparison
+
+```{r pressure, echo=FALSE, out.width="50%", out.height="auto"}
+knitr::include_graphics("./man/figures/LMD_workflow.png")
+```
 
@@
 
-You can install the released version of LMD from [CRAN](https://CRAN.R-project.org) with:
+LMD can be installed in R as follows:
 
 ``` r
-install.packages("LMD")
+install.packages("devtools")
+devtools::install_github("ruiqi0130/LocalizedMarkerDetector")
+
+library("LocalizedMarkerDetector")
 ```
 
-And the development version from [GitHub](https://github.com/) with:
+## Example tutorial
+Please check [LMD tutorial](https://ruiqi0130.github.io/LocalizedMarkerDetector/articles/).
 
-``` r
-# install.packages("devtools")
-devtools::install_github("ruiqi0130/LocalMarkerDetector")
-```
-## Example
+## References
+References of LMD functions can be found [here](https://ruiqi0130.github.io/LocalizedMarkerDetector/reference/index.html).
 
-This is a basic example which shows you how to solve a common problem:
-
-```{r example}
-library(LMD)
-## basic example code
-```
-
-What is special about using `README.Rmd` instead of just `README.md`? You can include R chunks like so:
-
-```{r cars}
-summary(cars)
-```
-
-You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date.
-`devtools::build_readme()` is handy for this.
-You could also use GitHub Actions to re-render `README.Rmd` every time you push.
-An example workflow can be found here: <https://github.com/r-lib/actions/tree/master/examples>.
-
-You can also embed plots, for example:
-
-```{r pressure, echo = FALSE}
-plot(pressure)
-```
-
-In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.
+<!-- Data used in this tutorial can be downloaded from [Tabula Muris](https://figshare.com/ndownloader/files/13092380). -->

### 2024-08-11 .build_test/.build_site.R

@@ -1,0 +1,36 @@
+library(devtools)
+library(pkgdown)
+
+setwd("/banach1/ruiqi/LocalizedMarkerDetector_package")
+
+# Create Template
+use_readme_rmd()
+# use_news_md()
+use_vignette("Example")
+
+# Build Package
+devtools::build_readme()
+Rcpp::compileAttributes()
+devtools::document()
+# load all function w/o install
+devtools::load_all(".") 
+
+
+# Render
+rmarkdown::render("vignettes/LMD_demo.Rmd")
+pkgdown::clean_site()
+
+# if (!dir.exists("docs/articles")) {
+#   dir.create("docs/articles", recursive = TRUE)
+# }
+# file.copy("vignettes/LMD_demo.html", "docs/articles/LMD_demo.html", overwrite = TRUE)
+
+pkgdown::build_site()
+
+
+# Install Package from local
+devtools::install() # install package
+
+# Install Package from github
+devtools::install_github("ruiqi0130/LocalizedMarkerDetector")
+

### 2024-08-11 .build_test/build_site.R

@@ -1,13 +1,0 @@
-require(devtools)
-require(pkgdown)
-
-setwd("../")
-
-use_readme_rmd()
-use_news_md()
-#use_vignette("Example") 
-devtools::document()
-
-pkgdown::clean_site()
-pkgdown::build_site()
-

### 2024-08-11 LMD-paper/analysis/mouse_bone_marrow_example.R

@@ -1,0 +1,266 @@
+setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
+source("../paths.R")
+library(LocalizedMarkerDetector)
+# Define package lists
+bioc_packages <- c("clusterProfiler", "AnnotationDbi", "ReactomePA", "org.Mm.eg.db","gprofiler2", "msigdbr")
+github_packages <- c("satijalab/seurat-wrappers");
+names(github_packages) = strsplit(github_packages, "/")[[1]][2]
+names(github_packages)[grepl("seurat-wrappers",names(github_packages))] = "SeuratWrappers"
+# Install packages
+sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
+sapply(1:length(github_packages), function(i) if (!requireNamespace(names(github_packages)[i], quietly = TRUE)) remotes::install_github(github_packages[i]))
+lapply(c(bioc_packages,names(github_packages)), require, character.only = TRUE)
+
+# Download/Load Data ===========
+dir.path <- dir.path0
+folder.path <- file.path(dir.path,"tabular_muris")
+dir.create(folder.path, recursive=T)
+tissue_download_link = c(
+  "marrow_facs" = "https://figshare.com/ndownloader/files/13092380",
+  "marrow_droplet" = "https://figshare.com/ndownloader/files/ 13089821"
+)
+for(tissue_name in names(tissue_download_link)){
+  file_name = paste0(tissue_name,".rds")
+  if (!file.exists(file.path(folder.path,file_name))) {
+    options(timeout=6000)
+    download.file(tissue_download_link[tissue_name],
+                  destfile = file.path(folder.path,file_name), method = 'libcurl')
+    # update version
+    tiss <- readRDS(file.path(folder.path,file_name))
+    tiss <- UpdateSeuratObject(tiss)
+    tiss <- RunALRA(tiss, assay = "RNA"); DefaultAssay(tiss) = "RNA"
+    saveRDS(tiss,file = file.path(folder.path,file_name))
+  }
+}
+
+# Prepare input data ===========
+tissue_name = "marrow_facs"
+tiss <- readRDS(file.path(folder.path,paste0(tissue_name,".rds")))
+DefaultAssay(tiss) <- "RNA"
+n_dim = 20
[…]

### 2024-08-11 LMD-paper/analysis/mouse_smoM2_example.R

@@ -1,0 +1,396 @@
+setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
+source("../paths.R")
+dir.path = file.path(dir.path0,"smom2")
+library(LocalizedMarkerDetector)
+
+# Define package lists
+bioc_packages <- c("clusterProfiler", "AnnotationDbi", "ReactomePA", "org.Mm.eg.db",
+                   "gprofiler2", "msigdbr")
+github_packages <- c("satijalab/seurat-wrappers");
+names(github_packages) = strsplit(github_packages, "/")[[1]][2]
+names(github_packages)[grepl("seurat-wrappers",names(github_packages))] = "SeuratWrappers"
+# Install packages
+sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
+sapply(1:length(github_packages), function(i) if (!requireNamespace(names(github_packages)[i], quietly = TRUE)) remotes::install_github(github_packages[i]))
+lapply(c(bioc_packages,names(github_packages)), require, character.only = TRUE)
+
+# Load Data --------------------
+folder.path <- file.path(dir.path,"raw_data","preprocess_data")
+sample_ls = unlist(lapply(c("E13.5", "E14.5"),function(i) paste0(i,"_", c("MUT", "CTL"))))
+sample_ls = paste0("smom2_dermal_",sample_ls)
+for(i in sample_ls){
+  assign(paste0("data_S_",i), 
+         readRDS(file.path(folder.path,sprintf("data_S_%s_v0.rds",i))))
+}
+# Process ------------
+folder.path <- file.path(dir.path,"process_data")
+dir.create(folder.path, recursive=T)
+lapply(sample_ls, function(sample_name){
+  data_S <- get(paste0("data_S_",sample_name))
+  data_S <- data_S %>% NormalizeData() %>% FindVariableFeatures() %>%
+    ScaleData(verbose = FALSE) %>% RunPCA(npcs = 50, verbose = FALSE)
+  min.pc = FindPC(srat = data_S)
+  cat("# of PC used: ", min.pc, "\n")
+  data_S <- RunUMAP(data_S, dims = 1:min.pc, seed.use = 42)
+  data_S <- data_S %>% FindNeighbors() %>% FindClusters(res = 1.5)
+  assign(paste0("data_S_",sample_name),data_S, envir = .GlobalEnv)
+})
+
[…]

### 2024-08-11 LMD-paper/analysis/mouse_smoM2_example_process_from_10x.R

@@ -1,0 +1,166 @@
+###### Preprocess all samples
+# by Rihao Qu
+# original file: /data/rq25/peggy_10x_mouse_smom2_202112/seurat_analysis/seurat.R
+# original raw data file: /data/rq25/peggy_10x_mouse_smom2_202112/
+
+#####
+##### Load required packages & functions for analysis
+#####
+library(Seurat)
+library(scales)
+library(dplyr)
+library(viridis)
+library(grid)
+Load10xData <- function(data_dir, sample_names, sub_rna_dir = "filtered_feature_bc_matrix"){
+  data_S_list <- list()
+  data_dir0 <- data_dir
+  if (TRUE){
+    for(i in sample_names){
+      data_dir <- file.path(data_dir0, i, sub_rna_dir)
+      print(data_dir)
+      data_S_list[[i]] <- CreateSeuratObject(Read10X(data.dir = data_dir), project = i)
+    }
+  }
+  if (TRUE){
+    if (length(grep("^mt",rownames(data_S_list[[1]]))) > 0) pattern_mt <- "^mt"
+    if (length(grep("^MT",rownames(data_S_list[[1]]))) > 0) pattern_mt <- "^MT"
+    if (length(grep("^hg19-MT",rownames(data_S_list[[1]]))) > 0) pattern_mt <- "^hg19-MT"
+    print(pattern_mt)
+    for(i in names(data_S_list)){
+      data_S_list[[i]][["percent.mt"]] <- PercentageFeatureSet(
+        object = data_S_list[[i]], pattern = pattern_mt
+      )
+    }
+  }
+  
+  data_S_list
+}
+MergeData <- function(data_S_list){
+  sample_names <- names(data_S_list)
+  n_sample <- length(sample_names)
+  data_S <- merge(
+    x = data_S_list[[1]], data_S_list[sample_names[2:n_sample]]
+  )
+  data_S
+}
+FilterCells <- function(data_S_list, ngene_lth = NULL, ngene_hth = NULL, mt_hth, ngene_lqth = NULL, ngene_hqth = NULL){
+  if (is.null(ngene_lth)){
+    for(i in names(data_S_list)){
+      data_S_list[[i]]$ngene_lth <- quantile(data_S_list[[i]]$nFeature_RNA,ngene_lqth)
+      data_S_list[[i]]$ngene_hth <- quantile(data_S_list[[i]]$nFeature_RNA,ngene_hqth)
[…]

### 2024-08-11 LMD-paper/benchmark/compare_method_azimuth.R

@@ -1,0 +1,201 @@
+cran_packages <- c("readxl","dplyr","BiocManager","RColorBrewer")
+bioc_packages <- c("Seurat")
+sapply(cran_packages, function(pkg) if(!requireNamespace(pkg, quietly = TRUE)){install.packages(pkg)})
+sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
+lapply(c(cran_packages,bioc_packages), require, character.only = TRUE)
+
+# Download/Load Data ===========
+setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
+source("../paths.R")
+
+dir.path <- dir.path0
+data_source = "azimuth"
+folder.path <- file.path(dir.path,data_source)
+dir.create(folder.path, recursive=T)
+tissue_download_link = c(
+  "human_kidney" = "https://seurat.nygenome.org/azimuth/demo_datasets/kidney_demo_stewart.rds",
+  "human_bone_marrow" = "https://seurat.nygenome.org/azimuth/demo_datasets/bmcite_demo.rds",
+  "human_pancreas" = "https://seurat.nygenome.org/azimuth/demo_datasets/enge.rds",
+  "mouse_motor_cortex" = "https://seurat.nygenome.org/azimuth/demo_datasets/allen_mop_2020.rds",
+  "human_lung_hlca" = "https://seurat.nygenome.org/hlca_ref_files/ts_opt.rds",
+  "human_motor_cortex" = "https://seurat.nygenome.org/azimuth/demo_datasets/allen_m1c_2019_ssv4.rds"
+)
+for(tissue_name in names(tissue_download_link)){
+  file_name = paste0(tissue_name,".rds")
+  if (!file.exists(file.path(folder.path,file_name))) {
+    options(timeout=6000)
+    download.file(tissue_download_link[tissue_name],
+                  destfile = file.path(folder.path,file_name), method = 'libcurl')
+    # update version
+    tiss <- readRDS(file.path(folder.path,file_name))
+    tiss <- UpdateSeuratObject(tiss)
+    saveRDS(tiss,file = file.path(folder.path,file_name))
+  }
+}
+
+# Prepare input data ===========
+data_source = "azimuth"
+tissue_name = names(tissue_download_link)[1]
+
+for(tissue_name in names(tissue_download_link)){
+dir.path <- dir.path0
+folder.path <- file.path(dir.path,data_source)
[…]

### 2024-08-11 LMD-paper/benchmark/run_methods_function.R

@@ -1,0 +1,230 @@
+# Install Marcopolo: pip install marcopolo-pytorch --upgrade
+# Install SEMITONES: pip install https://github.com/ohlerlab/SEMITONES/archive/master.zip
+# Install Hotspot: pip install git+https://github.com/yoseflab/Hotspot.git
+# Install singleCellHaystack: install.packages("singleCellHaystack")
+
+cran_packages <- c("reticulate","singleCellHaystack")
+github_packages <- c("immunogenomics/presto")
+sapply(cran_packages, function(pkg) if(!requireNamespace(pkg, quietly = TRUE)){install.packages(pkg)})
+# sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
+sapply(github_packages, function(pkg) if (!requireNamespace(strsplit(pkg, "/")[[1]][2], quietly = TRUE)) remotes::install_github(pkg))
+lapply(c(cran_packages), require, character.only = TRUE)
+
+#' LMD: Genes were ranked by the increasing order of Cumulative Diffusion-KL score
+library(LocalizedMarkerDetector)
+RunLMD <- function(dat, feature_space, dir.file){
+  res = LMD(dat,feature_space,max_time = 2^20,knn = 5)
+  marker = show_result_lmd(res)$'gene_table'
+  write.table(marker,file = dir.file,row.names = TRUE)
+}
+
+#' HVG
+RunHVG <- function(srat_obj, input_genes, dir.file){
+  DefaultAssay(srat_obj) <- "RNA"
+  marker <- subset(srat_obj,features = input_genes) %>% ScaleData(features = input_genes) %>% 
+    FindVariableFeatures(nfeatures = length(input_genes)) %>% VariableFeatures() %>% as.data.frame() %>% setNames(c("gene")) %>% mutate(rank = 1:n())
+  write.table(marker,file = dir.file,row.names = FALSE)
+}
+
+#' SinglecellHaystack: 
+#' https://alexisvdb.github.io/singleCellHaystack/articles/examples/a02_example_scRNAseq.html
+#' Genes were ranked based on the increasing order of log(adjusted P-value)
+RunHaystack <- function(dat, feature_space, dir.file = NULL){
+  set.seed(123)
+  res <- haystack(x = feature_space, expression = dat)
+  marker <- show_result_haystack(res.haystack = res)
[…]

### 2024-08-11 LMD-paper/benchmark/sensitivity_test.R

@@ -1,0 +1,95 @@
+cran_packages <- c("readxl","dplyr","BiocManager")
+bioc_packages <- c("Seurat")
+sapply(cran_packages, function(pkg) if(!requireNamespace(pkg, quietly = TRUE)){install.packages(pkg)})
+sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
+lapply(c(cran_packages,bioc_packages), require, character.only = TRUE)
+
+# Load Data =======
+setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
+source("../paths.R")
+
+dir.path <- dir.path0
+data_source = "tabular_muris"
+tissue_name = "marrow_facs"
+tiss <- readRDS(file.path(dir.path,data_source,paste0(tissue_name,".rds")))
+DefaultAssay(tiss) <- "RNA"
+dat = as.matrix(tiss[[DefaultAssay(tiss)]]@data)
+Gene_detected_count <- apply(dat > apply(dat,2,median),1,sum)
+selected_genes = (Gene_detected_count >= 10) & (Gene_detected_count <= ncol(dat) * 0.5)
+selected_genes = names(selected_genes)[selected_genes]
+dat = dat[selected_genes,,drop = FALSE]
+
+
+# kNN sensitivity Test --------
+folder.path = file.path(dir.path,data_source,"intermediate_data",
+                        tissue_name, "sensitivity_test")
+dir.create(folder.path,recursive = T)
+library(LocalizedMarkerDetector)
+# Use Default PCs
+n_dim = 20
+feature_space = Embeddings(tiss[["pca"]])[,1:n_dim]
+score_df = do.call(cbind,lapply(c(3:10,20,30,40,50),function(kNN){
+  try({
+    res = LMD(dat,feature_space,knn = kNN)
+    return(res$cumulative_score)
+  },silent = TRUE)
+}))
+colnames(score_df) = paste0("kNN",c(3:10,20,30,40,50))
+write.table(score_df,file = file.path(folder.path,paste0(tissue_name,"_different_knn.csv")),row.names = TRUE)
+
+# PC sensitivity Test --------
+folder.path = file.path(dir.path,data_source,"intermediate_data",
+                        tissue_name, "sensitivity_test")
+library(LocalizedMarkerDetector)
+# Generate PC embeddings
+DefaultAssay(tiss) = "RNA"
[…]

### 2024-08-11 LMD-paper/paper_figure_function.R

@@ -1,0 +1,71 @@
+add_newline_in_middle <- function(input_string, max_length) {
+  if (nchar(input_string) > max_length) {
+    # Find the approximate middle
+    middle <- ceiling(nchar(input_string) / 2)
+    
+    # Find the closest space to the middle
+    space_before <- regexpr(" ", substring(input_string, 1, middle), fixed = TRUE)
+    space_after <- regexpr(" ", substring(input_string, middle + 1), fixed = TRUE)
+    
+    if (space_before == -1 && space_after == -1) {
+      # No spaces found, cannot split without breaking words
+      return(input_string)
+    }
+    
+    if (space_before == -1) {
+      split_pos <- middle + space_after
+    } else if (space_after == -1) {
+      split_pos <- space_before
+    } else {
+      split_pos <- ifelse(middle - space_before < space_after, space_before, middle + space_after)
+    }
+    
+    # Insert newline at the split position
+    output_string <- paste0(
+      substring(input_string, 1, split_pos), "\n", substring(input_string, split_pos + 1)
+    )
+  } else {
+    output_string <- input_string
+  }
+  
+  return(output_string)
+}
+
+Visualize_jaccard_mtx <- function(jaccard_mtx, threshold = 0.4){
+  if(is.null(rownames(jaccard_mtx))|is.null(colnames(jaccard_mtx))){
+    stop("Please define celltype names & module names")
+  }
+  df = reshape2::melt(jaccard_mtx)
+  colnames(df)[1:2] = c("CellType","Module")
+  df$CellType = factor(df$CellType,levels = rownames(jaccard_mtx))
+  df$Module = factor(df$Module,levels = rev(colnames(jaccard_mtx)))
+  
+  p <- ggplot(df, aes(x=CellType, y=Module, fill=value)) +
+    geom_tile(color="white") +
+    scale_fill_gradientn(colors=colorRampPalette(rev(c("firebrick3", "white")))(99),limits = c(0,1), name="Jaccard Index") +
+    theme_minimal() +
+    theme(
+      axis.text.x = element_text(angle=45, hjust=1,size = 15),
+      axis.text.y = element_text(size = 15),
+      legend.title = element_text(size = 20),
+      legend.text = element_text(size = 15),
+      axis.title = element_blank(),
+      panel.grid.major = element_blank(),
+      panel.grid.minor = element_blank()
[…]

### 2024-08-11 LMD-paper/paths.R

@@ -1,0 +1,1 @@
+dir.path0 <- "/banach1/ruiqi/local_marker/LMD_data"

### 2024-08-11 R/AddModuleActivityScore.R

@@ -1,85 +1,0 @@
-#' Add module activity scores
-#'
-#' @param srat seurat object
-#' @param gene_partition vector; gene modules
-#' @param assay string; asssay used for calculated modulescore
-#' @param do_local_smooth Boolean; TRUE for smooth the module score for each cell by its neighborhoods
-#' @param knn integer; neighborhoods size for local smoothing
-#' @param major_vote integer; majority vote number for local smoothing
-#' @param nloop integer; number of sampling
-#' @param module_names string; prefix name for each gene module
-#' 
-#' 
-#'
-#' @return
-#' A Seurat object
-#'
-#' @export
-#'
-#' @examples
-#'
-#'
-
-
-
-AddModuleActivityScore <- function(srat, gene_partition, assay = "RNA", do_local_smooth = FALSE, knn = 10, major_vote = 6, nloop = 100, module_names = NULL){
-  # srat: seurat object
-  # gene_partition: gene modules
-  # assay: asssay used for calculated modulescore
-  # do_local_smooth: TRUE for smooth the module score for each cell by its neighborhoods
-  # knn: neighborhoods size for local smoothing
-  # major_vote: majority vote number for local smoothing
-  # nloop: number of sampling
-  # module_names: prefix name for each gene module
-  
-  
-  # adjust the format of gene_partition
-  gene_partition = setNames(as.character(gene_partition),names(gene_partition))
-  gene_partition = gene_partition[names(gene_partition) %in% rownames(srat)]
-  gene_partition = gene_partition[!is.na(gene_partition)]
-  gene_partition = as.factor(gene_partition)
-  dat = srat[[assay]]@data[names(gene_partition),]
-  if(do_local_smooth){
-    if ("pca" %in% names(srat@reductions)){
-      ndims = FindPC(srat)
-      feature_space <- Embeddings(srat, reduction = "pca")[,1:ndims]
-      cat(ncol(feature_space),"PC used for building graph for majority vote\n")
-    }else if ("lsi" %in% names(srat@reductions)){
-      ndims = 50
-      feature_space <- Embeddings(srat, reduction = "lsi")[,2:ndims]
-      cat(ncol(feature_space),"LSI used for building graph for majority vote\n")
-    }
-    else{
-      stop("Please RunPCA")
-    }
-    A = Symmetric_KNN_graph(knn = knn, feature_space = feature_space)$'adj_matrix'
[…]

### 2024-08-11 R/FeaturePlot_diffusion.R

@@ -1,91 +1,0 @@
-#' Visualize diffusion 
-#'
-#' @param coord matrix; 2D coordinate of cells for visualization
-#' @param init_state matrix; Row-wise normalized expression matrix
-#' @param P_ls matrix; Diffused transition matrix
-#' @param W matrix; Symmetry affinity matrix of kNN graph
-#' @param check_time vector; Select several time point for visualization, e.g. c(2,16,128)
-#' @param gene_name string; Select one gene for visualization
-#' @param gene_color string; The color of expression level for visualization
-#' 
-#' 
-#'
-#' @return
-#' A ggplot object
-#'
-#' @export
-#'
-#' @examples
-#'
-#'
-
-
-
-FeaturePlot_diffusion <- function(coord, init_state, P_ls = NULL, W = NULL, check_time = NULL, gene_name = NULL, gene_color = "blue"){
-  # coord: 2D coordinate of cells for visualization
-  # init_state: Row-wise normalized expression matrix
-  # P_ls: Diffused transition matrix
-  # W: Symmetry affinity matrix of kNN graph
-  # check_time: Select several time point for visualization, e.g. c(2,16,128)
-  # gene_name: Select one gene for visualization
-  # gene_color: The color of expression level for visualization
-  
-  init_state = as.matrix(init_state)
-  tmp <- if (!is.null(P_ls) && length(P_ls) > 0) {
-    as.matrix(P_ls[[1]])
-  } else if (!is.null(W)) {
-    W
-  } else {
-    NULL
-  }
-  if(is.null(tmp) & is.null(W)){stop("Cell Graph 'W' not found")}
-  graph_node_number = ncol(tmp); rm(tmp)
-  if((ncol(init_state)!=graph_node_number) & (nrow(init_state) == graph_node_number)){
-    init_state = t(init_state)
-  }
-  if(ncol(init_state)!=graph_node_number | nrow(init_state)!=1){stop("Unmatched dimension, can only take expression vector for one gene!")}
-  check_step = sort(unique(floor(log(check_time,base = 2))))
-  check_step = check_step[check_step >= 0]
-  
-  if(0 %in% check_time){
-    check_time = c(0,2^check_step)
-  }else{
-    check_time = 2^check_step
-  }
-  if(is.null(P_ls)){P_ls = Obtain_Pls(W,max(check_time))}
-  check_time = check_time[check_time %in% names(P_ls)]
[…]

### 2024-08-11 R/RcppExports.R

@@ -1,0 +1,15 @@
+# Generated by using Rcpp::compileAttributes() -> do not edit by hand
+# Generator token: 10BE3573-1514-4C36-9D1C-5A225CD40393
+
+fastKLMatrix <- function(x, y) {
+    .Call(`_LocalizedMarkerDetector_fastKLMatrix`, x, y)
+}
+
+fastKLVector <- function(x, y) {
+    .Call(`_LocalizedMarkerDetector_fastKLVector`, x, y)
+}
+
+fastMatMult <- function(A, B) {
+    .Call(`_LocalizedMarkerDetector_fastMatMult`, A, B)
+}
+

### 2024-08-11 R/Symmetric_KNN_graph.R

@@ -1,96 +1,0 @@
-#' Construct a symmetric KNN graph
-#'
-#' @param knn integer; knn for constructing the graph
-#' @param feature_space matrix; cell x coordinate (e.g. 20PCs) matrix
-#' @param adjust_by_MST boolean; TRUE for connecting disconnected components by Minimum Spanning Trees
-#' @param self-loop integer; weight for self connect (default 1)
-#'
-#'
-#' @return
-#' A list of objects
-#'
-#' @export
-#'
-#' @examples
-#'
-#'
-
-
-Symmetric_KNN_graph <- function(knn = 5, feature_space, adjust_by_MST = TRUE, self_loop = 1){
-  # knn: knn for constructing the graph
-  # feature_space: cell x coordinate (e.g. 20PCs) matrix
-  # adjust_by_MST: TRUE for connecting disconnected components by Minimum Spanning Trees
-  # self-loop: weight for self connect (default 1)
-  
-  cat("Constructing KNN graph\n")
-  knn_list <- FNN::get.knn(feature_space, k = knn, algorithm = "kd_tree")
-  Idx = knn_list$nn.index
-  A = Matrix::sparseMatrix(i = rep(1:nrow(Idx),each = knn),
-                           j = c(t(Idx)),
-                           dims = c(nrow(Idx), nrow(Idx)),
-                           x = 1)
-  rownames(A) = colnames(A) = rownames(feature_space)
-  
-  # remove orphans/singletons
-  res = findDisconnectedComponents(A)
-  filtered_components <- lapply(res$components, function(comp) if (length(comp) >= 10) comp else NULL)
-  filtered_components <- Filter(Negate(is.null), filtered_components)
-  filtered_node_names <- unlist(lapply(filtered_components, function(comp) names(comp)))
-  A = A[rownames(A)%in%filtered_node_names,
-        colnames(A)%in%filtered_node_names]
-  
-  # Connect Components Using MST Principles
-  if(adjust_by_MST){
-    while(length(filtered_components) > 1){
-      cat(length(filtered_components), " Disconnected Components, adjust by MST.\n")
-      edgesToAdd <- lapply(1:(length(filtered_components)-1), function(i) {
-        lapply((i + 1):length(filtered_components), function(j) {
-          findShortestEdge(filtered_components[[i]], filtered_components[[j]], feature_space)
-        })
-      })
[…]

### 2024-08-11 R/calculate_lmds.R

@@ -1,0 +1,512 @@
+# =====================================
+# Calculating LMD Score ===============
+# =====================================
+# Extrinsic Function: fast_get_lmds, LMD, show_result_lmd
+
+#' Compute Score Profile
+#'
+#' Computes various scores for state transitions based on different metrics such as KL divergence and entropy.
+#'
+#' @param state_0 matrix; the initial state matrix.
+#' @param state_t matrix; the state matrix at time t.
+#' @param state_inf matrix; the state matrix at equilibrium.
+#' @param P_diag_t matrix; the diagonal elements of the diffusion operator at time t.
+#' @param score_ls character vector; a list of scores to compute. Possible values are "score0", "max_score0", "score1", "delta_correction", and "entropy". Default is \code{c("score0", "delta_correction")}.
+#' \itemize{
+#'   \item "score0": KL divergence between the initial state (\code{state_0}) and the state at time t (\code{state_t}).
+#'   \item "max_score0": KL divergence between the initial state (\code{state_0}) and the equilibrium state (\code{state_inf}).
+#'   \item "score1": KL divergence between the state at time t (\code{state_t}) and the equilibrium state (\code{state_inf}).
+#'   \item "delta_correction": KL divergence between the initial state (\code{state_0}) and the diagonal elements of the diffusion operator at time t (\code{diag(P^t)}).
+#'   \item "entropy": Entropy of the state at time t (\code{state_t}).
+#' }
+#' @return A data frame containing the computed scores, with each column representing a score.
+#' 
+#' @keywords internal
+#' 
+#' @export
+get_score_profile <- function(state_0 = NULL, state_t = NULL, state_inf = NULL, P_diag_t = NULL,
+                              score_ls = c("score0","delta_correction")){
+  # score_ls = c("score0","score1","delta_correction","entropy")
+  df = do.call(cbind,lapply(score_ls,function(s){
+    if(s == "score0"){
+      # KL(p^0||p^t)
+      return(fastKLMatrix(state_0, state_t))
+    }
+    if(s == "max_score0"){
+      return(fastKLVector(state_0,state_inf))
+    }
+    if(s == "score1"){
+      # KL(p^t||p^inf)
+      return(fastKLVector(state_t, state_inf))
+    }
+    if(s == "delta_correction"){
+      # KL(p^0||p_delta^t)
+      return(fastKLVector(state_0, P_diag_t))
+    }
+    if(s == "entropy"){
+      # H(pt)
[…]

### 2024-08-11 R/calculate_module_score.R

@@ -1,0 +1,161 @@
+# ======================================
+# Obtain Modules Activity Score ========
+# ======================================
+# Extrinsic Function: AddModuleActivityScore
+
+Obtain_cell_partition <- function(expr_dat, gene_partition, 
+                                  cell_kNN_graph = NULL, major_vote = 5){
+  # Fit GMM
+  if(sum(!names(gene_partition) %in% rownames(expr_dat))){
+    stop("gene name doesn't match data")}
+  
+  meta_g = sapply(levels(gene_partition),function(topic){
+    if(sum(gene_partition == topic, na.rm = TRUE) > 1){
+      return(colMeans(expr_dat[names(gene_partition)[gene_partition == topic],,drop = FALSE]))
+    }else{
+      return(NULL)
+    }
+  })
+  rownames(meta_g) = colnames(expr_dat)
+  
+  # tic()
+  cell_block = apply(meta_g,2,function(vec){GMM_partition(vec)})
+  # toc()
+  # cell_block = as.data.frame(meta_g) %>% reframe(across(everything(), GMM_partition))
+  cell_block = as.matrix(cell_block) + 0
+  
+  # Local smooth
+  if(!is.null(cell_kNN_graph)){
+    cell_block = cell_kNN_graph %*% cell_block
+    cell_block = (cell_block >= major_vote) + 0
+  }
+  
+  colnames(cell_block) = levels(gene_partition)
+  rownames(cell_block) = colnames(expr_dat)
+  return(cell_block)
+}
+GMM_partition <- function(vector){
+  # cat("bottleneck\n")
+  i = 1
+  opt = ClusterR::Optimal_Clusters_GMM(as.matrix(vector),max_clusters = 10, criterion = "BIC", km_iter = 10, em_iter = 5, seed = i, plot_data = FALSE)
+  opt_num = which.min(opt)
+  gmm = ClusterR::GMM(as.matrix(vector),gaussian_comps = opt_num, km_iter = 10, em_iter = 5, seed = i)
+  labels = stats::predict(gmm,as.matrix(vector))
+  
+  
+  if(length(unique(labels))>2){
+    centroid = setNames(gmm$centroids[,1],1:nrow(gmm$centroids))
+    
+    # Merge each cluster into two groups using hclust
+    clusters = cutree(hclust(dist(centroid), method = "average"), k = 2)
+    
+    labels = labels %in% names(clusters)[clusters == max(clusters)]
+  }else{
[…]

### 2024-08-11 R/construct_cell_graph.R

@@ -1,0 +1,223 @@
+# =========================================
+# Construct Cell Graph ====================
+# =========================================
+# Extrinsic Function: ConstructKnnGraph, ConstructGaussianGraph, ConstructDiffusionOperators
+
+findDisconnectedComponents <- function(adj_matrix) {
+  # Create a graph from the adjacency matrix
+  g <- igraph::graph_from_adjacency_matrix(adj_matrix, mode = "undirected")
+  # Find the connected components
+  components <- igraph::components(g)
+  # Get the number of disconnected components
+  num_components <- components$no
+  # Get the corresponding nodes for each component
+  component_nodes <- split(igraph::V(g), components$membership)
+  list(number_of_components = num_components, components = component_nodes)
+}
+findShortestEdge <- function(component1, component2, data) {
+  # Calculate all pairwise distances between nodes in the two components
+  distances <- outer(component1, component2, Vectorize(function(x, y) dist(data[c(x,y), ])))
+  # Find the minimum distance and the corresponding nodes
+  min_distance_idx <- which(distances == min(distances), arr.ind = TRUE)
+  return(c(component1[min_distance_idx[1]], component2[min_distance_idx[2]]))
+}
+
+#' Construct a Symmetric KNN graph
+#'
+#' Constructs a symmetric KNN graph from a given feature space.
+#'
+#' @param knn integer; the number of nearest neighbors to consider for constructing the graph. Default is 5.
+#' @param feature_space matrix; a cell-by-coordinate matrix (e.g., 20 principal components).
+#' @param adjust_by_MST boolean; TRUE for connecting disconnected components using Minimum Spanning Trees (MST). Default is TRUE.
+#' @param self_loop integer; weight for self connections (default is 1).
+#'
+#' @return A list containing the following elements:
+#' \describe{
+#'   \item{graph}{matrix; symmetric KNN graph W, computed as \code{pmax(1, (A + A^T) / 2)}.}
+#'   \item{adj_matrix}{matrix; adjacency matrix A.}
+#'   \item{component}{list; disconnected components in the graph, each component is represented as a subgraph.}
+#' }
+#'
+#' @examples
[…]

### 2024-08-11 R/matrix_operation.R

@@ -1,0 +1,176 @@
+# ========================
+# Mtx operation===========
+# ========================
+
+# # Enable C++11
+# Sys.setenv("PKG_CXXFLAGS"="-std=c++11")
+# # Create the C++ function
+# # requireNamespace("Rcpp")
+# Rcpp::cppFunction('arma::mat fastMatMult(arma::mat A, arma::mat B) {
+#   arma::mat C = A * B;
+#   return C;
+# }', depends="RcppArmadillo")
+
+
+# Rcpp::sourceCpp(code='
+# #include <Rcpp.h>
+# using namespace Rcpp;
+# NumericVector calcKL(NumericVector p, NumericVector q) {
+#   double sum = 0;
+#   for (int j = 0; j < p.size(); ++j) {
+#     if(p[j] == 0) continue;
+#     double val = p[j] * log(p[j] / q[j]);
+#     if (NumericVector::is_na(val) || std::isinf(val)) val = 0.0;
+#     sum += val;
+#   }
+#   return NumericVector::create(sum);
+# }
+# 
+# // [[Rcpp::export]]
+# NumericVector fastKLMatrix(NumericMatrix x, NumericMatrix y) {
+#   int n = x.nrow();
+#   NumericVector out(n);
+# 
+#   for (int i = 0; i < n; ++i) {
+#     out[i] = calcKL(x.row(i), y.row(i))[0];
+#   }
+#   return out;
+# }
+# 
+# // [[Rcpp::export]]
+# NumericVector fastKLVector(NumericMatrix x, NumericVector y) {
+#   int n = x.nrow();
+#   NumericVector out(n);
+# 
+#   for (int i = 0; i < n; ++i) {
+#     out[i] = calcKL(x.row(i), y)[0];
+#   }
+#   return out;
+# }
+# ')
+
+#' Fast Matrix Multiplication
+#'
+#' This function multiplies two matrices using Armadillo.
+#'
+#' @param A A numeric matrix.
+#' @param B A numeric matrix.
+#' @return A numeric matrix which is the product of A and B.
+#' @export
+#' @useDynLib LocalizedMarkerDetector, .registration = TRUE
+#' @import Rcpp
+#' @importFrom Rcpp sourceCpp
+#' @name fastMatMult
+#' @keywords internal
+#' @examples
+#' A <- matrix(1:4, 2, 2)
+#' B <- matrix(5:8, 2, 2)
+#' fastMatMult(A, B)
+NULL
+
+#' Fast KL Divergence Calculation for Matrices
+#'
[…]

### 2024-08-11 R/utils.R

@@ -1,0 +1,74 @@
+# =======================================
+# Miscellaneous Test Function =============
+# =======================================
+# Extrinsic Function: FindPC
+
+#' Determine Number of Principal Components
+#'
+#' Defines the number of principal components (PCs) to retain based on the elbow plot criteria.
+#' The function uses two criteria: 
+#' 1) The point where the principal components only contribute 5% of the standard deviation and the principal components cumulatively contribute 90% of the standard deviation.
+#' 2) The point where the percent change in variation between consecutive PCs is less than 0.1%.
+#'
+#' @param srat A Seurat object containing the PCA reduction.
+#' @param reduction Character; the name of the reduction technique to use. Default is "pca".
+#'
+#' @return Integer; the number of principal components to retain.
+#'
+#' @references
+#' For more details on the elbow plot method, refer to the [Harvard Chan Bioinformatics Core scRNA-seq training](https://hbctraining.github.io/scRNA-seq/lessons/elbow_plot_metric.html).
+#'
+#' @keywords internal
+#' @export
+FindPC = function(srat, reduction = "pca"){
+  stdv <- srat[[reduction]]@stdev
+  sum.stdv <- sum(srat[[reduction]]@stdev)
+  percent.stdv <- (stdv / sum.stdv) * 100
+  cumulative <- cumsum(percent.stdv)
+  co1 <- which(cumulative > 90 & percent.stdv < 5)[1]
+  co2 <- sort(which((percent.stdv[1:length(percent.stdv) - 1] - 
+                       percent.stdv[2:length(percent.stdv)]) > 0.1), 
+              decreasing = T)[1] + 1
+  min.pc <- min(co1, co2)
+  return(min.pc)
+}
+
+
+generate_pseudo_gene <- function(data){
+  randseed = 233
+  set.seed(randseed)
+  seed <- sample(c(1:1e5),size=1e5)
+  
+  dat.detection = data > apply(data,2,median)
+  each.gene.detection = apply(dat.detection,1,sum)
+  
+  # sample genes based on the hist of each.gene.detection
+  xhist=hist(each.gene.detection,breaks = 1000,plot = FALSE)
+  each.gene.bin = ceiling(each.gene.detection / unique(diff(xhist$breaks)))
+  
+  # choose bins where each sample belongs to
[…]

### 2024-08-11 scripts/LMD_function.R

@@ -1,1423 +1,0 @@
-cran_packages <- c("igraph","ggplot2", "cowplot", "RColorBrewer", "scales",
-                   "data.table", "dplyr", "patchwork", "philentropy",
-                   "pheatmap", "ggplotify", "ggraph", 
-                   "ClusterR", "Rcpp", "RcppArmadillo", 
-                   "tictoc","svMisc","RSpectra","Matrix","usedist","dendextend",
-                   "progress","latex2exp","parallel","dbscan","Rtsne","ggrepel", "BiocManager", "remotes")
-bioc_packages <- c("Seurat")
-sapply(cran_packages, function(pkg) if(!requireNamespace(pkg, quietly = TRUE)){install.packages(pkg)})
-sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
-lapply(c(cran_packages,bioc_packages), require, character.only = TRUE)
-# Main Function For LMD: 
-# Construct Graph: Symmetric_KNN_graph, plot_knn_graph
-# Diffusion score: Obtain_Pls, Rowwise_normalize, FeaturePlot_diffusion, fast_get_lmds, Visualize_score_pattern, LMD, show_result_lmd, 
-# Gene modules: Calculate_distance, Obtain_gene_partition, Visualize_gene_heatmap
-# Cell module score: AddModuleActivityScore
-
-# ========================
-# Mtx operation===========
-# ========================
-# Enable C++11
-Sys.setenv("PKG_CXXFLAGS"="-std=c++11")
-# Create the C++ function
-cppFunction('arma::mat fastMatMult(arma::mat A, arma::mat B) {
-  arma::mat C = A * B;
-  return C;
-}', depends="RcppArmadillo")
-
-
-Rcpp::sourceCpp(code='
-#include <Rcpp.h>
-using namespace Rcpp;
-NumericVector calcKL(NumericVector p, NumericVector q) {
-  double sum = 0;
-  for (int j = 0; j < p.size(); ++j) {
-    if(p[j] == 0) continue;
-    double val = p[j] * log(p[j] / q[j]);
-    if (NumericVector::is_na(val) || std::isinf(val)) val = 0.0;
-    sum += val;
-  }
-  return NumericVector::create(sum);
-}
-
-// [[Rcpp::export]]
-NumericVector fastKLMatrix(NumericMatrix x, NumericMatrix y) {
-  int n = x.nrow();
-  NumericVector out(n);
-
[…]

### 2024-08-11 scripts/analysis/mouse_bone_marrow_example.R

@@ -1,265 +1,0 @@
-dir.path0 <- "/banach1/ruiqi/local_marker"
-source(file.path(dir.path0,"LocalMarkerDetector","LMD_function.R"))
-# Define package lists
-bioc_packages <- c("clusterProfiler", "AnnotationDbi", "ReactomePA", "org.Mm.eg.db","gprofiler2", "msigdbr")
-github_packages <- c("satijalab/seurat-wrappers");
-names(github_packages) = strsplit(github_packages, "/")[[1]][2]
-names(github_packages)[grepl("seurat-wrappers",names(github_packages))] = "SeuratWrappers"
-# Install packages
-sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
-sapply(1:length(github_packages), function(i) if (!requireNamespace(names(github_packages)[i], quietly = TRUE)) remotes::install_github(github_packages[i]))
-lapply(c(bioc_packages,names(github_packages)), require, character.only = TRUE)
-
-# Download/Load Data ===========
-dir.path <- file.path(dir.path0,"LMD_data")
-folder.path <- file.path(dir.path,"tabular_muris")
-dir.create(folder.path, recursive=T)
-tissue_download_link = c(
-  "marrow_facs" = "https://figshare.com/ndownloader/files/13092380",
-  "marrow_droplet" = "https://figshare.com/ndownloader/files/ 13089821"
-)
-for(tissue_name in names(tissue_download_link)){
-  file_name = paste0(tissue_name,".rds")
-  if (!file.exists(file.path(folder.path,file_name))) {
-    options(timeout=6000)
-    download.file(tissue_download_link[tissue_name],
-                  destfile = file.path(folder.path,file_name), method = 'libcurl')
-    # update version
-    tiss <- readRDS(file.path(folder.path,file_name))
-    tiss <- UpdateSeuratObject(tiss)
-    tiss <- RunALRA(tiss, assay = "RNA"); DefaultAssay(tiss) = "RNA"
-    saveRDS(tiss,file = file.path(folder.path,file_name))
-  }
-}
-
-# Prepare input data ===========
-tissue_name = "marrow_facs"
-tiss <- readRDS(file.path(folder.path,paste0(tissue_name,".rds")))
-DefaultAssay(tiss) <- "RNA"
-n_dim = 20
[…]

### 2024-08-11 scripts/analysis/mouse_smoM2_example.R

@@ -1,396 +1,0 @@
-dir.path0 <- "/banach1/ruiqi/local_marker"
-dir.path1 = "/banach1/ruiqi/Peggy_data/Peggy_scdata/240329_smom2"
-source(file.path(dir.path0,"LocalMarkerDetector","LMD_function.R"))
-
-# Define package lists
-bioc_packages <- c("clusterProfiler", "AnnotationDbi", "ReactomePA", "org.Mm.eg.db",
-                   "gprofiler2", "msigdbr")
-github_packages <- c("satijalab/seurat-wrappers");
-names(github_packages) = strsplit(github_packages, "/")[[1]][2]
-names(github_packages)[grepl("seurat-wrappers",names(github_packages))] = "SeuratWrappers"
-# Install packages
-sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
-sapply(1:length(github_packages), function(i) if (!requireNamespace(names(github_packages)[i], quietly = TRUE)) remotes::install_github(github_packages[i]))
-lapply(c(bioc_packages,names(github_packages)), require, character.only = TRUE)
-
-# Load Data --------------------
-dir.path = dir.path1
-folder.path <- file.path(dir.path,"raw_data","preprocess_data")
-sample_ls = unlist(lapply(c("E13.5", "E14.5"),function(i) paste0(i,"_", c("MUT", "CTL"))))
-sample_ls = paste0("smom2_dermal_",sample_ls)
-for(i in sample_ls){
-  assign(paste0("data_S_",i), 
-         readRDS(file.path(folder.path,sprintf("data_S_%s_v0.rds",i))))
-}
-# Process ------------
-folder.path <- file.path(dir.path,"process_data")
-dir.create(folder.path, recursive=T)
-lapply(sample_ls, function(sample_name){
-  data_S <- get(paste0("data_S_",sample_name))
-  data_S <- data_S %>% NormalizeData() %>% FindVariableFeatures() %>%
-    ScaleData(verbose = FALSE) %>% RunPCA(npcs = 50, verbose = FALSE)
-  min.pc = FindPC(srat = data_S)
-  cat("# of PC used: ", min.pc, "\n")
-  data_S <- RunUMAP(data_S, dims = 1:min.pc, seed.use = 42)
-  data_S <- data_S %>% FindNeighbors() %>% FindClusters(res = 1.5)
[…]

### 2024-08-11 scripts/benchmark/compare_method_azimuth.R

@@ -1,200 +1,0 @@
-cran_packages <- c("readxl","dplyr","BiocManager","RColorBrewer")
-bioc_packages <- c("Seurat")
-sapply(cran_packages, function(pkg) if(!requireNamespace(pkg, quietly = TRUE)){install.packages(pkg)})
-sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
-lapply(c(cran_packages,bioc_packages), require, character.only = TRUE)
-
-# Download/Load Data ===========
-dir.path0 <- "/banach1/ruiqi/local_marker"
-dir.path <- file.path(dir.path0,"LMD_data")
-consortium = "azimuth"
-folder.path <- file.path(dir.path,consortium)
-dir.create(folder.path, recursive=T)
-tissue_download_link = c(
-  "human_kidney" = "https://seurat.nygenome.org/azimuth/demo_datasets/kidney_demo_stewart.rds",
-  "human_bone_marrow" = "https://seurat.nygenome.org/azimuth/demo_datasets/bmcite_demo.rds",
-  "human_pancreas" = "https://seurat.nygenome.org/azimuth/demo_datasets/enge.rds",
-  "mouse_motor_cortex" = "https://seurat.nygenome.org/azimuth/demo_datasets/allen_mop_2020.rds",
-  "human_lung_hlca" = "https://seurat.nygenome.org/hlca_ref_files/ts_opt.rds",
-  "human_motor_cortex" = "https://seurat.nygenome.org/azimuth/demo_datasets/allen_m1c_2019_ssv4.rds"
-)
-for(tissue_name in names(tissue_download_link)){
-  file_name = paste0(tissue_name,".rds")
-  if (!file.exists(file.path(folder.path,file_name))) {
-    options(timeout=6000)
-    download.file(tissue_download_link[tissue_name],
-                  destfile = file.path(folder.path,file_name), method = 'libcurl')
-    # update version
-    tiss <- readRDS(file.path(folder.path,file_name))
-    tiss <- UpdateSeuratObject(tiss)
-    saveRDS(tiss,file = file.path(folder.path,file_name))
-  }
-}
-
-# Prepare input data ===========
-consortium = "azimuth"
-tissue_name = names(tissue_download_link)[1]
-
-for(tissue_name in names(tissue_download_link)){
-dir.path <- file.path(dir.path0,"LMD_data")
-folder.path <- file.path(dir.path,consortium)
[…]

### 2024-08-11 scripts/benchmark/compare_method_tabula_muris.R

@@ -1,260 +1,0 @@
-cran_packages <- c("readxl","dplyr","BiocManager","RColorBrewer")
-bioc_packages <- c("Seurat")
-sapply(cran_packages, function(pkg) if(!requireNamespace(pkg, quietly = TRUE)){install.packages(pkg)})
-sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
-lapply(c(cran_packages,bioc_packages), require, character.only = TRUE)
-
-# Download/Load Data ===========
-dir.path0 <- "/banach1/ruiqi/local_marker"
-dir.path <- dir.path0
-consortium = "tabular_muris"
-folder.path <- file.path(dir.path,"LMD_data",consortium)
-dir.create(folder.path, recursive=T)
-tissue_download_link = c(
-  "marrow_facs" = "https://figshare.com/ndownloader/files/13092380",
-  "pancreas_facs" = "https://figshare.com/ndownloader/files/13092386",
-  "lung_facs" = "https://figshare.com/ndownloader/files/13092194",
-  "marrow_droplet" = "https://figshare.com/ndownloader/files/ 13089821"
-)
-for(tissue_name in names(tissue_download_link)){
-  file_name = paste0(tissue_name,".rds")
-  if (!file.exists(file.path(folder.path,file_name))) {
-    options(timeout=6000)
-    download.file(tissue_download_link[tissue_name],
-                  destfile = file.path(folder.path,file_name), method = 'libcurl')
-    # update version
-    tiss <- readRDS(file.path(folder.path,file_name))
-    tiss <- UpdateSeuratObject(tiss)
-    saveRDS(tiss,file = file.path(folder.path,file_name))
-  }
-}
-
-# Prepare input data ===========
-consortium = "tabular_muris"
-tissue_name = names(tissue_download_link)[1]
-
-dir.path <- dir.path0
-folder.path <- file.path(dir.path,"LMD_data",consortium)
-tiss <- readRDS(file.path(folder.path,paste0(tissue_name,".rds")))
-DefaultAssay(tiss) <- "RNA"
-n_dim = 20
-feature_space = Embeddings(tiss[["pca"]])[,1:n_dim]
-dat = as.matrix(tiss[[DefaultAssay(tiss)]]@data)
-Gene_detected_count <- apply(dat > apply(dat,2,median),1,sum)
[…]

### 2024-08-11 scripts/benchmark/sensitivity_test.R

@@ -1,93 +1,0 @@
-cran_packages <- c("readxl","dplyr","BiocManager")
-bioc_packages <- c("Seurat")
-sapply(cran_packages, function(pkg) if(!requireNamespace(pkg, quietly = TRUE)){install.packages(pkg)})
-sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
-lapply(c(cran_packages,bioc_packages), require, character.only = TRUE)
-
-# Load Data =======
-dir.path0 <- "/banach1/ruiqi/local_marker"
-dir.path <- dir.path0
-consortium = "tabular_muris"
-tissue_name = "marrow_facs"
-tiss <- readRDS(file.path(dir.path,"LMD_data",consortium,paste0(tissue_name,".rds")))
-DefaultAssay(tiss) <- "RNA"
-dat = as.matrix(tiss[[DefaultAssay(tiss)]]@data)
-Gene_detected_count <- apply(dat > apply(dat,2,median),1,sum)
-selected_genes = (Gene_detected_count >= 10) & (Gene_detected_count <= ncol(dat) * 0.5)
-selected_genes = names(selected_genes)[selected_genes]
-dat = dat[selected_genes,,drop = FALSE]
-
-
-# kNN sensitivity Test --------
-folder.path = file.path(dir.path,"LMD_data",consortium,"intermediate_data",
-                        tissue_name, "sensitivity_test")
-dir.create(folder.path,recursive = T)
-source(file.path(dir.path,"LocalMarkerDetector","LMD_function.R"))
-# Use Default PCs
-n_dim = 20
-feature_space = Embeddings(tiss[["pca"]])[,1:n_dim]
-score_df = do.call(cbind,lapply(c(3:10,20,30,40,50),function(kNN){
-  try({
-    res = LMD(dat,feature_space,knn = kNN)
-    return(res$cumulative_score)
-  },silent = TRUE)
-}))
-colnames(score_df) = paste0("kNN",c(3:10,20,30,40,50))
-write.table(score_df,file = file.path(folder.path,paste0(tissue_name,"_different_knn.csv")),row.names = TRUE)
-
-# PC sensitivity Test --------
-folder.path = file.path(dir.path,"LMD_data",consortium,"intermediate_data",
-                        tissue_name, "sensitivity_test")
[…]

### 2024-08-11 src/RcppExports.cpp

@@ -1,0 +1,61 @@
+// Generated by using Rcpp::compileAttributes() -> do not edit by hand
+// Generator token: 10BE3573-1514-4C36-9D1C-5A225CD40393
+
+#include <RcppArmadillo.h>
+#include <Rcpp.h>
+
+using namespace Rcpp;
+
+#ifdef RCPP_USE_GLOBAL_ROSTREAM
+Rcpp::Rostream<true>&  Rcpp::Rcout = Rcpp::Rcpp_cout_get();
+Rcpp::Rostream<false>& Rcpp::Rcerr = Rcpp::Rcpp_cerr_get();
+#endif
+
+// fastKLMatrix
+NumericVector fastKLMatrix(NumericMatrix x, NumericMatrix y);
+RcppExport SEXP _LocalizedMarkerDetector_fastKLMatrix(SEXP xSEXP, SEXP ySEXP) {
+BEGIN_RCPP
+    Rcpp::RObject rcpp_result_gen;
+    Rcpp::RNGScope rcpp_rngScope_gen;
+    Rcpp::traits::input_parameter< NumericMatrix >::type x(xSEXP);
+    Rcpp::traits::input_parameter< NumericMatrix >::type y(ySEXP);
+    rcpp_result_gen = Rcpp::wrap(fastKLMatrix(x, y));
+    return rcpp_result_gen;
+END_RCPP
+}
+// fastKLVector
+NumericVector fastKLVector(NumericMatrix x, NumericVector y);
+RcppExport SEXP _LocalizedMarkerDetector_fastKLVector(SEXP xSEXP, SEXP ySEXP) {
+BEGIN_RCPP
+    Rcpp::RObject rcpp_result_gen;
+    Rcpp::RNGScope rcpp_rngScope_gen;
+    Rcpp::traits::input_parameter< NumericMatrix >::type x(xSEXP);
+    Rcpp::traits::input_parameter< NumericVector >::type y(ySEXP);
+    rcpp_result_gen = Rcpp::wrap(fastKLVector(x, y));
+    return rcpp_result_gen;
+END_RCPP
+}
+// fastMatMult
+arma::mat fastMatMult(arma::mat A, arma::mat B);
+RcppExport SEXP _LocalizedMarkerDetector_fastMatMult(SEXP ASEXP, SEXP BSEXP) {
+BEGIN_RCPP
+    Rcpp::RObject rcpp_result_gen;
+    Rcpp::RNGScope rcpp_rngScope_gen;
[…]

### 2024-08-11 src/fastKL.cpp

@@ -1,0 +1,34 @@
+#include <Rcpp.h>
+using namespace Rcpp;
+NumericVector calcKL(NumericVector p, NumericVector q) {
+  double sum = 0;
+  for (int j = 0; j < p.size(); ++j) {
+    if(p[j] == 0) continue;
+    double val = p[j] * log(p[j] / q[j]);
+    if (NumericVector::is_na(val) || std::isinf(val)) val = 0.0;
+    sum += val;
+  }
+  return NumericVector::create(sum);
+}
+
+// [[Rcpp::export]]
+NumericVector fastKLMatrix(NumericMatrix x, NumericMatrix y) {
+  int n = x.nrow();
+  NumericVector out(n);
+  
+  for (int i = 0; i < n; ++i) {
+    out[i] = calcKL(x.row(i), y.row(i))[0];
+  }
+  return out;
+}
+
+// [[Rcpp::export]]
+NumericVector fastKLVector(NumericMatrix x, NumericVector y) {
+  int n = x.nrow();
+  NumericVector out(n);
+  
+  for (int i = 0; i < n; ++i) {
+    out[i] = calcKL(x.row(i), y)[0];
+  }
+  return out;
+}

### 2024-08-11 src/fastMatMult.cpp

@@ -1,0 +1,8 @@
+#include <RcppArmadillo.h>
+// [[Rcpp::depends(RcppArmadillo)]]
+
+// [[Rcpp::export]]
+arma::mat fastMatMult(arma::mat A, arma::mat B) {
+  arma::mat C = A * B;
+  return C;
+}

### 2024-08-11 .build_test/.build_site.R
add_citation_adjust_figure

@@ -5,6 +5,7 @@ setwd("/banach1/ruiqi/LocalizedMarkerDetector_package")
 
 # Create Template
 use_readme_rmd()
+usethis::use_mit_license("Ruiqi Li")
 # use_news_md()
 use_vignette("Example")
 
@@ -17,7 +18,7 @@ devtools::load_all(".")
 
 
 # Render
-rmarkdown::render("vignettes/LMD_demo.Rmd")
+# rmarkdown::render("vignettes/LMD_demo.Rmd")
 pkgdown::clean_site()
 
 # if (!dir.exists("docs/articles")) {

### 2024-08-18 .build_test/.build_site.R
08/17

@@ -5,11 +5,11 @@ setwd("/banach1/ruiqi/local_marker/LocalizedMarkerDetector")
 
 # Create Template
 use_readme_rmd()
-usethis::use_mit_license("Ruiqi Li")
 # use_news_md()
 use_vignette("Example")
 
 # Build Package
+usethis::use_mit_license("Ruiqi Li")
 devtools::build_readme()
 Rcpp::compileAttributes()
 devtools::document()

### 2024-08-18 R/calculate_lmds.R

@@ -123,7 +123,7 @@ fast_calculate_score_profile <- function(W, max_time = 2^15,
 #' @export
 fast_calculate_score_profile_largeData <- function(W, max_time = 2^15, init_state,
                                                    score_ls = c("score0")){
-  cat("Calculate LMD score profile for large data\n")
+  cat("Calculate LMD score profile for large data...\n")
   if((ncol(W) != ncol(init_state)) & (ncol(W) == nrow(init_state))){
     init_state = t(init_state)
   }
@@ -133,7 +133,7 @@ fast_calculate_score_profile_largeData <- function(W, max_time = 2^15, init_stat
   final_state = rep(1/nrow(W),nrow(W)) # bi-stochastic
   
   # Calculate transition matrix
-  cat("Doubly Stochastic\n")
+  cat("Run doubly stochastic on affinity matrix...\n")
   P = Doubly_stochastic(W)
   # cat("Adjust self-loop weight")
   # diag(P) = 2/(min(rowSums(W)) - max(diag(W)))
@@ -228,7 +228,7 @@ fast_calculate_score_profile_highres <- function(W, max_time = 100, init_state,
   final_state = rep(1/nrow(W),nrow(W)) # bi-stochastic
   
   # Calculate transition matrix
-  cat("Doubly Stochastic\n")
+  cat("Run doubly stochastic on affinity matrix...\n")
   P = Doubly_stochastic(W)
   P = as.matrix(P)
   score_df = get_score_profile(state_0 = init_state,
@@ -380,7 +380,7 @@ LMD <- function(expression, feature_space, knn = 5,
   rho = rho[,colnames(W),drop = FALSE]
   rho = rho[which(apply(rho,1,function(x) sum(x>0) >= min_cell))
             ,,drop = FALSE] # sanity check & remove genes which express at less than 5 cells
-  cat(sprintf("Remove %d genes which express at less than %d cells\n",
+  cat(sprintf("Remove %d genes which express in less than %d cells\n",
               nrow(expression) - nrow(rho),min_cell))
   res = fast_get_lmds(W = W, max_time = max_time,
                       init_state = rho, 
@@ -450,7 +450,7 @@ Calculate_outgoing_weight <- function(W, max_time = 2^15, init_state){
   final_state = rep(1/nrow(W),nrow(W)) # bi-stochastic
   
   # Calculate transition matrix
-  cat("Doubly Stochastic\n")
[…]

### 2024-09-26 LMD-paper/paper_figure.Rmd
figure_change

@@
 library("ggplot2")
-library("ggsignif")
+# library("ggsignif")
 library("LocalizedMarkerDetector")
@@
 library("reshape2")
-wide_data = df_auc; wide_data$AUROC = round(wide_data$AUROC,3)
+wide_data = df_auc; 
+# wide_data$AUROC = round(wide_data$AUROC,3)
 wide_data <- reshape2::dcast(wide_data, gt_set + Tissue + data_source ~ Method, value.var = "AUROC")
@@
 wide_data[,1:2] = apply(wide_data[,1:2],2,function(x) {x[duplicated(x)] = "";x})
-wide_data$Tissue[17] = "Motor Cortex(Mouse)"
-wide_data$Tissue[13] = "Motor Cortex(Human)"
+wide_data$Tissue[which(wide_data$Tissue == "Mouse-Motor Cortex")] = "Motor Cortex(Mouse)"
+wide_data$Tissue[which(wide_data$Tissue == "Human-Motor Cortex")] = "Motor Cortex(Human)"
 wide_data$Tissue = gsub("Human-|Mouse-","",wide_data$Tissue)
-# calculate average
-avg_tmp = wide_data %>%
-    group_by(Criterion) %>%
-    summarize(across(3:(ncol(wide_data)-1), ~ round(mean(.x, na.rm = TRUE), 3)))
-avg_tmp$DataSource = c("Average",""); avg_tmp$Tissue = "-"
-wide_data = rbind(wide_data,avg_tmp)
+
+# calculate median rank
+ranked_data <- data.frame(t(apply(wide_data[,4:ncol(wide_data)],1,function(x) rank(-x, na.last = "keep"))), check.names = FALSE)
+ranked_data$Criterion = wide_data$Criterion
+median_rank = ranked_data %>% group_by(Criterion) %>% summarize(across(1:(ncol(ranked_data)-1), ~ median(.x, na.rm = TRUE)))
+median_rank$DataSource = c("Median Rank",""); median_rank$Tissue = "-"
+wide_data = rbind(wide_data,median_rank)
+
+# # calculate average
+# avg_tmp = wide_data %>%
+#     group_by(Criterion) %>%
+#     summarize(across(3:(ncol(wide_data)-1), ~ mean(.x, na.rm = TRUE)))
+# avg_tmp$DataSource = c("Average",""); avg_tmp$Tissue = "-"
+# wide_data = rbind(wide_data,avg_tmp)
+
+wide_data[,4:ncol(wide_data)] = round(wide_data[,4:ncol(wide_data)],3)
+
 bold_max <- function(vec) {
@@
 
-write.csv(wide_data, file = file.path(folder.path,"AUROC_table.csv"), row.names = FALSE)
-```
+# write.csv(wide_data, file = file.path(folder.path,"AUROC_table.csv"), row.names = FALSE)
+```
+## Colored Table figure
+```{r}
+library(tidyr)
+
+# separate wide_data into two by true_set
+df_head = wide_data %>% filter(Criterion == " CellMarkerDB") %>% select(c("DataSource","Tissue"))
+df_head[nrow(df_head),"DataSource"] = ""
+df_head[nrow(df_head),"Tissue"] = "Median Rank"
+df_head$DataSource = c(
+  rep(df_head$DataSource[1],3),
+  rep(df_head$DataSource[4],6),
+  df_head$DataSource[10])
+
+pl = lapply(1:2,function(i){
+trueset = unique(wide_data$Criterion)[i]
+wide_data_sub = wide_data %>% filter(Criterion == trueset) %>% select(all_of(unname(method_fullname))) %>% mutate(across(everything(), ~ as.numeric(gsub("\\\\textbf\\{(.*)\\}", "\\1", .)))) %>% mutate(df_head)
+
+ranked_data_sub = ranked_data %>% filter(Criterion == trueset) %>% select(all_of(unname(method_fullname))) %>% mutate(df_head[1:(nrow(df_head)-1),])
+ranked_data_sub = rbind(ranked_data_sub,wide_data_sub[nrow(wide_data_sub),])
+
+# Melt the data into long format for ggplot2
+df_long <- wide_data_sub %>%
+  pivot_longer(cols = all_of(unname(method_fullname)), names_to = "Method", values_to = "AUROC") %>%
+  left_join(ranked_data_sub %>%
+              pivot_longer(cols = all_of(unname(method_fullname)), names_to = "Method", values_to = "Rank"),
+            by = c("DataSource","Tissue","Method"))
+
+df_long$Method = factor(df_long$Method, levels = rev(colnames(wide_data)[4:10]))
+df_long$DataSource = factor(df_long$DataSource, levels = unique(df_long$DataSource))
+
+# Plot the heatmap
+p = ggplot(df_long, aes(x = Tissue, y = Method, fill = Rank)) +
+    geom_tile(color = "white") + 
+    scale_fill_distiller(palette = "YlOrBr", direction = -1, na.value = "white") +
+    # scale_fill_gradient(low = "brown", high = "white", na.value = "white") +
+    geom_text(aes(label = AUROC), size = 3) + 
+    theme_minimal() + facet_grid(~ DataSource, scales = "free",space = "free", switch = "x") + ggtitle(paste0("True markerset ",i)) + 
+    theme(plot.title = element_text(size = 15,face = "bold"),
+          axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
+          strip.placement = 'outside',
+          axis.title.x = element_blank(), 
+          axis.title.y = element_blank(),
+          panel.grid = element_blank(), 
+          strip.text = element_text(size = 10,face = "bold"),
+          axis.text.y = element_text(face = "bold", size = 10))
+
+if(i==2){p = p + theme(axis.text.y = element_blank())}
+
+p
+})
+
+p = wrap_plots(pl, nrow = 1) + plot_layout(guides = "collect")
+
+ggsave(filename = file.path(figure_path,"fig-known_marker_comparison.png"), plot = p, width = 12, height = 6)
+```
+
 # SupFig-RunTime
@@
 # })
-g = c("Cd19","Naaa","Mogat2","Grina","Abca13","Zf12","Il2rb","Phgdh","Vpreb2","Fyb")
+g = c("Cd19","Naaa","Mogat2","Grina","Abca13","Zf12","Il2rb","Phgdh","Vpreb2","Fyb") # <100&>500
 names(g) = rep(method_vs_set,each = 2)
@@
 ```
-
+## Curve Plot & Coverage rate
 ```{r}
@@
 ggsave(filename = file.path(figure_path,"supfig-densityindex.png"), plot = fig, width = 12, height = 5)
+```
+## Colored Table
+```{r}
+data_source = "tabular_muris"
+dir.path <- dir.path0
+# tissue_name = "marrow_facs" # marrow_facs_granulocyte
+
+pl = lapply(c("marrow_facs","marrow_facs_granulocyte"),function(tissue_name){
+score_measure = read.table(file.path(dir.path,"benchmark",data_source,paste0(tissue_name,"_DI.csv")),header = TRUE)
+score_measure$Method = ifelse(score_measure$Method %in% names(method_fullname),method_fullname[score_measure$Method],score_measure$Method)
+total_gene = score_measure %>% filter(Method == "All genes") %>%.$TopGene
+score_measure = score_measure %>%
+  filter(Method == "All genes") %>% 
+  slice(rep(1:n(), each = length(method_fullname))) %>% 
+  mutate(Method = method_fullname) %>%
+  bind_rows(score_measure) %>% filter(Method != "All genes") %>% filter(TopGene %in% c(100,300,500,1000,3000,5000,10000,total_gene)) %>% mutate(TopGene = as.factor(TopGene)) %>% mutate(Method = factor(Method,levels = rev(method_fullname)))
+
+levels(score_measure$TopGene)[nlevels(score_measure$TopGene)] = sprintf("All(%d)",total_gene)
+
+fig = ggplot(score_measure, aes(x = TopGene, y = Method, fill = DensityIndex)) + 
+  geom_tile(color = "white") +
+  scale_fill_gradient(low = "white",high = "red") +
+  geom_text(aes(label = round(DensityIndex,3)), size = 3) + theme_minimal() + xlab("# of Top Genes") + ylab("Method") + 
+  ggtitle(ifelse(tissue_name == "marrow_facs", "Bone Marrow","Bone Marrow (granulocyte)")) + 
+  theme(plot.title = element_text(size = 20, face = "bold"),
+          axis.text = element_text(size = 10),
+          axis.title = element_text(size = 10),
+          panel.grid = element_blank(),
+        axis.ticks = element_line())
+fig
+})
+
+fig = wrap_plots(pl,ncol = 1) + plot_annotation(tag_levels = "A") & theme(plot.tag = element_text(size = 20))
+ggsave(filename = file.path(figure_path,"supfig-densityindex.png"), plot = fig, width = 10, height = 10)
 ```
@@
 
+# SupFig - sample-specificity for each module
+```{r}
+df = read.csv(file.path(dir.path,"intermediate_data","cell_prop_in_module_table.csv"))
+df$gene_props = df$gene_props * 100
+df = df %>% filter(module=="Module12" | sample!="E14.5_WT") %>% filter(gene_props > 5 & gene_props < 50)
+p = ggplot(df,
+       aes(x=gene_props, y=cell_props, linetype = sample,color = sample)) +
+  geom_line() + facet_wrap(~module,scales = "free") +
+  labs(x = "% of genes", y = "Proportion of cells", title = "Expression of Module 12 genes in DC samples") + 
+  theme(legend.position="right",
+          plot.title = element_text(size = 10),
+          axis.title = element_text(size = 10),
+          axis.text = element_text(size = 5),
+          legend.title = element_text(size = 5),
+          legend.text = element_text(size = 5))
+ggsave(filename = file.path(figure_path,"supfig-smom2_sample_specificity.png"), plot = p, width = 4, height = 3)
+```
+
 # SupsupFig - Boxplot of ModuleActivityScore of active cells

### 2024-09-26 LMD-paper/benchmark/compare_method_tabula_muris.R

@@ -38,6 +38,16 @@ tissue_name = names(tissue_download_link)[1]
 dir.path <- dir.path0
 folder.path <- file.path(dir.path,data_source)
 tiss <- readRDS(file.path(folder.path,paste0(tissue_name,".rds")))
+
+# take subset of original dataset
+# tissue_name = paste0(tissue_name,"_granulocyte")
+if(tissue_name == "marrow_facs_granulocyte"){
+  tiss <- subset(tiss, cell_ontology_class == "granulocyte")
+  tiss <- tiss %>% NormalizeData() %>% FindVariableFeatures() %>%
+    ScaleData(verbose = FALSE) %>% RunPCA(npcs = 20, verbose = FALSE)
+  tiss = RunUMAP(tiss, dims = 1:20)
+}
+
 DefaultAssay(tiss) <- "RNA"
 n_dim = 20
 feature_space = Embeddings(tiss[["pca"]])[,1:n_dim]
@@ -79,9 +89,9 @@ for(method in method_ls){
   if(method == "semi"){
     RunSEMITONES(dat, feature_space, dir.file)
   }
-  # if(method == "marcopolo"){
-  #   RunMarcopolo(tiss, selected_genes, dir.file)
-  # }
+  if(method == "marcopolo"){
+    RunMarcopolo(tiss, selected_genes, dir.file)
+  }
   end.time <- Sys.time()
   time.taken <- end.time - start.time
   df_runtime[method,"Time"] = as.numeric(time.taken, units = "mins")
@@ -199,23 +209,25 @@ df_benchmark = read.table(file.path(folder.path.rank, paste0(tissue_name,"_bench
                           header = TRUE)
 rownames(df_benchmark) = df_benchmark$gene
 n_dim = 20
+low_qc = 1 # remove low-quality cells: cells express in less than N genes
+
+topn = c(seq(50,90,10),seq(100,900,100),seq(1000,3000,500),seq(4000,9000,1000),seq(10000,nrow(df_benchmark),2000))
+# topn = c(seq(20,90,10),seq(100,1000,100),seq(1000,3000,500),seq(4000,nrow(df_benchmark),1000))
 df_density_index =do.call(rbind,lapply(c(method_ls,"All genes"),function(method){
   if(method == "All genes"){
     top_gs = rownames(df_benchmark)
     topn = length(top_gs)
   }else{
     top_gs = rownames(df_benchmark)[order(df_benchmark[,method])]
[…]

### 2025-03-24 LMD-paper/paper_figure.Rmd
revision

@@
 # library("ggsignif")
+# remotes::install_github("KlugerLab/LocalizedMarkerDetector")
+# remotes::install_github("KlugerLab/GeneTrajectory")
 library("LocalizedMarkerDetector")
@@
 figure_path = file.path(dir.path0,"figures")
-dir.create(figure_path,recursive = TRUE)
-setwd(figure_path)
+if(!dir.exists(figure_path)){dir.create(figure_path,recursive = TRUE)}
 ```
@@
 tiss <- readRDS(file.path(folder.path,paste0(tissue_name,".rds")))
-cell_label = tiss$cell_ontology_class
-coldef = setNames(colorRampPalette(brewer.pal(12, "Paired"))(length(unique(cell_label))),unique(cell_label))
-p1 = DimPlot(tiss, group.by = "cell_ontology_class",label = FALSE, cols = coldef) + guides(color = guide_legend(ncol = 1,override.aes = list(size = 3))) + labs(title = NULL)
+unique_labels <- unique(tiss$cell_ontology_class)
+label_numbers <- setNames(seq_along(unique_labels), unique_labels)
+tiss$cell_number <- as.character(label_numbers[tiss$cell_ontology_class])
+Idents(tiss) <- tiss$cell_number 
+coldef <- setNames(colorRampPalette(brewer.pal(12, "Paired"))(length(label_numbers)), 
+                   label_numbers)
+p1 <- DimPlot(tiss, group.by = "cell_number", label = TRUE, cols = coldef) +
+  guides(color = guide_legend(ncol = 1, override.aes = list(size = 3))) +
+  scale_color_manual(values = coldef, labels = paste0(seq_along(unique_labels), " - ", unique_labels)) +  # Modify legend labels
+  labs(title = NULL)
 
@@
 tiss <- readRDS(file.path(folder.path,paste0(tissue_name,".rds")))
-cell_label = tiss$cell_ontology_class
-coldef = setNames(colorRampPalette(brewer.pal(12, "Paired"))(length(unique(cell_label))),unique(cell_label))
-p2 = DimPlot(tiss, group.by = "cell_ontology_class",label = FALSE, cols = coldef) + guides(color = guide_legend(ncol = 1,override.aes = list(size = 3))) + labs(title = NULL)
+unique_labels <- unique(tiss$cell_ontology_class)
+label_numbers <- setNames(seq_along(unique_labels), unique_labels)
+tiss$cell_number <- as.character(label_numbers[tiss$cell_ontology_class])
+Idents(tiss) <- tiss$cell_number 
+coldef <- setNames(colorRampPalette(brewer.pal(12, "Paired"))(length(label_numbers)), 
+                   label_numbers)
+p2 <- DimPlot(tiss, group.by = "cell_number", label = TRUE, cols = coldef) +
+  guides(color = guide_legend(ncol = 1, override.aes = list(size = 3))) +
+  scale_color_manual(values = coldef, labels = paste0(seq_along(unique_labels), " - ", unique_labels)) +  # Modify legend labels
+  labs(title = NULL)
 
 p = ((p1 / p2) + plot_annotation(tag_levels = 'A')) & theme(plot.tag = element_text(size = 20))
-ggsave(filename = file.path(figure_path,"supfig-bonemarrow.png"), plot = p, width = 8, height = 10, dpi = 300)
+ggsave(filename = file.path(figure_path,"supfig-bonemarrow.png"), plot = p, width = 9, height = 10, dpi = 300)
 ```
@@
 ```{r}
-method_ls = c("lmd","haystack","hotspot","marcopolo","semi","wilcox_no_filter","hvg")
-method_fullname = setNames(c("LMD","singleCellHaystack","Hotspot","Marcopolo","SEMITONES","Seurat v4","HVGs"),method_ls)
+method_ls = c("lmd","haystack","hotspot","marcopolo","semi","wilcox_no_filter","hvg", "cosg", "scmarker")
+method_fullname = setNames(c("LMD","singleCellHaystack","Hotspot","Marcopolo","SEMITONES","Seurat v4","HVGs", "COSG", "SCMarker"),method_ls)
 ```
@@
 dir.path <- dir.path0
-data_source = "tabular_muris"
-tissue_name = "marrow_facs"
+# data_source = "tabular_muris"
+# tissue_name = "marrow_facs"
 folder.path = file.path(dir.path,"benchmark")
@@
 # Define Full name
-tissue_ls = list("tabular_muris" = c("marrow_facs","pancreas_facs","lung_facs"),
+tissue_ls = list("tabular_muris" = c("marrow_facs","pancreas_facs","lung_facs","marrow_droplet"),
                  "azimuth" = c("human_kidney", "human_bone_marrow", "human_pancreas", "mouse_motor_cortex", "human_lung_hlca", "human_motor_cortex"))
@@
 names(tissue_full_name) = unlist(tissue_ls)
-tissue_full_name = unlist(lapply(1:length(tissue_full_name),function(i){
-  ori_name = names(tissue_full_name)[i]
-  if(grepl("_facs|mouse",ori_name)){
-    tmp_name = paste0("Mouse-\n",tissue_full_name[i])
-  }
-  else if(grepl("human",ori_name)){
-    tmp_name = paste0("Human-\n",tissue_full_name[i])
-  }
-  else{
-    tmp_name = tissue_full_name[i]
-  }
-  tmp_name
-}))
-names(tissue_full_name) = unlist(tissue_ls)
+tissue_full_name = unlist(tissue_full_name)
+tissue_full_name["human_motor_cortex"] = paste0(tissue_full_name["human_motor_cortex"],"(Human)")
+tissue_full_name["mouse_motor_cortex"] = paste0(tissue_full_name["mouse_motor_cortex"],"(Mouse)")
+
+# add technologies as suffix
+techs = c(rep("FACS",3),"Droplet","10x","Cite-seq","Smart-seq2","10x","10x","Smart-seq v4")
+tissue_full_name = setNames(paste0(tissue_full_name," ","(",techs,")"),names(tissue_full_name))
 ```
@@
                   df_all}))
+```
+## Louvain resolution for Cluster-dependent methods
+```{r}
+# Plot performance for each datasets
+method_subls = c("wilcox_no_filter","cosg")
+df_auc_sub = do.call(rbind,lapply(method_subls, function(method){
+  df = do.call(rbind,
+        lapply(1:length(tissue_ls),function(i){
+            df_all = NULL
+            for(j in tissue_ls[[i]]){
+                df = read.table(file.path(folder.path, names(tissue_ls)[i],method,paste0(j,"_auroc.csv")),header = TRUE)
+                df$'Tissue' = j
+                df_all = rbind(df_all,df)
+            }
+            df_all$'data_source' = names(tissue_ls)[i]
+            df_all}))
+  df$'Method' = method
+  df
+}))
+```
+```{r}
+# plot louvain-clustering vs. real celltype
+df_tmp = do.call(rbind,
+        lapply(1:length(tissue_ls),function(i){
+            df_all = NULL
+            for(j in tissue_ls[[i]]){
+                df = read.table(file.path(folder.path, names(tissue_ls)[i],paste0(j,"_louvain_clusters.csv")),header = TRUE)
+                df$'Tissue' = j
+                df_all = rbind(df_all,df)
+            }
+            df_all$'data_source' = names(tissue_ls)[i]
+            df_all}))
+df_tmp$Tissue = tissue_full_name[df_tmp$Tissue]
+df_tmp <- df_tmp %>%
+  mutate(Tissue = factor(Tissue, levels = unique(Tissue[rev(order(data_source))])))
+
+p = df_tmp %>% ggplot(aes(x = as.factor(res), y = ncluster, group = 1)) + 
+  geom_line() + geom_hline(aes(yintercept = ncelltype), color="red", linetype="dashed") + 
+    facet_wrap(~Tissue, scales = "free") + xlab("Resolution") + ggtitle("The number of Louvain clusters by different resolution")
+ggsave(filename = file.path(figure_path,"supfig-seurat_performance1.png"), 
+       plot = p, width = 10, height = 5)
+```
+```{r}
+# plot Seurat
+df_tmp = df_auc_sub %>% filter(gt_set %in% c("CellMarkerDB","Top100"))
+df_tmp$gt_set = as.factor(df_tmp$gt_set); levels(df_tmp$gt_set) = c("Trueset1","Trueset2")
+df_tmp$Tissue = tissue_full_name[df_tmp$Tissue]
+df_tmp <- df_tmp %>%
+  mutate(Tissue = factor(Tissue, levels = unique(Tissue[rev(order(data_source))])))
+
+pl = lapply(method_subls,function(method){
+  df_tmp %>% filter(Method == method) %>% ggplot(aes(x = as.factor(res), y = AUROC, group = gt_set, color = gt_set)) + 
+  geom_line() + 
+    facet_wrap(~Tissue) +
+    ggtitle(method_fullname[method]) + xlab("Resolution")
+})
+p = wrap_plots(pl, nrow = 2)
+ggsave(filename = file.path(figure_path,"supfig-seurat_performance.png"), 
+       plot = p, width = 16, height = 16)
+```
+```{r}
+# # Replace cluster-dependent method w/ resolution of maximum performance
+# df_auc_sub %>% 
+#   filter(gt_set %in% c("Top100", "CellMarkerDB")) %>% 
+#   group_by(res) %>% 
+#   summarize(median_AUROC = median(AUROC, na.rm = TRUE)) %>% 
+#   arrange(desc(median_AUROC))
+# 
+# default_res = 5
+# # df_auc_sub = df_auc_sub %>%
+# #     group_by(Tissue, gt_set, data_source, Method) %>%
+# #     filter(AUROC == max(AUROC))
+# df_auc_sub = df_auc_sub %>% filter(res == 5) %>% select(colnames(df_auc))
+# '%notin%' <- Negate('%in%')
+# df_auc = rbind(df_auc %>% filter(Method %notin% method_subls),df_auc_sub)
+```
+## Re-organize table
+```{r}
+# re-organize the df
+# Order Method based on mean AUROC
 df_auc$Method = method_fullname[df_auc$Method]
-df_auc = df_auc %>% filter(gt_set %in% c("CellMarkerDB","Top100"))
-# Order Method based on mean AUROC
+top_name = "Top100"
+df_auc = df_auc %>% filter(gt_set %in% c("CellMarkerDB",top_name))
 df_auc$Method <- with(df_auc, reorder(Method, -AUROC, FUN = median))
-df_auc$gt_set <- factor(df_auc$gt_set,levels = c("CellMarkerDB","Top100"))
+df_auc$gt_set <- factor(df_auc$gt_set,levels = c("CellMarkerDB",top_name))
 levels(df_auc$gt_set) = c("Criterion1-\n CellMarkerDB","Criterion2-\n MaxFC")
@@
 ```
+
 ## Figure
@@
 wide_data[,1:2] = apply(wide_data[,1:2],2,function(x) {x[duplicated(x)] = "";x})
-wide_data$Tissue[which(wide_data$Tissue == "Mouse-Motor Cortex")] = "Motor Cortex(Mouse)"
-wide_data$Tissue[which(wide_data$Tissue == "Human-Motor Cortex")] = "Motor Cortex(Human)"
-wide_data$Tissue = gsub("Human-|Mouse-","",wide_data$Tissue)
+
+wide_data[,4:ncol(wide_data)] = round(wide_data[,4:ncol(wide_data)],3)
 
 # calculate median rank
-ranked_data <- data.frame(t(apply(wide_data[,4:ncol(wide_data)],1,function(x) rank(-x, na.last = "keep"))), check.names = FALSE)
+ranked_data <- data.frame(t(apply(wide_data[,4:ncol(wide_data)],1,function(x) rank(-x, na.last = "keep", ties.method = "min"))), check.names = FALSE)
 ranked_data$Criterion = wide_data$Criterion
@@
 # wide_data = rbind(wide_data,avg_tmp)
-
-wide_data[,4:ncol(wide_data)] = round(wide_data[,4:ncol(wide_data)],3)
 
@@
 ```
-## Colored Table figure
+### Colored Table figure
 ```{r}
@@
 df_head$DataSource = c(
-  rep(df_head$DataSource[1],3),
-  rep(df_head$DataSource[4],6),
-  df_head$DataSource[10])
+  rep(df_head$DataSource[1],4),
+  rep(df_head$DataSource[5],6),
+  df_head$DataSource[nrow(df_head)])
 
@@
 
-df_long$Method = factor(df_long$Method, levels = rev(colnames(wide_data)[4:10]))
+df_long$Method = factor(df_long$Method, levels = rev(colnames(wide_data)[-c(1:3)]))
 df_long$DataSource = factor(df_long$DataSource, levels = unique(df_long$DataSource))
+tissue_order = unique(df_long$Tissue[order(df_long$DataSource)]); tissue_order[1:4] = tissue_order[c(2:4,1)]
+df_long <- df_long %>%
+  mutate(Tissue = factor(Tissue, levels = tissue_order))
 
@@
     geom_tile(color = "white") + 
-    scale_fill_distiller(palette = "YlOrBr", direction = -1, na.value = "white") +
+    scale_fill_distiller(palette = "YlOrBr", direction = -1, na.value = "white", limits = c(1, max(df_long$Rank, na.rm = TRUE)), 
+                         breaks = seq(1, max(df_long$Rank, na.rm = TRUE))) +
     # scale_fill_gradient(low = "brown", high = "white", na.value = "white") +
@@
 p = wrap_plots(pl, nrow = 1) + plot_layout(guides = "collect")
-
-ggsave(filename = file.path(figure_path,"fig-known_marker_comparison.png"), plot = p, width = 12, height = 6)
+```
+```{r}
+ggsave(filename = file.path(figure_path,"fig-known_marker_comparison.png"), plot = p, width = 13, height = 8)
 ```
@@
 ```{r}
-df_runtime = df_runtime %>% filter(Method == "LMD") %>% arrange(Time, nCells, nGenes) %>% select(c("Time","nCells","nGenes")) %>% mutate(Time = round(Time, 2))
-colnames(df_runtime)[1] = "Time (min)"
+df_runtime = df_runtime %>% filter(Method == "LMD") %>% arrange(Time, nCells, nGenes) %>% select(c("data_source","Tissue","nCells","ncluster","nGenes","Time")) %>% mutate(Time = round(Time, 2))
+colnames(df_runtime)[grep("Time",colnames(df_runtime))] = "Time (min)"
 write.csv(df_runtime, file = file.path(folder.path,"RunTime_table.csv"), row.names = FALSE)
@@
 # Fig - Gene Ranking Comparisons between LMD and Other Methods
+```{r}
+method_vs = "lmd"
+method_vs_set = c("wilcox_no_filter","cosg","hotspot","marcopolo","scmarker","semi","haystack")
+```
+## Selected genes Table
+### Bone marrow facs
 ```{r}
@@
 tiss = readRDS(file.path(dir.path,data_source, paste0(tissue_name,".rds")))
-method_vs = "lmd"
-method_vs_set = c("wilcox_no_filter","hotspot","marcopolo","semi","haystack")
+
 ```
@@
 # })
-g = c("Cd19","Naaa","Mogat2","Grina","Abca13","Zf12","Il2rb","Phgdh","Vpreb2","Fyb") # <100&>500
+# g = c("Cd19","Naaa","Mogat2","Grina","Abca13","Zf12","Il2rb","Phgdh","Vpreb2","Fyb") # <100&>500 # before revision
+
+g = c("Hp","Cyp11a1",#wilcox
+      "1810033B17Rik","Pltp",#cosg
+      "Mogat2","Grina",#hotspot
+      "Abca13","Zf12",#marcopolo
+      "Pilra","Rgs14",#scmarker
+      "Il2rb","Phgdh",#semi
+      "Vpreb2","Fyb"#schaystack
+      ) # <100&>500
+
+df_label = df_label_all[g,]
+df_label = df_label[order(df_label$lmd),]
+colnames(df_label) = method_fullname[colnames(df_label)]
+write.csv(df_label,file = file.path(dir.path,"benchmark", paste0(tissue_name,"_benchmark_rank_table_selectedgenes.csv")),row.names = TRUE)
+```
+### human kidney 10x
+```{r}
+dir.path <- dir.path0
+data_source = "azimuth"
+tissue_name = "human_kidney"
+df_benchmark = read.table(file.path(dir.path,"benchmark", data_source, paste0(tissue_name,"_benchmark_rank_table.csv")),header = TRUE)
+rownames(df_benchmark) = df_benchmark$gene
+tiss = readRDS(file.path(dir.path,data_source, paste0(tissue_name,".rds")))
+tiss = RunUMAP(tiss, dims = 1:20, seed.use = 42)
+```
+```{r}
+# Plot DimPlot
+cell_label = tiss$cell_ontology_class
+coldef = setNames(colorRampPalette(brewer.pal(12, "Paired"))(length(unique(cell_label))),unique(cell_label))
+p1 = DimPlot(tiss, group.by = "cell_ontology_class",label = FALSE, cols = coldef) + guides(color = guide_legend(ncol = 1,override.aes = list(size = 3))) + labs(title = NULL)
+```
+```{r}
+sub = do.call(c,lapply(method_vs_set,function(method){
+  sub = c(which(df_benchmark[,method] <= 100 & df_benchmark[,method_vs] > 500),
+          which(df_benchmark[,method] > 500 & df_benchmark[,method_vs] <= 100) )
+  sub
+}))
+df_label_all = df_benchmark[unique(sub),c(method_vs,method_vs_set)]
+# kidney
+g = c("SPINK8","SPP1",#wilcox
+      "RAMP2","KIAA0101",#cosg
+      "SAA2","GSTA1",#hotspot
+      "MEIS2","AIM1",#scmarker
+      "MUC16","ATF3",#semi
+      "CEACAM6","COX7A1"#schaystack
+      )
+
+df_label = df_label_all[g,]
+df_label = df_label[order(df_label$lmd),]
+colnames(df_label) = method_fullname[colnames(df_label)]
+write.csv(df_label,file = file.path(dir.path,"benchmark", paste0(tissue_name,"_benchmark_rank_table_selectedgenes.csv")),row.names = TRUE)
+```
+## FeaturePlot
+```{r}
 names(g) = rep(method_vs_set,each = 2)
+unified_range <- range(FetchData(tiss, vars = g), na.rm = TRUE)
+
 pl = lapply(1:length(g),function(i){
@@
   color_label = ifelse(df_label_all[x,method_vs] > df_label_all[x,names(g)[i]],"blue","red")
-  FeaturePlot(tiss,features = x,order =TRUE) + labs(color = "Expression") + labs(title = x,subtitle = subtitle) + 
+  FeaturePlot(tiss,features = x,order =TRUE,pt.size = 1) + labs(color = "Expression") + scale_color_gradientn(colors = c("lightgrey", "blue"), 
+                          limits = unified_range) + labs(title = x,subtitle = subtitle) + 
     theme(
@@
 
-fig = wrap_plots(pl,nrow = 2)
-
-ggsave(filename = file.path(figure_path,"fig-ranking_comparison2.png"), plot = fig, width = 25, height = 10)
-```
+fig = wrap_plots(pl,nrow = 2) + plot_layout(guides = "collect")
+
+# ggsave(filename = file.path(figure_path,"fig-ranking_comparison2.png"), plot = fig, width = 35, height = 10)
+ggsave(filename = file.path(figure_path,"Supfig-ranking_comparison2.png"), plot = fig, width = 32, height = 10)
+```
+## RankComparisonPlot
 ```{r}
@@
 
-ggsave(filename = file.path(figure_path,"fig-ranking_comparison1.png"), plot = p, width = 25, height = 6)
+# ggsave(filename = file.path(figure_path,"fig-ranking_comparison1.png"), plot = p, width = 35, height = 6)
+ggsave(filename = file.path(figure_path,"Supfig-ranking_comparison1.png"), plot = p, width = 32, height = 6)
 ```
@@
 # Set the color code for all methods
-method_ls = c("lmd","haystack","hotspot","marcopolo","semi","wilcox_no_filter","hvg")
-method_fullname = setNames(c("LMD","singleCellHaystack","Hotspot","Marcopolo","SEMITONES","Seurat v4","HVGs"),method_ls)
 coldef = setNames(
@@
 total_gene = score_measure %>% filter(Method == "All genes") %>%.$TopGene
+scmarker = score_measure %>% filter(Method == "SCMarker")
 score_measure = score_measure %>%
@@
   mutate(Method = method_fullname) %>%
-  bind_rows(score_measure) %>% filter(Method != "All genes") %>% filter(TopGene %in% c(100,300,500,1000,3000,5000,10000,total_gene)) %>% mutate(TopGene = as.factor(TopGene)) %>% mutate(Method = factor(Method,levels = rev(method_fullname)))
-
+  bind_rows(score_measure) %>% filter(Method != "All genes") 
+if(tissue_name == "marrow_facs"){
+  score_measure = score_measure %>% filter(TopGene %in% c(100,300,500,1000,3000,5000,10000,total_gene))
+}else{
+  score_measure = score_measure %>% filter(TopGene %in% c(20,50,100,500,1000,5000,10000,total_gene))
+}
+score_measure = score_measure %>% mutate(TopGene = as.factor(TopGene)) %>% mutate(Method = factor(Method,levels = rev(method_fullname)))
 levels(score_measure$TopGene)[nlevels(score_measure$TopGene)] = sprintf("All(%d)",total_gene)
+
+numeric_levels = as.numeric(levels(score_measure$TopGene)[!grepl("All", levels(score_measure$TopGene))])
+closest_level <- numeric_levels[which.min(abs(numeric_levels - scmarker$TopGene))]
+score_measure = rbind(score_measure,scmarker %>% mutate(TopGene = factor(closest_level, levels = levels(score_measure$TopGene))))
+levels(score_measure$Method)[grep("SCMarker",levels(score_measure$Method))] = paste0(scmarker$Method," (",scmarker$TopGene," genes)")
 
@@
   geom_text(aes(label = round(DensityIndex,3)), size = 3) + theme_minimal() + xlab("# of Top Genes") + ylab("Method") + 
-  ggtitle(ifelse(tissue_name == "marrow_facs", "Bone Marrow","Bone Marrow (granulocyte)")) + 
+  ggtitle(ifelse(tissue_name == "marrow_facs", "Bone Marrow(FACS)","Bone Marrow(FACS) (granulocyte)")) + 
   theme(plot.title = element_text(size = 20, face = "bold"),
@@
 # SupFig - Sensitivity Test
+```{r}
+dir.path <- dir.path0
+data_source = "tabular_muris"
+tissue_name = "marrow_facs"
+```
 ## kNN
@@
   geom_line() +
-  geom_point() + labs(x = "# of kNN used", y = "Jaccard Index\n (with kNN5)",color = "Top Genes") + geom_vline(xintercept = 3, color="red", linetype="dashed") + 
+  geom_point() + labs(x = "# of kNN used", y = "Jaccard Index\n (vs. kNN5)",color = "Top Genes") + geom_vline(xintercept = "5", color="red", linetype="dashed") + 
   scale_y_continuous(limits = c(0,1)) +
@@
   geom_line() +
-  geom_point() + labs(x = "# of PC used", y = "Jaccard Index\n (with 20PC)", color = "Top Genes") + geom_vline(xintercept = 2, color="red", linetype="dashed") + 
+  geom_point() + labs(x = "# of PC used", y = "Jaccard Index\n (vs. 20PC)", color = "Top Genes") + geom_vline(xintercept = "20", color="red", linetype="dashed") + 
   scale_y_continuous(limits = c(0,1)) +
@@
     axis.line = element_line(colour = "black"))
-
-# fig = fig3 + fig4 + plot_annotation(title = "Sensitivity Test: Different PC", theme = theme(plot.title = element_text(size = 20)))
-# ggsave(filename = file.path(figure_path,"supfig-sensitivitypc.png"), plot = fig, width = 16, height = 5)
-```
-```{r}
-p = (fig1 + fig2) / (fig3 + fig4) + plot_layout(guides = "collect") + plot_annotation(tag_level = 'A') & theme(plot.tag = element_text(face="bold", size = 20)) 
-ggsave(filename = file.path(figure_path,"supfig-sensitivity.png"), plot = p, width = 16, height = 10)
+```
+
+## Cell number
+```{r}
+folder.path = file.path(dir.path,data_source,"intermediate_data", tissue_name, "sensitivity_test")
+test_para = "cellprop"
+
+## Rank Stability
+df = read.table(file.path(folder.path,sprintf("%s_jaccard_ind_%s.csv",tissue_name,test_para)),header = TRUE)
+df[, c("cellprop","rep")] = do.call(rbind,stringr::str_extract_all(df$cellprop, "(?<=_)\\d+(\\.\\d+)?"))
+df$Top_Genes = as.factor(df$Top_Genes)
+df_summary <- df %>%
+  group_by(cellprop, Top_Genes) %>%
+  summarise(
+    median_Jaccard = median(JaccardIndex),
+    lower = median_Jaccard - sd(JaccardIndex) / sqrt(n()),  # Standard error (SE)
+    upper = median_Jaccard + sd(JaccardIndex) / sqrt(n())
+  )
+
+fig5 = ggplot(data = df_summary, aes(x = cellprop, y = median_Jaccard, group = Top_Genes, color = Top_Genes)) +
+  geom_line() +
+  geom_point() + 
+  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) + labs(x = "Cell number downsampling proprotion", y = "Jaccard Index\n (vs.
+Original Data)", color = "Top Genes") + geom_vline(xintercept = "20", color="red", linetype="dashed") +
+  scale_y_continuous(limits = c(0,1)) +
+  theme(
+    legend.title = element_text(size = 20),
+    legend.text = element_text(size = 15),
+    axis.title.x = element_text(size = 20),
+    axis.title.y = element_text(size = 20),
+    axis.text.x = element_text(size = 15),
+    axis.text.y = element_text(size = 15),
+    panel.grid = element_blank(),
+    panel.background = element_blank(),
+    axis.line = element_line(colour = "black"))
+
+# AUC stability
+df = read.table(file.path(folder.path,sprintf("%s_auroc_%s.csv",tissue_name,test_para)),header = TRUE)
+df = df %>% filter(gt_set %in% c("Top100","CellMarkerDB"))
+df$gt_set = factor(df$gt_set)
+df[, c("cellprop","rep")] = do.call(rbind,stringr::str_extract_all(df$Parameter, "(?<=_)\\d+(\\.\\d+)?"))
+df_summary <- df %>% select(!Parameter) %>%
+  group_by(cellprop, gt_set) %>%
+  summarise(
+    median_auroc = median(AUROC),
+    lower = median_auroc - sd(AUROC) / sqrt(n()),  # Standard error (SE)
+    upper = median_auroc + sd(AUROC) / sqrt(n())
+  )
+
+fig6 = ggplot(data = df_summary, aes(x = cellprop, y = median_auroc, group = gt_set, color = gt_set)) +
+  geom_line() +
+  geom_point() + geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) + 
+  labs(x = "Cell number downsampling proprotion", y = "AUROC", color = "Gold-Standard Genesets") + 
+  scale_y_continuous(limits = c(0,1)) +
+  theme(
+    legend.title = element_text(size = 20),
+    legend.text = element_text(size = 15),
+    axis.title.x = element_text(size = 20),
+    axis.title.y = element_text(size = 20),
+    axis.text.x = element_text(size = 15),
+    axis.text.y = element_text(size = 15),
+    panel.grid = element_blank(),
+    panel.background = element_blank(),
+    axis.line = element_line(colour = "black"))
+
+fig = fig5 + fig6
+ggsave(filename = file.path(figure_path,"supfig-sensitivity-umi.png"), plot = fig, width = 16, height = 5)
+```
+
+## Data sparsity (UMI)
+```{r}
+folder.path = file.path(dir.path,data_source,"intermediate_data", tissue_name, "sensitivity_test")
+test_para = "umi"
+
+## Rank Stability
+df = read.table(file.path(folder.path,sprintf("%s_jaccard_ind_%s.csv",tissue_name,test_para)),header = TRUE)
+df[, c("umi","rep")] = do.call(rbind,stringr::str_extract_all(df$umi, "(?<=_)\\d+(\\.\\d+)?"))
+df$Top_Genes = as.factor(df$Top_Genes)
+df$umi = as.factor(as.numeric(df$umi))
+df_summary <- df %>%
+  group_by(umi, Top_Genes) %>%
+  summarise(
+    median_Jaccard = median(JaccardIndex),
+    lower = median_Jaccard - sd(JaccardIndex) / sqrt(n()),  # Standard error (SE)
+    upper = median_Jaccard + sd(JaccardIndex) / sqrt(n())
+  )
+
+fig7 = ggplot(data = df_summary, aes(x = umi, y = median_Jaccard, group = Top_Genes, color = Top_Genes)) +
+  geom_line() +
+  geom_point() + 
+  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) + labs(x = "UMI downsampling proprotion", y = "Jaccard Index\n (vs.
+Original Data)", color = "Top Genes") + geom_vline(xintercept = "20", color="red", linetype="dashed") +
+  scale_y_continuous(limits = c(0,1)) +
+  theme(
+    legend.title = element_text(size = 20),
+    legend.text = element_text(size = 15),
+    axis.title.x = element_text(size = 20),
+    axis.title.y = element_text(size = 20),
+    axis.text.x = element_text(size = 15,angle = 45),
+    axis.text.y = element_text(size = 15),
+    panel.grid = element_blank(),
+    panel.background = element_blank(),
+    axis.line = element_line(colour = "black"))
+
+# AUC stability
+df = read.table(file.path(folder.path,sprintf("%s_auroc_%s.csv",tissue_name,test_para)),header = TRUE)
+df = df %>% filter(gt_set %in% c("Top100","CellMarkerDB"))
+df$gt_set = factor(df$gt_set)
+df[, c("umi","rep")] = do.call(rbind,stringr::str_extract_all(df$Parameter, "(?<=_)\\d+(\\.\\d+)?"))
+df$umi = as.factor(as.numeric(df$umi))
+df_summary <- df %>% select(!Parameter) %>%
+  group_by(umi, gt_set) %>%
+  summarise(
+    median_auroc = median(AUROC),
+    lower = median_auroc - sd(AUROC) / sqrt(n()),  # Standard error (SE)
+    upper = median_auroc + sd(AUROC) / sqrt(n())
+  )
+
+fig8 = ggplot(data = df_summary, aes(x = umi, y = median_auroc, group = gt_set, color = gt_set)) +
+  geom_line() +
+  geom_point() + geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) + 
+  labs(x = "UMI downsampling proprotion", y = "AUROC", color = "Gold-Standard Genesets") + 
+  scale_y_continuous(limits = c(0,1)) +
+  theme(
+    legend.title = element_text(size = 20),
+    legend.text = element_text(size = 15),
+    axis.title.x = element_text(size = 20),
+    axis.title.y = element_text(size = 20),
+    axis.text.x = element_text(size = 15,angle = 45),
+    axis.text.y = element_text(size = 15),
+    panel.grid = element_blank(),
+    panel.background = element_blank(),
+    axis.line = element_line(colour = "black"))
+
+fig = fig7 + fig8
+ggsave(filename = file.path(figure_path,"supfig-sensitivity-umi.png"), plot = fig, width = 16, height = 5)
+```
+
+## Binary
+```{r}
+folder.path = file.path(dir.path,data_source,"intermediate_data", tissue_name, "sensitivity_test")
+test_para = "binary"
+
+## Rank Stability
+df = read.table(file.path(folder.path,sprintf("%s_jaccard_ind_%s.csv",tissue_name,test_para)),header = TRUE)
+df$Top_Genes = as.factor(df$Top_Genes)
+p1 = ggplot(data = df %>% filter(get(test_para) %in% c("binary")), 
+            aes(x = Top_Genes, y = JaccardIndex, group = 1)) +
+  geom_line() + labs(x = "Top Genes", y = "Jaccard Index") + ylim(0,1) + 
+  theme_minimal() + ggtitle("Jaccard Index of Binary scale vs. Original scale") +
+  theme(
+    title = element_text(size = 20),
+    legend.title = element_text(size = 20),
+    legend.text = element_text(size = 15),
+    axis.title.x = element_text(size = 20),
+    axis.title.y = element_text(size = 20),
+    axis.text.x = element_text(size = 15),
+    axis.text.y = element_text(size = 15),
+    # panel.grid = element_blank(),
+    panel.background = element_blank(),
+    axis.line = element_line(colour = "black"))
+
+df = read.table(file.path(folder.path,sprintf("%s_auroc_%s.csv",tissue_name,test_para)),header = TRUE)
+df = df %>% filter(gt_set %in% c("Top100","CellMarkerDB"))
+df$gt_set = factor(df$gt_set)
+
+p2 = ggplot(data = df %>% filter(gt_set %in% c("CellMarkerDB","Top100")) %>% mutate(Parameter = as.factor(Parameter)), 
+            aes(x = gt_set, y = AUROC, group = Parameter, fill = Parameter)) +
+  geom_bar(stat = "identity", position = "dodge") +  
+  labs(x = "Gold-Standard Geneset", fill = "Scales") + ylim(0,1) +  
+  theme_minimal() + ggtitle("AUROC of two scales") +
+  theme(
+    title = element_text(size = 20),
+    legend.title = element_text(size = 20),
+    legend.text = element_text(size = 15),
+    axis.title.x = element_text(size = 20),
+    axis.title.y = element_text(size = 20),
+    axis.text.x = element_text(size = 15),
+    axis.text.y = element_text(size = 15),
+    # panel.grid = element_blank(),
+    panel.background = element_blank(),
+    axis.line = element_line(colour = "black"))
+
+ggsave(filename = file.path(figure_path,"supfig-sensitivity-binary.png"), plot = p1 + p2, width = 18, height = 10)
+```
+
+## Alternative graphs
+```{r}
+folder.path = file.path(dir.path,data_source,"intermediate_data", tissue_name, "sensitivity_test")
+test_para = "graph"
+
+## Rank Stability
+df = read.table(file.path(folder.path,sprintf("%s_jaccard_ind_%s.csv",tissue_name,test_para)),header = TRUE)
+df$Top_Genes = as.factor(df$Top_Genes)
+
+p1 = ggplot(data = df %>% filter(get(test_para) != "kNN"), 
+            aes(x = Top_Genes, y = JaccardIndex, group = graph, color = graph)) +
+  geom_line() + labs(x = "Top Genes", y = "Jaccard Index", color = "Graphs") + ylim(0,1) + 
+  theme_minimal() + ggtitle("Jaccard Index (vs. kNN graph)") +
+  theme(
+    title = element_text(size = 20),
+    legend.title = element_text(size = 20),
+    legend.text = element_text(size = 15),
+    axis.title.x = element_text(size = 20),
+    axis.title.y = element_text(size = 20),
+    axis.text.x = element_text(size = 15),
+    axis.text.y = element_text(size = 15),
+    # panel.grid = element_blank(),
+    panel.background = element_blank(),
+    axis.line = element_line(colour = "black"))
+
+df = read.table(file.path(folder.path,sprintf("%s_auroc_%s.csv",tissue_name,test_para)),header = TRUE)
+df = df %>% filter(gt_set %in% c("Top100","CellMarkerDB"))
+df$gt_set = factor(df$gt_set)
+
+p2 = ggplot(data = df %>% filter(gt_set %in% c("CellMarkerDB","Top100")) %>% mutate(Parameter = as.factor(Parameter)), 
+            aes(x = gt_set, y = AUROC, group = Parameter, fill = Parameter)) +
+  geom_bar(stat = "identity", position = "dodge") +  
+  labs(x = "Gold-Standard Geneset", fill = "Graphs") + ylim(0,1) +  
+  theme_minimal() + ggtitle("AUROC of different graphs") +
+  theme(
+    title = element_text(size = 20),
+    legend.title = element_text(size = 20),
+    legend.text = element_text(size = 15),
+    axis.title.x = element_text(size = 20),
+    axis.title.y = element_text(size = 20),
+    axis.text.x = element_text(size = 15),
+    axis.text.y = element_text(size = 15),
+    # panel.grid = element_blank(),
+    panel.background = element_blank(),
+    axis.line = element_line(colour = "black"))
+
+ggsave(filename = file.path(figure_path,"supfig-sensitivity-graph.png"), plot = p1 + p2, width = 18, height = 10)
+```
+
+## Ambient RNA
+```{r}
+folder.path = file.path(dir.path,"simulated_pancreas", "sensitivity_test")
+test_para = "ambient"
+
+## Rank Stability
+df = read.table(file.path(folder.path,sprintf("jaccard_ind_%s.csv",test_para)),header = TRUE)
+df$Top_Genes = as.factor(df$Top_Genes)
+
+p1 = ggplot(data = df %>% filter(get(test_para) != "original"), 
+            aes(x = Top_Genes, y = JaccardIndex, group = get(test_para), color = get(test_para))) +
+  geom_line() + labs(x = "Top Genes", y = "Jaccard Index", color = "Contamination level") + ylim(0,1) + 
+  theme_minimal() + ggtitle("Jaccard Index (vs. uncontaminated)") +
+  theme(
+    title = element_text(size = 20),
+    legend.title = element_text(size = 15),
+    legend.text = element_text(size = 10),
+    axis.title.x = element_text(size = 20),
+    axis.title.y = element_text(size = 20),
+    axis.text.x = element_text(size = 15),
+    axis.text.y = element_text(size = 15),
+    # panel.grid = element_blank(),
+    legend.position = "bottom",
+    panel.background = element_blank(),
+    axis.line = element_line(colour = "black"))
+
+contaminated_genes = readRDS(file.path(dir.path,"simulated_pancreas","gene500.rds"))
+
+# Spearman correlation of the rank of 500 contaminated genes (scatterplot)
+df = reshape2::melt(score_ranked %>% mutate(gene = row.names(score_ranked)) %>% filter(gene %in% contaminated_genes),id = c("original","gene"),variable.name = "level",value.name = "rank")
+cor_results <- df %>%
+  group_by(level) %>%
+  summarize(spearman = cor(rank, original, method = "spearman")) %>%
+  mutate(label = paste0("Spearman: ", round(spearman, 4)))
+
+p2 = ggplot(df, aes(x = original, y = rank)) +
+  geom_point() + 
+  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "blue") +  
+  facet_wrap(~ level) + 
+  geom_text(data = cor_results, aes(x = Inf, y = Inf, label = label),
+            hjust = 1.2, vjust = 1.2, size = 4, color = "red", fontface = "bold") + 
+  theme_minimal() +
+  labs(title = "Rank Comparison for 500 contaminated genes",
+       x = "Rank (uncontaminated data)",
+       y = "Rank (contaminated data)",
+       color = "Contamination Level") +  
+  theme_minimal() + 
+  theme(
+    title = element_text(size = 20),
+    legend.title = element_text(size = 20),
+    strip.text = element_text(size = 15),
+    legend.text = element_text(size = 15),
+    axis.title.x = element_text(size = 20),
+    axis.title.y = element_text(size = 20),
+    axis.text.x = element_text(size = 15),
+    axis.text.y = element_text(size = 15),
+    # panel.grid = element_blank(),
+    panel.background = element_blank(),
+    axis.line = element_line(colour = "black"))
+
+ggsave(filename = file.path(figure_path,"supfig-sensitivity-ambient.png"), plot = p1 + p2 + plot_layout(widths = c(1, 2)), width = 18, height = 6)
+```
+
+## Graph Coarse-graining
+```{r}
+data_source = "azimuth"
+tissue_name = "human_bone_marrow"
+
+folder.path = file.path(dir.path,data_source,"intermediate_data", tissue_name, "sensitivity_test")
+test_para = "cg"
+
+## Rank Stability
+df = read.table(file.path(folder.path,sprintf("jaccard_ind_%s.csv",test_para)),header = TRUE)
+df$Top_Genes = as.factor(df$Top_Genes)
+df$Ncell = as.factor(df$Ncell)
+
+p1 = ggplot(data = df, aes(x = Ncell, y = JaccardIndex, group = Top_Genes, color = Top_Genes)) +
+  geom_line() + geom_point() + labs(x = "Ncell for coarse-grained cell graph", y = "Jaccard Index\n (vs.
+Original graph)", color = "Top Genes") + geom_vline(xintercept = "30651", color="red", linetype="dashed") +
+  scale_y_continuous(limits = c(0,1)) +
+  theme(
+    legend.title = element_text(size = 20),
+    legend.text = element_text(size = 15),
+    axis.title.x = element_text(size = 20),
+    axis.title.y = element_text(size = 20),
+    axis.text.x = element_text(size = 15, angle = 45),
+    axis.text.y = element_text(size = 15),
+    panel.grid = element_blank(),
+    panel.background = element_blank(),
+    axis.line = element_line(colour = "black"))
+
+# AUC stability
+df = read.table(file.path(folder.path,sprintf("auroc_%s.csv",test_para)),header = TRUE)
+df = df %>% filter(gt_set %in% c("Top100","CellMarkerDB"))
+df$Parameter = as.factor(df$Parameter)
+df$gt_set = factor(df$gt_set)
+
+p2 = ggplot(data = df, aes(x = Parameter, y = AUROC, group = gt_set, color = gt_set)) +
+  geom_line() +
+  geom_point()  + 
+  labs(x = "Ncell for coarse-grained cell graph", y = "AUROC", color = "Gold-Standard Genesets") + 
+  scale_y_continuous(limits = c(0,1)) +
+  theme(
+    legend.title = element_text(size = 20),
+    legend.text = element_text(size = 15),
+    axis.title.x = element_text(size = 20),
+    axis.title.y = element_text(size = 20),
+    axis.text.x = element_text(size = 15, angle = 45),
+    axis.text.y = element_text(size = 15),
+    panel.grid = element_blank(),
+    panel.background = element_blank(),
+    axis.line = element_line(colour = "black"))
+
+df = read.table(file.path(folder.path,sprintf("runtime_%s.csv",test_para)),header = TRUE)
+df$nCells = as.factor(df$nCells)
+p3 = ggplot(data = df, aes(x = nCells, y = Time_min, group = 1)) + geom_line() +
+  geom_point()  + 
+  labs(x = "Ncell for coarse-grained cell graph", y = "RunTime (min)") +
+  theme(
+    legend.title = element_text(size = 20),
+    legend.text = element_text(size = 15),
+    axis.title.x = element_text(size = 20),
+    axis.title.y = element_text(size = 20),
+    axis.text.x = element_text(size = 15),
+    axis.text.y = element_text(size = 15),
+    panel.grid = element_blank(),
+    panel.background = element_blank(),
+    axis.line = element_line(colour = "black"))
+
+p = ((p1 + p2) / (p3 + plot_spacer())) + plot_annotation(tag_level = 'A') & theme(plot.tag = element_text(face="bold", size = 20)) 
+ggsave(filename = file.path(figure_path,"supfig-sensitivity-cg.png"), plot = p, width = 16, height = 10)
+```
+
+## Overall Plot
+```{r}
+p = ((fig1 + fig2) / (fig3 + fig4) / (fig5 + NoLegend() + fig6 + NoLegend()) / (fig7+ NoLegend() + fig8+ NoLegend())) + plot_layout(guides = "collect") + plot_annotation(tag_level = 'A') & theme(plot.tag = element_text(face="bold", size = 20))
+ggsave(filename = file.path(figure_path,"supfig-sensitivity.png"), plot = p, width = 16, height = 20)
+```
+
+## Cell Type Unbalanced/Connectivity
+```{r}
+data_source = "tabular_muris"
+tissue_name = "marrow_facs"
+dir.path <- dir.path0
+folder.path = file.path(dir.path,data_source)
+
+tiss <- readRDS(file.path(folder.path,paste0(tissue_name,".rds")))
+
+df_benchmark = read.table(file.path(dir.path,"benchmark", data_source, paste0(tissue_name,"_benchmark_rank_table.csv")),header = TRUE)
+rownames(df_benchmark) = df_benchmark$gene
+df_benchmark = df_benchmark[order(df_benchmark$lmd),]
+celldb_marker = read.table(file.path(folder.path,"ground_truth_geneset", paste0(tissue_name,"_ground_truth_c2.txt")))[,1]
+max_logfc = read.table(file.path(folder.path,"ground_truth_geneset", paste0(tissue_name,"_ground_truth_c1.txt"))) %>% rownames()
+
+dat_binary = tiss@assays$RNA@counts > 0
+
+gene_list = lapply(c(seq(100,2000,100),1277),function(x)df_benchmark$gene[1:x])
+names(gene_list) = paste0("lmd_top_",c(seq(100,2000,100),1277))
+max_logfc_list = lapply(c(seq(100,2000,100),1277),function(x) max_logfc[1:x])
+names(max_logfc_list) = paste0("maxfc_top_",c(seq(100,2000,100),1277))
+gene_list = c(gene_list,max_logfc_list,list("CellMarkerDB_1277_genes" = celldb_marker))
+
+cell_label = tiss$cell_ontology_class
+coldef = setNames(colorRampPalette(brewer.pal(12, "Paired"))(length(unique(cell_label))),unique(cell_label))
+
+df_prop = do.call(cbind,lapply(names(gene_list),function(top_gene){
+  genes = gene_list[[top_gene]]
+  nFeatures = colSums(dat_binary)
+  nFeatures_lmd = colSums(dat_binary[genes,])
+  prop = nFeatures_lmd/length(genes)
+  prop
+}))
+colnames(df_prop) = names(gene_list)
+tiss = AddMetaData(tiss,data.frame(df_prop))
+
+pl <- lapply(colnames(df_prop),function(top_gene){
+  # FeaturePlot(tiss, features = top_gene, order = TRUE) + scale_color_gradient(low = "lightgrey", high = "blue", limits = c(0,1))
+  Idents(tiss) = tiss$cell_ontology_class
+  VlnPlot(tiss, features = top_gene, cols = coldef) + ylim(0,1)
+})
+p = wrap_plots(pl, nrow = 5) & NoAxes()
+ggsave(filename = "1.png",p,width = 15,height = 15)
+
+p = (wrap_plots(pl, nrow = 5) & theme(legend.position = "bottom",axis.text.x = element_blank()) ) + plot_layout(guides = "collect")
+ggsave(filename = "1.png",p,width = 28, height = 12)
+
+VlnPlot(tiss, features = c("lmd_top_1300", "CellMarkerDB_1277_genes"), 
+        group.by = "cell_ontology_class", 
+        stack = TRUE) + xlab("Proportion") + 
+        theme(legend.position = "none") + 
+  coord_cartesian(xlim = c(0,0.45))
 ```

### 2025-03-24 .build_test/.build_site.R

@@ -29,7 +29,7 @@ dir_copy("articles", "docs/articles", overwrite = TRUE)
 # file.copy("articles/LMD_cross_comparison_demo.html", "docs/articles/LMD_cross_comparison_demo.html", overwrite = TRUE)
 
 pkgdown::build_site()
-
+devtools::check(vignettes = FALSE)
 
 # Install Package from local
 devtools::install() # install package

### 2025-03-24 LMD-paper/analysis/mouse_smoM2_example_process_from_10x.R

@@ -122,7 +122,7 @@ data_S <- data_S %>% NormalizeData() %>%
 data_S <- RunPCA(data_S, npcs = 30, verbose = F)
 # plot(data_S@reductions$pca@stdev)
 data_S <- RunUMAP(data_S, dims = 1:30)
-DimPlot(data_S, reduction = "umap", group.by = "orig.ident", shuffle = TRUE)
+DimPlot(data_S, reduction = "umap", group.by = c("orig.ident","Phase"), shuffle = TRUE)
 
 #####
 ##### Clustering
@@ -137,7 +137,7 @@ DimPlot(data_S, reduction = "umap", label = T, group.by = "cluster", split.by = 
 #####
 ##### Extract Dermal Cells
 #####
-FeaturePlot(data_S,label = TRUE,features = c("Dkk1","Dkk2","Lef1","Ptch1","Sox2","Sox18"),order = T,ncol = 3) & NoAxes()
+FeaturePlot(data_S,label = TRUE,features = c("Dkk1","Dkk2","Lef1","Ptch1","Sox2","Sox18", "Col1a1", "Col1a2", "Pdgfra"),order = T,ncol = 3) & NoAxes()
 data_S_dermal <- subset(data_S, cluster %in% c(0:3, 6:7))
 data_S_dermal <- data_S_dermal %>% NormalizeData() %>% 
   FindVariableFeatures() %>% ScaleData() %>%
@@ -145,8 +145,7 @@ data_S_dermal <- data_S_dermal %>% NormalizeData() %>%
   RunUMAP(dims = 1:30) %>% FindNeighbors(dims = 1:30) %>% 
   FindClusters(resolution = 0.3)
 data_S_dermal$cluster <- data_S_dermal$RNA_snn_res.0.3 #integrated
-DimPlot(data_S_dermal, reduction = "umap", label = T, group.by = "cluster")
-DimPlot(data_S_dermal, reduction = "umap", label = T, group.by = "orig.ident")
+DimPlot(data_S_dermal, reduction = "umap", label = T, group.by = c("cluster","orig.ident"))
 
 #####
 ##### Separate into different conditions & Save Results

### 2025-03-24 LMD-paper/benchmark/compare_method_azimuth.R

@@ -3,6 +3,7 @@ bioc_packages <- c("Seurat")
 sapply(cran_packages, function(pkg) if(!requireNamespace(pkg, quietly = TRUE)){install.packages(pkg)})
 sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
 lapply(c(cran_packages,bioc_packages), require, character.only = TRUE)
+source(file.path(dirname(rstudioapi::getActiveDocumentContext()$path),"run_methods_function.R"))
 
 # Download/Load Data ===========
 setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
@@ -11,7 +12,7 @@ source("../paths.R")
 dir.path <- dir.path0
 data_source = "azimuth"
 folder.path <- file.path(dir.path,data_source)
-dir.create(folder.path, recursive=T)
+if(!dir.exists(folder.path)){dir.create(folder.path, recursive=T)}
 tissue_download_link = c(
   "human_kidney" = "https://seurat.nygenome.org/azimuth/demo_datasets/kidney_demo_stewart.rds",
   "human_bone_marrow" = "https://seurat.nygenome.org/azimuth/demo_datasets/bmcite_demo.rds",
@@ -19,49 +20,77 @@ tissue_download_link = c(
   "mouse_motor_cortex" = "https://seurat.nygenome.org/azimuth/demo_datasets/allen_mop_2020.rds",
   "human_lung_hlca" = "https://seurat.nygenome.org/hlca_ref_files/ts_opt.rds",
   "human_motor_cortex" = "https://seurat.nygenome.org/azimuth/demo_datasets/allen_m1c_2019_ssv4.rds"
+  # "human_pbmc" = "https://www.dropbox.com/scl/fi/1fbur4p7law980iptl5zy/pbmc_10k_v3_filtered_feature_bc_matrix.h5?rlkey=3if8o9hxxisk62u7b2bqmcsr5&e=3&st=2go4s4gz&dl=1"
 )
 for(tissue_name in names(tissue_download_link)){
   file_name = paste0(tissue_name,".rds")
   if (!file.exists(file.path(folder.path,file_name))) {
-    options(timeout=6000)
-    download.file(tissue_download_link[tissue_name],
-                  destfile = file.path(folder.path,file_name), method = 'libcurl')
-    # update version
-    tiss <- readRDS(file.path(folder.path,file_name))
-    tiss <- UpdateSeuratObject(tiss)
[…]

### 2025-03-24 LMD-paper/benchmark/compare_method_smom2.R

@@ -1,0 +1,131 @@
+cran_packages <- c("readxl","dplyr","BiocManager","RColorBrewer")
+bioc_packages <- c("Seurat")
+sapply(cran_packages, function(pkg) if(!requireNamespace(pkg, quietly = TRUE)){install.packages(pkg)})
+sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
+lapply(c(cran_packages,bioc_packages), require, character.only = TRUE)
+source(file.path(dirname(rstudioapi::getActiveDocumentContext()$path),"run_methods_function.R"))
+
+# Download/Load Data ===========
+setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
+source("../paths.R")
+
+# Load data ===========
+dir.path <- dir.path0
+data_source = "smom2"
+folder.path <- file.path(dir.path,data_source)
+sample_ls = unlist(lapply(c("E13.5", "E14.5"),function(i) paste0(i,"_", c("MUT", "CTL"))))
+sample_ls = paste0("smom2_dermal_",sample_ls)
+
+data_S_ls = lapply(sample_ls, function(sample_name){
+  tiss <- readRDS(file.path(folder.path,"process_data",paste0("data_S_",sample_name,".rds")))
+  tiss
+})
+names(data_S_ls) = sample_ls
+
+# Run ===========
+method_ls = c("lmd","hvg", "wilcox_no_filter",
+              "haystack","hotspot",
+              "cosg","scmarker","semi")
+
+lapply(sample_ls, function(sample_name){
+  # Prepare input data --------
+  tiss = data_S_ls[[sample_name]]
+  DefaultAssay(tiss) <- "RNA"
+  cat(sample_name,":",FindPC(tiss),"PC\n")
+  dat = as.matrix(tiss[["RNA"]]@data)
+  Gene_detected_count <- apply(dat > apply(dat,2,median),1,sum)
+  selected_genes = names(Gene_detected_count)[(Gene_detected_count >= 10) & (Gene_detected_count <= ncol(dat) * 0.5)]
+  dat = dat[selected_genes,,drop = FALSE]
+  feature_space = Embeddings(tiss[["pca"]])[,1:FindPC(srat = tiss)]
+  
+  # Run Each Method and save results ========
+  dir.path <- dir.path0
+  folder.path = file.path(dir.path,"benchmark",data_source)
+  if(!file.exists(folder.path)){
+    dir.create(folder.path, recursive = T)
+  }
[…]

### 2025-03-24 LMD-paper/benchmark/compare_method_tabula_muris.R

@@ -3,6 +3,7 @@ bioc_packages <- c("Seurat")
 sapply(cran_packages, function(pkg) if(!requireNamespace(pkg, quietly = TRUE)){install.packages(pkg)})
 sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
 lapply(c(cran_packages,bioc_packages), require, character.only = TRUE)
+source(file.path(dirname(rstudioapi::getActiveDocumentContext()$path),"run_methods_function.R"))
 
 # Download/Load Data ===========
 setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
@@ -11,43 +12,71 @@ source("../paths.R")
 dir.path <- dir.path0
 data_source = "tabular_muris"
 folder.path <- file.path(dir.path,data_source)
-dir.create(folder.path, recursive=T)
+if(!dir.exists(folder.path)){
+  dir.create(folder.path, recursive=T)
+}
+
 tissue_download_link = c(
   "marrow_facs" = "https://figshare.com/ndownloader/files/13092380",
   "pancreas_facs" = "https://figshare.com/ndownloader/files/13092386",
   "lung_facs" = "https://figshare.com/ndownloader/files/13092194",
-  "marrow_droplet" = "https://figshare.com/ndownloader/files/ 13089821"
+  "marrow_droplet" = "https://figshare.com/ndownloader/files/13089821"
 )
-for(tissue_name in names(tissue_download_link)){
+tissue_ls = names(tissue_download_link)
+for(tissue_name in tissue_ls){
   file_name = paste0(tissue_name,".rds")
   if (!file.exists(file.path(folder.path,file_name))) {
     options(timeout=6000)
     download.file(tissue_download_link[tissue_name],
                   destfile = file.path(folder.path,file_name), method = 'libcurl')
     # update version
-    tiss <- readRDS(file.path(folder.path,file_name))
+    load(file.path(folder.path,file_name))
     tiss <- UpdateSeuratObject(tiss)
     saveRDS(tiss,file = file.path(folder.path,file_name))
   }
 }
 
-# Prepare input data ===========
-data_source = "tabular_muris"
-tissue_name = names(tissue_download_link)[1]
-
-dir.path <- dir.path0
-folder.path <- file.path(dir.path,data_source)
[…]

### 2025-03-24 LMD-paper/benchmark/graph_coarse_grain_test.R

@@ -1,0 +1,142 @@
+cran_packages <- c("readxl","dplyr","BiocManager")
+bioc_packages <- c("Seurat")
+sapply(cran_packages, function(pkg) if(!requireNamespace(pkg, quietly = TRUE)){install.packages(pkg)})
+sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
+if (!requireNamespace("GeneTrajectory", quietly = TRUE)) devtools::install_github("KlugerLab/GeneTrajectory")
+lapply(c(cran_packages,bioc_packages), require, character.only = TRUE)
+library(LocalizedMarkerDetector)
+library(GeneTrajectory)
+source(file.path(dirname(rstudioapi::getActiveDocumentContext()$path),"run_methods_function.R"))
+setwd("/banach1/ruiqi/local_marker/LocalizedMarkerDetector")
+devtools::load_all(".") 
+
+# mask the LocalizedMarkerDetector/GeneTrajectory function for testing purpose
+coarse.grain <- function (cell.embedding, gene.expression, graph.dist, N = 1000, 
+                          random.seed = 1) 
+{
+  message("Run k-means clustering")
+  set.seed(random.seed)
+  km.res <- stats::kmeans(cell.embedding, N, iter.max = 50)
+  message("Coarse-grain matrices")
+  KNN.membership.mat <- matrix(0, nrow = N, ncol = nrow(cell.embedding))
+  for (i in 1:ncol(KNN.membership.mat)) {
+    KNN.membership.mat[km.res$cluster[i], i] <- 1
+  }
+  KNN.membership.mat <- KNN.membership.mat/apply(KNN.membership.mat, 
+                                                 1, sum)
+  gene.expression.updated <- gene.expression %*% t(biclust::binarize(KNN.membership.mat, 0))
+  graph.dist.updated <- KNN.membership.mat %*% graph.dist %*%
+    t(KNN.membership.mat)
+  
+  colnames(graph.dist.updated) = rownames(graph.dist.updated) = paste0("cg",1:ncol(graph.dist.updated))
+  colnames(gene.expression.updated) = colnames(graph.dist.updated)
+  output <- list()
+  output[["gene.expression"]] <- gene.expression.updated
+  output[["graph"]] <- graph.dist.updated
+  output
+  return(output)
+}
+
[…]

### 2025-03-24 LMD-paper/benchmark/graph_disconnectivity_test.R

@@ -1,0 +1,241 @@
+cran_packages <- c("readxl","dplyr","BiocManager")
+bioc_packages <- c("Seurat")
+sapply(cran_packages, function(pkg) if(!requireNamespace(pkg, quietly = TRUE)){install.packages(pkg)})
+sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
+lapply(c(cran_packages,bioc_packages), require, character.only = TRUE)
+library(LocalizedMarkerDetector)
+source(file.path(dirname(rstudioapi::getActiveDocumentContext()$path),"run_methods_function.R"))
+
+# mask the LocalizedMarkerDetector function for testing purpose
+findDisconnectedComponents <- function(adj_matrix) {
+  # Create a graph from the adjacency matrix
+  g <- igraph::graph_from_adjacency_matrix(adj_matrix, mode = "undirected")
+  # Find the connected components
+  components <- igraph::components(g)
+  # Get the number of disconnected components
+  num_components <- components$no
+  # Get the corresponding nodes for each component
+  component_nodes <- split(igraph::V(g), components$membership)
+  list(number_of_components = num_components, components = component_nodes)
+}
+findShortestEdge <- function(component1, component2, data) {
+  # Calculate all pairwise distances between nodes in the two components
+  distances <- outer(component1, component2, Vectorize(function(x, y) dist(data[c(x,y), ])))
+  # Find the minimum distance and the corresponding nodes
+  min_distance_idx <- which(distances == min(distances), arr.ind = TRUE)
+  return(list(from = component1[min_distance_idx[1]], to = component2[min_distance_idx[2]], distance = min(distances)))
+}
+ConstructKnnGraph_from_embeddings <- function(knn = 5, feature_space, self_loop = 1){
+  cat("Constructing KNN graph\n")
+  knn_list <- FNN::get.knn(feature_space, k = knn, algorithm = "kd_tree")
+  Idx = knn_list$nn.index
+  A = Matrix::sparseMatrix(i = rep(1:nrow(Idx),each = knn),
+                           j = c(t(Idx)),
+                           dims = c(nrow(Idx), nrow(Idx)),
+                           x = 1)
[…]

### 2025-03-24 LMD-paper/benchmark/run_methods_function.R

@@ -4,11 +4,12 @@
 # Install singleCellHaystack: install.packages("singleCellHaystack")
 
 cran_packages <- c("reticulate","singleCellHaystack")
-github_packages <- c("immunogenomics/presto")
+github_packages <- c("immunogenomics/presto","KChen-lab/SCMarker")
 sapply(cran_packages, function(pkg) if(!requireNamespace(pkg, quietly = TRUE)){install.packages(pkg)})
-# sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
 sapply(github_packages, function(pkg) if (!requireNamespace(strsplit(pkg, "/")[[1]][2], quietly = TRUE)) remotes::install_github(pkg))
-lapply(c(cran_packages), require, character.only = TRUE)
+if (!requireNamespace("COSG", quietly = TRUE)) remotes::install_github("genecell/COSGR")
+lapply(c(cran_packages, "presto","COSG","SCMarker"), require, character.only = TRUE)
+
 
 #' LMD: Genes were ranked by the increasing order of Cumulative Diffusion-KL score
 library(LocalizedMarkerDetector)
@@ -145,11 +146,78 @@ hs_results = hs.compute_autocorrelations()
   write.table(marker,file = dir.file,row.names = TRUE)
 }
 
-#' Seurat v4 Wilcoxon unfiltered: Genes were ranked by their lowest adjusted P-value in any cluster (increasing order), if two genes had the same, prioritize genes with larger FC.
-RunSeuratv4 <- function(srat_obj, input_genes, feature_space, dir.file){
+#' ## SCMarker
+#' https://github.com/KChen-lab/SCMarker
+#' No rank is provided, only provide a set of genes
+RunSCMarker <- function(dat, feature_space, dir.file = NULL){
+  res = ModalFilter(data=dat,geneK=10,cellK=10,width=2)
+  res = GeneFilter(res)
+  res = getMarker(res, k = 300, n = 30)
+  marker = data.frame(gene = rownames(dat), row.names = rownames(dat))
+  marker[res$marker,"rank"] = 1
+  marker$rank[is.na(marker$rank)] = nrow(marker)
+  
+  if(is.null(dir.file)){return(marker)}else{
+    write.table(marker,file = dir.file,row.names = TRUE)
+  }
+}
+
+#' ## COSG
[…]

### 2025-03-24 LMD-paper/benchmark/sensitivity_test.R

@@ -3,6 +3,11 @@ bioc_packages <- c("Seurat")
 sapply(cran_packages, function(pkg) if(!requireNamespace(pkg, quietly = TRUE)){install.packages(pkg)})
 sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
 lapply(c(cran_packages,bioc_packages), require, character.only = TRUE)
+library(LocalizedMarkerDetector)
+source(file.path(dirname(rstudioapi::getActiveDocumentContext()$path),"run_methods_function.R"))
+
+setwd("/banach1/ruiqi/local_marker/LocalizedMarkerDetector")
+devtools::load_all(".")
 
 # Load Data =======
 setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
@@ -10,7 +15,7 @@ source("../paths.R")
 
 dir.path <- dir.path0
 data_source = "tabular_muris"
-tissue_name = "marrow_facs"
+tissue_name = "marrow_facs" # pancreas_facs, marrow_droplet
 tiss <- readRDS(file.path(dir.path,data_source,paste0(tissue_name,".rds")))
 DefaultAssay(tiss) <- "RNA"
 dat = as.matrix(tiss[[DefaultAssay(tiss)]]@data)
@@ -19,12 +24,20 @@ selected_genes = (Gene_detected_count >= 10) & (Gene_detected_count <= ncol(dat)
 selected_genes = names(selected_genes)[selected_genes]
 dat = dat[selected_genes,,drop = FALSE]
 
-
-# kNN sensitivity Test --------
+# Sensitivity Test ---------
 folder.path = file.path(dir.path,data_source,"intermediate_data",
                         tissue_name, "sensitivity_test")
-dir.create(folder.path,recursive = T)
-library(LocalizedMarkerDetector)
+if(!dir.exists(folder.path)){
+  dir.create(folder.path,recursive = T)
+}
+test_para_ls = c("pc","knn","cellprop","umi",
+                 "binary","graph","ambient")
+base_ls = c("ori_20PC","kNN5","cell_prop_1_1",
+            "umi_prop_1_1","original","kNN","original")
+names(base_ls) = test_para_ls
+
+## kNN --------
+test_para = test_para_ls[2]
 # Use Default PCs
 n_dim = 20
 feature_space = Embeddings(tiss[["pca"]])[,1:n_dim]
[…]

### 2025-03-24 R/calculate_lmds.R

@@ -164,7 +164,7 @@ fast_calculate_score_profile_largeData <- function(W, max_time = 2^15, init_stat
       break
     }
     # Check the sparsity of P, if too dense, transfer it to dense matrix
-    if(is_sparse_matrix(P) && (min(rowSums(P == 0)/ncol(P)) < 0.9)){
+    if(is_sparse_matrix(P) && (min(Matrix::rowSums(P == 0)/ncol(P)) < 0.9)){
       P = as.matrix(P)
     }
     # cat("Graph dyadic\n")
@@ -369,12 +369,18 @@ fast_get_lmds <- function(W, max_time = 2^15, init_state, P_ls = NULL, correctio
 #' @export
 LMD <- function(expression, feature_space, knn = 5, 
                 kernel = FALSE, max_time = 2^20, adjust_bridge = TRUE, self_loop = 1,
-                score_correction = FALSE, largeData = TRUE, highres = FALSE, min_cell = 5){
+                score_correction = FALSE, largeData = TRUE, highres = FALSE, min_cell = 5, kernel_used = NULL){
   if(any(colnames(expression) != rownames(feature_space))){stop("Cells in expression mtx and feature space don't match.")}
-  if(kernel){
-    W = ConstructGaussianGraph(knn = knn, feature_space = feature_space, alpha = 1, coef = 2, epsilon = 1e-3, self_loop = self_loop)$'graph'
+  if(!kernel){
+    W = ConstructKnnGraph(knn = knn, feature_space = feature_space, adjust_disconnection = adjust_bridge, self_loop = self_loop)$'graph'
   }else{
-    W = ConstructKnnGraph(knn = knn, feature_space = feature_space, adjust_by_MST = adjust_bridge, self_loop = self_loop)$'graph'
+    if(kernel_used == "SNN"){
+      W = ConstructSNNGraph(knn = knn, feature_space = feature_space, self_loop = self_loop)$'graph'
+    }else if(kernel_used == "Gaussian"){
+      W = ConstructGaussianGraph(knn = knn, feature_space = feature_space, alpha = 1, coef = 2, epsilon = 1e-3, self_loop = self_loop)$'graph'
+    }else{
+      stop("Please provide graph type: SNN or Gaussian")
+    }
   }
   rho = RowwiseNormalize(expression)
   rho = rho[,colnames(W),drop = FALSE]
[…]

### 2025-03-24 R/construct_cell_graph.R

@@ -19,7 +19,7 @@ findShortestEdge <- function(component1, component2, data) {
   distances <- outer(component1, component2, Vectorize(function(x, y) dist(data[c(x,y), ])))
   # Find the minimum distance and the corresponding nodes
   min_distance_idx <- which(distances == min(distances), arr.ind = TRUE)
-  return(c(component1[min_distance_idx[1]], component2[min_distance_idx[2]]))
+  return(list(from = component1[min_distance_idx[1]], to = component2[min_distance_idx[2]], distance = min(distances)))
 }
 
 #' Construct a Symmetric KNN graph
@@ -28,7 +28,7 @@ findShortestEdge <- function(component1, component2, data) {
 #'
 #' @param knn integer; the number of nearest neighbors to consider for constructing the graph. Default is 5.
 #' @param feature_space matrix; a cell-by-coordinate matrix (e.g., 20 principal components).
-#' @param adjust_by_MST boolean; TRUE for connecting disconnected components using Minimum Spanning Trees (MST). Default is TRUE.
+#' @param adjust_disconnection boolean; TRUE for fully connecting disconnected components. Default is TRUE.
 #' @param self_loop integer; weight for self connections (default is 1).
 #'
 #' @return A list containing the following elements:
@@ -45,7 +45,8 @@ findShortestEdge <- function(component1, component2, data) {
 #' adj_matrix <- result$adj_matrix
 #' 
 #' @export
-ConstructKnnGraph <- function(knn = 5, feature_space, adjust_by_MST = TRUE, self_loop = 1){
+ConstructKnnGraph <- function(knn = 5, feature_space, 
+                              adjust_disconnection = TRUE, self_loop = 1){
   cat("Constructing KNN graph\n")
   knn_list <- FNN::get.knn(feature_space, k = knn, algorithm = "kd_tree")
   Idx = knn_list$nn.index
@@ -62,17 +63,41 @@ ConstructKnnGraph <- function(knn = 5, feature_space, adjust_by_MST = TRUE, self
   filtered_node_names <- unlist(lapply(filtered_components, function(comp) names(comp)))
   A = A[rownames(A)%in%filtered_node_names,
         colnames(A)%in%filtered_node_names]
+  if(dim(A)[1] < nrow(feature_space)){
+    cat("Remove ", nrow(feature_space) - dim(A)[1], " singleton cells.\n")
+  }
   
[…]

### 2025-04-21 README.Rmd
add_citation_adjust_figure
250421

@@
 ---
-
-<!-- README.md is generated from README.Rmd. Please edit that file -->
 
@@
 
-```{r pressure, echo=FALSE, out.width="50%", out.height="auto"}
+```{r pressure, echo=FALSE, out.width="100%", out.height="auto"}
 knitr::include_graphics("./man/figures/LMD_workflow.png")
@@
 
+### Docker installation
+
+Alternatively, we provide an option of installing LocalizedMarkerDetector with the accompanied lmd docker environment. The pre-built docker image can be downloaded using
+
+``` bash
+docker pull ruiqi0130/lmd:lastest
+```
+
+and to run the docker container:
+
+``` bash
+docker run -it --rm -e USER=john -e USERID=$(id -u) --name container_name -p 8029:8787  -v /data/:/data/   docker.io/jyc34/bipca-experiment:lastest
+```
+
+Here, change `/data:/data to `<your_local_data_directory>:/data`.
+A rstudio will be on port 8029.
+If you would like to link a local bipca installation (for package development, for instance), you can write -v <your_local_bipca_directory:/bipca.
+
 ## Example tutorial

### 2025-04-21 .build_test/.build_site.R

@@ -3,19 +3,25 @@ library(pkgdown)
 
 setwd("/banach1/ruiqi/local_marker/LocalizedMarkerDetector")
 
+# Create docker image
+renv::init()
+renv::snapshot()
+
 # Create Template
 use_readme_rmd()
 # use_news_md()
-use_vignette("Example")
+use_vignette("Example") # create a vignette template called Example
 
 # Build Package
 usethis::use_mit_license("Ruiqi Li")
+# renders README.Rmd -> README.md
 devtools::build_readme()
+# Generates the RcppExports.R and updates NAMESPACE with exported C++ functions
 Rcpp::compileAttributes()
-devtools::document()
+# Generates documentation (.Rd files) from Roxygen comments and updates the NAMESPACE file accordingly.
+devtools::document() 
 # load all function w/o install
 devtools::load_all(".") 
-
 
 # Render
 # rmarkdown::render("vignettes/LMD_demo.Rmd")
@@ -29,6 +35,7 @@ dir_copy("articles", "docs/articles", overwrite = TRUE)
 # file.copy("articles/LMD_cross_comparison_demo.html", "docs/articles/LMD_cross_comparison_demo.html", overwrite = TRUE)
 
 pkgdown::build_site()
+
 devtools::check(vignettes = FALSE)
 
 # Install Package from local

### 2025-04-21 LMD-paper/analysis/mouse_bone_marrow_example.R

@@ -1,266 +1,0 @@
-setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
-source("../paths.R")
-library(LocalizedMarkerDetector)
-# Define package lists
-bioc_packages <- c("clusterProfiler", "AnnotationDbi", "ReactomePA", "org.Mm.eg.db","gprofiler2", "msigdbr")
-github_packages <- c("satijalab/seurat-wrappers");
-names(github_packages) = strsplit(github_packages, "/")[[1]][2]
-names(github_packages)[grepl("seurat-wrappers",names(github_packages))] = "SeuratWrappers"
-# Install packages
-sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
-sapply(1:length(github_packages), function(i) if (!requireNamespace(names(github_packages)[i], quietly = TRUE)) remotes::install_github(github_packages[i]))
-lapply(c(bioc_packages,names(github_packages)), require, character.only = TRUE)
-
-# Download/Load Data ===========
-dir.path <- dir.path0
-folder.path <- file.path(dir.path,"tabular_muris")
-dir.create(folder.path, recursive=T)
-tissue_download_link = c(
-  "marrow_facs" = "https://figshare.com/ndownloader/files/13092380",
-  "marrow_droplet" = "https://figshare.com/ndownloader/files/ 13089821"
-)
-for(tissue_name in names(tissue_download_link)){
-  file_name = paste0(tissue_name,".rds")
-  if (!file.exists(file.path(folder.path,file_name))) {
-    options(timeout=6000)
-    download.file(tissue_download_link[tissue_name],
-                  destfile = file.path(folder.path,file_name), method = 'libcurl')
-    # update version
-    tiss <- readRDS(file.path(folder.path,file_name))
-    tiss <- UpdateSeuratObject(tiss)
-    tiss <- RunALRA(tiss, assay = "RNA"); DefaultAssay(tiss) = "RNA"
-    saveRDS(tiss,file = file.path(folder.path,file_name))
-  }
-}
-
-# Prepare input data ===========
-tissue_name = "marrow_facs"
-tiss <- readRDS(file.path(folder.path,paste0(tissue_name,".rds")))
-DefaultAssay(tiss) <- "RNA"
-n_dim = 20
[…]

### 2025-04-21 LMD-paper/analysis/mouse_smoM2_example.R

@@ -1,430 +1,0 @@
-setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
-source("../paths.R")
-dir.path = file.path(dir.path0,"smom2")
-library(LocalizedMarkerDetector)
-
-# Define package lists
-bioc_packages <- c("clusterProfiler", "AnnotationDbi", "ReactomePA", "org.Mm.eg.db",
-                   "gprofiler2", "msigdbr")
-github_packages <- c("satijalab/seurat-wrappers");
-names(github_packages) = strsplit(github_packages, "/")[[1]][2]
-names(github_packages)[grepl("seurat-wrappers",names(github_packages))] = "SeuratWrappers"
-# Install packages
-sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
-sapply(1:length(github_packages), function(i) if (!requireNamespace(names(github_packages)[i], quietly = TRUE)) remotes::install_github(github_packages[i]))
-lapply(c(bioc_packages,names(github_packages)), require, character.only = TRUE)
-
-# Load Data --------------------
-folder.path <- file.path(dir.path,"raw_data","preprocess_data")
-sample_ls = unlist(lapply(c("E13.5", "E14.5"),function(i) paste0(i,"_", c("MUT", "CTL"))))
-sample_ls = paste0("smom2_dermal_",sample_ls)
-for(i in sample_ls){
-  assign(paste0("data_S_",i), 
-         readRDS(file.path(folder.path,sprintf("data_S_%s_v0.rds",i))))
-}
-# Process ------------
-folder.path <- file.path(dir.path,"process_data")
-dir.create(folder.path, recursive=T)
-lapply(sample_ls, function(sample_name){
-  data_S <- get(paste0("data_S_",sample_name))
-  data_S <- data_S %>% NormalizeData() %>% FindVariableFeatures() %>%
-    ScaleData(verbose = FALSE) %>% RunPCA(npcs = 50, verbose = FALSE)
-  min.pc = FindPC(srat = data_S)
-  cat("# of PC used: ", min.pc, "\n")
-  data_S <- RunUMAP(data_S, dims = 1:min.pc, seed.use = 42)
-  data_S <- data_S %>% FindNeighbors() %>% FindClusters(res = 1.5)
-  assign(paste0("data_S_",sample_name),data_S, envir = .GlobalEnv)
-})
-
[…]

### 2025-04-21 LMD-paper/benchmark/compare_method_azimuth.R

@@ -1,320 +1,0 @@
-cran_packages <- c("readxl","dplyr","BiocManager","RColorBrewer")
-bioc_packages <- c("Seurat")
-sapply(cran_packages, function(pkg) if(!requireNamespace(pkg, quietly = TRUE)){install.packages(pkg)})
-sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
-lapply(c(cran_packages,bioc_packages), require, character.only = TRUE)
-source(file.path(dirname(rstudioapi::getActiveDocumentContext()$path),"run_methods_function.R"))
-
-# Download/Load Data ===========
-setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
-source("../paths.R")
-
-dir.path <- dir.path0
-data_source = "azimuth"
-folder.path <- file.path(dir.path,data_source)
-if(!dir.exists(folder.path)){dir.create(folder.path, recursive=T)}
-tissue_download_link = c(
-  "human_kidney" = "https://seurat.nygenome.org/azimuth/demo_datasets/kidney_demo_stewart.rds",
-  "human_bone_marrow" = "https://seurat.nygenome.org/azimuth/demo_datasets/bmcite_demo.rds",
-  "human_pancreas" = "https://seurat.nygenome.org/azimuth/demo_datasets/enge.rds",
-  "mouse_motor_cortex" = "https://seurat.nygenome.org/azimuth/demo_datasets/allen_mop_2020.rds",
-  "human_lung_hlca" = "https://seurat.nygenome.org/hlca_ref_files/ts_opt.rds",
-  "human_motor_cortex" = "https://seurat.nygenome.org/azimuth/demo_datasets/allen_m1c_2019_ssv4.rds"
-  # "human_pbmc" = "https://www.dropbox.com/scl/fi/1fbur4p7law980iptl5zy/pbmc_10k_v3_filtered_feature_bc_matrix.h5?rlkey=3if8o9hxxisk62u7b2bqmcsr5&e=3&st=2go4s4gz&dl=1"
-)
-for(tissue_name in names(tissue_download_link)){
-  file_name = paste0(tissue_name,".rds")
-  if (!file.exists(file.path(folder.path,file_name))) {
-    if(grepl(".rds",tissue_download_link[tissue_name])){
-      options(timeout=6000)
-      download.file(tissue_download_link[tissue_name],
-                    destfile = file.path(folder.path,file_name), method = 'libcurl')
-      # update version
[…]

### 2025-04-21 LMD-paper/benchmark/compare_method_smom2.R

@@ -1,131 +1,0 @@
-cran_packages <- c("readxl","dplyr","BiocManager","RColorBrewer")
-bioc_packages <- c("Seurat")
-sapply(cran_packages, function(pkg) if(!requireNamespace(pkg, quietly = TRUE)){install.packages(pkg)})
-sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
-lapply(c(cran_packages,bioc_packages), require, character.only = TRUE)
-source(file.path(dirname(rstudioapi::getActiveDocumentContext()$path),"run_methods_function.R"))
-
-# Download/Load Data ===========
-setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
-source("../paths.R")
-
-# Load data ===========
-dir.path <- dir.path0
-data_source = "smom2"
-folder.path <- file.path(dir.path,data_source)
-sample_ls = unlist(lapply(c("E13.5", "E14.5"),function(i) paste0(i,"_", c("MUT", "CTL"))))
-sample_ls = paste0("smom2_dermal_",sample_ls)
-
-data_S_ls = lapply(sample_ls, function(sample_name){
-  tiss <- readRDS(file.path(folder.path,"process_data",paste0("data_S_",sample_name,".rds")))
-  tiss
-})
-names(data_S_ls) = sample_ls
-
-# Run ===========
-method_ls = c("lmd","hvg", "wilcox_no_filter",
-              "haystack","hotspot",
-              "cosg","scmarker","semi")
-
-lapply(sample_ls, function(sample_name){
-  # Prepare input data --------
-  tiss = data_S_ls[[sample_name]]
-  DefaultAssay(tiss) <- "RNA"
-  cat(sample_name,":",FindPC(tiss),"PC\n")
-  dat = as.matrix(tiss[["RNA"]]@data)
-  Gene_detected_count <- apply(dat > apply(dat,2,median),1,sum)
-  selected_genes = names(Gene_detected_count)[(Gene_detected_count >= 10) & (Gene_detected_count <= ncol(dat) * 0.5)]
-  dat = dat[selected_genes,,drop = FALSE]
-  feature_space = Embeddings(tiss[["pca"]])[,1:FindPC(srat = tiss)]
-  
-  # Run Each Method and save results ========
-  dir.path <- dir.path0
-  folder.path = file.path(dir.path,"benchmark",data_source)
-  if(!file.exists(folder.path)){
-    dir.create(folder.path, recursive = T)
-  }
[…]

### 2025-04-21 LMD-paper/benchmark/graph_disconnectivity_test.R

@@ -1,241 +1,0 @@
-cran_packages <- c("readxl","dplyr","BiocManager")
-bioc_packages <- c("Seurat")
-sapply(cran_packages, function(pkg) if(!requireNamespace(pkg, quietly = TRUE)){install.packages(pkg)})
-sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
-lapply(c(cran_packages,bioc_packages), require, character.only = TRUE)
-library(LocalizedMarkerDetector)
-source(file.path(dirname(rstudioapi::getActiveDocumentContext()$path),"run_methods_function.R"))
-
-# mask the LocalizedMarkerDetector function for testing purpose
-findDisconnectedComponents <- function(adj_matrix) {
-  # Create a graph from the adjacency matrix
-  g <- igraph::graph_from_adjacency_matrix(adj_matrix, mode = "undirected")
-  # Find the connected components
-  components <- igraph::components(g)
-  # Get the number of disconnected components
-  num_components <- components$no
-  # Get the corresponding nodes for each component
-  component_nodes <- split(igraph::V(g), components$membership)
-  list(number_of_components = num_components, components = component_nodes)
-}
-findShortestEdge <- function(component1, component2, data) {
-  # Calculate all pairwise distances between nodes in the two components
-  distances <- outer(component1, component2, Vectorize(function(x, y) dist(data[c(x,y), ])))
-  # Find the minimum distance and the corresponding nodes
-  min_distance_idx <- which(distances == min(distances), arr.ind = TRUE)
-  return(list(from = component1[min_distance_idx[1]], to = component2[min_distance_idx[2]], distance = min(distances)))
-}
-ConstructKnnGraph_from_embeddings <- function(knn = 5, feature_space, self_loop = 1){
-  cat("Constructing KNN graph\n")
-  knn_list <- FNN::get.knn(feature_space, k = knn, algorithm = "kd_tree")
-  Idx = knn_list$nn.index
-  A = Matrix::sparseMatrix(i = rep(1:nrow(Idx),each = knn),
-                           j = c(t(Idx)),
-                           dims = c(nrow(Idx), nrow(Idx)),
-                           x = 1)
[…]

### 2025-04-21 LMD-paper/benchmark/run_methods_function.R

@@ -1,339 +1,0 @@
-# Install Marcopolo: pip install marcopolo-pytorch --upgrade
-# Install SEMITONES: pip install https://github.com/ohlerlab/SEMITONES/archive/master.zip
-# Install Hotspot: pip install git+https://github.com/yoseflab/Hotspot.git
-# Install singleCellHaystack: install.packages("singleCellHaystack")
-
-cran_packages <- c("reticulate","singleCellHaystack")
-github_packages <- c("immunogenomics/presto","KChen-lab/SCMarker")
-sapply(cran_packages, function(pkg) if(!requireNamespace(pkg, quietly = TRUE)){install.packages(pkg)})
-sapply(github_packages, function(pkg) if (!requireNamespace(strsplit(pkg, "/")[[1]][2], quietly = TRUE)) remotes::install_github(pkg))
-if (!requireNamespace("COSG", quietly = TRUE)) remotes::install_github("genecell/COSGR")
-lapply(c(cran_packages, "presto","COSG","SCMarker"), require, character.only = TRUE)
-
-
-#' LMD: Genes were ranked by the increasing order of Cumulative Diffusion-KL score
-library(LocalizedMarkerDetector)
-RunLMD <- function(dat, feature_space, dir.file){
-  res = LMD(dat,feature_space,max_time = 2^20,knn = 5)
-  marker = show_result_lmd(res)$'gene_table'
-  write.table(marker,file = dir.file,row.names = TRUE)
-}
-
-#' HVG
-RunHVG <- function(srat_obj, input_genes, dir.file){
-  DefaultAssay(srat_obj) <- "RNA"
-  marker <- subset(srat_obj,features = input_genes) %>% ScaleData(features = input_genes) %>% 
-    FindVariableFeatures(nfeatures = length(input_genes)) %>% VariableFeatures() %>% as.data.frame() %>% setNames(c("gene")) %>% mutate(rank = 1:n())
-  write.table(marker,file = dir.file,row.names = FALSE)
-}
-
-#' SinglecellHaystack: 
-#' https://alexisvdb.github.io/singleCellHaystack/articles/examples/a02_example_scRNAseq.html
-#' Genes were ranked based on the increasing order of log(adjusted P-value)
-RunHaystack <- function(dat, feature_space, dir.file = NULL){
-  set.seed(123)
-  res <- haystack(x = feature_space, expression = dat)
-  marker <- show_result_haystack(res.haystack = res)
[…]

### 2025-04-21 LMD-paper/benchmark/sensitivity_test.R

@@ -1,294 +1,0 @@
-cran_packages <- c("readxl","dplyr","BiocManager")
-bioc_packages <- c("Seurat")
-sapply(cran_packages, function(pkg) if(!requireNamespace(pkg, quietly = TRUE)){install.packages(pkg)})
-sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
-lapply(c(cran_packages,bioc_packages), require, character.only = TRUE)
-library(LocalizedMarkerDetector)
-source(file.path(dirname(rstudioapi::getActiveDocumentContext()$path),"run_methods_function.R"))
-
-setwd("/banach1/ruiqi/local_marker/LocalizedMarkerDetector")
-devtools::load_all(".")
-
-# Load Data =======
-setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
-source("../paths.R")
-
-dir.path <- dir.path0
-data_source = "tabular_muris"
-tissue_name = "marrow_facs" # pancreas_facs, marrow_droplet
-tiss <- readRDS(file.path(dir.path,data_source,paste0(tissue_name,".rds")))
-DefaultAssay(tiss) <- "RNA"
-dat = as.matrix(tiss[[DefaultAssay(tiss)]]@data)
-Gene_detected_count <- apply(dat > apply(dat,2,median),1,sum)
-selected_genes = (Gene_detected_count >= 10) & (Gene_detected_count <= ncol(dat) * 0.5)
-selected_genes = names(selected_genes)[selected_genes]
-dat = dat[selected_genes,,drop = FALSE]
-
-# Sensitivity Test ---------
-folder.path = file.path(dir.path,data_source,"intermediate_data",
-                        tissue_name, "sensitivity_test")
-if(!dir.exists(folder.path)){
-  dir.create(folder.path,recursive = T)
-}
-test_para_ls = c("pc","knn","cellprop","umi",
-                 "binary","graph","ambient")
-base_ls = c("ori_20PC","kNN5","cell_prop_1_1",
-            "umi_prop_1_1","original","kNN","original")
-names(base_ls) = test_para_ls
-
-## kNN --------
-test_para = test_para_ls[2]
-# Use Default PCs
-n_dim = 20
-feature_space = Embeddings(tiss[["pca"]])[,1:n_dim]
[…]

### 2025-04-21 LMD-paper/paper_figure_function.R

@@ -1,71 +1,0 @@
-add_newline_in_middle <- function(input_string, max_length) {
-  if (nchar(input_string) > max_length) {
-    # Find the approximate middle
-    middle <- ceiling(nchar(input_string) / 2)
-    
-    # Find the closest space to the middle
-    space_before <- regexpr(" ", substring(input_string, 1, middle), fixed = TRUE)
-    space_after <- regexpr(" ", substring(input_string, middle + 1), fixed = TRUE)
-    
-    if (space_before == -1 && space_after == -1) {
-      # No spaces found, cannot split without breaking words
-      return(input_string)
-    }
-    
-    if (space_before == -1) {
-      split_pos <- middle + space_after
-    } else if (space_after == -1) {
-      split_pos <- space_before
-    } else {
-      split_pos <- ifelse(middle - space_before < space_after, space_before, middle + space_after)
-    }
-    
-    # Insert newline at the split position
-    output_string <- paste0(
-      substring(input_string, 1, split_pos), "\n", substring(input_string, split_pos + 1)
-    )
-  } else {
-    output_string <- input_string
-  }
-  
-  return(output_string)
-}
-
-Visualize_jaccard_mtx <- function(jaccard_mtx, threshold = 0.4){
-  if(is.null(rownames(jaccard_mtx))|is.null(colnames(jaccard_mtx))){
-    stop("Please define celltype names & module names")
-  }
-  df = reshape2::melt(jaccard_mtx)
-  colnames(df)[1:2] = c("CellType","Module")
-  df$CellType = factor(df$CellType,levels = rownames(jaccard_mtx))
-  df$Module = factor(df$Module,levels = rev(colnames(jaccard_mtx)))
-  
-  p <- ggplot(df, aes(x=CellType, y=Module, fill=value)) +
-    geom_tile(color="white") +
-    scale_fill_gradientn(colors=colorRampPalette(rev(c("firebrick3", "white")))(99),limits = c(0,1), name="Jaccard Index") +
-    theme_minimal() +
-    theme(
-      axis.text.x = element_text(angle=45, hjust=1,size = 15),
-      axis.text.y = element_text(size = 15),
-      legend.title = element_text(size = 20),
-      legend.text = element_text(size = 15),
-      axis.title = element_blank(),
-      panel.grid.major = element_blank(),
-      panel.grid.minor = element_blank()
[…]

### 2025-04-21 LMD-paper/paths.R

@@ -1,1 +1,0 @@
-dir.path0 <- "/banach1/ruiqi/local_marker/LMD_data"

### 2025-04-21 R/calculate_lmds.R

@@ -360,28 +360,46 @@ fast_get_lmds <- function(W, max_time = 2^15, init_state, P_ls = NULL, correctio
 #' @param largeData logical; if TRUE, uses functions optimized for large matrix multiplication. Default is TRUE.
 #' @param highres logical; if TRUE, uses fine time scales (0, 1, 2, 3, ...). If FALSE, uses dyadic time scales (0, 2, 4, ...). Default is FALSE.
 #' @param min_cell integer; removes genes expressing in fewer than this number of cells. Default is 5.
-#'
+#' @param kernel_used character; kernel type used when \code{kernel = TRUE}. Options are \code{"Gaussian"} or \code{"SNN"}.
+#' @param alpha numeric; exponent used in the alpha-decaying Gaussian kernel
+#' @param coef numeric; coefficient for adaptive bandwidth. Default is 1.
+#' @param epsilon numeric; threshold below which edge weights are set to zero to sparsify the graph. Default is 1e-3.
+#' @param graph.affinity matrix; user-supplied cell-cell affinity matrix. If provided, \code{feature_space} and graph construction parameters are ignored.
+#' 
+#' 
 #' @return A list containing:
 #' \item{score_profile}{data frame; the computed score profile for different diffusion times.}
 #' \item{cumulative_score}{numeric vector; the LMD score for each gene.}
 #'
 #'
 #' @export
-LMD <- function(expression, feature_space, knn = 5, 
+LMD <- function(expression, feature_space = NULL, knn = 5, 
                 kernel = FALSE, max_time = 2^20, adjust_bridge = TRUE, self_loop = 1,
-                score_correction = FALSE, largeData = TRUE, highres = FALSE, min_cell = 5, kernel_used = NULL){
-  if(any(colnames(expression) != rownames(feature_space))){stop("Cells in expression mtx and feature space don't match.")}
-  if(!kernel){
-    W = ConstructKnnGraph(knn = knn, feature_space = feature_space, adjust_disconnection = adjust_bridge, self_loop = self_loop)$'graph'
-  }else{
-    if(kernel_used == "SNN"){
-      W = ConstructSNNGraph(knn = knn, feature_space = feature_space, self_loop = self_loop)$'graph'
-    }else if(kernel_used == "Gaussian"){
-      W = ConstructGaussianGraph(knn = knn, feature_space = feature_space, alpha = 1, coef = 2, epsilon = 1e-3, self_loop = self_loop)$'graph'
[…]

### 2025-04-23 README.Rmd
250422

--- 2024-08-11 README.Rmd
@@
 install.packages("devtools")
-devtools::install_github("ruiqi0130/LocalizedMarkerDetector")
+devtools::install_github("KlugerLab/LocalizedMarkerDetector")
 
@@
 ## Example tutorial
-Please check [LMD tutorial](https://ruiqi0130.github.io/LocalizedMarkerDetector/articles/).
+Please check [LMD tutorial](https://KlugerLab.github.io/LocalizedMarkerDetector/articles/).
 
 ## References
-References of LMD functions can be found [here](https://ruiqi0130.github.io/LocalizedMarkerDetector/reference/index.html).
+References of LMD functions can be found [here](https://KlugerLab.github.io/LocalizedMarkerDetector/reference/index.html).
 

### 2025-04-23 .build_test/.build_site.R

@@ -1,46 +1,0 @@
-library(devtools)
-library(pkgdown)
-
-setwd("/banach1/ruiqi/local_marker/LocalizedMarkerDetector")
-
-# Create docker image
-renv::init()
-renv::snapshot()
-
-# Create Template
-use_readme_rmd()
-# use_news_md()
-use_vignette("Example") # create a vignette template called Example
-
-# Build Package
-usethis::use_mit_license("Ruiqi Li")
-# renders README.Rmd -> README.md
-devtools::build_readme()
-# Generates the RcppExports.R and updates NAMESPACE with exported C++ functions
-Rcpp::compileAttributes()
-# Generates documentation (.Rd files) from Roxygen comments and updates the NAMESPACE file accordingly.
-devtools::document() 
-# load all function w/o install
-devtools::load_all(".") 
-
-# Render
-# rmarkdown::render("vignettes/LMD_demo.Rmd")
-pkgdown::clean_site()
-
-# if (!dir.exists("docs/articles")) {
-#   dir.create("docs/articles", recursive = TRUE)
-# }
-library(fs)
-dir_copy("articles", "docs/articles", overwrite = TRUE)
-# file.copy("articles/LMD_cross_comparison_demo.html", "docs/articles/LMD_cross_comparison_demo.html", overwrite = TRUE)
-
-pkgdown::build_site()
-
-devtools::check(vignettes = FALSE)
-
-# Install Package from local
-devtools::install() # install package
-
-# Install Package from github
-devtools::install_github("ruiqi0130/LocalizedMarkerDetector")
-

### 2025-04-29 README.Rmd
update readme

@@
 
+## Docker Installation
+
+Alternatively, we provide a Docker environment with LMD and all its dependencies pre-installed. The pre-built Docker image can be downloaded using:
+
+``` bash
+docker pull ruiqi0130/lmd-rstudio:lastest
+```
+
+and to run the docker container:
+
+``` bash
+docker run -it --rm --name lmd -p 8029:8787  -v /data/:/data/   docker.io/ruiqi0130/lmd-rstudio:latest
+```
+
+Here, change `/data:/data` to `<your_local_data_directory>:/data`. A rstudio will be on port 8029.
+
 ## Example tutorial
